### 2019-12-08 presentations/2019-12-09/presentation.tex
Preparing presentation for tomorrow 1

\documentclass[10pt]{beamer}

%packages
\usepackage{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,tabu}
\usepackage{color}
\usepackage{tikz}
\usetikzlibrary{matrix,chains,positioning,decorations.pathreplacing,arrows}
\usetikzlibrary{fit,shapes}
\usetikzlibrary{calc,shadings}
\usepackage{pgfplots}
\usepackage{colortbl}
\usepackage{eurosym}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{tabularx}

%definitions
\usepackage{algorithm,algorithmic}

%theme
\usetheme{Dresden}
\usecolortheme{rose}
\useoutertheme{tree}

%environments
\newenvironment{ExampleGer}{\begin{exampleblock}{Beispiel}}{\end{exampleblock}}

\newenvironment{customlegend}[1][]{%
	\begingroup
	% inits/clears the lists (which might be populated from previous
	% axes):
	\csname pgfplots@init@cleared@structures\endcsname
	\pgfplotsset{#1}%
}{%
	% draws the legend:
	\csname pgfplots@createlegend\endcsname
	\endgroup
}%

%definitions
\def\addlegendimage{\csname pgfplots@addlegendimage\endcsname}
% definition to insert numbers
\pgfkeys{/pgfplots/number in legend/.style={%
		/pgfplots/legend image code/.code={%
			\node at (0.295,-0.0225){#1};
		},%
	},
}

%general
\title{Combining bottom up and top down approaches for Drug-Target-Interaction prediction}
\author{Tilman Hinnerichs}
\institute{BORG - KAUST}
\date{December 09, 2019}

%presentation
\expandafter\def\expandafter\insertshorttitle\expandafter{%
	\insertshorttitle\hspace{5cm}%
	\insertframenumber\,/\,\inserttotalframenumber
}

\definecolor{myblue}{RGB}{80,80,160}
\definecolor{mygreen}{RGB}{80,160,80}
\definecolor{myturk}{RGB}{160,160,160}

\begin{document}

\begin{frame}
	\titlepage
\end{frame}

\begin{frame}{Outline}
	\setbeamertemplate{section in toc}[sections numbered]
	\tableofcontents
\end{frame}

\section{Problem Description}
\begin{frame}{Problem Description}
	Prediction over bipartite graph:

	\begin{tikzpicture}[thick,
	every node/.style={draw,circle},
	fsnode/.style={fill=myblue},
	fs1node/.style={fill=myturk},
	ssnode/.style={fill=mygreen},
	%every fit/.style={ellipse,draw,inner sep=-1pt,text width=2cm},
	%->,shorten >= 3pt,shorten <= 3pt
	]

	% the vertices of U
	\begin{scope}[start chain=going below,node distance=5mm]
	\foreach \i in {1,2,...,3}
	\node[fsnode,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of U1
	\begin{scope}[node distance=5mm]
	\foreach \i in {4,5}
	\node[fs1node,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of V
	\begin{scope}[xshift=4cm,yshift=-0.25cm,start chain=going below,node distance=7mm]
	\foreach \i in {1,2,...,4}
	\node[ssnode,on chain] (s\i) [label=right: T\i] {};
	\end{scope}

	% the set U
	%\node [myblue, fit=(f1) (f3)] {};
	\node [white,fit=(f1) (f5),label=above:$Drugs$] {};

	% the set V
	\node [white,fit=(s1) (s4),label=above:$Proteins$] {};

	% the edges
	\draw (f1) -- (s1);
	\draw (s1) -- (f2);
	\draw (f2) -- (s2);
	\draw (s2) -- (f3);
	\draw (s3) -- (f3);
	\draw (f3) -- (s4);
	\draw (s4) -- (f5);
	\draw (f5) -- (s1);
	\end{tikzpicture}

\end{frame}

\section{Recent approaches}
\begin{frame}{Classification of recent approaches\footnote{Chen Wang et al., Briefings in Bioinformatics, 2018}\footnote{Yu Ding, Briefings in Bioinformatics, 2019}}
	\begin{table}
		\begin{tabularx}{\textwidth}{|>{\setlength\hsize{.5\hsize}\setlength\linewidth{\hsize}}X|>{\setlength\hsize{1.25\hsize}\setlength\linewidth{\hsize}}X|>{\setlength\hsize{1.25\hsize}\setlength\linewidth{\hsize}}X|}
			\hline
			&Drugs&Protein\\
			\hline
			bottom-up&
			\begin{itemize}
				\item GCN over molecules
				\item drug similarity
			\end{itemize}&
			\begin{itemize}
				\item secondary structure prediction
				\item contact prediction
				\item convolution over amino acid sequences
			\end{itemize}\\
			\hline
			top-down&
			\begin{itemize}
				\item network approaches
				\item drug similarity
			\end{itemize}&
			\begin{itemize}
				\item protein similarity
			\end{itemize}\\
			\hline
		\end{tabularx}
	\end{table}	
\end{frame}

\begin{frame}{Problems of recent approaches}
	Main issues:
	\begin{itemize}
		\item lack of generalization
		\item Usually only top-down or bottom-up
	\end{itemize}
	\pause
	\begin{tikzpicture}[thick,
	every node/.style={draw,circle},
	fsnode/.style={fill=myblue},
	fs1node/.style={fill=myturk},
	ssnode/.style={fill=mygreen},
	%every fit/.style={ellipse,draw,inner sep=-1pt,text width=2cm},
	%->,shorten >= 3pt,shorten <= 3pt
	]

	% the vertices of U
	\begin{scope}[start chain=going below,node distance=5mm]
	\foreach \i in {1,2,...,3}
	\node[fsnode,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of U1
	\begin{scope}[node distance=5mm]
	\foreach \i in {4,5}
	\node[fs1node,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of V
	\begin{scope}[xshift=4cm,yshift=-0.25cm,start chain=going below,node distance=7mm]
	\foreach \i in {1,2,...,4}
	\node[ssnode,on chain] (s\i) [label=right: T\i] {};
	\end{scope}

	% the set U
	%\node [myblue, fit=(f1) (f3)] {};
	\node [white,fit=(f1) (f5),label=above:$Drugs$] {};

	% the set V
	\node [white,fit=(s1) (s4),label=above:$Proteins$] {};

	% the edges
	\draw (f1) -- (s1);
	\draw (s1) -- (f2);
	\draw (f2) -- (s2);
	\draw (s2) -- (f3);
	\draw (s3) -- (f3);
	\draw (f3) -- (s4);
	\draw (s4) -- (f5);
	\draw (f5) -- (s1);
	\end{tikzpicture}

\end{frame}

\section{A medium approach to DTI prediction}

\begin{frame}{Full citations}
	\footnotesize
	\begin{itemize}
		\item Yu Ding, Hong Wang, Hewei Zheng, Lianzong Wang, Guosi Zhang, Jiaxin Yang, Xiaoyan Lu, Yu Bai, Haotian Zhang, Jing Li, Wenyan Gao, Fukun Chen, Shui Hu, Jingqi Wu, Liangde Xu, Evaluation of drug efficacy based on the spatial position comparison of drug–target interaction centers, Briefings in Bioinformatics, , bbz024, https://doi.org/10.1093/bib/bbz024
		\item Chen Wang, Lukasz Kurgan, Review and comparative assessment of similarity-based methods for prediction of drug–protein interactions in the druggable human proteome, Briefings in Bioinformatics, , bby069, https://doi.org/10.1093/bib/bby069
	\end{itemize}
\end{frame}

\end{document}

### 2019-12-08 src/DTI_predictor.py
Running GCN node embedding 1

@@ -59,15 +59,15 @@ def missing_target_predictor(results_filename='../results/results_log',
     print("Scaling data ...")
     # side_effect_features = np.tile(DTI_data_preparation.get_side_effect_similarity_feature_list(drug_list), (len(protein_list),1))
     # side_effect_features = side_effect_features.reshape((len(protein_list), len(drug_list), 1430))
-    # DDI_features = np.tile(DTI_data_preparation.get_DDI_feature_list(drug_list), (len(protein_list),1))
-    # DDI_features = DDI_features.reshape((len(protein_list), len(drug_list), len(drug_list)))
+    DDI_features = np.tile(DTI_data_preparation.get_DDI_feature_list(drug_list), (len(protein_list),1))
+    DDI_features = DDI_features.reshape((len(protein_list), len(drug_list), len(drug_list)))
     print("Finished.\n")
 
     print("Get DTIs")
     PPI_dti_features = DTI_data_preparation.get_PPI_dti_feature_list(drug_list, protein_list)
 
-    # y_dti_data = DTI_data_preparation.get_DTIs(drug_list=drug_list, protein_list=protein_list)
-    # y_dti_data = y_dti_data.reshape((len(protein_list), len(drug_list)))
+    y_dti_data = DTI_data_preparation.get_DTIs(drug_list=drug_list, protein_list=protein_list)
+    y_dti_data = y_dti_data.reshape((len(protein_list), len(drug_list)))
     print("Finished.\n")
     print("Finished loading data.\n")
 
@@ -84,6 +84,9 @@ def missing_target_predictor(results_filename='../results/results_log',
         if embedding_method=='graphsage':
             embedding_layer_sizes[-1] = embedding_output_size
             generator = GraphSAGENodeGenerator(G, embedding_batch_size, num_samples)
+        elif embedding_method == 'gcn' or embedding_method == 'chebyshev' or \
+                embedding_method == 'sgc' or embedding_method == 'self_loops':
+            generator = FullBatchNodeGenerator(G, method='gcn')
         else:
             print("No valid embedding method chosen.")
             raise Exception
@@ -155,10 +158,19 @@ def missing_target_predictor(results_filename='../results/results_log',
             train_gen = generator.flow(protein_list[train], PPI_dti_features[train], shuffle=True)
 
             if embedding_method == 'graphsage':
-                embedding_layer = GraphSAGE(layer_sizes=embedding_layer_sizes,
-                                            generator=train_gen,
-                                            bias=True,
[…]

### 2019-12-08 presentations/2019-12-09/presentation.tex
Rerun different embeddings 1, added presentation

@@
 \definecolor{mygreen}{RGB}{80,160,80}
-\definecolor{myturk}{RGB}{160,160,160}
+\definecolor{myturk}{RGB}{240,60,60}
 
@@
 	fsnode/.style={fill=myblue},
-	fs1node/.style={fill=myturk},
+	fs1node/.style={fill=myblue},
 	ssnode/.style={fill=mygreen},
@@
 \section{Recent approaches}
-\begin{frame}{Classification of recent approaches\footnote{Chen Wang et al., Briefings in Bioinformatics, 2018}\footnote{Yu Ding, Briefings in Bioinformatics, 2019}}
+\begin{frame}{Classification of recent approaches\footnote{Chen Wang et al., Briefings in Bioinformatics, 2018}\footnote{Yu Ding et al., Briefings in Bioinformatics, 2019}}
 	\begin{table}
@@
 	\begin{itemize}
-		\item lack of generalization
+		\item Lack ability to generalize or are unable to spot small differences
 		\item Usually only top-down or bottom-up
+		\item Not making use of interaction networks
 	\end{itemize}
@@
 	fs1node/.style={fill=myturk},
+	trainlegendnode/.style={fill=myblue},
+	testlegendnode/.style={fill=myturk},
 	ssnode/.style={fill=mygreen},
@@
 
+	% legend
+	\begin{scope}[xshift=7cm,yshift=-0.25cm,start chain=going below,node distance=7mm]
+	\foreach \i in {1}
+	\node[trainlegendnode,on chain] (trainlegend\i) [label=right: Train] {};
+	\end{scope}
+
+	\begin{scope}[xshift=7cm,yshift=-0.25cm, node distance=1mm]
+	\foreach \i in {1}
+	\node[testlegendnode,on chain] (testlegend\i) [label=right: Test] {};
+	\end{scope}
+
 	% the set U
@@
 
+\begin{frame}{}
+
+\end{frame}
+
 \section{A medium approach to DTI prediction}

### 2019-12-08 src/DTI_predictor.py

@@ -96,6 +96,8 @@ def missing_target_predictor(results_filename='../results/results_log',
                                                sparse=False,
                                                teleport_probability=0.1
                                                )
+        elif embedding_method == 'appnp':
+            generator = FullBatchNodeGenerator(G, method="gcn", sparse=True)
         else:
             print("No valid embedding method chosen.")
             raise Exception
@@ -198,11 +200,23 @@ def missing_target_predictor(results_filename='../results/results_log',
                     dropout=0.5
                 )
             elif embedding_method == 'ppnp':
-                ppnp = PPNP(layer_sizes=embedding_layer_sizes,
-                            activations=['relu']*len(embedding_layer_sizes),
-                            generator=generator,
-                            dropout=0.5,
-                            kernel_regularizer=regularizers.l2(0.001))
+                embedding_layer = PPNP(
+                    layer_sizes=embedding_layer_sizes,
+                    activations=['relu']*len(embedding_layer_sizes),
+                    generator=generator,
+                    dropout=0.5,
+                    kernel_regularizer=regularizers.l2(0.001)
+                )
+            elif embedding_method == 'appnp':
+                embedding_layer = APPNP(
+                    layer_sizes=embedding_layer_sizes,
+                    activations=['relu'] * len(embedding_layer_sizes),
+                    bias=True,
+                    generator=generator,
+                    teleport_probability=0.1,
+                    dropout=0.5,
+                    kernel_regularizer=regularizers.l2(0.001)
+                )
 
             x_inp, x_out = embedding_layer.build() if embedding_method not in ['gcn', 'gat', 'ppnp'] else embedding_layer.node_model()
 
@@ -461,7 +475,8 @@ if __name__ == '__main__':
     # missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, num_samples=[100, 100], embedding_layer_sizes= [128, 64], embedding_output_size=64)
     # missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, num_samples=[200, 100], embedding_layer_sizes= [128, 64], embedding_output_size=128)
 
[…]

### 2019-12-08 presentations/2019-12-09/presentation.tex
Complete hmmer compuatation with biggies

@@
 
-\begin{frame}{}
-
+\section{A medium approach to DTI prediction}
+\begin{frame}{How to bring both approaches together?}
+	\begin{enumerate}
+		\item For each drug, filter possible targets according to protein based measurement
+		\item Bring those together with the interaction networks for the prediction
+	\end{enumerate}
 \end{frame}
 
-\section{A medium approach to DTI prediction}
+\subsection{Filtering possible targets}
+\begin{frame}{How to build such a filter?}
+	STITCH database got 
+	\begin{itemize}
+		\item 390 000 chemicals, and
+		\item 3.6 million proteins
+		\item from over 2000 organisms
+	\end{itemize}
+	\pause 
+	\vspace{0.5cm}
+	For each drug, find a motif in the target amino acid sequences and query that against all possible proteins and use the results as features for the prediction
+\end{frame}
+
+\begin{frame}{How to find motifs?}
+	\begin{enumerate}
+		\item Build multi sequence alignment with alignment tool of choice over non-human proteins
+			\begin{itemize}
+				\item FAMSA, MAFFT, Kalign, MSA Props, Muscle, \dots
+			\end{itemize}
+		\item Build representation of that alignment
+			\begin{itemize}
+				\item Hidden Markov model with HMMER
+				\item PWM
+			\end{itemize}
+		\item Query representation against database
+	\end{enumerate}
+	\pause
+	\vspace{0.5cm}
+	This is very expensive to compute \dots (10000 hours for alignments of 614 drugs + IBEX queue times for such big jobs)
+\end{frame}
+
+\begin{frame}{}
+\end{frame}
 

### 2019-12-08 src/alignment_pipeline.py

@@ -461,8 +461,8 @@ if __name__ == '__main__':
 
     run_hmm_pipeline(min_score=700,
                      alignment_method='famsa',
-                     workers=4,
-                     threads_per_worker=10)
+                     workers=1,
+                     threads_per_worker=40)
 
 
 

### 2019-12-08 src/run_job_pipeline.py

@@ -55,6 +55,6 @@ def submit_jobscript_n_times(n):
 if __name__ == '__main__':
     # run_jobs(parts=60, amount=646)
     cancel_jobs()
-    submit_jobscript_n_times(50)
+    # submit_jobscript_n_times(50)
 
     pass

### 2019-12-08 src/alignment_pipeline.py
Complete hmmer compuatation with biggies 2

@@ -351,7 +351,7 @@ def run_hmm_pipeline(min_score=700,
         alignment_file = alignment_path + drug_name + "_"+alignment_method+"_aligned_"+str(min_score)+"_min_score.afa"
         hmmbuild_file = hmmbuild_target_path + drug_name + "_"+alignment_method+"_aligned_"+str(min_score)+"_min_score.hmm"
 
-        if os.path.exists(hmmbuild_file) and os.stat(hmmbuild_file).st_size != 0:
+        if os.path.exists(hmmbuild_file) and os.stat(hmmbuild_file).st_size == 0:
             return
 
         print("Building Hidden Markov Model ...")
@@ -364,11 +364,10 @@ def run_hmm_pipeline(min_score=700,
         # "--"+rel_weight_method+" "+\
 
         print(command)
-        subprocess.call(command, shell=True)
+        # subprocess.call(command, shell=True)
         print("Finished.\n")
 
         # Parallel(n_jobs=10)(delayed(hmm_build)(filename) for filename in [file for file in os.listdir(alignment_path) if file.startswith("CID")])
-        '''
         # run hmm search on previously computed hidden markov model over all sequences
         # hmmsearch params
         max_flag = False
@@ -413,7 +412,6 @@ def run_hmm_pipeline(min_score=700,
                         split_list.append(ele)
                 protein_id = split_list[8].strip()
                 predicted_targets_filehandler.write(protein_id+'\n')
-        '''
 
     q = queue.Queue()
     files = os.listdir(alignment_path)
@@ -461,8 +459,8 @@ if __name__ == '__main__':
 
     run_hmm_pipeline(min_score=700,
                      alignment_method='famsa',
-                     workers=1,
-                     threads_per_worker=40)
+                     workers=5,
+                     threads_per_worker=8)
 
 
 

### 2019-12-08 src/run_job_pipeline.py
Complete hmmer compuatation with biggies 3

@@ -54,7 +54,7 @@ def submit_jobscript_n_times(n):
 
 if __name__ == '__main__':
     # run_jobs(parts=60, amount=646)
-    cancel_jobs()
-    # submit_jobscript_n_times(50)
+    # cancel_jobs()
+    submit_jobscript_n_times(20)
 
     pass

### 2019-12-08 presentations/2019-12-09/presentation.tex
Complete hmmer compuatation with biggies 3

@@
 \expandafter\def\expandafter\insertshorttitle\expandafter{%
-	\insertshorttitle\hspace{5cm}%
+	\insertshorttitle\hspace{4.5cm}%
 	\insertframenumber\,/\,\inserttotalframenumber
@@
 	\end{enumerate}
+	\pause
+	Two hypotheses to be tested:
+	\begin{itemize}
+		\item Interaction only network sufficient?
+		\item Does filter method increase performance?
+	\end{itemize}
 \end{frame}
@@
 	\vspace{0.5cm}
-	This is very expensive to compute \dots (10000 hours for alignments of 614 drugs + IBEX queue times for such big jobs)
+	This is quite expensive to compute \dots (8000 hours for alignments of 614 drugs + IBEX queue times for such big jobs)
 \end{frame}
 
-\begin{frame}{}
+\begin{frame}{Other features for the model}
+	\begin{itemize}
+		\item Protein-protein-interaction network from STRING (only human)
+		\item Drug-drug-interaction network from Boyce\footnote{Boyce et al., 2015} and Drugbank
+		\item Side-effect data for drugs from SIDER, annotated with MedDRA hierarchy (semantic similarity)
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Features in context}
+	\begin{tikzpicture}[thick,
+	every node/.style={draw,circle},
+	fsnode/.style={fill=myblue},
+	fs1node/.style={fill=myturk},
+	drughierarchynode/.style={fill=myturk},
+	druglegendnode/.style={fill=myblue},
+	hierarchylegendnode/.style={fill=myturk},
+	targetlegendnode/.style={fill=mygreen},
+	ssnode/.style={fill=mygreen},
+	%every fit/.style={ellipse,draw,inner sep=-1pt,text width=2cm},
+	%->,shorten >= 3pt,shorten <= 3pt
+	]
+
+	% the vertices of U
+	\begin{scope}[xshift=2.25cm,yshift=-3cm,start chain=going right,node distance=5mm]
+	\foreach \i in {1,2,...,6}
+	\node[fsnode,on chain] (f\i) {};
+	\end{scope}
+
+	\begin{scope}[xshift=3.5cm,yshift=-2.5cm,start chain=going right,node distance=5mm]
+	\foreach \i in {7,8,...,10}
+	\node[fsnode,on chain] (f\i) {};
+	\end{scope}
+
+	% Hierarchy nodes
+	\begin{scope}[xshift=2cm,yshift=-4.2cm,start chain=going right,node distance=7mm]
+	\foreach \i in {1,2,...,6}
+	\node[drughierarchynode,on chain] (dh\i) {};
+	\end{scope}
+
+	% the vertices of V
+	\begin{scope}[xshift=2.5cm,yshift=0cm,start chain=going right,node distance=7mm]
+	\foreach \i in {1,2,...,5}
+	\node[ssnode,on chain] (s\i) {};
+	\end{scope}
+	\begin{scope}[xshift=3cm,yshift=-1cm,start chain=going right,node distance=7mm]
+	\foreach \i in {6,7,...,9}
+	\node[ssnode,on chain] (s\i) {};
+	\end{scope}
+
+	% legend
+	\begin{scope}[xshift=9cm,yshift=0cm, node distance=1mm]
+	\foreach \i in {1}
+	\node[hierarchylegendnode] (hlegend\i) [label=right: Side effects] {};
+	\end{scope}
+	\begin{scope}[xshift=9cm,yshift=-0.5cm, node distance=1mm]
+	\foreach \i in {1}
+	\node[druglegendnode] (dlegend\i) [label=right: Drugs] {};
+	\end{scope}
+	\begin{scope}[xshift=9cm,yshift=-1cm, node distance=1mm]
+	\foreach \i in {1}
+	\node[targetlegendnode] (tlegend\i) [label=right: Targets] {};
+	\end{scope}
+
+	\node [white, fit=(dh3) (dh4),label=below:$\textbf{\dots}$] {};
+	\node [white, fit=(f3),label=below:$\textbf{\dots}$] {};
+	\node [white,fit=(f7), label=above:$\textbf{\dots}$] {};
+
+	\node [white,fit=(s5), label=right:$\approx17000$] {};
+	\node [white,fit=(f10), label=right:$\approx600$] {};
+	\node [white,fit=(dh6), label=right:$\approx1400$] {};
+
+	% the set U
+	%\node [myblue, fit=(f1) (f3)] {};
+
+	% DTI edges
+	\draw (f1) -- (s6);
+	\draw (f2) -- (s6);
+
+	% PPI edges
+	\draw (s1) -- (s6);
+	\draw (s1) -- (s2);
+	\draw (s1) -- (s6);
+	\draw (s1) -- (s7);
+	\draw (s2) -- (s6);
+	\draw (s4) -- (s9);
+	\draw (s3) -- (s7);
+	\draw (s4) -- (s8);
+	\draw (s4) -- (s7);
+	\draw (s2) -- (s9);
+	\draw (s5) -- (s8);
+
+	% DDI edges
+	\draw (f3) -- (f7);
+	\draw (f4) -- (f7);
+	\draw (f5) -- (f7);
+	\draw (f3) -- (f4);
+	\draw (f5) -- (f10);
+	\draw (f2) -- (f7);
+
+	% Side effect edges
+	\draw (f1) -- (dh1);
+	\draw (f2) -- (dh1);
+
+	\end{tikzpicture}
+\end{frame}
+
+\begin{frame}{How to learn a sufficient embedding?}
+	\begin{itemize}
+		\item Use Filtered results as node features
+		\item Build a node classification algorithm
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{First results}
+	\begin{figure}
+		\small
+		\centering
+		\begingroup
+		\def\arraystretch{1.2}
+		\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
+			\hline
+			Approach  & \multicolumn{2}{c}{Layers} & \multicolumn{3}{|c|}{Acceptor} & \multicolumn{3}{c|}{Donor} \\
+			\cline{2-9}
+			&Conv. & Others & Acc. & Prec. & Rec. & Acc. & Prec. & Rec. \\
+			\hline
+			CNN DPDB & 4 & 5 & 94.4 & 95.4 & 94.6 & 94.9 & 94.4 & 94.7 \\
+			CNN DPDB & 4 & 7 & 93.5 & 93.3 & 94.5 & 94.0 & 94.0 & 93.3 \\
+			CNN DPDB & 6 & 5 & 94.0 & 93.9 & 94.9 & 94.2 & 95.4 & 91.6 \\
+			CNN DPDB & 6 & 5 & 94.4 & 97.0 & 93.8 & 95.2 & 96.5 & 93.7 \\
+			CNN DPDB & 2 & 4 & 94.3 & 95.6 & 94.3 & 95.3 & 96.9 & 94.4 \\
+			SpliceRover & 4 & 2 & 96.1 & 93.9 & 97.4 & 95.4 & 95.6 & 96.7 \\
+			Splice2Deep & - & - & 95.2 &  -- & 94.9 & 95.6 & -- & 98.8\\
+
+			\hline  
+		\end{tabular}
+		\endgroup
+	\end{figure}
 \end{frame}
@@
 		\item Chen Wang, Lukasz Kurgan, Review and comparative assessment of similarity-based methods for prediction of drug–protein interactions in the druggable human proteome, Briefings in Bioinformatics, , bby069, https://doi.org/10.1093/bib/bby069
+		\item Toward a complete dataset of drug–drug interaction information from publicly available sources, SerkanAyvazaJohnHornbOktieHassanzadehcQianZhudJohannStaneNicholas P.TatonettifSantiagoVilarfMathiasBrochhausengMatthiasSamwaldhMajidRastegar-MojaradiMichelDumontierjRichard D.Boyce, 2015
+
 	\end{itemize}

### 2019-12-08 src/DTI_predictor.py

@@ -88,8 +88,10 @@ def missing_target_predictor(results_filename='../results/results_log',
             generator = GraphSAGENodeGenerator(G, embedding_batch_size, num_samples)
         elif embedding_method == 'hinsage':
             generator = HinSAGENodeGenerator(G, embedding_batch_size, num_samples)
-        elif embedding_method in ['gcn', 'chebyshev', 'sgc', 'gat']:
+        elif embedding_method in ['gcn', 'chebyshev', 'gat']:
             generator = FullBatchNodeGenerator(G, method=embedding_method)
+        elif embedding_method == 'sgc':
+            generator = FullBatchNodeGenerator(G, method='sgc', k=2)
         elif embedding_method == 'ppnp':
             generator = FullBatchNodeGenerator(G,
                                                method="ppnp",
@@ -217,8 +219,17 @@ def missing_target_predictor(results_filename='../results/results_log',
                     dropout=0.5,
                     kernel_regularizer=regularizers.l2(0.001)
                 )
+            elif embedding_method == 'sgc':
+                embedding_layer = GCN(
+                    layer_sizes=[embedding_output_size],
+                    generator=generator,
+                    bias=True,
+                    dropout=0.5,
+                    activations=["softmax"],
+                    kernel_regularizer=regularizers.l2(5e-4),
+                )
 
-            x_inp, x_out = embedding_layer.build() if embedding_method not in ['gcn', 'gat', 'ppnp'] else embedding_layer.node_model()
+            x_inp, x_out = embedding_layer.build() if embedding_method not in ['gcn', 'gat', 'ppnp', 'appnp', 'sgc'] else embedding_layer.node_model()
 
             prediction = layers.Dense(units=PPI_dti_features.shape[1], activation="linear")(x_out)
 
@@ -253,7 +264,7 @@ def missing_target_predictor(results_filename='../results/results_log',
 
         # if plot:
             # dti_utils.plot_history(history)
-        if embedding_method in ['gcn', 'gat', 'ppnp']:
+        if embedding_method in ['gcn', 'gat', 'ppnp', 'appnp', 'sgc']:
             protein_node_embeddings = protein_node_embeddings.reshape(protein_node_embeddings.shape[1],
                                                                       protein_node_embeddings.shape[2])
 
@@ -477,6 +488,7 @@ if __name__ == '__main__':
 
[…]

### 2019-12-08 src/alignment_pipeline.py

@@ -364,9 +364,10 @@ def run_hmm_pipeline(min_score=700,
         # "--"+rel_weight_method+" "+\
 
         print(command)
-        # subprocess.call(command, shell=True)
+        subprocess.call(command, shell=True)
         print("Finished.\n")
 
+        '''
         # Parallel(n_jobs=10)(delayed(hmm_build)(filename) for filename in [file for file in os.listdir(alignment_path) if file.startswith("CID")])
         # run hmm search on previously computed hidden markov model over all sequences
         # hmmsearch params
@@ -378,9 +379,6 @@ def run_hmm_pipeline(min_score=700,
         hmmsearch_file = hmm_search_results_path + drug_name + "_" + alignment_method + "_aligned_" + str(min_score) + "_min_score.out"
 
         command = "hmmsearch " + \
-                  ("--max " if max_flag else "") + \
-                  "--nonli " + \
-                  "--nontextw "+\
                   "--cpu " + str(threads_per_worker) + " " + \
                   hmmbuild_file + " " + all_prots_fasta_filename + " > " + hmmsearch_file
 
@@ -412,6 +410,7 @@ def run_hmm_pipeline(min_score=700,
                         split_list.append(ele)
                 protein_id = split_list[8].strip()
                 predicted_targets_filehandler.write(protein_id+'\n')
+        '''
 
     q = queue.Queue()
     files = os.listdir(alignment_path)
@@ -459,8 +458,8 @@ if __name__ == '__main__':
 
     run_hmm_pipeline(min_score=700,
                      alignment_method='famsa',
-                     workers=5,
-                     threads_per_worker=8)
+                     workers=1,
+                     threads_per_worker=40)
 
 
 

### 2019-12-08 src/jobscript.sh

@@ -4,7 +4,7 @@
 #SBATCH -J Hmmer
 #SBATCH -o jobscript_outputs/Hmmer.%J.out
 #SBATCH -e jobscript_outputs/Hmmer.%J.err
-#SBATCH --time=3-00:00:00
+#SBATCH --time=1-00:00:00
 #SBATCH --mem=240G
 #SBATCH --constraint=[intel]
 #SBATCH --cpus-per-task=40

### 2019-12-08 src/alignment_pipeline.py
Complete hmmer compuatation with biggies 4

@@ -351,7 +351,7 @@ def run_hmm_pipeline(min_score=700,
         alignment_file = alignment_path + drug_name + "_"+alignment_method+"_aligned_"+str(min_score)+"_min_score.afa"
         hmmbuild_file = hmmbuild_target_path + drug_name + "_"+alignment_method+"_aligned_"+str(min_score)+"_min_score.hmm"
 
-        if os.path.exists(hmmbuild_file) and os.stat(hmmbuild_file).st_size == 0:
+        if os.path.exists(hmmbuild_file) and os.stat(hmmbuild_file).st_size != 0:
             return
 
         print("Building Hidden Markov Model ...")

### 2019-12-08 src/jobscript.sh

@@ -4,7 +4,7 @@
 #SBATCH -J Hmmer
 #SBATCH -o jobscript_outputs/Hmmer.%J.out
 #SBATCH -e jobscript_outputs/Hmmer.%J.err
-#SBATCH --time=1-00:00:00
+#SBATCH --time=3-00:00:00
 #SBATCH --mem=240G
 #SBATCH --constraint=[intel]
 #SBATCH --cpus-per-task=40

### 2019-12-08 src/run_job_pipeline.py

@@ -54,7 +54,7 @@ def submit_jobscript_n_times(n):
 
 if __name__ == '__main__':
     # run_jobs(parts=60, amount=646)
-    # cancel_jobs()
-    submit_jobscript_n_times(20)
+    cancel_jobs()
+    submit_jobscript_n_times(4)
 
     pass

### 2019-12-08 src/alignment_pipeline.py
Complete hmmer compuatation with biggies 5

@@ -351,7 +351,7 @@ def run_hmm_pipeline(min_score=700,
         alignment_file = alignment_path + drug_name + "_"+alignment_method+"_aligned_"+str(min_score)+"_min_score.afa"
         hmmbuild_file = hmmbuild_target_path + drug_name + "_"+alignment_method+"_aligned_"+str(min_score)+"_min_score.hmm"
 
-        if os.path.exists(hmmbuild_file) and os.stat(hmmbuild_file).st_size != 0:
+        if os.path.exists(hmmbuild_file) and os.stat(hmmbuild_file).st_size == 0:
             return
 
         print("Building Hidden Markov Model ...")
@@ -364,10 +364,9 @@ def run_hmm_pipeline(min_score=700,
         # "--"+rel_weight_method+" "+\
 
         print(command)
-        subprocess.call(command, shell=True)
+        # subprocess.call(command, shell=True)
         print("Finished.\n")
 
-        '''
         # Parallel(n_jobs=10)(delayed(hmm_build)(filename) for filename in [file for file in os.listdir(alignment_path) if file.startswith("CID")])
         # run hmm search on previously computed hidden markov model over all sequences
         # hmmsearch params
@@ -410,7 +409,6 @@ def run_hmm_pipeline(min_score=700,
                         split_list.append(ele)
                 protein_id = split_list[8].strip()
                 predicted_targets_filehandler.write(protein_id+'\n')
-        '''
 
     q = queue.Queue()
     files = os.listdir(alignment_path)
@@ -458,8 +456,8 @@ if __name__ == '__main__':
 
     run_hmm_pipeline(min_score=700,
                      alignment_method='famsa',
-                     workers=1,
-                     threads_per_worker=40)
+                     workers=5,
+                     threads_per_worker=8)
 
 
 

### 2019-12-08 src/run_job_pipeline.py

@@ -54,7 +54,7 @@ def submit_jobscript_n_times(n):
 
 if __name__ == '__main__':
     # run_jobs(parts=60, amount=646)
-    cancel_jobs()
-    submit_jobscript_n_times(4)
+    # cancel_jobs()
+    submit_jobscript_n_times(20)
 
     pass

### 2019-12-08 presentations/2019-12-09/presentation.tex
Complete hmmer compuatation with biggies 6

@@
 		\item Use Filtered results as node features
-		\item Build a node classification algorithm
+		\item Run different models for classification of proteins/nodes in the graph:
+		\begin{itemize}
+			\item GraphSage
+			\item HinSage
+			\item Attri2Vec
+			\item Graph Attention Network (GAT)
+			\item Graph Convolutional Network (GCN) for node classification
+			\item GCNs for graph classification of induced subgraph
+			\item SimpleGCN
+			\item Personalized Propagation of Neural Predictions (PPNP)
+			\item Approximate Personalized Propagation of Neural Predictions (APPNP)
+		\end{itemize}
 	\end{itemize}

### 2019-12-08 src/alignment_pipeline.py

@@ -372,7 +372,7 @@ def run_hmm_pipeline(min_score=700,
         # hmmsearch params
         max_flag = False
 
-        all_prots_fasta_filename = "../data/STITCH_data/9606.protein.sequences.v10.fa"
+        all_prots_fasta_filename = "../data/STITCH_data/9606.protein.sequences.v11.0.fa"
 
         hmm_search_results_path = '../data/hmm_search_results/'
         hmmsearch_file = hmm_search_results_path + drug_name + "_" + alignment_method + "_aligned_" + str(min_score) + "_min_score.out"

### 2019-12-08 src/run_job_pipeline.py

@@ -55,6 +55,6 @@ def submit_jobscript_n_times(n):
 if __name__ == '__main__':
     # run_jobs(parts=60, amount=646)
     # cancel_jobs()
-    submit_jobscript_n_times(20)
+    submit_jobscript_n_times(30)
 
     pass

### 2019-12-09 presentations/2019-12-09/presentation.tex
Added latest version of presentation
UPdated presentation, running graphsage again 1

@@
 \begin{frame}{How to bring both approaches together?}
+	\begin{exampleblock}{Algorithm}
 		\begin{enumerate}
@@
 		\end{enumerate}
+	\end{exampleblock}
 	\pause
@@
 	\begin{figure}
-		\small
 		\centering
 		\begingroup
+
+		\footnotesize
 		\def\arraystretch{1.2}
-		\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
+		\begin{tabular}{|r|r|r|r|r|r|r|r|}
 			\hline
-			Approach  & \multicolumn{2}{c}{Layers} & \multicolumn{3}{|c|}{Acceptor} & \multicolumn{3}{c|}{Donor} \\
-			\cline{2-9}
-			&Conv. & Others & Acc. & Prec. & Rec. & Acc. & Prec. & Rec. \\
+			\multicolumn{4}{|c|}{Used data}&layer sizes&model&Acc&AUC\\
+			%\hline
+			PPI&DDI&SE&HMM&&&&\\
 			\hline
-			CNN DPDB & 4 & 5 & 94.4 & 95.4 & 94.6 & 94.9 & 94.4 & 94.7 \\
-			CNN DPDB & 4 & 7 & 93.5 & 93.3 & 94.5 & 94.0 & 94.0 & 93.3 \\
-			CNN DPDB & 6 & 5 & 94.0 & 93.9 & 94.9 & 94.2 & 95.4 & 91.6 \\
-			CNN DPDB & 6 & 5 & 94.4 & 97.0 & 93.8 & 95.2 & 96.5 & 93.7 \\
-			CNN DPDB & 2 & 4 & 94.3 & 95.6 & 94.3 & 95.3 & 96.9 & 94.4 \\
-			SpliceRover & 4 & 2 & 96.1 & 93.9 & 97.4 & 95.4 & 95.6 & 96.7 \\
-			Splice2Deep & - & - & 95.2 &  -- & 94.9 & 95.6 & -- & 98.8\\
-
+			1&1&0&0&641&graphsage&88.97&66.20\\
+			1&1&0&0&641&graphsage&80.28&70.51\\
+			1&1&0&0&641&graphsage&87.44&67.80\\
+			1&1&0&0&641&graphsage&86.46&68.39\\
+			1&1&0&0&641&graphsage&88.84&66.62\\
+			1&1&1&0&641&graphsage&93.34&62.42\\
+			1&1&1&0&641&graphsage&73.20&62.53\\
+			1&1&1&0&641&graphsage&93.27&64.81\\
+			1&1&1&0&641&graphsage&74.28&59.18\\
+			1&1&1&0&641&graphsage&72.40&63.93\\
+			1&1&1&0&641&gcn&88.61&67.27\\
+			1&1&1&0&641&gat&82.65&71.69\\
+			1&1&1&0&641&hinsage&83.18&70.33\\
+			1&1&1&0&641&ppnp&89.80&67.78\\
+			1&1&1&0&641&appnp&84.35&71.82\\
+			1&1&1&0&641&sgc&86.50&68.45\\
+			1&1&1&0&641&gcn&59.84&50.10\\
+			1&1&1&0&641&gcn&97.11&56.02\\
+			1&1&1&0&641&gat&59.77&50.06\\
+			1&1&1&0&641&gat&55.48&55.93\\
+			1&1&1&0&641&hinsage&59.77&50.06\\
+			1&1&1&0&641&hinsage&94.06&63.52\\
+			1&1&1&0&641&ppnp&21.54&50.15\\
+			1&1&1&0&641&ppnp&77.98&55.85\\
+			1&1&1&0&641&appnp&40.71&50.13\\
+			1&1&1&0&641&sgc&2.27&50.18\\
 			\hline  
@@
 	\end{figure}
+\end{frame}
+
+\begin{frame}{Future Work}
+	\begin{itemize}
+		\item Prove improvement of generalization
+		\item Test performance against recent approaches
+		\item Tweak model properly
+		\item If performance and generalization is good enough:
+		\begin{itemize}
+			\item Test some predicted interactions in vitro
+			\item Use DTI for Side-effect-phenotype mapping
+		\end{itemize}
+	\end{itemize}
 \end{frame}

### 2019-12-09 src/DTI_predictor.py

@@ -389,7 +389,7 @@ def missing_target_predictor(results_filename='../results/results_log',
     print("Mean f1-score:", np.mean(cv_scores['f1-score']))
 
     with open(file=results_table_filename, mode='a') as f:
-        print("1" +"\t"+ "1" +"\t"+ "1" +"\t"+ "0" +"\t"+
+        print("1" +"\t"+ "0" +"\t"+ "1" +"\t"+ "0" +"\t"+
               str(len(protein_list)) +"\t"+ str(len(drug_list)) +"\t"+
               str(nb_epochs) +"\t"+
               str(embedding_layer_sizes) + "\t" +
@@ -486,9 +486,11 @@ if __name__ == '__main__':
     # missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, num_samples=[100, 100], embedding_layer_sizes= [128, 64], embedding_output_size=64)
     # missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, num_samples=[200, 100], embedding_layer_sizes= [128, 64], embedding_output_size=128)
 
-    missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[32,32], embedding_output_size=64, embedding_method='gcn')
-    missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[32,32], embedding_output_size=64, embedding_method='gat')
-    missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[32], num_samples=[100], embedding_output_size=64, embedding_method='hinsage')
-    missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[64, 64, 64], embedding_output_size=128, embedding_method='ppnp')
-    missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[64, 64, 64], embedding_output_size=128, embedding_method='appnp')
-    missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[64, 64, 64], embedding_output_size=128, embedding_method='sgc')
+    # missing_target_predictor(batch_size=10000, nb_epochs=20, plot=True, embedding_layer_sizes=[32,32], embedding_output_size=64, embedding_method='gcn')
[…]

### 2019-12-09 src/my_test.py

@@ -113,13 +113,36 @@ def merge_protein_to_subgraph_dicts(start):
     print("Finished.\n")
     '''
 
+def prepare_results():
+    filename = "../results/results_table"
+    with open(file=filename, mode='r') as f:
+        for line in f:
+            split_line = line.strip().split('\t')
+            split_line.pop(6)
+            split_line.pop(-1)
+            if len(split_line) == 9:
+                split_line.insert(-2, 'graphsage')
+            else:
+                split_line.pop(-4)
+
+            split_line.pop(4)
+            split_line.pop(5)
+
+            split_line[-1] = "%.2f" % float(split_line[-1])
+            split_line[-2] = "%.2f" % float(split_line[-2])
+            print('&'.join(split_line) + "\\\\")
+
+
+
 if __name__ == '__main__':
     # test_target_subset()
 
     # run_PPI_parallel()
 
-    _, start = sys.argv
-    merge_protein_to_subgraph_dicts(start)
+    # _, start = sys.argv
+    # merge_protein_to_subgraph_dicts(start)
+
+    prepare_results()
 
     pass
 

### 2019-12-09 presentations/2019-12-09/presentation.tex
Added presentation updates, ideas and TODO, Rerunning alignment for all drugs in DDI graph 1

@@
 			\hline
-			1&1&0&0&641&graphsage&88.97&66.20\\
-			1&1&0&0&641&graphsage&80.28&70.51\\
-			1&1&0&0&641&graphsage&87.44&67.80\\
-			1&1&0&0&641&graphsage&86.46&68.39\\
-			1&1&0&0&641&graphsage&88.84&66.62\\
-			1&1&1&0&641&graphsage&93.34&62.42\\
-			1&1&1&0&641&graphsage&73.20&62.53\\
-			1&1&1&0&641&graphsage&93.27&64.81\\
-			1&1&1&0&641&graphsage&74.28&59.18\\
-			1&1&1&0&641&graphsage&72.40&63.93\\
-			1&1&1&0&641&gcn&88.61&67.27\\
-			1&1&1&0&641&gat&82.65&71.69\\
-			1&1&1&0&641&hinsage&83.18&70.33\\
-			1&1&1&0&641&ppnp&89.80&67.78\\
-			1&1&1&0&641&appnp&84.35&71.82\\
-			1&1&1&0&641&sgc&86.50&68.45\\
-			1&1&1&0&641&gcn&59.84&50.10\\
-			1&1&1&0&641&gcn&97.11&56.02\\
-			1&1&1&0&641&gat&59.77&50.06\\
-			1&1&1&0&641&gat&55.48&55.93\\
-			1&1&1&0&641&hinsage&59.77&50.06\\
-			1&1&1&0&641&hinsage&94.06&63.52\\
-			1&1&1&0&641&ppnp&21.54&50.15\\
-			1&1&1&0&641&ppnp&77.98&55.85\\
-			1&1&1&0&641&appnp&40.71&50.13\\
-			1&1&1&0&641&sgc&2.27&50.18\\
+			1&1&0&0&[32, 32]&graphsage&80.28&70.51\\
+			1&1&0&0&[64]&hinsage&83.18&70.33\\
+			1&1&0&0&[32, 64]&gcn&88.61&67.27\\
+			1&1&0&0&[32, 64]&gat&82.65&71.69\\
+			1&1&0&0&[64, 64, 128]&ppnp&89.80&67.78\\
+			1&1&0&0&[64, 64, 128]&appnp&84.35&71.82\\
+
+			\hline  
+		\end{tabular}
+		\endgroup
+	\end{figure}
+\end{frame}
+
+\begin{frame}{First results}
+	\begin{figure}
+		\centering
+		\begingroup
+
+		\footnotesize
+		\def\arraystretch{1.2}
+		\begin{tabular}{|r|r|r|r|r|r|r|r|}
+			\hline
+			\multicolumn{4}{|c|}{Used data}&layer sizes&model&Acc&AUC\\
+			%\hline
+			PPI&DDI&SE&HMM&&&&\\
+			\hline
+			1&1&1&0&[32, 64]&graphsage&93.27&64.81\\
+			1&1&1&0&[64]&hinsage&94.06&63.52\\
+			1&1&1&0&[32, 64]&gcn&97.11&56.02\\
+			1&1&1&0&[32, 64]&gat&59.77&50.06\\
+			1&1&1&0&[64, 64, 128]&ppnp&77.98&55.85\\
+			1&1&1&0&[64, 64, 128]&appnp&94.43&66.43\\
+			1&1&1&0&[64, 64, 128]&sgc&86.50&68.45\\
+			1&0&1&0&[64]&hinsage&59.77&50.06\\
+			1&0&1&0&[32, 64]&gcn&59.84&50.10\\
+			1&0&1&0&[32, 64]&gat&55.48&55.93\\
+			1&0&1&0&[64, 64, 128]&ppnp&21.54&50.15\\
+			1&0&1&0&[64, 64, 128]&appnp&40.71&50.13\\
+			1&0&1&0&[64, 64, 128]&sgc&2.27&50.18\\
+
 			\hline  
@@
 
+\begin{frame}
+	\begin{figure}
+		\includegraphics[width=\linewidth]{../../models/graphsage_model.png}
+		\caption{Graphsage model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[height=0.9\textheight]{../../models/hinsage_model.png}
+		\caption{Hinsage model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[width=\linewidth]{../../models/gcn_model.png}
+		\caption{GCN model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[width=\linewidth]{../../models/gat_model.png}
+		\caption{GAT model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[height=0.9\textheight]{../../models/ppnp_model.png}
+		\caption{PPNP model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[height=0.9\textheight]{../../models/appnp_model.png}
+		\caption{APPNP model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[width=\linewidth]{../../models/sgc_model.png}
+		\caption{SGC model}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[height=0.9\textheight]{../../models/overall_model.png}
+		\caption{Overall model}
+	\end{figure}
+\end{frame}
+
 \end{document}

### 2019-12-09 src/DTI_data_preparation.py

@@ -3,6 +3,8 @@ import networkx as nx
 
 import pickle
 
+import os
+
 import PPI_utils
 import DDI_utils
 import similarity_measurement
@@ -154,6 +156,12 @@ def get_annotated_PPI_graph():
         PPI_graph.node[protein]['node_feature'] = node_feature_dict[protein]
 
     return PPI_graph
+
+def write_drug_to_HMM_filtered_targets_dict():
+    predicted_targets_dir = "../data/predicted_targets/"
+
+    files = os.listdir(predicted_targets_dir)
+
 
 def test():
     # print("DTI", len(get_human_proteins()))

### 2019-12-09 src/alignment_pipeline.py

@@ -16,6 +16,7 @@ from Bio import SeqIO
 from Bio.Align.Applications import ClustalOmegaCommandline, MuscleCommandline, MafftCommandline, MSAProbsCommandline, TCoffeeCommandline
 
 from similarity_measurement import *
+import DDI_utils
 
 
 def write_pruned_SeqIO_fasta_dict():
@@ -284,7 +285,9 @@ def run_MSA(min_score=800,
 
     files = None
     if human_only:
-        drug_list = get_SIDER_Boyce_Drubank_drug_intersection()
+        merged_DDI_graph = DDI_utils.get_merged_DDI_graph()
+        drug_list = list(merged_DDI_graph.nodes())
+        # drug_list = get_SIDER_Boyce_Drubank_drug_intersection()
         # drug_list = get_SIDER_drug_list()
         files = [drug_name+"_fasta_" + str(min_score) + "_min_score.fasta" for drug_name in drug_list]
         shuffle(files)
@@ -440,7 +443,6 @@ if __name__ == '__main__':
 
     # create_fasta_files(min_score=700)
 
-    '''
     args = sys.argv + ['', '']
     start = args[1]
     end = args[2]
@@ -452,12 +454,13 @@ if __name__ == '__main__':
             start=start,
             end=end,
             human_only=True)
-    '''
 
+    '''
     run_hmm_pipeline(min_score=700,
                      alignment_method='famsa',
                      workers=5,
                      threads_per_worker=8)
+    '''
 
 
 

### 2019-12-09 src/jobscript.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
 #SBATCH -N 1
 #SBATCH --partition=batch
-#SBATCH -J Hmmer
-#SBATCH -o jobscript_outputs/Hmmer.%J.out
-#SBATCH -e jobscript_outputs/Hmmer.%J.err
+#SBATCH -J MSA
+#SBATCH -o jobscript_outputs/MSA.%J.out
+#SBATCH -e jobscript_outputs/MSA.%J.err
 #SBATCH --time=3-00:00:00
 #SBATCH --mem=240G
 #SBATCH --constraint=[intel]

### 2019-12-09 src/my_test.py

@@ -126,7 +126,7 @@ def prepare_results():
                 split_line.pop(-4)
 
             split_line.pop(4)
-            split_line.pop(5)
+            split_line.pop(4)
 
             split_line[-1] = "%.2f" % float(split_line[-1])
             split_line[-2] = "%.2f" % float(split_line[-2])

### 2019-12-09 src/alignment_pipeline.py
Running hmm search again 1

@@ -443,6 +443,7 @@ if __name__ == '__main__':
 
     # create_fasta_files(min_score=700)
 
+    '''
     args = sys.argv + ['', '']
     start = args[1]
     end = args[2]
@@ -454,13 +455,12 @@ if __name__ == '__main__':
             start=start,
             end=end,
             human_only=True)
-
     '''
+
     run_hmm_pipeline(min_score=700,
                      alignment_method='famsa',
                      workers=5,
                      threads_per_worker=8)
-    '''
 
 
 

### 2019-12-09 src/jobscript.sh
Running hmm search again 2

@@ -1,9 +1,9 @@
 #!/bin/bash
 #SBATCH -N 1
 #SBATCH --partition=batch
-#SBATCH -J MSA
-#SBATCH -o jobscript_outputs/MSA.%J.out
-#SBATCH -e jobscript_outputs/MSA.%J.err
+#SBATCH -J Hmmer
+#SBATCH -o jobscript_outputs/Hmmer.%J.out
+#SBATCH -e jobscript_outputs/Hmmer.%J.err
 #SBATCH --time=3-00:00:00
 #SBATCH --mem=240G
 #SBATCH --constraint=[intel]

### 2019-12-09 src/run_job_pipeline.py

@@ -55,6 +55,6 @@ def submit_jobscript_n_times(n):
 if __name__ == '__main__':
     # run_jobs(parts=60, amount=646)
     # cancel_jobs()
-    submit_jobscript_n_times(30)
+    submit_jobscript_n_times(15)
 
     pass

### 2020-04-09 paper/draft.tex
Added paper draft with bullet points 1

\documentclass[]{article}
\usepackage[left=3cm,right=3cm,top=1.5cm,bottom=2cm,includeheadfoot]{geometry} 
\usepackage{babel}
\usepackage{hyperref}
\usepackage{mwe}
\usepackage[markcase=noupper
]{scrlayer-scrpage}
\usepackage{amsmath}
\usepackage{amssymb}

\KOMAoptions{
	headsepline = true
}
\ihead{Tilman Hinnerichs}
\ohead{Combining bottom-up and top-down approaches}
\cfoot*{\pagemark}
%opening
\title{ECCB paper: Combining bottom-up and top-down approaches for drug-target-interaction prediction}
\author{Tilman Hinnerichs\\
\href{mailto:tilman@hinnerichs.com}{tilman@hinnerichs.com}}
\date{}
\pagestyle{headings}

% Build subsubsubsection
\usepackage{titlesec}

\titleclass{\subsubsubsection}{straight}[\subsection]

\newcounter{subsubsubsection}[subsubsection]
\renewcommand\thesubsubsubsection{\thesubsubsection.\arabic{subsubsubsection}}
\renewcommand\theparagraph{\thesubsubsubsection.\arabic{paragraph}} % optional; useful if paragraphs are to be numbered

\titleformat{\subsubsubsection}
{\normalfont\normalsize\bfseries}{\thesubsubsubsection}{1em}{}
\titlespacing*{\subsubsubsection}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{5}{\z@}%
	{3.25ex \@plus1ex \@minus.2ex}%
	{-1em}%
	{\normalfont\normalsize\bfseries}}
\renewcommand\subparagraph{\@startsection{subparagraph}{6}{\parindent}%
	{3.25ex \@plus1ex \@minus .2ex}%
	{-1em}%
	{\normalfont\normalsize\bfseries}}
\def\toclevel@subsubsubsection{4}
\def\toclevel@paragraph{5}
\def\toclevel@paragraph{6}
\def\l@subsubsubsection{\@dottedtocline{4}{7em}{4em}}
\def\l@paragraph{\@dottedtocline{5}{10em}{5em}}
\def\l@subparagraph{\@dottedtocline{6}{14em}{6em}}
\makeatother

\setcounter{secnumdepth}{4}
\setcounter{tocdepth}{4}

%\usepackage{fancyhdr}

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}

% Algorithm
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\begin{document}

	\title{ \normalsize \textsc{MPI: MiS}
		\\ [0.5cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{Combining bottom-up and top-down approaches for drug-target-interaction prediction}}
		\HRule{0.5pt} \\ [0.5cm]
		\normalsize }

\maketitle

\section{Recent work}
	\begin{itemize}
		\item Write about how to find suitable representations for both drugs and proteins
		\item Recent papers (see survey paper from briefings in Bioinformatics) $\rightarrow$ use either bottom up or top down
		\item best bottom up and top down results come from which papers?
		\item What are problems of recent approaches?
		\begin{itemize}
			\item Lack ability to generalize (bottom-up) or are unable to spot small differences (top-down)
			\item only top down or bottom-up
			\item not making use of interaction networks
		\end{itemize}
		\item They all do split over drugs and not over proteins $\rightarrow$ why is that bad? How can that be omitted? What consequences does that have for new drug-target pairs? 
		\begin{itemize}
			\item its feasible to engineer sophisticated representations for drugs, but not really for the proteins, both in bottom-up and in top-down approaches
			\item Thus, we will focus on building proper representations for the proteins
		\end{itemize}
		\item Additionally, PPI-graphs have been used in recent works to do \dots, but have found no application in dti prediction. This also applies for DDI-graphs With more and more uprising Graph Learning approaches, we can learn more sophisticated representations and test more complicated hypothesis.

	\end{itemize}

\section{How to combine mutual exclusive approaches}
\subsection{Problem description}
	\begin{itemize}
		\item What is dti prediction? $\rightarrow$ problem description
		\item Emphasize the importance of ability to generalize
		\item Issues with data in general $\rightarrow$ Open vs. closed world assumption. (To specific? Can we address this issue?)
		\item 
	\end{itemize}
\subsection{Issues to be solved}
	\begin{itemize}
		\item We are testing two hypothesis at the same time:
			\begin{itemize}
				\item Can we use PPI and DDI graphs to learn a protein representation that generalizes better?
				\item Can the built filter increase the performance of that very method? (Find better formulation)
			\end{itemize}
	\end{itemize}
\subsection{Role of interaction networks}
\subsubsection{DDI}
	\begin{itemize}
		\item What are DDI networks? What are they able to tell?
		\item Why are they important?
		\item Where do they come from? $\rightarrow$ Boyce et al., Maybe repeat some ideas from their paper
	\end{itemize}
\subsubsection{PPI}
	\begin{itemize}
		\item What are PPI networks? What can they tell? What does a link in the graph mean? Why would that be important to our problem? (Role of pathways)
		\item Recent work on PPIs? Add some ideas from them.
		\item What are we actually trying to find? (Regions within PPI network that are of interest for drug, \textbf{OR} patterns in the PPI graph that are of interest for the drug)
	\end{itemize}
\subsection{Finding node features for the PPI graph}
\subsubsection{Encoding of number of neighbours}
	\begin{itemize}
		\item Encode number of neighbours in node features $\rightarrow$ Many neighbours $\implies$ well studied and possibly many targets
	\end{itemize}
\subsubsection{How to build a drug specific filter (optional)}
	\begin{itemize}
		\item STITCH got 390.000 chemicals, 3.6 million proteins from over 2000 organisms
		\item for each drug find a motif in the target amino acid squences and query against all proteins and use the results as features for the prediction
		\item Why do need those filters? Why should they work (Discuss with Robert)? How do they help our prediction? 
		\item write down algorithm that is used: alignment $\rightarrow$ HMM $\rightarrow$ query
	\end{itemize}
\subsection{Features for drugs}
	Similarity scores over MedDRA data. 
	\begin{itemize}
		\item Why can they be useful? How did we obtain them?
	\end{itemize}

\section{Models}
	Transductive vs inductive models
	\begin{itemize}
		\item When is what kind of model better for the model? For which one of the two hypothesis? (transductive for patterns and inductive for regions of interest)
		\item What architectures are used? (Plain convolution for patterns and Attention based fo regions)
		\item Maybe some images of model structure
	\end{itemize}

\section{Experiments}
None yet

\section{Discussion}

\section{Future Work}

\end{document}

### 2020-04-09 src/run_job_pipeline.py

@@ -73,6 +73,8 @@ def submit_gpu_job(num_proteins=-1,
 #SBATCH --gres=gpu:v100:{num_gpus}
 #SBATCH --mem-per-gpu=60G
 #SBATCH --constraint=[gpu]
+#SBATCH --sockets-per-node=1
+#SBATCH --gpus-per-socket={num_gpus}
 #SBATCH --cpus-per-gpu=6
 
 #run the application:
@@ -83,7 +85,7 @@ conda activate ~/.conda/envs/dti/
 module load cuda/10.0.130
 
 export CUDA_VISIBLE_DEVICES={vis_dev}
-python3 torch_dti_predictor.py '''.format(jobname=jobname, days=str(days), num_gpus=str(num_gpus), vis_dev=str(list(range(num_gpus)))[1:-1])
+python3 torch_dti_predictor.py '''.format(jobname=jobname, days=str(days), num_gpus=str(num_gpus), vis_dev=str(list(range(num_gpus)))[1:-1].replace(' ', ''))
     preface_script += "--num_proteins {num_prots} ".format(num_prots=str(num_proteins))
     if epochs:
         preface_script += "--num_epochs={num_epochs} ".format(num_epochs=str(epochs))

### 2020-04-19 paper/main.tex
Added paper template, rerun gpu jobs with higher batch size 1

\documentclass{bioinfo}
\copyrightyear{2015} \pubyear{2015}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Subject Section}

\title[short Title]{This is a title}
\author[Sample \textit{et~al}.]{Corresponding Author\,$^{\text{\sfb 1,}*}$, Co-Author\,$^{\text{\sfb 2}}$ and Co-Author\,$^{\text{\sfb 2,}*}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
$^{\text{\sf 2}}$Department, Institution, City, Post Code,
Country.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:} Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text
Text Text Text Text Text.\\
\textbf{Results:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text\\
\textbf{Availability:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text Text  Text\\
\textbf{Contact:} \href{name@bio.com}{name@bio.com}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

\maketitle

\section{Introduction}

The challenge of

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text. Figure~\ref{fig:01}
shows that the above method  Text Text Text Text  Text Text Text
Text Text Text  Text Text.  \citep{Bag01} wants to know about
{\ldots}{\ldots} text follows.
\begin{equation}
\sum \text{\it x}+ \text{\it y} =\text{\it Z}\label{eq:01}\vspace*{-10pt}
\end{equation}

%\enlargethispage{12pt}

\section{Approach}

Equation~(\ref{eq:01}) Text Text Text Text Text Text  Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
.....

\begin{methods}
\section{Methods}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text
text\vspace*{1pt}

\begin{itemize}
\item for bulleted list, use itemize
\item for bulleted list, use itemize
\item for bulleted list, use itemize\vspace*{1pt}
\end{itemize}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text Text Text Text Text Text Text Text Text Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text\vadjust{\newpage}.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text Text Text Text Text Text Text Text Text Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

\subsection{This is subheading}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text
text. Text Text Text Text Text Text  Text Text Text Text Text Text
Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

\subsubsection{This is subsubheading}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text\break
text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text
textText Text Text Text Text Text  Text Text Text Text Text Text
Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

\enlargethispage{6pt}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\begin{table}[!t]
\processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
head2 & head3 & head4\\\midrule
row1 & row1 & row1 & row1\\
row2 & row2 & row2 & row2\\
row3 & row3 & row3 & row3\\
row4 & row4 & row4 & row4\\\botrule
\end{tabular}}{This is a footnote}
\end{table}

\end{methods}

\begin{figure}[!tpb]%figure1
\fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
\end{minipage}}
%\centerline{\includegraphics{fig01.eps}}
\caption{Caption, caption.}\label{fig:01}
\end{figure}

%\begin{figure}[!tpb]%figure2
%%\centerline{\includegraphics{fig02.eps}}
%\caption{Caption, caption.}\label{fig:02}
%\end{figure}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\subsection{Test1}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text.

Table~\ref{Tab:01} shows that Text Text Text Text Text  Text Text
Text Text Text Text. Figure~2\vphantom{\ref{fig:02}} shows that
the above method Text Text. Text Text Text  Text Text Text Text
Text Text. Figure~2\vphantom{\ref{fig:02}} shows that the above
method Text Text. Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
Text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
%     as it may ignore the figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

(Table~\ref{Tab:01}) Text Text Text Text Text Text  Text Text Text
Text Text Text Text Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text
Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about  text text text text

\begin{enumerate}
\item this is item, use enumerate
\item this is item, use enumerate
\item this is item, use enumerate
\end{enumerate}

Text Text Text Text Text Text Text Text Text Text Text Text Text
Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows\vadjust{\pagebreak} that the
above method  Text Text Text Text Text Text Text Text Text Text
Text Text.  \citealp{Boffelli03} might want to know about text
text text text Text Text Text Text Text Text  Text Text Text Text
Text Text Text Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text Text Text Text Text Text\break Text.

Text Text Text Text Text Text  Text Text Text Text Text Text Text
Text Text  Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text\vspace*{-10pt}

\section*{Acknowledgements}

Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text
text text text\vspace*{-12pt}

\section*{Funding}

This work has been supported by the... Text Text  Text Text.\vspace*{-12pt}

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
%\bibliography{Document}

\begin{thebibliography}{}

\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.

\bibitem[Bag {\it et~al}., 2001]{Bag01}
Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.

\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
in nigral dopaminergic neurnol cell: correlation with the known
pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
Res.}, \textbf{110}(Suppl. 1), 76--84.

\bibitem[Lehmann, 1986]{Leh86}
Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.

\bibitem[Crenshaw and Jones, 2003]{Cre03}
Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
cancer management: one tumor, one chip. \textit{Bioinformatics},
doi:10.1093/bioinformatics/btn000.

\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
???--???.

\bibitem[Bardet, 1920]{Bar20}
Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
polydactylie et retinite pigmentaire (contribution a l'etude des
formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
institution, Paris, France.

\end{thebibliography}
\end{document}

### 2020-04-19 src/run_job_pipeline.py

@@ -156,17 +156,17 @@ if __name__ == '__main__':
     cancel_jobs()
     # 'ChebConv','GraphConv', 'TAGConv', 'ARMAConv', 'SGConv', 'FeaStConv'
     for arch in ['GCNConv','SAGEConv', 'GATConv']:
-        submit_gpu_job(epochs=30, batch_size=32, days=2, arch=arch, mode='protein_drughub')
-        submit_gpu_job(epochs=30, batch_size=32, days=2, arch=arch, mode='drug_drughub')
+        submit_gpu_job(epochs=30, batch_size=80, days=2, arch=arch, mode='protein_drughub')
+        submit_gpu_job(epochs=30, batch_size=80, days=2, arch=arch, mode='drug_drughub')
 
         for fold in range(1, 3):
             # submit_gpu_job(epochs=30, batch_size=32, days=2, arch=arch, mode='drug', fold=fold)
 
-            submit_gpu_job(epochs=30, batch_size=32, days=2, arch=arch, fold=fold)
+            submit_gpu_job(epochs=30, batch_size=80, days=2, arch=arch, fold=fold)
             # submit_gpu_job(num_proteins=4000, epochs=30, batch_size=64, arch=arch)
             # submit_gpu_job(num_proteins=1000, epochs=30, batch_size=256, arch=arch)
 
-            submit_gpu_job(epochs=30, batch_size=32, days=2, arch='Res'+arch, fold=fold)
+            submit_gpu_job(epochs=30, batch_size=80, days=2, arch='Res'+arch, fold=fold)
             # submit_gpu_job(num_proteins=4000, epochs=30, batch_size=64, arch='Res'+arch)
             # submit_gpu_job(num_proteins=1000, epochs=30, batch_size=256, arch='Res'+arch)
 

### 2020-04-19 src/torch_dti_predictor.py

@@ -121,9 +121,9 @@ def transductive_missing_target_predictor(config,
             loss = train(model=model, device=device, train_loader=train_loader, optimizer=optimizer, epoch=epoch, weight_dict=weight_dict)
             print('Train loss:', loss)
 
-            if epoch%config.num_epochs == 0:
+            if epoch%10 == 0:
                 print('Predicting for validation data...')
-                file='../results/full_interactions_results_' +config.arch+'_'+ str(num_proteins) + '_prots_'+str(config.num_epochs)+'_epochs'
+                file='../results/full_interactions_results_' +config.arch+'_'+ str(num_proteins) + '_prots_'+str(epoch)+'_epochs'
                 with open(file=file, mode='a') as f:
                     train_labels, train_predictions = predicting(model, device, train_loader)
                     print('Train:', 'Acc, ROC_AUC, f1, matthews_corrcoef',
@@ -269,9 +269,9 @@ def transductive_missing_drug_predictor(config,
                          weight_dict=weight_dict)
             print('Train loss:', loss)
 
-            if epoch % config.num_epochs == 0:
+            if epoch % 10 == 0:
                 print('Predicting for validation data...')
-                file = '../results/full_interactions_drug_split_results_' + config.arch + '_' + str(num_proteins) + '_prots_' + str(config.num_epochs) + '_epochs'
+                file = '../results/full_interactions_drug_split_results_' + config.arch + '_' + str(num_proteins) + '_prots_' + str(epoch) + '_epochs'
                 with open(file=file, mode='a') as f:
                     train_labels, train_predictions = predicting(model, device, train_loader)
                     print('Train:', 'Acc, ROC_AUC, f1, matthews_corrcoef',
@@ -412,9 +412,9 @@ def test_predictor_on_drughub_protein_data(config):
                      weight_dict=weight_dict)
         print('Train loss:', loss)
 
-        if epoch % config.num_epochs == 0:
+        if epoch % 10 == 0:
             print('Predicting for validation data...')
-            file = '../results/drughub_protein_split_' + config.arch + '_prots_' + str(config.num_epochs) + '_epochs'
+            file = '../results/drughub_protein_split_' + config.arch + '_prots_' + str(epoch) + '_epochs'
             with open(file=file, mode='a') as f:
                 train_labels, train_predictions = predicting(model, device, train_loader)
                 print('Train:', 'Acc, ROC_AUC, f1, matthews_corrcoef',
[…]

### 2020-04-19 paper/main.tex
Submitting smaller jobs 1

@@
 
-The challenge of
+\cite{Boffelli03}
 
@@
 %
-%\bibliography{Document}
+\bibliography{Document}
 
-\begin{thebibliography}{}
-
-\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
-Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
-
-\bibitem[Bag {\it et~al}., 2001]{Bag01}
-Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.
-
-\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
-Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
-in nigral dopaminergic neurnol cell: correlation with the known
-pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
-Res.}, \textbf{110}(Suppl. 1), 76--84.
-
-\bibitem[Lehmann, 1986]{Leh86}
-Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.
-
-\bibitem[Crenshaw and Jones, 2003]{Cre03}
-Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
-cancer management: one tumor, one chip. \textit{Bioinformatics},
-doi:10.1093/bioinformatics/btn000.
-
-\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
-Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
-(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
-???--???.
-
-\bibitem[Bardet, 1920]{Bar20}
-Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
-polydactylie et retinite pigmentaire (contribution a l'etude des
-formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
-institution, Paris, France.
-
-\end{thebibliography}
 \end{document}

### 2020-04-19 src/run_job_pipeline.py

@@ -153,20 +153,20 @@ if __name__ == '__main__':
     # cancel_jobs()
     # submit_jobscript_n_times(50)
 
-    cancel_jobs()
+    # cancel_jobs()
     # 'ChebConv','GraphConv', 'TAGConv', 'ARMAConv', 'SGConv', 'FeaStConv'
     for arch in ['GCNConv','SAGEConv', 'GATConv']:
         # submit_gpu_job(epochs=6, batch_size=80, days=2, arch=arch, mode='protein_drughub')
         # submit_gpu_job(epochs=6, batch_size=80, days=2, arch=arch, mode='drug_drughub')
 
-        for fold in range(1, 2):
+        for fold in range(1, 3):
             # submit_gpu_job(epochs=30, batch_size=32, days=2, arch=arch, mode='drug', fold=fold)
 
-            submit_gpu_job(epochs=6, batch_size=160, days=2, arch=arch, fold=fold, num_gpus=8)
+            submit_gpu_job(epochs=10, batch_size=8, days=2, arch=arch, fold=fold, num_gpus=4)
             # submit_gpu_job(num_proteins=4000, epochs=30, batch_size=64, arch=arch)
             # submit_gpu_job(num_proteins=1000, epochs=30, batch_size=256, arch=arch)
 
-            submit_gpu_job(epochs=6, batch_size=160, days=2, arch='Res'+arch, fold=fold, num_gpus=8)
+            submit_gpu_job(epochs=10, batch_size=8, days=2, arch='Res'+arch, fold=fold, num_gpus=4)
             # submit_gpu_job(num_proteins=4000, epochs=30, batch_size=64, arch='Res'+arch)
             # submit_gpu_job(num_proteins=1000, epochs=30, batch_size=256, arch='Res'+arch)
 

### 2020-04-19 src/torch_networks.py

@@ -23,48 +23,48 @@ class TemplateSimpleNet(torch.nn.Module):
 
         # GCN layers
         if 'GCNConv' in conv_method:
-            self.conv1 = nn.GCNConv(num_features, num_features*8, cached=False)
-            self.conv2 = nn.GCNConv(num_features*8, num_features*64, cached=False)
-            self.conv3 = nn.GCNConv(num_features*64, num_features*256, cached=False)
+            self.conv1 = nn.GCNConv(num_features, num_features*4, cached=False)
+            self.conv2 = nn.GCNConv(num_features*4, num_features*16, cached=False)
+            self.conv3 = nn.GCNConv(num_features*16, num_features*128, cached=False)
         elif 'ChebConv' in conv_method:
-            self.conv1 = nn.ChebConv(num_features, num_features*8, 3)
-            self.conv2 = nn.ChebConv(num_features*8, num_features*64, 3)
-            self.conv3 = nn.ChebConv(num_features*64, num_features*256, 3)
+            self.conv1 = nn.ChebConv(num_features, num_features*4, 3)
+            self.conv2 = nn.ChebConv(num_features*4, num_features*16, 3)
+            self.conv3 = nn.ChebConv(num_features*16, num_features*128, 3)
         elif 'SAGEConv' in conv_method:
             self.conv1 = nn.SAGEConv(num_features, num_features*8)
-            self.conv2 = nn.SAGEConv(num_features*8, num_features*64)
-            self.conv3 = nn.SAGEConv(num_features*64, num_features*256)
+            self.conv2 = nn.SAGEConv(num_features*4, num_features*64)
+            self.conv3 = nn.SAGEConv(num_features*16, num_features*128)
         elif 'GraphConv' in conv_method:
             self.conv1 = nn.GraphConv(num_features, num_features*8)
-            self.conv2 = nn.GraphConv(num_features*8, num_features*64)
-            self.conv3 = nn.GraphConv(num_features*64, num_features*256)
+            self.conv2 = nn.GraphConv(num_features*4, num_features*64)
+            self.conv3 = nn.GraphConv(num_features*16, num_features*128)
         elif 'GATConv' in conv_method:
-            self.conv1 = nn.GATConv(num_features, num_features*8, heads=5)
-            self.conv2 = nn.GATConv(num_features*8, num_features*64, heads=5)
-            self.conv3 = nn.GATConv(num_features*64, num_features*256, heads=5)
[…]

### 2020-04-19 paper/main.tex
First paragraphs of paper

@@
 \documentclass{bioinfo}
+\usepackage[english]{babel}
 \copyrightyear{2015} \pubyear{2015}
+\bibliographystyle{plainnat}
+
+\usepackage{natbib}
 
@@
 
-\title[short Title]{This is a title}
-\author[Sample \textit{et~al}.]{Corresponding Author\,$^{\text{\sfb 1,}*}$, Co-Author\,$^{\text{\sfb 2}}$ and Co-Author\,$^{\text{\sfb 2,}*}$}
+\title[short Title]{Combining Bottom-up and top-down approaches for drug-target-interaction prediction}
+\author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
 \address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
@@
 Text Text Text Text Text Text Text Text Text Text Text Text Text Text  Text\\
-\textbf{Contact:} \href{name@bio.com}{name@bio.com}\\
-\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
+\textbf{Contact:} \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}\\
+\textbf{Supplementary information:}10264703 Supplementary data are available at \textit{Bioinformatics}
 online.}
@@
 
-\cite{Boffelli03}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text. Figure~\ref{fig:01}
-shows that the above method  Text Text Text Text  Text Text Text
-Text Text Text  Text Text.  \citep{Bag01} wants to know about
-{\ldots}{\ldots} text follows.
-\begin{equation}
-\sum \text{\it x}+ \text{\it y} =\text{\it Z}\label{eq:01}\vspace*{-10pt}
-\end{equation}
+\cite{Survey2018}\\
+In history, traditional remedies, that were known for their medicinal properties lead to drugs by extraction of the functional ingredients.
+Alternatively, characteristics and features of potential drugs were detected by accident.
+More recently, biological drug targets were found \textit{in silico} through discovery of suitable computational predictors.\\
+The challenge of accurately predicting drug-target-interactions (DTI) has shown its importance in the fields of drug repurposing and repositioning, and in the exploration of novel drugs and their interaction partners.
+Knowledge about those links between compounds and their target proteins help in an array of medical and pharmaceutical studies.
+Additionally, those associations can be utilized to identify disease specific targets, leading to desirable therapeutic effects.\\
+With the rapidly growing field of various machine learning approaches and their application to bioscientifical problems in the realm of bioinformatics, rapid advances were made.
+Almost all state of the art models for drug-target-interaction prediction were based on the usage of neural networks with increasing size.\\
+Approaches on this rather sophisticated problem are either based on top-down or network approaches (\textbf{CITATION}), or bottom-up and molecular approaches (\textbf{CITATION}).
+However, both kinds contain some problems, that are not solvable within their own system.
+Thus, bottom-up approaches share the lack of ability to generalize, which we will show in later sections.
+This
 
@@
 
-Equation~(\ref{eq:01}) Text Text Text Text Text Text  Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-.....
-
-\begin{methods}
 \section{Methods}
 
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text
-text\vspace*{1pt}
-
-\begin{itemize}
-\item for bulleted list, use itemize
-\item for bulleted list, use itemize
-\item for bulleted list, use itemize\vspace*{1pt}
-\end{itemize}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
-Text Text Text Text Text Text Text Text Text Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text\vadjust{\newpage}.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
-Text Text Text Text Text Text Text Text Text Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-\subsection{This is subheading}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text
-text. Text Text Text Text Text Text  Text Text Text Text Text Text
-Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-\subsubsection{This is subsubheading}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text\break
-text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text
-textText Text Text Text Text Text  Text Text Text Text Text Text
-Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-\enlargethispage{6pt}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-\begin{table}[!t]
-\processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
-head2 & head3 & head4\\\midrule
-row1 & row1 & row1 & row1\\
-row2 & row2 & row2 & row2\\
-row3 & row3 & row3 & row3\\
-row4 & row4 & row4 & row4\\\botrule
-\end{tabular}}{This is a footnote}
-\end{table}
-
-\end{methods}
-
-\begin{figure}[!tpb]%figure1
-\fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
-235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
-\end{minipage}}
-%\centerline{\includegraphics{fig01.eps}}
-\caption{Caption, caption.}\label{fig:01}
-\end{figure}
-
-%\begin{figure}[!tpb]%figure2
-%%\centerline{\includegraphics{fig02.eps}}
-%\caption{Caption, caption.}\label{fig:02}
-%\end{figure}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-\subsection{Test1}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
 \section{Discussion}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text.
-
-Table~\ref{Tab:01} shows that Text Text Text Text Text  Text Text
-Text Text Text Text. Figure~2\vphantom{\ref{fig:02}} shows that
-the above method Text Text. Text Text Text  Text Text Text Text
-Text Text. Figure~2\vphantom{\ref{fig:02}} shows that the above
-method Text Text. Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method Text
-Text.
 
@@
 
-(Table~\ref{Tab:01}) Text Text Text Text Text Text  Text Text Text
-Text Text Text Text Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text  Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about  text text text text
-
-\begin{enumerate}
-\item this is item, use enumerate
-\item this is item, use enumerate
-\item this is item, use enumerate
-\end{enumerate}
-
-Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows\vadjust{\pagebreak} that the
-above method  Text Text Text Text Text Text Text Text Text Text
-Text Text.  \citealp{Boffelli03} might want to know about text
-text text text Text Text Text Text Text Text  Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text Text Text Text Text Text Text  Text Text.
-\citealp{Boffelli03} might want to know about text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text\break Text.
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text
-Text Text  Text Text Text Text Text Text.
-Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
-Text Text Text\vspace*{-10pt}
+\vspace*{-10pt}
 
@@
 
-Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text
-text text text\vspace*{-12pt}
+\vspace*{-12pt}
 
@@
 %
-\bibliography{Document}
+\bibliography{citations}
 

### 2020-04-19 paper/main.tex
2nd paragraph for introduction

@@
 Additionally, those associations can be utilized to identify disease specific targets, leading to desirable therapeutic effects.\\
-With the rapidly growing field of various machine learning approaches and their application to bioscientifical problems in the realm of bioinformatics, rapid advances were made.
+With the rapidly growing field of machine learning approaches and their application to bioscientifical problems in the realm of bioinformatics, different kinds of data, such as long DNA sequences could be utilized for feature generation, while rapid advances were made.
 Almost all state of the art models for drug-target-interaction prediction were based on the usage of neural networks with increasing size.\\
-Approaches on this rather sophisticated problem are either based on top-down or network approaches (\textbf{CITATION}), or bottom-up and molecular approaches (\textbf{CITATION}).
-However, both kinds contain some problems, that are not solvable within their own system.
-Thus, bottom-up approaches share the lack of ability to generalize, which we will show in later sections.
-This
+Only recently, the technique of graph learning was introduced by (T.
+Kipf et.
+al) through graph convolution algorithms, and improved and altered under usage of attention mechanisms (GAT), random walks (SAGE, APPNP, PPNP) and mixtures of both (copy pytorch geometric citations).
+While based on diverse systems, they can be relevant for testing distinct hypothesis for given graphs.
+While convolutional filters are suitable for finding patterns among the the given graph, attention mechanisms are more relevant for discovery of important regions within.
+Lately, graph learning approaches found application for computing compound representations for DTI prediction.\\
+Approaches on this rather sophisticated problem can divided into top-down or network approaches (\textbf{CITATION}), and bottom-up or molecular approaches (\textbf{CITATION}).
+Top-down approaches take advantage of other data such as diseases (CITATION), side effects, knowledge graphs or ontologies, in order to learn representations for both compound and protein.
+On the other hand, bottom-up aspirations attempt to learn from chemical properties.
+For drugs, molecular structure (CITATION GraphDTA), molecular fingerprints, similarity to other drugs (See Bioinf Survey), and other molecular features.
+On the protein side, secondary structure prediction (CITATION), contact prediction (CITATION), or simply convolution over the amino acid sequences served to obtain a feature for the given proteins.
+However, both kinds contain and share some problems, that are not solvable within themselves.
+Thus, bottom-up approaches share the lack of ability to generalize, which we will show in later sections, and usually focus on engineering sophisticated features for the drugs, while neglecting to formulate meaningful features on the protein side.
+Top-down approaches lack the ability to spot small differences to cope with small differences within the drug structure and rely heavily on given data for the considered drug-target pair.
+The latter is not suitable for predictions on novel or unseen compounds, as e.g., data on side effects or its impact on diseases is seldom given for novel drugs.\\
+In order to design such a feature for proteins and drugs, respectively, we make use of the interaction networks for both proteins and compounds.
+Drug-drug interaction networks were introduced by (Boyce et al.
+CITATION) have have been used for $\dots$ (CITATION).
+Drug-drug interaction networks may give a hint on common targeted pathways or $\dots$ (See Boyce paper what they could be used for).
+As an additional compound feature we will use semantic side effect similarity, which we will discuss later on.\\
+Protein-protein interaction networks have shown great results in $\dots$(CITATION, copy some stuff and ideas from them ).
 

### 2020-04-20 paper/main.tex
Starting Approach paragraph and added citations for introduction part

@@
 In history, traditional remedies, that were known for their medicinal properties lead to drugs by extraction of the functional ingredients.
-Alternatively, characteristics and features of potential drugs were detected by accident.
-More recently, biological drug targets were found \textit{in silico} through discovery of suitable computational predictors.\\
+Alternatively, characteristics and features of potential drugs were detected by accident like in the case of penicillin.
+More recently, biological drug targets can be found \textit{in silico} through discovery of suitable computational predictors.\\
 The challenge of accurately predicting drug-target-interactions (DTI) has shown its importance in the fields of drug repurposing and repositioning, and in the exploration of novel drugs and their interaction partners.
@@
 Almost all state of the art models for drug-target-interaction prediction were based on the usage of neural networks with increasing size.\\
-Only recently, the technique of graph learning was introduced by (T.
-Kipf et.
-al) through graph convolution algorithms, and improved and altered under usage of attention mechanisms (GAT), random walks (SAGE, APPNP, PPNP) and mixtures of both (copy pytorch geometric citations).
+Only recently, the technique of graph learning was introduced by \citet{GCNConv} through graph convolution algorithms, and improved and altered under usage of different kernels(\cite{ChebConv}, \cite{ARMAConv}) attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of both (\cite{SAGEConv}).
 While based on diverse systems, they can be relevant for testing distinct hypothesis for given graphs.
@@
 In order to design such a feature for proteins and drugs, respectively, we make use of the interaction networks for both proteins and compounds.
-Drug-drug interaction networks were introduced by (Boyce et al.
-CITATION) have have been used for $\dots$ (CITATION).
-Drug-drug interaction networks may give a hint on common targeted pathways or $\dots$ (See Boyce paper what they could be used for).
+Drug-drug interaction networks were introduced and standardized by \cite{Boyce2015} and have been used for clinical decision support (\cite{Scheife2015}).
+Drug-drug interaction networks may give a hint on common targeted pathways.
 As an additional compound feature we will use semantic side effect similarity, which we will discuss later on.\\
-Protein-protein interaction networks have shown great results in $\dots$(CITATION, copy some stuff and ideas from them ).
+Protein-protein interaction networks have shown great results in $\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context for molecular system biology.
+However, these contexts were never applied to the problem of drug-target-interaction prediction.
+Thus we formalized our hypotheses over these interaction graphs and will test them in the following chapters.
 
 %\enlargethispage{12pt}
+
+\section{Approach}
+As shown, recent work lacks the ability to combine both top-down and bottom-up approaches for their predictions, however performing quite well on their datasets.
+When building the train-test split over compound-protein pairs, there are the following options:
+\begin{enumerate}
+	\item Build split over drugs
+	\item Build split over drug-target pairs
+	\item Build split over proteins
+\end{enumerate} 
+In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
+The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
+However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier.
+Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE).
+When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
+Hence, when entirely new molecules are seen, the models perform much worse.
+\\
+The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.\\
+As mentioned in the introduction it is quite difficult to learn suitable features from proteins.
+In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features.
+\\
+Thus, building splitting over proteins is the most challenging of the three options.
+
+In this work we are testing the following hypotheses:
+\begin{enumerate}
+	\item Can we build a best performing model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
+	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
+\end{enumerate}
+We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account. For a
 

### 2020-04-20 paper/main.tex
Updated Approach and Method chapter

@@
 \section{Approach}
+\subsection{Choosing appropriate train-test splits}
 As shown, recent work lacks the ability to combine both top-down and bottom-up approaches for their predictions, however performing quite well on their datasets.
-When building the train-test split over compound-protein pairs, there are the following options:
+Drug-target interaction prediction is the task to accurately predict whether for a given drug and a given protein there is a biological interaction within humans.
+However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
 \begin{enumerate}
@@
 \\
-Thus, building splitting over proteins is the most challenging of the three options.
+Thus, building splitting over proteins is the most challenging of the three options. \\
+
+\subsection{Tested hypotheses}
 
@@
 \begin{enumerate}
-	\item Can we build a best performing model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
+	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
 	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
 \end{enumerate}
-We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account. For a
+We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account.
+Thus, we propose a novel approach to combine those mutual exclusive attempts, through the usage of interaction networks, similarity and molecular features.
+Additionally, we test the latter by building a simple molecular DTI predictor and enhance it under usage of the interaction networks.\\
 
-\section{Approach}
+For the bottom-up approach we build a that only relies only on molecular features, which we will discuss in more detail in the following methods chapter.
+For the combination of both approaches we now attach the predictions to the protein-protein interaction graph as node features for future graph learning steps.
+In this graph we tried to find both patterns and regions for each drug that could be of interest through application of different graph convolutional layers, which in return represent the feature for each protein.
+Representing the drug we take the drug-drug interaction graph and the semantic similarity over side effects which we will explain in the following paragraphs.
 
 \section{Methods}
+\subsection{Data}
+The data for the different parts of this model were obtained from different sources.
+Starting with the protein-protein interactions, we fetched 11.574 proteins with over 170.000 links from STRING (\cite{STRINGv10}).
+As STRING provides statistical scores for each association, we filtered them as advised by a threshold of 700.
+For the drug-drug interaction we retrieved the drugs from \cite{Boyce2015}.
+Side effects were obtained from MedDRA database (\cite{MedDRA}) that hierarchically orders side effects.
+Links between drugs and side effects were fetched from \textit{SIDER} (\cite{SIDER}) side effects database.
+The intersection between these sources yielded 641 common drugs for usage in our data.
+For the drug-target interactions themselves, we fetched
+137.000 links from STITCH database (\cite{STITCHv5}). We provide links and methods to and for the necessary data in the provided Github repository.\\
+
+\subsection{Models}
+The used model consists of two separate models, that help to fuse together the two methods:
+\begin{enumerate}
+	\item The molecular predictor
+	\item The interaction network based predictor
+\end{enumerate}
+We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins.
+Regarding proteins, we used the pretrained feature generator from \textit{DeepGoPlus} (\citep{DeepGoPlus}) that was originally designed for protein function prediction and is regarded as state of the art for this purpose.
+For drugs we used a pretrained fingerprint model from SMILES transformer (\cite{SmilesTransformer}), that provides a simple and fast method to compute fingerprints through autoencoder models.
+The encodings from these two models were funneled into a simple deep neural network (TABLE) with few fully connected.
+\\
+The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE).
+Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug.
+Thus, given a compound-target pair the nodes of the PPI graph now hold bottom-up features for
 

### 2020-04-20 paper/main.tex
Update Methods paragraph 1

@@
 \copyrightyear{2015} \pubyear{2015}
-\bibliographystyle{plainnat}
-
 \usepackage{natbib}
+\bibliographystyle{natbib.bst}
 
@@
 Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug.
-Thus, given a compound-target pair the nodes of the PPI graph now hold bottom-up features for
+Thus, given a compound-target pair, the nodes of the PPI graph now hold bottom-up features, which can now be processed by the graph algorithms.
+\\
+The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node.
+In contrast to learning over whole graphs we perform node classification within the graph.
+\\
+The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
+For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy.
+Concerning this hierarchy, drug-drug similarity is computed by the Resnik similarity (\cite{Resnik1995}).
+For the given compound we take the corresponding row of this symmetric similarity matrix.
+\\
+
+Thereby, we concatenate these three features together and funnel them into another deep neural network.
 

### 2020-04-22 paper/main.tex
Added Figures 1

@@
 
-\title[short Title]{Combining Bottom-up and top-down approaches for drug-target-interaction prediction}
+\title[short Title]{Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
 \author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
@@
 For drugs we used a pretrained fingerprint model from SMILES transformer (\cite{SmilesTransformer}), that provides a simple and fast method to compute fingerprints through autoencoder models.
-The encodings from these two models were funneled into a simple deep neural network (TABLE) with few fully connected.
+The encodings from these two models were funneled into a simple deep neural network (see Figure~\ref{fig:01}) with few fully connected.
 \\
@@
 Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug.
-Thus, given a compound-target pair, the nodes of the PPI graph now hold bottom-up features, which can now be processed by the graph algorithms.
+Thus, given a compound-target pair, the nodes of the PPI graph now hold bottom-up features, which can now be processed by the graph learning algorithms.
 \\
@@
 
-Thereby, we concatenate these three features together and funnel them into another deep neural network.
+Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
+
+\begin{figure}[!tpb]%figure1
+
+	\centerline{\includegraphics[width=0.5\textwidth]{figures/MolecularPredictor.png}}
+	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:01}
+\end{figure}
+
+\begin{figure}[!tpb]%figure2
+
+\centerline{\includegraphics[width=0.5\textwidth]{figures/InteractionNetwork.png}}
+\caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.
+Protein and drug features are represented by blue and red, respectively.}\label{fig:02}
+\end{figure}
 

### 2020-04-22 paper/main.tex
Updated methods part 1

@@
 In contrast to learning over whole graphs we perform node classification within the graph.
+These layers are either graph convolutional layers, that learn a certain kernel over the graph, or attention based.
+Different layers of both and other types such as were tested.
 \\
@@
 Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
+This network finally yields our prediction.
+We hereby perform splits over both drugs and proteins, in order to test and show the discrepancy and increasing difficulty.\\ Implementation was done in PyTorch (\cite{Pytorch}) and is available on Github under \href{github.com/thinnerichs/KAUST-dti-metabol}{github.com/thinnerichs/KAUST-dti-metabol}.
+Graph learning methods were build with help of PyTorch-Geometric (\cite{PytorchGeometric}), a geometric deep learning extension library for PyTorch, that recently got a lot of attention in the machine learning community.
+This library gives the potential to use many state of the art graph learning mechanisms, such as plain but effective graph convolution (\citet{GCNConv}), Chebychev kernels (\cite{ChebConv}), ARMA kernels (\cite{ARMAConv}), translation-invariant operators (\cite{FeaStConv}), attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of the latter two (\cite{SAGEConv}).
+The performance of these various layer types were tested for this particular problem, as discussed in the results section.
 
 \begin{figure}[!tpb]%figure1
-
-	\centerline{\includegraphics[width=0.5\textwidth]{figures/MolecularPredictor.png}}
+	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
 	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:01}
@@
 \begin{figure}[!tpb]%figure2
-
-\centerline{\includegraphics[width=0.5\textwidth]{figures/InteractionNetwork.png}}
+\centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
 \caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.

### 2020-04-22 paper/main.tex
Small proof read 1

@@
 \maketitle
-
 \section{Introduction}
@@
 Almost all state of the art models for drug-target-interaction prediction were based on the usage of neural networks with increasing size.\\
-Only recently, the technique of graph learning was introduced by \citet{GCNConv} through graph convolution algorithms, and improved and altered under usage of different kernels(\cite{ChebConv}, \cite{ARMAConv}) attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of both (\cite{SAGEConv}).
+Only recently, the technique of graph learning was introduced by \citet{GCNConv} through graph convolution algorithms, and improved and altered under usage of different kernels(\cite{ChebConv}, \cite{ARMAConv}), attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of both (\cite{SAGEConv}).
 While based on diverse systems, they can be relevant for testing distinct hypothesis for given graphs.
@@
 
-%\enlargethispage{12pt}
+\enlargethispage{12pt}
 
@@
 
-For the bottom-up approach we build a that only relies only on molecular features, which we will discuss in more detail in the following methods chapter.
+For the bottom-up approach we build a model that only relies on molecular features, which we will discuss in more detail in the following methods chapter.
 For the combination of both approaches we now attach the predictions to the protein-protein interaction graph as node features for future graph learning steps.
@@
 Regarding proteins, we used the pretrained feature generator from \textit{DeepGoPlus} (\citep{DeepGoPlus}) that was originally designed for protein function prediction and is regarded as state of the art for this purpose.
-For drugs we used a pretrained fingerprint model from SMILES transformer (\cite{SmilesTransformer}), that provides a simple and fast method to compute fingerprints through autoencoder models.
+For drugs we used a pretrained fingerprint model from \textit{SMILES transformer} (\cite{SmilesTransformer}), that provides a simple and fast method to compute fingerprints through autoencoder models.
 The encodings from these two models were funneled into a simple deep neural network (see Figure~\ref{fig:01}) with few fully connected.

### 2020-09-16 src/torch_networks.py
Update network for more expressivity 1

@@ -477,6 +477,8 @@ class QuickTemplateNodeFeatureNet(torch.nn.Module):
         # batch_size = drug_feature.size(0)
 
         PPI_x = F.relu(self.linear1(PPI_x))
+        PPI_x = F.relu(self.linear2(PPI_x))
+        PPI_x = F.relu(self.linear3(PPI_x))
         PPI_x = F.relu(self.linear4(PPI_x))
 
 

### 2021-01-05 paper/main.tex
comments on into

@@
 \appnotes{Manuscript Category}
+
+\renewcommand{\cite}{\citep}
+\usepackage{todonotes}
 
@@
 \cite{Survey2018}\\
-In history, traditional remedies, that were known for their medicinal properties lead to drugs by extraction of the functional ingredients.
-Alternatively, characteristics and features of potential drugs were detected by accident like in the case of penicillin.
-More recently, biological drug targets can be found \textit{in silico} through discovery of suitable computational predictors.\\
-The challenge of accurately predicting drug-target-interactions (DTI) has shown its importance in the fields of drug repurposing and repositioning, and in the exploration of novel drugs and their interaction partners.
-Knowledge about those links between compounds and their target proteins help in an array of medical and pharmaceutical studies.
-Additionally, those associations can be utilized to identify disease specific targets, leading to desirable therapeutic effects.\\
-With the rapidly growing field of machine learning approaches and their application to bioscientifical problems in the realm of bioinformatics, different kinds of data, such as long DNA sequences could be utilized for feature generation, while rapid advances were made.
-Almost all state of the art models for drug-target-interaction prediction were based on the usage of neural networks with increasing size.\\
-Only recently, the technique of graph learning was introduced by \citet{GCNConv} through graph convolution algorithms, and improved and altered under usage of different kernels(\cite{ChebConv}, \cite{ARMAConv}), attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of both (\cite{SAGEConv}).
-While based on diverse systems, they can be relevant for testing distinct hypothesis for given graphs.
-While convolutional filters are suitable for finding patterns among the the given graph, attention mechanisms are more relevant for discovery of important regions within.
-Lately, graph learning approaches found application for computing compound representations for DTI prediction.\\
-Approaches on this rather sophisticated problem can divided into top-down or network approaches (\textbf{CITATION}), and bottom-up or molecular approaches (\textbf{CITATION}).
-Top-down approaches take advantage of other data such as diseases (CITATION), side effects, knowledge graphs or ontologies, in order to learn representations for both compound and protein.
-On the other hand, bottom-up aspirations attempt to learn from chemical properties.
-For drugs, molecular structure (CITATION GraphDTA), molecular fingerprints, similarity to other drugs (See Bioinf Survey), and other molecular features.
-On the protein side, secondary structure prediction (CITATION), contact prediction (CITATION), or simply convolution over the amino acid sequences served to obtain a feature for the given proteins.
-However, both kinds contain and share some problems, that are not solvable within themselves.
-Thus, bottom-up approaches share the lack of ability to generalize, which we will show in later sections, and usually focus on engineering sophisticated features for the drugs, while neglecting to formulate meaningful features on the protein side.
-Top-down approaches lack the ability to spot small differences to cope with small differences within the drug structure and rely heavily on given data for the considered drug-target pair.
-The latter is not suitable for predictions on novel or unseen compounds, as e.g., data on side effects or its impact on diseases is seldom given for novel drugs.\\
-In order to design such a feature for proteins and drugs, respectively, we make use of the interaction networks for both proteins and compounds.
-Drug-drug interaction networks were introduced and standardized by \cite{Boyce2015} and have been used for clinical decision support (\cite{Scheife2015}).
-Drug-drug interaction networks may give a hint on common targeted pathways.
-As an additional compound feature we will use semantic side effect similarity, which we will discuss later on.\\
+In history, traditional remedies, that were known for their medicinal
+properties lead to drugs by extraction of the functional
+ingredients. Alternatively, characteristics and features of potential
+drugs were detected by accident like in the case of penicillin. More
+recently, biological drug targets can be found \textit{in silico}
+through discovery of suitable computational predictors.
+
+The challenge of accurately predicting drug-target-interactions (DTI)
+has shown its importance in the fields of drug repurposing and
+repositioning, and in the exploration of novel drugs and their
+interaction partners. Knowledge about those links between compounds
+and their target proteins help in an array of medical and
+pharmaceutical studies. Additionally, those associations can be
+utilized to identify disease specific targets, leading to desirable
+therapeutic effects.
+
+With the rapidly growing field of machine learning approaches and
+their application to bioscientifical problems in the realm of
+bioinformatics, different kinds of data, such as long DNA sequences
+could be utilized for feature generation, while rapid advances were
+made. Almost all state of the art models for drug-target-interaction
+prediction were based on the usage of neural networks with increasing
+size.\todo{Needs ref, or reformulate}
+
+Only recently, the technique of graph learning was introduced by
+\citet{GCNConv} through graph convolution algorithms, and improved and
+altered under usage of different kernels \cite{ChebConv, ARMAConv},
+attention mechanisms \cite{GATConv}, random walks \cite{APPNPConv},
+and mixtures of both \cite{SAGEConv}. While based on diverse systems,
+they can be relevant for testing distinct hypothesis for given
+graphs. While convolutional filters are suitable for finding patterns
+among the the given graph, attention mechanisms are more relevant for
+discovery of important regions within. Lately, graph learning
+approaches found application for computing compound representations
+for DTI prediction.
+
+Approaches on this rather sophisticated\todo{Try to avoid}
+problem\todo{Need to state the problem clearly here or at end of prev
+  paragraph} can divided into top-down or network approaches
+(\textbf{CITATION}), and bottom-up or molecular approaches
+(\textbf{CITATION}). Top-down approaches take advantage of other data
+such as diseases (CITATION), side effects, knowledge graphs or
+ontologies, in order to learn representations for both compound and
+protein. \todo[inline]{We call these approaches ``top-down'' because
+  they start with the observable characteristics induced by a drug and
+  infer the targets based on the likely molecular mechanisms that
+  result in these phenotypes.} On the other hand, bottom-up approaches
+attempt to learn from chemical properties of proteins or drugs to
+infer candidate drug--target interactions. For drugs, molecular
+structure (CITATION GraphDTA), molecular fingerprints, similarity to
+other drugs (See Bioinf Survey), and other molecular features may be
+used. On the protein side, secondary structure prediction (CITATION),
+contact prediction (CITATION), or simply\todo{Don't use ``simply''.}
+convolution over the amino acid sequences can be used to obtain a
+feature representation for a given proteins. However, both bottom-up
+and top-down approaches to drug--target interaction prediction
+\todo[inline]{replace: ``contain and share some problems'' with
+  something like ``have some limitations''} that are not solvable within
+themselves.
+\todo[inline]{Following is not sufficiently precise; here, you need to
+clearly state the challenges faced by both approaches, ideally with
+references.}
+Thus, bottom-up approaches share the lack of ability to
+generalize, which we will show in later sections, and usually focus on
+engineering sophisticated features for the drugs, while neglecting to
+formulate meaningful features on the protein side. Top-down approaches
+lack the ability to spot small differences to cope with small
+differences within the drug structure and rely heavily on given data
+for the considered drug-target pair. The latter is not suitable for
+predictions on novel or unseen compounds, as e.g., data on side
+effects or its impact on diseases is seldom given for novel drugs.
+
+In order to design such a feature for proteins and drugs,
+respectively, we make use of the interaction networks for both
+proteins and compounds. Drug-drug interaction networks were introduced
+and standardized by \citet{Boyce2015} and have been used for clinical
+decision support \cite{Scheife2015}. Drug-drug interaction networks
+may give a hint on common targeted pathways. As an additional compound
+feature we will use semantic side effect similarity, which we will
+discuss later on\todo{Generally, try to avoid pointers to ``later''.}.
+
 Protein-protein interaction networks have shown great results in $\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context for molecular system biology.
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2021-01-05 paper/main.tex
comments on into
Updated paper

--- 2020-04-22 paper/main.tex
@@
 The challenge of accurately predicting drug-target-interactions (DTI) has shown its importance in the fields of drug repurposing and repositioning, and in the exploration of novel drugs and their interaction partners.
-Knowledge about those links between compounds and their target proteins help in an array of medical and pharmaceutical studies.
+Knowledge about those links between compounds and their target proteins aid an array of medical and pharmaceutical studies.
 Additionally, those associations can be utilized to identify disease specific targets, leading to desirable therapeutic effects.\\
@@
 While convolutional filters are suitable for finding patterns among the the given graph, attention mechanisms are more relevant for discovery of important regions within.
-Lately, graph learning approaches found application for computing compound representations for DTI prediction.\\
+Lately, graph learning approaches found application for computing compound representations for DTI prediction.\textbf{CITATION}\\
 Approaches on this rather sophisticated problem can divided into top-down or network approaches (\textbf{CITATION}), and bottom-up or molecular approaches (\textbf{CITATION}).
-Top-down approaches take advantage of other data such as diseases (CITATION), side effects, knowledge graphs or ontologies, in order to learn representations for both compound and protein.
-On the other hand, bottom-up aspirations attempt to learn from chemical properties.
-For drugs, molecular structure (CITATION GraphDTA), molecular fingerprints, similarity to other drugs (See Bioinf Survey), and other molecular features.
-On the protein side, secondary structure prediction (CITATION), contact prediction (CITATION), or simply convolution over the amino acid sequences served to obtain a feature for the given proteins.
+Top-down approaches take advantage of other data such as diseases (CITATION), side effects, knowledge graphs or ontologies, in order to learn representations for either compound or protein.
+On the other hand, bottom-up methods attempt to learn from chemical properties.
+For drugs, molecular structure (CITATION GraphDTA), molecular fingerprints, similarity to other drugs \cite{Survey2018}, and other molecular features.
+On the protein side, secondary structure prediction (CITATION), contact prediction (CITATION), or simply convolution over the amino acid sequences served to obtain a molecular feature for the given proteins.
 However, both kinds contain and share some problems, that are not solvable within themselves.
-Thus, bottom-up approaches share the lack of ability to generalize, which we will show in later sections, and usually focus on engineering sophisticated features for the drugs, while neglecting to formulate meaningful features on the protein side.
-Top-down approaches lack the ability to spot small differences to cope with small differences within the drug structure and rely heavily on given data for the considered drug-target pair.
-The latter is not suitable for predictions on novel or unseen compounds, as e.g., data on side effects or its impact on diseases is seldom given for novel drugs.\\
-In order to design such a feature for proteins and drugs, respectively, we make use of the interaction networks for both proteins and compounds.
-Drug-drug interaction networks were introduced and standardized by \cite{Boyce2015} and have been used for clinical decision support (\cite{Scheife2015}).
-Drug-drug interaction networks may give a hint on common targeted pathways.
-As an additional compound feature we will use semantic side effect similarity, which we will discuss later on.\\
-Protein-protein interaction networks have shown great results in $\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context for molecular system biology.
-However, these contexts were never applied to the problem of drug-target-interaction prediction.
-Thus we formalized our hypotheses over these interaction graphs and will test them in the following chapters.
+Thus, bottom-up approaches share the lack of ability to generalize by dearth to inference on unknown chemical structures.
+Furthermore, usually focus on engineering expressive and sophisticated features on the compound side, while relying on less advanced features on the protein side\cite{Survey2018}.
+On the other hand, top-down approaches are unable to spot e.g.
+small non-functional alterations of the compound, and are hence unable to infer on minor differences and rely heavily on rich neighbourhood data for the considered drug-target pair.
+However, both types of concepts have shown their significance in a variety of tasks, while a combination of both has hardly been done.
+\\
+
+In order to combine both methods, we will use protein-protein interaction networks, which have shown great results in $\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context for molecular system biology.
+We will utilize this context hypothesis for deducing on the hypothesis of locality of protein functionality.
+We will therefor combine molecular representations for both drugs and proteins with features obtained from the Human Phenotype Ontology (HPO) over PPI graphs under usage of graph neural networks.
 
@@
 
-\section{Approach}
-\subsection{Choosing appropriate train-test splits}
-As shown, recent work lacks the ability to combine both top-down and bottom-up approaches for their predictions, however performing quite well on their datasets.
-Drug-target interaction prediction is the task to accurately predict whether for a given drug and a given protein there is a biological interaction within humans.
+\section{Methods}
+\subsection{Problem Description}
+The issue of predicting drug-target interactions can be described quite easily: For a given drug and a given protein we want to forecast whether those interact or not.
+We hereby do not differentiate between activation and inhibition.
+
+\begin{itemize}
+	\item how did we model it
+\end{itemize}
+
+\subsection{Choosing a train-test splitting scheme}
+In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
+Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
 However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
+
 \begin{enumerate}
@@
 \end{enumerate}
+
 In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
-However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier.
-Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE).
-When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
-Hence, when entirely new molecules are seen, the models perform much worse.
+As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous results.
+This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire.
+The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
+Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive.
+As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes.
 \\
-The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.\\
-As mentioned in the introduction it is quite difficult to learn suitable features from proteins.
-In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features.
-\\
-Thus, building splitting over proteins is the most challenging of the three options. \\
+
+\subsection{Does protein functio localize on the PPI graph?}
+
+\subsection{Combining mutually exclusive approaches}
 

### 2021-01-09 paper/main.tex
Update paper draft and TODO list for upcoming chapters

--- 2021-01-05 paper/main.tex
@@
 
-\section{Approach}
-\subsection{Choosing appropriate train-test splits}
-As shown, recent work lacks the ability to combine both top-down and bottom-up approaches for their predictions, however performing quite well on their datasets.
-Drug-target interaction prediction is the task to accurately predict whether for a given drug and a given protein there is a biological interaction within humans.
+\section{Methods}
+\subsection{Datasets}
+The data for the different parts of this model were obtained from various sources.
+Starting with the protein-protein interactions, we fetched 11.574 proteins with over 170.000 links from STRING (\cite{STRINGv10}).
+For the drug-target interactions themselves, we fetched
+137.000 links from STITCH database (\cite{STITCHv5}).
+As both STRING and STITCH provide probability scores for each association, we filtered them as advised by a threshold of 700, thus only obtaining likely interactions.\\
+For the ontology segment we utilized PhenomeNET \citep{PhenomeNET2011}, a collection of various ontologies such as Human Phenotype Ontology \citep{HPO2018}, Gene Ontology (\citet{GOoriginal2000} \citet{GOrecent2020}), Mammalian Phenotype Ontology \citep{MP2009} and numerous others.
+\todo{which ontologies to cite}
+Side effects and their links to drugs were obtained Side Effect Resource (SIDER)\citep{SIDER} and structured according MedDRA database \citep{MedDRA}.
+They were mapped to PhenomeNET with aid of \textit{Phenomebrowser.net}, which provides a SPARQL query endpoint for the mentioned resources.
+\\
+
+We additionally obtained molecular structure based features for drugs from \textit{SmilesTransformer}\citep{SmilesTransformer} and proteins from \textit{DeepGOPlus}\citep{DeepGoPlus}.\\
+
+Eventually the intersection between these resources yielded \todo{num drugs} drugs and \todo{num proteins} human proteins for the training phase.
+We provide links to and methods for the necessary data in the provided Github repository.\\
+
+\subsection{Problem Description}
+The issue of predicting drug-target interactions can be described quite briefly: For a given drug and a given protein we want to forecast whether those interact or not.
+We hereby do not differentiate between activation and inhibition, and do not erect statements on the strength of the bond.\\
+
+\subsection{Model} 
+In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each.
+In the top-down section, we used \textit{DL2vec}\citep{DL2vec2020} to obtain ontology based representations.
+Hereby, DL2vec constructs a graph by introducing vertices and edges for each ontology class and axiom, respectively, followed by random walks starting from each entity.
+These walks are eventually learned on using a Word2vec \citep{Word2vec2013} model.
+Thus, we pick up rich, neighbourhood focused representations for each entity, which has shown great results for representing protein function and phenotypes.
+\\
+
+As we want to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
+On the other hand we build a deep neural network for the molecular structure based features.
+\\
+
+\subsubsection{Graph convolutional layers}
+These molecular and ontology based sub-models were added to a larger graph convolutional model.
+As described before, graph convolution has shown significant performance increase in a variety of tasks.
+While there are various methods out there we will only introduce the mose basic one here.
+A graph convolutional layer w.r.t.
+\citet{GCNConv} hereby consists of a learnable weight matrix followed by an aggregation step, formalized by
+
+\begin{equation}
+	\mathbf{X}^{\prime} = \mathbf{\hat{D}}^{-1/2} \mathbf{\hat{A}}
+	\mathbf{\hat{D}}^{-1/2} \mathbf{X} \mathbf{\Theta}
+\end{equation}
+
+where for a given graph $G=(V,E)$, $\hat{A} = A + I$ denotes the adjacency matrix with added self-loops for each vertex, $D$ is described by $\hat{D}_{ii} = \sum_{j=0} \hat{A}_{ij}$, a diagonal matrix displaying the degree of each node, and $\Theta$ denotes the learnable weight matrix.
+Added self-loops enforce that each node representation is directly dependent on its own preceding one.
+Notably, the number of graph convolutional layers stacked equals the radius of relevant nodes for each vertex within the graph.\\
+
+The update rule for each individual node is denoted by
+
+\begin{equation}
+	\mathbf{x}^{\prime}_i = \mathbf{\Theta} \sum^{N}_{j}
+	\frac{1}{\sqrt{\hat{d}_j \hat{d}_i}} \mathbf{x}_j
+\end{equation}
+
+where both $\hat{d}_i, \hat{d}_j$ are dependent on the edge weights $e_{ij}$ of the graph.
+With simple, single valued edge weights such as $e_{ij}=1 \text{ }\forall (i,j)\in E$, all $\hat{d}_i$ reduce to $d_i$, i.e.
+the degree of each vertex $i$.
+\\
+
+While in this initial formulation the node-wise update step is defined by the sum over all neighbouring node representations, we are able to alter this formulation for a more sophisticated message passing scheme.
+As described we are able to rearrange the order of activation function, aggregation and linear neural layer with this formulation as proposed by \citet{GENConv2020}:
+
+\begin{equation}
+	\mathbf{x}_i^{\prime} = \mathrm{MLP} \left( \mathbf{x}_i +
+	\mathrm{AGG} \left( \left\{
+	\mathrm{ReLU} \left( \mathbf{x}_j + \mathbf{e_{ji}} \right) +\epsilon
+	: j \in \mathcal{N}(i) \right\} \right)
+	\right)
+\end{equation}
+
+While the reordering is merely import for numerical stability, the authors claim that this alteration of the original formulation eases the vanishing gradient problem for deeper convolutional networks.
+\\
+
+\subsection{Choosing a train-test splitting scheme}
+In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
+Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
 However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
+
 \begin{enumerate}
@@
 \end{enumerate} 
+In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
+As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous results.
+This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire.
+The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
+Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive.
+As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes.
+\\
+
 In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
@@
 \section{Methods}
-\subsection{Data}
-The data for the different parts of this model were obtained from different sources.
-Starting with the protein-protein interactions, we fetched 11.574 proteins with over 170.000 links from STRING (\cite{STRINGv10}).
-As STRING provides statistical scores for each association, we filtered them as advised by a threshold of 700.
-For the drug-drug interaction we retrieved the drugs from \cite{Boyce2015}.
-Side effects were obtained from MedDRA database (\cite{MedDRA}) that hierarchically orders side effects.
-Links between drugs and side effects were fetched from \textit{SIDER} (\cite{SIDER}) side effects database.
-The intersection between these sources yielded 641 common drugs for usage in our data.
-For the drug-target interactions themselves, we fetched
-137.000 links from STITCH database (\cite{STITCHv5}). We provide links and methods to and for the necessary data in the provided Github repository.\\
 

### 2021-01-09 paper/main.tex
Update method part of paper, continue results

@@
 \section{Methods}
+\subsection{Problem Description}
+The issue of predicting drug-target interactions can be described quite briefly: For a given drug and a given protein we want to forecast whether those interact or not.
+Additionally, we presume the closed world assumption, that all drug-protein pairs without an interaction, do not have one eventually.
+We also do not differentiate between activation and inhibition, and do not erect statements on the strength of the bond, thus leading to a binary classification task.\\
+
 \subsection{Datasets}
 The data for the different parts of this model were obtained from various sources.
-Starting with the protein-protein interactions, we fetched 11.574 proteins with over 170.000 links from STRING (\cite{STRINGv10}).
+Starting with the protein-protein interactions, we fetched \text{num proteins} human proteins with over 170.000 links from STRING (\cite{STRINGv10}).
 For the drug-target interactions themselves, we fetched
@@
 
-Eventually the intersection between these resources yielded \todo{num drugs} drugs and \todo{num proteins} human proteins for the training phase.
+For comparative evaluation we used the gold standard dataset introduced by \citet{Yaminishi2008}, which includes both drug-target interaction pairs and side effects.\\
+
+Eventually, we only considered proteins that had at least one link in either STITCH or STRING, and drugs with at least one side effect and one existing target.
+Thus, the intersection between these resources yielded \todo{num drugs} drugs and \todo{num proteins} human proteins for the training phase.
 We provide links to and methods for the necessary data in the provided Github repository.\\
 
-\subsection{Problem Description}
-The issue of predicting drug-target interactions can be described quite briefly: For a given drug and a given protein we want to forecast whether those interact or not.
-We hereby do not differentiate between activation and inhibition, and do not erect statements on the strength of the bond.\\
+\subsection{Evaluation and metrics}
+As the number of drug-targets are sparsely given w.r.t.
+to the number of both drugs and proteins considered, the resulting training, validation and testing datasets are highly imbalanced.
+As there are only $~22.000$ links in the considered subset the ratio
+\begin{equation*}
+	\frac{\#dti\_links}{\#drugs \cdot \#proteins} \approx 360,
+\end{equation*}
 
-\subsection{Model} 
+consequently needing compensation in the computed loss function and appropriate metrics for the evaluation.\\
+
+On that account, we weighted all positive drug-protein pair samples with this ratio by introducing the following loss function w.r.t. to binary cross-entropy:
+\begin{equation}
+	l(x,y) = - w \left[ y \cdot \log x + (1 - y) \cdot \log (1 - x) \right]
+\end{equation}
+for given prediction $x$ and target $y$.
+We average this loss among all drug-protein pairs in the training set, leading to a stable environment for the used optimizing scheme \textit{Adam} \citep{Adam2014}.
+We implemented a 5-fold cross validation among the proteins as justified in the results section.
+Furthermore, we  used early stopping to detect plateaus in the training process.
+\\
+
+To assess each model, we computed the \textit{area under receiver operating characteristic}-score (AUROC), \textit{F1-score} (F1) and \textit{Matthews correlation coefficient} (MCC) for each considered hyperparameter configuration setting.\\
+
+Avoiding validation overfitting, we eventually tested the best performing model on the Yamanishi dataset as presented and discussed.
+Furthermore, we compare our results on this very dataset with various other cutting edge approaches to DTI-prediction.\\
+
+\subsection{Model} \todo{speaking about the model structure without introducing it at all feels very odd.}
 In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each.
@@
 
-As we want to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
-On the other hand we build a deep neural network for the molecular structure based features.
+As we wanted to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
+On the other hand we built a deep neural network for the molecular structure based features.
 \\
@@
 \\
+
+\subsubsection{Hyperparameter tuning}
+To find the best hyperparameter configuration for the proposed model we performed a grid search, to find the most expressive and non-redundant representation.
+Therefor, we pretrained the bottom-up and the top-down model separately and aimed at best performing models w.r.t.
+previously described metrics.
+Embedding sizes, depth of the neural network, optimizer, learning rate and layer types were found from an extensive, manual grid search.
 

### 2021-01-11 paper/main.tex
Added figures to paper

@@
 The data for the different parts of this model were obtained from various sources.
-Starting with the protein-protein interactions, we fetched \text{num proteins} human proteins with over 170.000 links from STRING (\cite{STRINGv10}).
+Starting with the protein-protein interactions, we fetched $\approx11000$ human proteins with over 170.000 links from STRING (\cite{STRINGv10}).
 For the drug-target interactions themselves, we fetched
@@
 Eventually, we only considered proteins that had at least one link in either STITCH or STRING, and drugs with at least one side effect and one existing target.
-Thus, the intersection between these resources yielded \todo{num drugs} drugs and \todo{num proteins} human proteins for the training phase.
+Thus, the intersection between these resources yielded 1160 drugs and 6680 human proteins for the training phase.
 We provide links to and methods for the necessary data in the provided Github repository.\\
@@
 
-\subsection{Model} \todo{speaking about the model structure without introducing it at all feels very odd.}
+\subsection{Model}
 In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each.
@@
 Thus, we pick up rich, neighbourhood focused representations for each entity, which has shown great results for representing protein function and phenotypes.
+The overall structure of the ontology can be seen in Figure~\ref{fig:Onto}.\\
+
+\begin{figure}[!tpb]%figure1
+	\centerline{\includegraphics[width=0.5\textwidth]{figures/drug_protein_ontology_network.png}}
+	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}\label{fig:Onto}
+\end{figure}
+
+As we wanted to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
+On the other hand we built a deep neural network for the molecular structure based features, also benefiting from the siamese network architecture.
+Therefore, the precomputed embeddings are run through a deep neural network transformation (DNNT) which also reduces the eventual representation size for drugs and proteins separately.
+An example structure for both types of features can be found in Figure~\ref{fig:SiameseNetwork}.\\
+
+\begin{figure}[!tpb]%figure1
+	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
+	\caption{Siamese network applied to molecular and DL2vec features, utilizing deep neural network transformations (DNNT).
+The similarity function $\otimes$ yields the similarity between both transformed embeddings e.g.
+by computing the cosine similarity.}\label{fig:SiameseNetwork}
+\end{figure}
+
+\subsubsection{Graph convolutional layers}
+These molecular and ontology based sub-models were added to a larger graph convolutional model, based on the protein-protein interaction (PPI) graph.
+The PPI dataset is represented by a graph $G=(V,E)$, where each protein is represented by a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$ symbolizes an interaction between two proteins.
+Additionally, we introduce mapping $x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node feature $x_v = x(v)$, where $d$ denotes the dimensionality of the node features.
 \\
 
-As we wanted to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
-On the other hand we built a deep neural network for the molecular structure based features.
-\\
-
-\subsubsection{Graph convolutional layers}
-These molecular and ontology based sub-models were added to a larger graph convolutional model.
 As described before, graph convolution has shown significant performance increase in a variety of tasks.
-While there are various methods out there we will only introduce the mose basic one here.
+While there are various methods out there we will only introduce the most basic one here.
 A graph convolutional layer w.r.t.
@@
 the degree of each vertex $i$.
+We will denote this type of graph convolutional neural layers with \textsc{GCNConv}.
 \\
@@
 While in this initial formulation the node-wise update step is defined by the sum over all neighbouring node representations, we are able to alter this formulation for a more sophisticated message passing scheme.
-As described we are able to rearrange the order of activation function, aggregation and linear neural layer with this formulation as proposed by \citet{GENConv2020}:
+As described we are able to rearrange the order of activation function $\sigma$, aggregation $\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this formulation as proposed by \citet{GENConv2020}:
 
@@
 	\mathrm{AGG} \left( \left\{
-	\mathrm{ReLU} \left( \mathbf{x}_j + \mathbf{e_{ji}} \right) +\epsilon
+	\mathrm{\sigma} \left( \mathbf{x}_j + \mathbf{e_{ji}} \right) +\epsilon
 	: j \in \mathcal{N}(i) \right\} \right)
@@
 
+where we will generally only consider $\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$.
+We will denote this generalized layer type as \textsc{GENConv}, following the notation of \citet{PytorchGeometric}.
 While the reordering is merely import for numerical stability, the authors claim that this alteration of the original formulation eases the vanishing gradient problem for deeper convolutional networks.
+Additionally, the authors revise the aggregation function by generalizing it, allowing more sophisticated weighting functions such as learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming signals for each vertex, substituting the previously mentioned averaging step.
+Hence, while \textsc{GCNConv} suffers from both vanishing gradients and signal fading for large scale, highly connected graphs, each propagation step in \textsc{GENConv} emphasizes strong signals with values close to $0$ and $1$.
 \\
+
+We employ another mechanism to avoid redundancy and fading signals in stacked graph convolutional networks, which was introduced by \citet{DeepGCN2019} and refined in \citet{DeeperGCN2020}.
+The authors propose a residual network architecture and normalization scheme, solving these issues in a variety of graph predicting tasks.
+This structure is depicted in Figure~\ref{fig:ResGraphConvBlocks} resulting in reusable residual blocks, which can be stacked multiple times, thereby not losing focus of each node neighbourhood.
+\\
+
+ \begin{figure}[!tpb]%figure1
+ 	\centerline{\includegraphics[width=0.2\textwidth]{figures/ResGraphConvBlocks.png}}
+ 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
+ 	\label{fig:ResGraphConvBlocks}
+ \end{figure}
 
@@
 Embedding sizes, depth of the neural network, optimizer, learning rate and layer types were found from an extensive, manual grid search.
+
+\section{Results}
+
+\subsection{High level description}
+To obtain a prediction for a given drug-protein pair, we first need to pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.
+These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.
+Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs.
+The overall architecture can be seen in Figure~\ref{fig:Fullmodel}.
+
+\begin{figure}[!tpb]%figure1
+	\centerline{\includegraphics[width=0.33\textwidth]{figures/full_model.png}}
+	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
+The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
+}
+	\label{fig:Fullmodel}
+\end{figure}
 
@@
 	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
-	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:01}
+	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
 \end{figure}

### 2021-01-12 src/DTI_data_preparation.py
Retest on NeoDTI data 1

@@ -373,7 +373,7 @@ def get_yamanishi_data(original_drug_list, original_protein_list):
     yamanishi_protein_mapping = PPI_utils.get_protein_Yamanishi_to_STITCH_mapping()
 
 
-    path = '../data/Yamanishi_data/'
+    path = '../data/NeoDTI_data/'
 
     drug_list = []
     with open(file=path + 'drug.txt', mode='r') as f:
@@ -401,16 +401,6 @@ def get_yamanishi_data(original_drug_list, original_protein_list):
 
     drug_indices = [drug_list.index(drug) for drug in drug_list if drug!=None and not drug.startswith('1')]
     protein_indices = list(set([protein_list.index(protein) for protein in protein_list if protein in original_protein_list]))
-
-    filename = '../../drug_indices.csv'
-    with open(file=filename, mode='w') as f:
-        for index in drug_indices:
-            f.write(str(index)+'\n')
-
-    filename = '../../prot_indices.csv'
-    with open(file=filename, mode='w') as f:
-        for index in protein_indices:
-            f.write(str(index)+'\n')
 
     drug_list = np.array(drug_list)
     protein_list = np.array(protein_list)

### 2021-01-17 src/DL2vec/jobscript.sh
update jobscript for rerunning drug DL2vec

@@ -14,8 +14,8 @@ module load gcc/6.4.0
 conda activate ~/.conda/envs/dti/
 module load groovy/3.0.6
 
-python runDL2vec.py -embedsize 200 -ontology ../../data/drug_indications/UMLS.owl -associations ../../data/drug_indications/drug_indication_association_file -outfile ../../data/drug_indications/embedding_model -entity_list ../../data/drug_indications/drug_indication_entity_list -num_workers 40
+# python runDL2vec.py -embedsize 200 -ontology ../../data/drug_indications/UMLS.owl -associations ../../data/drug_indications/drug_indication_association_file -outfile ../../data/drug_indications/embedding_model -entity_list ../../data/drug_indications/drug_indication_entity_list -num_workers 40
 
-# python runDL2vec.py -embedsize 200 -num_workers 40 -ontology ../../data/PhenomeNET_data/phenomenet.owl -associations ../../data/PhenomeNET_data/drug_association_file -outfile ../../data/PhenomeNET_data/drug_embedding_model -entity_list ../../data/PhenomeNET_data/drug_entity_list
+python runDL2vec.py -embedsize 200 -num_workers 40 -ontology ../../data/PhenomeNET_data/phenomenet.owl -associations ../../data/PhenomeNET_data/drug_association_file -outfile ../../data/PhenomeNET_data/drug_embedding_model -entity_list ../../data/PhenomeNET_data/drug_entity_list -file_prefix drug_
 
 

### 2021-01-17 paper/main.tex
Retest on NeoDTI data 1

@@
 The data for the different parts of this model were obtained from various sources.
-Starting with the protein-protein interactions, we fetched $\approx11000$ human proteins with over 170.000 links from STRING (\cite{STRINGv10}).
+Starting with the protein-protein interactions, we fetched $\approx11000$ human proteins with over 170.000 links from STRING \citep{STRINGv10}.
 For the drug-target interactions themselves, we fetched
-137.000 links from STITCH database (\cite{STITCHv5}).
+137.000 links from STITCH database \citep{STITCHv5}.
 As both STRING and STITCH provide probability scores for each association, we filtered them as advised by a threshold of 700, thus only obtaining likely interactions.\\
@@
 \\
-
-We additionally obtained molecular structure based features for drugs from \textit{SmilesTransformer}\citep{SmilesTransformer} and proteins from \textit{DeepGOPlus}\citep{DeepGoPlus}.\\
 
@@
 \subsection{Model}
-In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each.
+In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each individually.
+Hence, we assemble rich molecular structure based features for drugs from \textit{SmilesTransformer}\citep{SmilesTransformer} and proteins from \textit{DeepGOPlus}\citep{DeepGoPlus}.
+\textit{SmilesTransformer} introduces an autoencoder, learning over the SMILES strings and thus the molecular organization of each drug in an unsupervised manner.
+On the other hand, we take advantage of the pretrained models of \textit{DeepGOPlus}, obtaining features from the proteins amino acid sequence, showing significant performance in the field of protein function prediction.
+Thus, both embeddings seem to suitably supplement the  following ontology based represenations.\\
+
 In the top-down section, we used \textit{DL2vec}\citep{DL2vec2020} to obtain ontology based representations.
@@
 \subsection{High level description}
-To obtain a prediction for a given drug-protein pair, we first need to pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.
+To obtain a prediction for a given drug-protein pair, we first build DL2vec representations for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
+Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.
 These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.

### 2021-01-25 paper/main.tex
Added results for last 2 months

@@
 
+\section{Findings}
+
+\subsection{Deification of our method}
+\begin{itemize}
+	\item we built protein function and ontology based features based on DL2vec
+	\item Ontology derived protein function focused features are highly predictive for dtis
+	\item We built a versatile template for various features to test localization on the PPI graph
+	\item normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
+	\item protein functions localize on the PPI graph, while molecular features don't
+	\item all AUROC in \% AUROC score on STITCH
+
+	\begin{tabular}{|l|r|r|}
+		\hline
+		DNNT model&Without graph model&With graph model\\
+		\hline
+		MolPred&$69$&$69$\\
+		PhenomeNETPred&$88$&$92$\\
+		\hline
+		MolPred + PhenomeNETPred & $89$ & $93$\\
+		\hline
+	\end{tabular}
+	\item microAUC for MolPred + PhenomeNETPred on graph is about $93+-$
+	\item and on yamanishi dataset
+
+	\begin{tabular}{|l|r|r|}
+		\hline
+		DNNT model&Without graph model&With graph model\\
+		\hline
+		PhenomeNETPred&$83$&$84$\\
+		\hline
+		MolPred + PhenomeNETPred & $83$ & $84.5$\\
+		\hline
+	\end{tabular}
+	\item MicroAUC is about $83$
+\end{itemize}
+
+\subsection{How to insult other methods}
+\begin{itemize}
+	\item Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
+	\item Running split over proteins is harder than, drug and drug protein pair split (see below table)
+	\item this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
+	\item using indications is like cheating, as not applicable for searching new drugs
+	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
+	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
+	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
+
+	\begin{tabular}{|l|l|r|r|r|}
+		\hline
+		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
+		\hline
+		DTINet&DP pairs&$91$&$84.1$&$67.2$\\
+		DTIGEMS+&DP pairs&$93$& $72.2$& $67.8$ \\
+		DTI-CDF&Proteins&$83$&$83$&$79$\\
+		GNN-DTI&&&&\\
+		\hline
+		MolTrans&&&&\\
+		GraphData&&&&\\
+		\hline
+	\end{tabular}
+	\item A naive predictor (ranking proteins) and predict each drug similarly achieves cutting edge performance ($87.5$ AUROC for whole dataset, $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No prot focused microAUC possible.
+$\rightarrow$ hub proteins
+	\item Yamanishi Dataset is only partially suitable, if everybody just derives a suitable subset (DTIGEMS)
+\end{itemize}
+
 \subsection{Choosing a train-test splitting scheme}

### 2021-01-26 paper/main.tex
Update DTA prediction scores table

@@
 
-	\begin{tabular}{|l|l|r|r|r|}
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
 		\hline
@@
 		DTI-CDF&Proteins&$83$&$83$&$79$\\
-		GNN-DTI&&&&\\
-		\hline
-		MolTrans&&&&\\
-		GraphData&&&&\\
 		\hline
@@
 	\item Yamanishi Dataset is only partially suitable, if everybody just derives a suitable subset (DTIGEMS)
+	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap
+
 \end{itemize}
+
+\begin{figure}[!tpb]%figure1
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
+		\hline
+		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
+		\hline
+		DeepDTI&Drugs&$87.6$&$75.9$&$70.1$\\
+		DeepDTA&DP pairs&87.6&$76.7$&$69.4$\\
+		DeepConv-DTI&DP pairs&88.3&$76.6$&$73.0$\\
+		MolTrans&DP pairs&89.5&$77.0$&$74.0$\\
+
+		\hline
+	\end{tabular}
+\end{figure}
 

### 2021-01-27 paper/main.tex
.

@@
 \subsection{Problem Description}
-The issue of predicting drug-target interactions can be described quite briefly: For a given drug and a given protein we want to forecast whether those interact or not.
-Additionally, we presume the closed world assumption, that all drug-protein pairs without an interaction, do not have one eventually.
-We also do not differentiate between activation and inhibition, and do not erect statements on the strength of the bond, thus leading to a binary classification task.\\
+The issue of predicting drug-target interactions can be described
+quite briefly: For a given drug and a given protein we want to
+determine whether those interact or not.  We do not differentiate
+between types of interaction such as activation and inhibition, and do
+not predict the strength of the interaction.  If we additionally make
+the closed world assumption, i.e., assume that our knowledge is
+complete and all drug--protein pairs without a known interaction do
+not interact, we can formulate the problem as a binary classification
+task.
 
 \subsection{Datasets}
-The data for the different parts of this model were obtained from various sources.
-Starting with the protein-protein interactions, we fetched $\approx11000$ human proteins with over 170.000 links from STRING \citep{STRINGv10}.
-For the drug-target interactions themselves, we fetched
-137.000 links from STITCH database \citep{STITCHv5}.
-As both STRING and STITCH provide probability scores for each association, we filtered them as advised by a threshold of 700, thus only obtaining likely interactions.\\
+The data for the different parts of this model were obtained from
+various sources. Starting with the protein-protein interactions, we
+fetched $\approx 11000$\todo{precise number} human proteins with over 170.000 links from
+STRING \citep{STRINGv10}. For the drug-target interactions themselves,
+we fetched 137.000 links from STITCH database \citep{STITCHv5}. As
+both STRING and STITCH provide probability scores for each
+association, we filtered them as advised by a threshold of 700, thus
+only obtaining likely interactions.
+
 For the ontology segment we utilized PhenomeNET \citep{PhenomeNET2011}, a collection of various ontologies such as Human Phenotype Ontology \citep{HPO2018}, Gene Ontology (\citet{GOoriginal2000} \citet{GOrecent2020}), Mammalian Phenotype Ontology \citep{MP2009} and numerous others.
@@
 $\rightarrow$ hub proteins
-	\item Yamanishi Dataset is only partially suitable, if everybody just derives a suitable subset (DTIGEMS)
+	\item Yamanishi Dataset is only partially suitable \todo{for
+            comparing results}}, if everybody just derives a suitable subset (DTIGEMS)
 	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap

### 2021-01-27 paper/main.tex
.

@@
 \subsection{Choosing a train-test splitting scheme}
+
+\todo[inline]{List a number of general challenges first, then refer
+  back to them: very similar drugs are designed to target the same
+  proteins, minor variations, leading to a kind of ``hidden''
+  duplication of drugs in the training/testing set; some proteins act
+  like hubs and interact with many drugs (making them easy and
+  reliable predictors); shared indications of drugs are usually
+  because they target the same proteins by design leading to some
+  trivial predictions; some proteins are more ``valuable'' targets due
+  to their disease involvement (such as cancer, meaning more drugs are
+  developed for targeting those); }
+
 In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
@@
 In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous results.
-This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire.
+As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results.
+This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}.
 The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
@@
 When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
-Hence, when entirely new molecules are seen, the models perform much worse.
-\\
-The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.\\
+Hence, when entirely new molecules are seen, the models perform much worse
+
+The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
+
 As mentioned in the introduction it is quite difficult to learn suitable features from proteins.

### 2021-01-27 paper/main.tex
.

@@
   to their disease involvement (such as cancer, meaning more drugs are
-  developed for targeting those); }
+  developed for targeting those). If the aim is to find a drug for a
+  given protein, the challenge is to develop a training (and
+  evaluation) scheme that does not simply overfit to these biases.}
 

### 2021-01-27 paper/main.tex
.

@@
 \subsection{High level description}
+Our aim is to develop and evaluate a computational model to predict
+drug--target interactions. Specifically, given a protein, we are
+interested in identifying the drugs that may target the protein, so as
+to reflect the case where a drug is desired to target a
+disease-associated protein.
+
 To obtain a prediction for a given drug-protein pair, we first build DL2vec representations for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
@@
 
+\todo[inline]{Relate these to the issues listed above:}
 \begin{enumerate}

### 2021-01-29 src/kalign2/config.h
Purged src dir

@@ -1,67 +1,0 @@
-/* config.h.  Generated by configure.  */
-/* config.h.in.  Generated from configure.ac by autoheader.  */
-
-/* Define to 1 if you have the <getopt.h> header file. */
-#define HAVE_GETOPT_H 1
-
-/* Define to 1 if you have the <inttypes.h> header file. */
-#define HAVE_INTTYPES_H 1
-
-/* Define to 1 if your system has a GNU libc compatible `malloc' function, and
-   to 0 otherwise. */
-#define HAVE_MALLOC 1
-
-/* Define to 1 if you have the <memory.h> header file. */
-#define HAVE_MEMORY_H 1
-
-/* Define to 1 if your system has a GNU libc compatible `realloc' function,
-   and to 0 otherwise. */
-#define HAVE_REALLOC 1
-
-/* Define to 1 if you have the <stdint.h> header file. */
-#define HAVE_STDINT_H 1
-
-/* Define to 1 if you have the <stdlib.h> header file. */
-#define HAVE_STDLIB_H 1
-
-/* Define to 1 if you have the <strings.h> header file. */
-#define HAVE_STRINGS_H 1
-
-/* Define to 1 if you have the <string.h> header file. */
-#define HAVE_STRING_H 1
-
-/* Define to 1 if you have the <sys/stat.h> header file. */
-#define HAVE_SYS_STAT_H 1
-
-/* Define to 1 if you have the <sys/types.h> header file. */
-#define HAVE_SYS_TYPES_H 1
-
-/* Define to 1 if you have the <unistd.h> header file. */
-#define HAVE_UNISTD_H 1
-
-/* Define to the address where bug reports for this package should be sent. */
-#define PACKAGE_BUGREPORT "BUG-REPORT-ADDRESS"
-
-/* Define to the full name of this package. */
-#define PACKAGE_NAME "FULL-PACKAGE-NAME"
-
-/* Define to the full name and version of this package. */
-#define PACKAGE_STRING "FULL-PACKAGE-NAME VERSION"
-
-/* Define to the one symbol short name of this package. */
-#define PACKAGE_TARNAME "full-package-name"
-
-/* Define to the version of this package. */
-#define PACKAGE_VERSION "VERSION"
-
-/* Define to 1 if you have the ANSI C header files. */
-#define STDC_HEADERS 1
-
-/* Define to empty if `const' does not conform to ANSI C. */
-/* #undef const */
-
[…]

### 2021-01-29 src/kalign2/kalign2.h

@@ -1,417 +1,0 @@
-/*
-	kalign2.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <float.h>
-
-#define INFTY FLT_MAX
-#define FLOATINFTY FLT_MAX
-
-#define NODESIZE 16
-
-
-#ifdef MEMORY 
-#define tmalloc malloc
-#endif
-
-
-extern unsigned int numseq;
-extern unsigned int numprofiles;
-extern float gpo;
-extern float gpe;
-extern float tgpe;
-
-struct feature_matrix{
-	float** m;
-	int mdim;
-	int stride;
-};
-
-struct utype_ufeat{
-	struct feature *t;
-	struct feature *f;
-};
-
-struct parameters{
-	char **infile;
-	char *input;
-	char *outfile;
-	char* format;
-	//int reformat;
-	char* feature_type;
-	char* alignment_type;
-	char* feature_mode;
-	char* distance;
-	char* tree;
-	char* sort;
-	char* sub_matrix;
-	char* print_tree;
-	char* print_svg_tree;
-	float gpo;
-	float gpe;
-	float tgpe;
-	float secret;
-	float zlevel;
-	float same_feature_score;
-	float diff_feature_score;
-	
-	int reformat;
-	int id;
-	int aa;
-	int alter_gaps;
-	int ntree;
-	int help_flag;
-	int quiet;
-	
-	int dna;
-	float alter_range;
-	int alter_weight;
-	float internal_gap_weight;
-	int smooth_window;
-	float gap_inc;
[…]

### 2021-01-29 src/kalign2/kalign2_advanced_gaps.c

@@ -1,1011 +1,0 @@
-/*
-	kalign2_advanced_gaps.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_advanced_gaps.h"
-
-
-int** advanced_hirschberg_alignment(struct alignment* aln,int* tree,float**submatrix, int** map,int window,float strength,float internal_gap_weight)
-{
-	struct hirsch_mem* hm = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	float** profile = 0;
-
-	profile = malloc(sizeof(float*)*numprofiles);
-	for ( i = 0;i< numprofiles;i++){
-		profile[i] = 0;
-	}
-
-	map = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i < numprofiles;i++){
-		map[i] = 0;
-	}
-	
-	hm = hirsch_mem_alloc(hm,1024);
-
-	fprintf(stderr,"\nAlignment:\n");
-
-	for (i = 0; i < (numseq-1);i++){
-		a = tree[i*3];
-		b = tree[i*3+1];
-		c = tree[i*3+2];
[…]

### 2021-01-29 src/kalign2/kalign2_advanced_gaps.h

@@ -1,34 +1,0 @@
-/*
-	kalign2_advanced_gaps.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-void advanced_smooth_gaps(float* prof,int len,int window,float strength);
-float* advanced_make_profile(float* prof, int* seq,int len,float** subm);
-float* advanced_update(const float* profa,const float* profb,float* newp,int* path,int sipa,int sipb,float internal_gap_weight);
-
-
-int* advanced_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm, int* hirsch_path);
-int* advanced_hirsch_align_two_pp_vector(const float* prof1,const float* prof2,struct hirsch_mem* hm,int* hirsch_path,float input_states[],int old_cor[]);
-struct states* advanced_foward_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-struct states* advanced_backward_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm);

### 2021-01-29 src/kalign2/kalign2_alignment_types.c

@@ -1,888 +1,0 @@
-/*
-	kalign2_alignment_types.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-int** default_alignment(struct alignment* aln,int* tree,float**submatrix, int** map)
-{
-	struct dp_matrix *dp = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	float** profile = 0;
-	float* profa = 0;
-	float* profb = 0;
-	
-	profile = malloc(sizeof(float*)*numprofiles);
-	for ( i = 0;i< numprofiles;i++){
-		profile[i] = 0;
-	}
-
-	map = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i < numprofiles;i++){
-		map[i] = 0;
-	}
-	
-	
-	dp = dp_matrix_alloc(dp,511,511);
-	
-	fprintf(stderr,"\nAlignment:\n");
-	
-	//c = numseq;
-	for (i = 0; i < (numseq-1);i++){
-		a = tree[i*3];
-		b = tree[i*3+1];
-		c = tree[i*3+2];
[…]

### 2021-01-29 src/kalign2/kalign2_conservation.c

@@ -1,57 +1,0 @@
-/*
-	kalign2_conservation.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include <math.h>
-
-#include "kalign2.h"
-
-void entrophy(int* prof,int len)
-{
-	int i,j;
-	float shannon = 0.0;
-	float log_two = log(2);
-	fprintf(stderr,"%f\n",log_two);
-	for ( i = 0; i < len; i++){
-		shannon = 0.0;
-		//prof[3] = 10;
-		//prof[23] += 10;
-		for ( j = 0; j < 23;j++){
-			if(prof[j]){
-			shannon += (float)prof[j]* log((float)prof[j]/(float)prof[23])/log_two;
-		
-		//	fprintf(stderr,"%f += %d/%d * %f\n",shannon,prof[j],prof[23],log((float)prof[j]/(float)prof[23])/log_two);
-			}
-		}
-		fprintf(stderr,"%f	",shannon);
-		
-		if (prof[23] < 23){
-			shannon = -shannon / (log((float)prof[23])/log_two);
-		}else{
-			shannon = -shannon / (log((float)23)/log_two);
-		}
-		fprintf(stderr,"%f\n",shannon);
-		prof+=64;
-	}
-}

### 2021-01-29 src/kalign2/kalign2_distance_calculation.c

@@ -1,1001 +1,0 @@
-/*
-	kalign2_distance_calculation.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-float** protein_pairwise_alignment_distance(struct alignment* aln,float** dm,struct parameters* param,float**subm, int nj)
-{
-	int i,j,c;
-	int * path = 0;
-	int len_a = 0;
-	int len_b = 0;
-	struct dp_matrix *dp = 0;
-	int a,b;
-	
-	
-	fprintf(stderr,"Distance Calculation:\n");
-	
-	b = (numseq*(numseq-1))/2;
-	a = 1;	
-	
-	
-	
-	dp = dp_matrix_alloc(dp,511,511);
-	
-	if (nj){
-		dm = malloc (sizeof(float*)*numprofiles);
-		for (i = numprofiles;i--;){
-			dm[i] = malloc (sizeof (float)*(numprofiles));
-			for (j = numprofiles;j--;){
-				dm[i][j] = 0.0f;
-			}
-		}
-	}else{
-		dm = malloc (sizeof(float*)*numseq);
-		for (i = numseq;i--;){
[…]

### 2021-01-29 src/kalign2/kalign2_dp.c

@@ -1,3347 +1,0 @@
-/*
-	kalign2_dp.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-int* f_only_pp_dyn(int* path, struct dp_matrix *dp,const float* fprof1,const float* fprof2,const int len_a,const int len_b,int fdim,int stride)
-{
-//	unsigned int freq[26];
-	
-	struct states* s = 0;
-	char** trace = 0;
-	char* tracep = 0;
-	register float pa = 0;
-	register float pga = 0;
-	register float pgb = 0;
-	register float ca = 0;
-	register int i = 0;
-	register int j = 0;
-	register int c = 0;
-	register int f = 0;
-
-	s = dp->s;
-	
-	trace = dp->tb;
-
-	trace[len_a][len_b] = 32;
-
-	fprof1 += len_a * stride;
[…]

### 2021-01-29 src/kalign2/kalign2_feature.c

@@ -1,1403 +1,0 @@
-/*
-	kalign2_feature.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_feature.h"
-
-static int stride;
-static int dim;
-static int gpo_pos;
-static int gpe_pos;
-static int tgpe_pos;
-
-int** feature_hirschberg_alignment(struct alignment* aln,int* tree,float**submatrix, int** map,struct feature_matrix* fm)
-{
-	struct hirsch_mem* hm = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	float** profile = 0;
-	
-	stride = (26+fm->mdim)*2 + 3; 
-	dim = 26+fm->mdim;
-	gpo_pos = (dim << 1) + 0;
-	gpe_pos = (dim << 1) + 1;
-	tgpe_pos = (dim << 1) + 2;
-
-	profile = malloc(sizeof(float*)*numprofiles);
-	for ( i = 0;i< numprofiles;i++){
-		profile[i] = 0;
-	}
-
-	map = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i < numprofiles;i++){
[…]

### 2021-01-29 src/kalign2/kalign2_feature.h

@@ -1,36 +1,0 @@
-/*
-	feature.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include <string.h>
-
-float* feature_hirschberg_update(const float* profa,const float* profb,float* newp,int* path,int sipa,int sipb);
-float* make_unified_profile(float* prof,struct alignment* aln, int num,float** subm,struct feature_matrix* fm);
-void set_unified_gap_penalties(float* prof,int len,int nsip);
-
-int* feature_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm, int* hirsch_path);
-int* feature_hirsch_align_two_pp_vector(const float* prof1,const float* prof2,struct hirsch_mem* hm,int* hirsch_path, float input_states[],int old_cor[]);
-struct states* feature_foward_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-struct states* feature_backward_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-

### 2021-01-29 src/kalign2/kalign2_hirschberg.c

@@ -1,2948 +1,0 @@
-/*
-	kalign2_hirschberg.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_hirschberg.h"
-#define MAX(a, b) (a > b ? a : b)
-#define MAX3(a,b,c) MAX(MAX(a,b),c)
-//#include <emmintrin.h>
-
-int** hirschberg_alignment(struct alignment* aln,int* tree,float**submatrix, int** map,int window,float strength)
-{
-	struct hirsch_mem* hm = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	float** profile = 0;
-
-	profile = malloc(sizeof(float*)*numprofiles);
-	for ( i = 0;i< numprofiles;i++){
-		profile[i] = 0;
-	}
-
-	map = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i < numprofiles;i++){
-		map[i] = 0;
-	}
-	
-	hm = hirsch_mem_alloc(hm,1024);
-
-	fprintf(stderr,"\nAlignment:\n");
-
[…]

### 2021-01-29 src/kalign2/kalign2_hirschberg.h

@@ -1,44 +1,0 @@
-/*
-	kalign2_hirschberg.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-int* hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm, int* hirsch_path);
-struct states* foward_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-struct states* backward_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-int* hirsch_align_two_pp_vector(const float* prof1,const float* prof2,struct hirsch_mem* hm,int* hirsch_path,float input_states[],int old_cor[]);
-
-struct states* backward_hirsch_pp_dynm(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-
-struct states* backward_hirsch_pp_dyn0(const float* prof1,const float* prof2,struct hirsch_mem* hm);
-
-int* hirsch_ps_dyn(const float* prof1,const int* seq2,struct hirsch_mem* hm, int* hirsch_path,int sip);
-struct states* foward_hirsch_ps_dyn(const float* prof1,const int* seq2,struct hirsch_mem* hm,int sip);
-struct states* backward_hirsch_ps_dyn(const float* prof1,const int* seq2,struct hirsch_mem* hm,int sip);
-int* hirsch_align_two_ps_vector(const float* prof1,const int* seq2,struct hirsch_mem* hm,int* hirsch_path,float input_states[],int old_cor[],int sip);
-
-
[…]

### 2021-01-29 src/kalign2/kalign2_hirschberg_dna.c

@@ -1,2011 +1,0 @@
-/*
-	kalign2_hirschberg_dna.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-
-#include "kalign2.h"
-#include "kalign2_hirschberg_dna.h"
-#define MAX(a, b) (a > b ? a : b)
-#define MAX3(a,b,c) MAX(MAX(a,b),c)
-
-
-
-int** dna_alignment(struct alignment* aln,int* tree,float**submatrix, int** map,float strength)
-{
-	struct hirsch_mem* hm = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	float** profile = 0;
-
-	profile = malloc(sizeof(float*)*numprofiles);
-	for ( i = 0;i< numprofiles;i++){
-		profile[i] = 0;
-	}
-
-	map = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i < numprofiles;i++){
-		map[i] = 0;
-	}
-	
-	hm = hirsch_mem_alloc(hm,1024);
-	fprintf(stderr,"\nAlignment:\n");
[…]

### 2021-01-29 src/kalign2/kalign2_hirschberg_dna.h

@@ -1,39 +1,0 @@
-/*
-	kalign2_hirschberg_dna.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-int* hirsch_dna_ss_dyn(float**subm, const int* seq1,const int* seq2,struct hirsch_mem* hm, int* hirsch_path);
-int* hirsch_align_two_dna_ss_vector(float**subm,const int* seq1,const int* seq2,struct hirsch_mem* hm,int* hirsch_path,float input_states[],int old_cor[]);
-struct states* foward_hirsch_dna_ss_dyn(float**subm,const int* seq1,const int* seq2,struct hirsch_mem* hm);
-struct states* backward_hirsch_dna_ss_dyn(float**subm,const int* seq1,const int* seq2,struct hirsch_mem* hm);
-
-int* hirsch_dna_ps_dyn(const float* prof1,const int* seq2,struct hirsch_mem* hm, int* hirsch_path,int sip);
-int* hirsch_align_two_dna_ps_vector(const float* prof1,const int* seq2,struct hirsch_mem* hm,int* hirsch_path,float input_states[],int old_cor[],int sip);
-struct states* foward_hirsch_dna_ps_dyn(const float* prof1,const int* seq2,struct hirsch_mem* hm,int sip);
-struct states* backward_hirsch_dna_ps_dyn(const float* prof1,const int* seq2,struct hirsch_mem* hm,int sip);
-
-int* hirsch_dna_pp_dyn(const float* prof1,const float* prof2,struct hirsch_mem* hm, int* hirsch_path);
[…]

### 2021-01-29 src/kalign2/kalign2_hirschberg_large.c

@@ -1,2508 +1,0 @@
-/*
-	kalign2_hirschberg_large.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_hirschberg_large.h"
-#define MAX(a, b) (a > b ? a : b)
-#define MAX3(a,b,c) MAX(MAX(a,b),c)
-//#include <emmintrin.h>
-
-float local_gpo;
-float local_gpe;
-float local_tgpe;
-
-int** hirschberg_large_alignment(struct alignment* aln,int* tree,float**submatrix, int** map,int window,float strength)
-{
-	struct hirsch_large_mem* hm = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	
-
-	
-	float** profile = 0;
-	
-	float** subm = 0;
-	subm = malloc(sizeof(float*)*32);
-	for(i = 0; i < 32;i++){
-		subm[i] = malloc(sizeof(float)*32);
-		for (j = 0; j < 32;j++){
-			subm[i][j] = (float)submatrix[i][j];
-		}
[…]

### 2021-01-29 src/kalign2/kalign2_hirschberg_large.h

@@ -1,71 +1,0 @@
-/*
-	kalign2_hirschberg_large.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-struct large_states{
-	float a;
-	float ga;
-	float gb;
-	float x;
-};
-
-
-
-struct hirsch_large_mem{
-	struct large_states* f;
-	struct large_states* b;
-	int starta;
-	int startb;
-	int enda;
-	int endb;
-	int size;
-	int len_a;
-	int len_b;
-};
-
-int* hirsch_large_pp_dyn(const float* prof1,const float* prof2,struct hirsch_large_mem* hm, int* hirsch_path);
-struct large_states* foward_large_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_large_mem* hm);
-struct large_states* backward_large_hirsch_pp_dyn(const float* prof1,const float* prof2,struct hirsch_large_mem* hm);
-int* hirsch_large_align_two_pp_vector(const float* prof1,const float* prof2,struct hirsch_large_mem* hm,int* hirsch_path,float input_states[],int old_cor[]);
-
-int* hirsch_large_ps_dyn(const float* prof1,const int* seq2,struct hirsch_large_mem* hm, int* hirsch_path,int sip);
-struct large_states* foward_large_hirsch_ps_dyn(const float* prof1,const int* seq2,struct hirsch_large_mem* hm,int sip);
-struct large_states* backward_large_hirsch_ps_dyn(const float* prof1,const int* seq2,struct hirsch_large_mem* hm,int sip);
[…]

### 2021-01-29 src/kalign2/kalign2_inferface.c

@@ -1,428 +1,0 @@
-/*
-	kalign2_interface.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include <string.h>
-
-#include <getopt.h>
-
-//#include <unistd.h>
-#include "kalign2.h"
-#include "kalign2_input.h"
-
-void parameter_message(struct parameters* param)
-{
-        if(param->infile[0]){
-                if(param->dna){
-                        fprintf(stderr,"Aligning %d RNA/DNA sequences from file %s with these parameters:\n",numseq,param->infile[0]);
-                }else{
-                        fprintf(stderr,"Aligning %d protein sequences from file %s with these parameters:\n",numseq,param->infile[0]);
-                }
-        }else{
-                if(param->dna){
-                        fprintf(stderr,"Aligning %d RNA/DNA sequences with these parameters:\n",numseq);
-                }else{
-                        fprintf(stderr,"Aligning %d protein sequences with these parameters:\n",numseq);
-                }
-                
-        }
-       // fprintf(stderr,"	Distance estimation method:   %s\n",param->distance);
-	//fprintf(stderr,"	Tree building method:         %s\n",param->tree);
-	//fprintf(stderr,"	Ntree:                        %d\n",param->ntree);
-        //fprintf(stderr,"        %0.8f	gap open penalty\n",(float)gpo/5);
-	 fprintf(stderr,"        %0.8f	gap open penalty\n",(float)gpo);
-        //fprintf(stderr,"        %0.8f	gap extension\n",(float)gpe/10);
[…]

### 2021-01-29 src/kalign2/kalign2_input.c

@@ -1,1846 +1,0 @@
-/*
-	kalign2_input.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_input.h"
-
-#include <ctype.h>
-
-
-struct alignment* detect_and_read_sequences(struct alignment* aln,struct parameters* param)
-{
-	
-	int feature = 0;
-	char **input = 0;
-	unsigned short int* input_type = 0;
-	unsigned short int* input_numseq = 0;
-	
-	int num_input = 0;
-	int i = 0;
-	int j = 0;
-	int c = 0;
-	int free_read = 1;
-	while(free_read == 1 || param->infile[i]){
-		num_input++;
-		i++;
-		free_read = 0;
-	}
-	numseq = 0;
-
-	
-	input = malloc(sizeof(char*) * num_input);
-	input_type = malloc(sizeof(unsigned short int) * num_input);
-	input_numseq = malloc(sizeof(unsigned short int) * num_input);
-	
-	for (i = 0; i < num_input;i++){
-		input[i] = 0;
-		input_type[i] = 0;
-		input_numseq[i] = 0;
-	}
-
-	free_read = 0;
-	
-	if(param->quiet){
-		c = 1;
-	}else{
-		c = 0;
-	}
-	
-	
-	for (i = c; i < num_input;i++){
-		if(!param->infile[i]){
-			fprintf(stderr,"reading from STDIN: ");
[…]

### 2021-01-29 src/kalign2/kalign2_input.h

@@ -1,123 +1,0 @@
-/*
-	kalign2_input.h 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-
-#include <unistd.h>
-#include <string.h>
-#include <ctype.h>
-
-#define SEEK_START 0
-#define SEEK_END 2
-
-struct alignment* read_sequences(struct alignment* aln,char* string);
-struct alignment* read_sequences_from_swissprot(struct alignment* aln,char* string);
-struct alignment* read_sequences_uniprot_xml(struct alignment* aln,char* string);
-struct alignment* read_sequences_macsim_xml(struct alignment* aln,char* string);
-struct feature* read_ft(struct feature* ft,char* p);
-struct alignment* read_sequences_clustal(struct alignment* aln,char* string);
-struct alignment* read_sequences_stockholm(struct alignment* aln,char* string);
-
-struct alignment* read_alignment(struct alignment* aln,char* string);
-struct alignment* read_alignment_from_swissprot(struct alignment* aln,char* string);
-struct alignment* read_alignment_uniprot_xml(struct alignment* aln,char* string);
-struct alignment* read_alignment_macsim_xml(struct alignment* aln,char* string);
-struct feature* read_ft(struct feature* ft,char* p);
-struct alignment* read_alignment_clustal(struct alignment* aln,char* string);
-struct alignment* read_alignment_stockholm(struct alignment* aln,char* string);
-
-char* get_input_into_string(char* string,char* infile);
-
-
-
[…]

### 2021-01-29 src/kalign2/kalign2_main.c

@@ -1,310 +1,0 @@
-/*
-	kalign2_main.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-
-#include "kalign2.h"
-
-
-
-unsigned int numseq = 0;
-unsigned int numprofiles = 0;
-float gpo = 0;
-float gpe = 0;
-float tgpe = 0;
-
-int main(int argc,char **argv)
-{
-	int i;
-	int* tree = 0;
-	int a, b, c;
-	
-	struct alignment* aln = 0;
-	struct parameters* param = 0;
-	struct aln_tree_node* tree2 = 0;
-	
-	param = malloc(sizeof(struct parameters));
-	
-	param =  interface(param,argc,argv);
-	
-	aln = detect_and_read_sequences(aln,param);
-	
-	if(param->ntree > numseq){
-		param->ntree = numseq;
-	}
-
-	//DETECT DNA
-	if(param->dna == -1){
-		for (i = 0; i < numseq;i++){
-			param->dna = byg_detect(aln->s[i],aln->sl[i]);
-			if(param->dna){
-				break;
-			}
-		}
-	}
-	//param->dna = 0;
-	//fprintf(stderr,"DNA:%d\n",param->dna);
-	//exit(0);
-	
-	if(param->dna == 1){
-		//brief sanity check...
-		for (i = 0; i < numseq;i++){
-			if(aln->sl[i] < 6){
[…]

### 2021-01-29 src/kalign2/kalign2_mem.c

@@ -1,291 +1,0 @@
-/*
-	kalign2_mem.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-void freesimpletree(struct tree_node* p)
-{
-	if(p->left){
-		freesimpletree(p->left);
-	}
-	if(p->right){
-		freesimpletree(p->right);
-	}
-	free(p);
-}
-
-void free_real_tree(struct aln_tree_node* p)
-{
-	int i = 0;
-	while(p->links[i]){
-		free_real_tree(p->links[i]);
-		i++;	
-	}
-	free(p->internal_lables);
-	free(p->links);
-	free(p);
-}
-
-
-void free_feature_matrix(struct feature_matrix* fm)
-{
-	int i;
-	for (i = 0;i < fm->mdim;i++){
-		free(fm->m[i]);
-	}
-	free(fm->m);
-	free(fm);
-}
-
-
-void free_utf(struct utype_ufeat* utf)
-{
-	free_ft(utf->t);
-	free_ft(utf->f);
-	free(utf);
-}
-
-/*#ifndef MEMORY 
-void* malloc(int size)
-{
-	void* p;
[…]

### 2021-01-29 src/kalign2/kalign2_misc.c

@@ -1,743 +1,0 @@
-/*
-	kalign2_misc.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-
-
-#include <string.h>
-#include <ctype.h>
-#include "kalign2.h"
-
-void print_tree(struct aln_tree_node* p,struct alignment* aln,char* outfile)
-{
-	FILE *fout = NULL;
-	if ((fout = fopen(outfile, "w")) == NULL){
-		fprintf(stderr,"can't open output\n");
-		exit(0);
-	}
-	//fprintf(stderr,"\n\n%s\n",outfile);
-	/*if(byg_start("xml",outfile) != -1){
-		fprintf(fout,"<?xml version=\"1.0\" encoding=\"UTF-8\"?> <phyloxml xmlns:xsi=\"http://www.w3.org/2001/XMLSchema-instance\" xsi:noNamespaceSchemaLocation=\"http://www.phyloxml.org/schema/phyloxml.xsd\"><phylogeny>\n");
-		
-		print_phyloxml_tree(p,aln,fout);
-		
-		fprintf(fout,"</phylogeny></phyloxml>\n");
-		
-	}else{*/
[…]

### 2021-01-29 src/kalign2/kalign2_output.c

@@ -1,662 +1,0 @@
-/*
-	kalign2_output.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_output.h"
-
-void output(struct alignment* aln,struct parameters* param)
-{
-	if(!param->format){
-		fasta_output(aln,param->outfile);
-	}else{
-		if (byg_start(param->format,"alnALNclustalCLUSTALclustalwCLUSTALWclustalWClustalW") != -1){
-			aln_output(aln,param);
-		}else if (byg_start(param->format,"msfMSFgcgGCGpileupPILEUP") != -1){
-			msf_output(aln,param->outfile);
-		}else if (byg_start(param->format,"eclu") != -1){
-			clustal_output(aln,param->outfile);
-		}else if (byg_start("macsim",param->format) != -1){
-			macsim_output(aln,param->outfile,param->infile[0]);
-		}else{
-			fasta_output(aln,param->outfile);
-		}
-	}
-	free_param(param);
-}
-
[…]

### 2021-01-29 src/kalign2/kalign2_output.h

@@ -1,39 +1,0 @@
-/*
-	kalign2_output.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include <ctype.h>
-
-void aln_output(struct alignment* aln,struct parameters* param);
-void msf_output(struct alignment* aln,char* outfile);
-void fasta_output(struct alignment* aln,char* outfile);
-void clustal_output(struct alignment* aln,char* outfile);
-void macsim_output(struct alignment* aln,char* outfile,char* infile);
-
-struct names* get_meaningful_names(struct alignment* aln,int id);
-
-
-
-
-

### 2021-01-29 src/kalign2/kalign2_profile.c

@@ -1,1082 +1,0 @@
-/*
-	kalign2_profile.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-
-/*
-void add_feature_information_from_alignment(int* path,int* fprof1,int* fprof2,int weight)
-{
- 	int i = 0;
- 	int j = 0;
- 	int c = 1;
-	while(path[c] != 3){		
-		if (!path[c]){
-			fprof1[i] +=1;
-			fprof1[i+1] +=weight;
-			fprof2[j] +=1;
-			fprof2[j+1] +=weight;
-			i+=2;
-			j+=2;	
-		}
-		if (path[c] & 1){
-			j+=2;
-		}
-		if (path[c] & 2){
-			i+=2;		
-		}
-		c++;
-	}
-	free(path);
-}*/
-
-
-float* update2(const float* profa, const float* profb,float* newp,int* path,int sipa,int sipb,float internal_gap_weight)
-{
-	int i,c;
-	int* gap_len = 0;
-	int gap_cost = 0;
-	
-	gap_len = malloc(sizeof(int)* (path[0]+1));
-	gap_len[0] = 0;
[…]

### 2021-01-29 src/kalign2/kalign2_profile_alignment.c

@@ -1,579 +1,0 @@
-/*
-	kalign2_profile_alignment.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-#include "kalign2_profile_alignment.h"
-#include "kalign2_hirschberg.h"
-
-void profile_alignment_main(struct alignment* aln,struct parameters* param,float** submatrix)
-{
-	float** dm = 0;
-	int* tree = 0;
-	struct aln_tree_node* tree2 = 0;
-	int i,j;
-	int tmp_numseq;
-	int tmp_numprofiles;
-	
-	local_numseq = 0;
-	local_numprofiles = 0;
-	
-	//determine number of profiles that were inputted....
-	
-	while(aln->sl[local_numseq+numseq]){
-		local_numseq++;
-	}
-	
-	local_numprofiles = (local_numseq << 1) - 1;
-	//fprintf(stderr,"%d	%d\n",local_numseq,local_numprofiles);
-	
-	for (i = 0;i < numseq;i++){
[…]

### 2021-01-29 src/kalign2/kalign2_profile_alignment.h

@@ -1,40 +1,0 @@
-/*
-	kalign2_profile_alignment.h
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-int local_numseq;
-int local_numprofiles;
-
-int* assign_gap_codes(int* seq,int len);
-
-
-int is_member(struct alignment* aln,int test);
-
-float** dna_profile_distance(struct alignment* aln,float** dm,struct parameters* param, int nj);
-float** protein_profile_wu_distance(struct alignment* aln,float** dm,struct parameters* param, int nj);
-
-
-int** hirschberg_profile_alignment(struct alignment* aln,int* tree,float**submatrix, int** map);
-float* make_profile_from_alignment(float* prof, int num,struct alignment* aln,float** subm);
-
-

### 2021-01-29 src/kalign2/kalign2_simple_gaps.c

@@ -1,869 +1,0 @@
-/*
-	kalign2_simple_gaps.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-int** simple_hirschberg_alignment(struct alignment* aln,int* tree,int**submatrix, int** map)
-{
-	struct hirsch_mem* hm = 0;
-	int i,j,g,a,b,c;
-	int len_a;
-	int len_b;
-	int** profile = 0;
-
-	profile = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i< numprofiles;i++){
-		profile[i] = 0;
-	}
-
-	map = malloc(sizeof(int*)*numprofiles);
-	for ( i = 0;i < numprofiles;i++){
-		map[i] = 0;
-	}
-	
-	hm = hirsch_mem_alloc(hm,1024);
-
-	fprintf(stderr,"\nAlignment:\n");
-
-	for (i = 0; i < (numseq-1);i++){
-		a = tree[i*3];
-		b = tree[i*3+1];
-		c = tree[i*3+2];
[…]

### 2021-01-29 src/kalign2/kalign2_stats.c

@@ -1,23 +1,0 @@
-#include "kalign2.h"
-
-
-void stats(struct alignment* aln)
-{
-	int i,j;
-	float 
-	
-	for (i = 0; i < numseq-1;i++){
-		for (j = i + 1; j < numseq;j++){
-			for (j = 0; j < aln->sl[f];j++){
-		}
-		f = aln->nsip[i];
-		fprintf(stdout,">%s\n",aln->sn[f]);
-		c = 0;
-		for (j = 0; j < aln->sl[f];j++){
-			tmp = aln->s[f][j];]
-			aln->seq[f][j]
-		}
-	}
-}
-
-

### 2021-01-29 src/kalign2/kalign2_string_matching.c

@@ -1,174 +1,0 @@
-/*
-	kalign2_string_matching.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include <string.h>
-
-int byg_detect(int* text,int n)
-{
-	int Tc;
-	int i  = 0;
-	int s = 0;
-	int T[256];
-	for (i = 0;i < 256;i++){ 
-		T[i] = 0; 
-	}
-	int mb = 1;
-	//char *unique_aa = "EFILPQXZ";//permissiv
-	//ABCDEFGHIJKLMNOPQRSTUVWXYZ
-	char *unique_aa = "BDEFHIJKLMNOPQRSVWYZ";//restrictive
-	int aacode[26] = {0,1,2,3,4,5,6,7,8,-1,9,10,11,12,23,13,14,15,16,17,17,18,19,20,21,22};
-	for (i= 0;i < 20;i++){
-		T[(int)aacode[unique_aa[i]-65]] |= 1;
-	}
-	for (i = 0;i < n;i++){
-	//	fprintf(stderr,"%d\n",text[i]);
-		if(text[i] != -1){
-			s <<= 1;
-			s |= 1;
-			Tc = T[text[i]];
-			s &= Tc;
-			if(s & mb){
-				return 0;
-			}
-		}
-	}
-	return 1;
-}
-
[…]

### 2021-01-29 src/kalign2/kalign2_tree.c

@@ -1,872 +1,0 @@
-/*
-	kalign2_tree.c 
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-struct aln_tree_node* real_upgma(float **dm,int ntree)
-{
-	int i,j;
-	int *as = 0;
-	float max;
-	int node_a = 0;
-	int node_b = 0;
-	int cnode = numseq;
-	
-	struct aln_tree_node** tree = 0;
-	struct aln_tree_node* tmp = 0;
-	
-	as = malloc(sizeof(int)*numseq);
-	for (i = numseq; i--;){
-		as[i] = i+1;
-	}
-	
-	tree = malloc(sizeof(struct aln_tree_node*)*numseq);
-	for (i=0;i < numseq;i++){
-		tree[i] = malloc(sizeof(struct aln_tree_node));
-		tree[i]->done = 1;
-		tree[i]->num = i;
-		tree[i]->path = 0;
-		tree[i]->profile = 0;
-		tree[i]->seq = 0;//seq[i];
-		tree[i]->len = 0;//len[i]; 
-		/*
-		Needs to be +2 because:
[…]

### 2021-01-29 src/kalign2/kalign2_upgma.c

@@ -1,283 +1,0 @@
-/*
-	kalign2_upgma.c
-	
-	Released under GPL - see the 'COPYING' file   
-	
-	Copyright (C) 2006 Timo Lassmann <timolassmann@gmail.com>
-	
-	This program is free software; you can redistribute it and/or modify
-	it under the terms of the GNU General Public License as published by
-	the Free Software Foundation; either version 2 of the License, or
-	any later version.
-
-	This program is distributed in the hope that it will be useful,
-	but WITHOUT ANY WARRANTY; without even the implied warranty of
-	MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-	GNU General Public License for more details.
-
-	You should have received a copy of the GNU General Public License
-	along with this program; if not, write to the Free Software
-	Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-    
-	Please send bug reports, comments etc. to:
-	timolassmann@gmail.com
-*/
-
-#include "kalign2.h"
-
-struct aln_tree_node* real_upgma(float **dm)
-{
-	int i,j;
-	int *as = 0;
-	float max;
-	int node_a = 0;
-	int node_b = 0;
-	int cnode = numseq;
-	
-	struct aln_tree_node** tree = 0;
-	struct aln_tree_node* tmp = 0;
-	
-	as = malloc(sizeof(int)*numseq);
-	for (i = numseq; i--;){
-		as[i] = i+1;
-	}
-	
-	tree = malloc(sizeof(struct aln_tree_node*)*numseq);
-	for (i=0;i < numseq;i++){
-		tree[i] = malloc(sizeof(struct aln_tree_node));
-		tree[i]->done = 1;
-		tree[i]->num = i;
-		tree[i]->path = 0;
-		tree[i]->profile = 0;
-		tree[i]->seq = 0;//seq[i];
-		tree[i]->len = 0;//len[i]; 
-		/*
-		Needs to be +2 because:
-		at n = 3 is is possible to get a perfectly balanced binary tree with 4 sequences at intermediate nodes
-		*/
[…]

### 2021-01-29 src/keras_dgl/utils.py

@@ -1,267 +1,0 @@
-from __future__ import print_function
-
-import scipy.sparse as sp
-import numpy as np
-from scipy.sparse.linalg.eigen.arpack import eigsh, ArpackNoConvergence
-
-
-def encode_onehot(labels):
-    classes = set(labels)
-    classes_dict = {c: np.identity(len(classes))[i, :] for i, c in enumerate(classes)}
-    labels_onehot = np.array(list(map(classes_dict.get, labels)), dtype=np.int32)
-    return labels_onehot
-
-
-def load_data(path="data/cora/", dataset="cora"):
-    """Load citation network dataset (cora only for now)"""
-    print('Loading {} dataset...'.format(dataset))
-
-    idx_features_labels = np.genfromtxt("{}{}.content".format(path, dataset), dtype=np.dtype(str))
-    features = sp.csr_matrix(idx_features_labels[:, 1:-1], dtype=np.float32)
-    labels = encode_onehot(idx_features_labels[:, -1])
-
-    # build graph
-    idx = np.array(idx_features_labels[:, 0], dtype=np.int32)
-    idx_map = {j: i for i, j in enumerate(idx)}
-    edges_unordered = np.genfromtxt("{}{}.cites".format(path, dataset), dtype=np.int32)
-    edges = np.array(list(map(idx_map.get, edges_unordered.flatten())),
-                     dtype=np.int32).reshape(edges_unordered.shape)
-    adj = sp.coo_matrix((np.ones(edges.shape[0]), (edges[:, 0], edges[:, 1])),
-                        shape=(labels.shape[0], labels.shape[0]), dtype=np.float32)
-
-    # build symmetric adjacency matrix
-    adj = adj + adj.T.multiply(adj.T > adj) - adj.multiply(adj.T > adj)
-
-    # features = normalize_features(features)
-    # adj = normalize_adj(adj + sp.eye(adj.shape[0]))
-
-    print('Dataset has {} nodes, {} edges, {} features.'.format(adj.shape[0], edges.shape[0], features.shape[1]))
-
-    return features.todense(), adj, labels
-
-
-def load_data_attention(path="data/cora/", dataset="cora"):
-    """Load citation network dataset (cora only for now)"""
-    print('Loading {} dataset...'.format(dataset))
-
[…]

### 2021-01-29 paper/main.tex

@@
 drug--target interactions. Specifically, given a protein, we are
-interested in identifying the drugs that may target the protein, so as
-to reflect the case where a drug is desired to target a
-disease-associated protein.
+interested in identifying the drugs that may target the protein; this
+scenario is motivated by the need to identify drugs that can target a
+specific protein that may be associated with a disease or abnormal
+phenotypes.
+Our aim is further to combine two different types of features:
+structural information based on protein structure and the
 
-To obtain a prediction for a given drug-protein pair, we first build DL2vec representations for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
+To obtain a prediction for a given drug-protein pair\todo{should this
+  be about pairs, or rather something like ``given a protein, we aim
+  to find a drug...''? I.e., find a consistent way to separate
+  yourself language-wise from a naive prediction of pairs.}, we first
+build DL2vec representations\todo{In results: talk about what they are
+supposed to DO (``embeddings/representations of the phenotypes and
+functions...''), not how you implement them (you could use something
+else, for example).} for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
 Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.

### 2021-01-29 src/MSAProbs-0.9.7/MSAProbs/MSADef.h
Purge src dir 2

@@ -1,26 +1,0 @@
-#ifndef _MSA_DEF_H
-#define _MSA_DEF_H
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <float.h>
-#include <math.h>
-
-#ifdef _OPENMP
-#include <omp.h>
-#endif
-
-//maximum number
-#define MAX_INT_NUM				0x7FFFFFFF
-#define MAX_FLOAT_NUM			FLT_MAX
-#define INT_MULTIPLY			1000
-
-#define SUBMATRIX_INT_SCALE		100
-
-//a tree node is a leaf or a node
-enum {
-	NONE, NODE, LEAF
-};
-
-#endif
-

### 2021-01-29 src/MSAProbs-0.9.7/MSAProbs/MSAReadMatrix.cpp

@@ -1,174 +1,0 @@
-/***********************************************
- * # Copyright 2009-2010. Liu Yongchao
- * # Contact: Liu Yongchao, School of Computer Engineering,
- * #			 Nanyang Technological University.
- * # Emails:	 liuy0039@ntu.edu.sg; nkcslyc@hotmail.com
- * #
- * # GPL version 3.0 applies.
- * #
- * ************************************************/
-
-#include <string.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "MSAReadMatrix.h"
-
-#define TRACE 0
-
-////////////////////////////////////////////////////////////
-// extern variables for scoring matrix data
-////////////////////////////////////////////////////////////
-extern float g_gap_open1, g_gap_open2, g_gap_ext1, g_gap_ext2;
-extern char *aminos, *bases, matrixtype[20];
-extern int subst_index[26];
-
-extern double sub_matrix[26][26];
-
-extern float TEMPERATURE;
-extern int MATRIXTYPE;
-
-extern float GAPOPEN;
-extern float GAPEXT;
-
-typedef struct {
-	char input[30];
-	int matrix;
-	int N;
-	float T;
-	float beta;
-	char opt;			//can be 'P' or 'M'
-	float gapopen;
-	float gapext;
-} argument_decl;
-
-//argument support
-extern argument_decl argument;
-
-/////////////////////////////////////////////////////////
-//sets substitution matrix type
-////////////////////////////////////////////////////////
-void setmatrixtype(int le) {
-	switch (le) {
-	case 160:
-		strcpy(matrixtype, "gonnet_160");
-		break;
-	case 4:
-		strcpy(matrixtype, "nuc_simple");
-		break;
-	default:
-		strcpy(matrixtype, "CUSTOM");
-		break;
-
-	};
-
-}
-
-///////////////////////////////////////////////////////////////////
-//sets matrix flag
-///////////////////////////////////////////////////////////////////
-inline int matrixtype_to_int() {
-
-	if (!strcmp(matrixtype, "nuc_simple"))
-		return 4;
-	else if (!strcmp(matrixtype, "gonnet_160"))
-		return 160;
-	else
-		return 1000;
-
-}
-
-/////////////////////////////////////////////////////////////////
-//
[…]

### 2021-01-29 src/MSAProbs-0.9.7/MSAProbs/SafeVector.h

@@ -1,65 +1,0 @@
-/////////////////////////////////////////////////////////////////
-// SafeVector.h
-//
-// STL vector with array bounds checking.  To enable bounds
-// checking, #define ENABLE_CHECKS.
-/////////////////////////////////////////////////////////////////
-
-#ifndef SAFEVECTOR_H
-#define SAFEVECTOR_H
-
-#include <cassert>
-#include <vector>
-using namespace std;
-
-/////////////////////////////////////////////////////////////////
-// SafeVector
-//
-// Class derived from the STL std::vector for bounds checking.
-/////////////////////////////////////////////////////////////////
-
-template<class TYPE>
-class SafeVector: public std::vector<TYPE> {
-public:
-
-	// miscellaneous constructors
-	SafeVector() :
-			std::vector<TYPE>() {
-	}
-	SafeVector(size_t size) :
-			std::vector<TYPE>(size) {
-	}
-	SafeVector(size_t size, const TYPE &value) :
-			std::vector<TYPE>(size, value) {
-	}
-	SafeVector(const SafeVector &source) :
-			std::vector<TYPE>(source) {
-	}
-
-#ifdef ENABLE_CHECKS
-
-	// [] array bounds checking
-	TYPE &operator[](int index) {
-		assert (index >= 0 && index < (int) size());
-		return std::vector<TYPE>::operator[] ((size_t) index);
-	}
-
-	// [] const array bounds checking
-	const TYPE &operator[] (int index) const {
-		assert (index >= 0 && index < (int) size());
-		return std::vector<TYPE>::operator[] ((size_t) index);
-	}
-
-#endif
-
-};
-
-// some commonly used vector types
-typedef SafeVector<int> VI;
-typedef SafeVector<VI> VVI;
-typedef SafeVector<VVI> VVVI;
-typedef SafeVector<float> VF;
-typedef SafeVector<VF> VVF;
-typedef SafeVector<VVF> VVVF;
-
-#endif

### 2021-01-29 src/MSAProbs-0.9.7/MSAProbs/SparseMatrix.h

@@ -1,266 +1,0 @@
-/////////////////////////////////////////////////////////////////
-// SparseMatrix.h
-//
-// Sparse matrix computations
-/////////////////////////////////////////////////////////////////
-
-#ifndef SPARSEMATRIX_H
-#define SPARSEMATRIX_H
-
-#include <iostream>
-
-using namespace std;
-
-const float POSTERIOR_CUTOFF = 0.01;         // minimum posterior probability
-// value that is maintained in the
-// sparse matrix representation
-
-typedef pair<int, float> PIF;                 // Sparse matrix entry type
-//   first --> column
-//   second --> value
-
-/////////////////////////////////////////////////////////////////
-// SparseMatrix
-//
-// Class for sparse matrix computations
-/////////////////////////////////////////////////////////////////
-
-class SparseMatrix {
-
-	int seq1Length, seq2Length;                     // dimensions of matrix
-	VI rowSize;                              // rowSize[i] = # of cells in row i
-	SafeVector<PIF> data;                           // data values
-	SafeVector<SafeVector<PIF>::iterator> rowPtrs; // pointers to the beginning of each row
-
-	/////////////////////////////////////////////////////////////////
-	// SparseMatrix::SparseMatrix()
-	//
-	// Private constructor.
-	/////////////////////////////////////////////////////////////////
-
-	SparseMatrix() {
-	}
-
-public:
-
-	/////////////////////////////////////////////////////////////////
-	// SparseMatrix::SparseMatrix()
-	//
-	// Constructor.  Builds a sparse matrix from a posterior matrix.
-	// Note that the expected format for the posterior matrix is as
-	// a (seq1Length+1) x (seq2Length+1) matrix where the 0th row
-	// and 0th column are ignored (they should contain all zeroes).
-	/////////////////////////////////////////////////////////////////
-
-	SparseMatrix(int seq1Length, int seq2Length, const VF &posterior) :
-			seq1Length(seq1Length), seq2Length(seq2Length) {
-
-		int numCells = 0;
-
-		assert(seq1Length > 0);
-		assert(seq2Length > 0);
-
-		// calculate memory required; count the number of cells in the
-		// posterior matrix above the threshold
[…]

### 2021-01-29 paper/main.tex
.

@@
 \subtitle{Subject Section}
+\newcommand{\name}{Voodoo }
 
-\title[short Title]{Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
+\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
 \author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
@@
 
-\subsection{High level description}
-Our aim is to develop and evaluate a computational model to predict
-drug--target interactions. Specifically, given a protein, we are
-interested in identifying the drugs that may target the protein; this
-scenario is motivated by the need to identify drugs that can target a
+\subsection{\name identifies drugs that target a protein}
+
+We developed \name as a computational model to predict drug--target
+interactions. Specifically, given a protein, \name will identify and
+rank drugs that likely target this protein. \name combines two 
+types of features: structural information for drugs and proteins that
+can be used to determine if they may physically interact, and
+
+This scenario is motivated by the need to identify drugs that can target a
 specific protein that may be associated with a disease or abnormal
@@
 	\item Yamanishi Dataset is only partially suitable \todo{for
-            comparing results}}, if everybody just derives a suitable subset (DTIGEMS)
+            comparing results}, if everybody just derives a suitable subset (DTIGEMS)
 	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap

### 2021-01-29 paper/main.tex
Update splitting scheme section 1

--- 2021-01-27 paper/main.tex
@@
 	\centerline{\includegraphics[width=0.5\textwidth]{figures/drug_protein_ontology_network.png}}
-	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}\label{fig:Onto}
+	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}
+	\label{fig:Onto}
 \end{figure}
@@
 The similarity function $\otimes$ yields the similarity between both transformed embeddings e.g.
-by computing the cosine similarity.}\label{fig:SiameseNetwork}
+by computing the cosine similarity.}
+	\label{fig:SiameseNetwork}
 \end{figure}
@@
 Added self-loops enforce that each node representation is directly dependent on its own preceding one.
-Notably, the number of graph convolutional layers stacked equals the radius of relevant nodes for each vertex within the graph.\\
+Notably, the number of graph convolutional layers stacked equals the radius of relevant nodes for each vertex within the graph.
+\\
 
@@
 Hence, while \textsc{GCNConv} suffers from both vanishing gradients and signal fading for large scale, highly connected graphs, each propagation step in \textsc{GENConv} emphasizes strong signals with values close to $0$ and $1$.
-\\
+Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
+Note additionally, that the resulting information hold no information on their own connectivity.\\
+
+\begin{figure}[!tpb]%figure1
+	\centerline{\includegraphics[width=0.5\columnwidth]{figures/ResGraphConvBlocks.png}}
+	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
+	\label{fig:ResGraphConvBlocks}
+\end{figure}
 
@@
 \\
-
- \begin{figure}[!tpb]%figure1
- 	\centerline{\includegraphics[width=0.2\textwidth]{figures/ResGraphConvBlocks.png}}
- 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
- 	\label{fig:ResGraphConvBlocks}
- \end{figure}
 
@@
 
-\subsection{High level description}
+\subsection{Choosing a cross-validation splitting scheme}
+As in machine learning inference is derived from the underlying data, models and data are naturally and intrinsically linked.
+Thus, the more we understand about the pitfalls and biases within the data, the more we can try to bypass these difficulties.
+We will hereby abbreviate, e.g., a cross validation splitting scheme as \glqq split\grqq{} determining the train, validation and test subset of a given dataset.
+\\
+Especially within DTI prediction, one can observe elemental and basic skews \citep{Pahikkala2014}.
+(1.) Frequently, novel drugs are designed by altering non-functional components of a drug, leading to two and more very similar drugs, designed to target the same proteins (\textbf{CITATION?}).
+This bias can lead to \glqq hidden duplicates\grqq{} that can distribute among the train-test split, skewing the ultimate predictive performance.
+(2.) Additionally, within the realm of DTI prediction, some proteins, which we will denote as \textit{hub proteins}, have significantly more known interactions with drugs than others.
+In the considered subset of STITCH database, $5\%$ of the proteins had $40\%$ of the interactions, with similar biases for Yamanishi and Drugbank (\citet{Drugbank2007} and \citet{Drugbank2017}) datasets, making them easy and reliable predictors.
+(3.) The second distortion could be based on a publication bias for more \glqq valuable\grqq{} proteins due to their disease involvement, such as e.g., cancer, leading to more research and eventually more interacting drugs regarding those.
+Thus if we aim to find a drug for a given protein, the challenge is to develop a training and evaluation scheme, that does not simply overfit to those biases.\\
+
+In general, when concerning cross-validation splitting schemes for DTI prediction, there are the following three options:
+\begin{enumerate}
+	\item Build split over drugs
+	\item Build split over drug-target pairs
+	\item Build split over proteins
+\end{enumerate}
+where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
+Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
+\\
+The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.
+Thus, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
+Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions, while in the third scheme the number of interactions of the test proteins is not known.
+\\
+
+Considering the publications of the last
+
+In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
+Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
+However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
+
+\todo[inline]{Relate these to the issues listed above:}
+\begin{enumerate}
+	\item Build split over drugs
+	\item Build split over drug-target pairs
+	\item Build split over proteins
+\end{enumerate} 
+In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
+As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results.
+This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}.
+The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
+Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive.
+As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes.
+\\
+
+In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
+The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
+However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier.
+Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE).
+When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
+Hence, when entirely new molecules are seen, the models perform much worse
+
+The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
+
+As mentioned in the introduction it is quite difficult to learn suitable features from proteins.
+In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features.
+\\
+Thus, building splitting over proteins is the most challenging of the three options. \\
+
+\subsection{Baseline model}
 Our aim is to develop and evaluate a computational model to predict
 drug--target interactions. Specifically, given a protein, we are
-interested in identifying the drugs that may target the protein, so as
-to reflect the case where a drug is desired to target a
-disease-associated protein.
+interested in identifying the drugs that may target the protein; this
+scenario is motivated by the need to identify drugs that can target a
+specific protein that may be associated with a disease or abnormal
+phenotypes.
+Our aim is further to combine two different types of features:
+structural information based on protein structure and molecular drug composition, and their respective observable characteristics. \\
 
-To obtain a prediction for a given drug-protein pair, we first build DL2vec representations for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
-Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.
+Before introducing the core of our eventual model, we propose a suitable naive baseline model in order to analyze and understand the existing approaches and this novel method for drug target interaction prediction.
+For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, we construct an interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
+\begin{equation*}
+	M_{ij} = \begin{cases}
+		1 & \text{if } (d_i, p_j)\in Int\\
+		0 & otherwise
+	\end{cases}
+\end{equation*}
+describing for all drug--protein whether there is a known interaction or not. We now rank all proteins $p_j\in P$ descending by their number of drug interactors, characterized by 
+\begin{equation*}
+	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=0}^{|D|}M_{ij}
+\end{equation*}
+by summing over the columns of $M_{ij}$ and ranking these sums.
+We now finish our baseline predictor $P_k$ by predicting all drugs to interact with the top $k$ targets, denoted by $TopK(P)$ w.r.t. $M_{ij}$ from the previously introduced ranking, formalized by
+
+\begin{equation*}
+	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
+		1 & \text{ if }p_j \in TopK(P)\\
+		0 & otherwise
+	\end{cases}
+\end{equation*}
+with the only hyperparameter $k$.
+Note particularly, that the consequent prediction of $P_k$ is not dependent on the considered drug $d_i$, and will thus predict all drugs similarly for a given protein $p_j$.
+Subsequently, this naive predictor is not yielding any valuable information on the individual interactions with drugs for a given protein.
+Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
+
+\subsection{High level description}
+
+For a given protein, we want to find all suitable interacting drugs, \textbf{among a set of known drugs}.
+To obtain a prediction for a given drug-protein pair, we first
+build representations embodying both phenotypes and functions for both drugs and proteins, while also preparing structural, molecular representations, e.g., utilizing DL2vec over PhenomeNET and the prebuilt features from SMILES transformer and DeepGOPlus.
+Note that this approach can be easily generalized, profiting from other protein function and phenotype representation methods.
+\\
+Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers (DNNT) for both methods, and both drugs and proteins, individually.
 These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.
 Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs.
-The overall architecture can be seen in Figure~\ref{fig:Fullmodel}.
+The overall architecture is depicted in Figure~\ref{fig:Fullmodel}.\\
 
-\begin{figure}[!tpb]%figure1
-	\centerline{\includegraphics[width=0.33\textwidth]{figures/full_model.png}}
+\begin{figure}[!tpb]
+	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
 	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
@@
 \end{figure}
+
+Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model.
+Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
+By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
 
@@
 \end{figure}
-
-\subsection{Choosing a train-test splitting scheme}
-
-\todo[inline]{List a number of general challenges first, then refer
-  back to them: very similar drugs are designed to target the same
-  proteins, minor variations, leading to a kind of ``hidden''
-  duplication of drugs in the training/testing set; some proteins act
-  like hubs and interact with many drugs (making them easy and
-  reliable predictors); shared indications of drugs are usually
-  because they target the same proteins by design leading to some
-  trivial predictions; some proteins are more ``valuable'' targets due
-  to their disease involvement (such as cancer, meaning more drugs are
-  developed for targeting those). If the aim is to find a drug for a
-  given protein, the challenge is to develop a training (and
-  evaluation) scheme that does not simply overfit to these biases.}
-
-In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
-Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
-However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
-
-\todo[inline]{Relate these to the issues listed above:}
-\begin{enumerate}
-	\item Build split over drugs
-	\item Build split over drug-target pairs
-	\item Build split over proteins
-\end{enumerate} 
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results.
-This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}.
-The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
-Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive.
-As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes.
-\\
-
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
-However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier.
-Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE).
-When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
-Hence, when entirely new molecules are seen, the models perform much worse
-
-The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
-
-As mentioned in the introduction it is quite difficult to learn suitable features from proteins.
-In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features.
-\\
-Thus, building splitting over proteins is the most challenging of the three options. \\
 

### 2021-01-29 src/results_evaluation.py

@@ -120,13 +120,25 @@ def test_Yamanishi_AUC():
         for line in f:
             dti_matrix.append(list(map(int, list(line.strip().replace(' ', '')))))
     dti_matrix = np.array(dti_matrix)
+
+    '''
+    path = '../results/DTIGEMS/'
+    filename = path + 'ic_admat_dgc.txt'
+
+    dti_matrix = np.genfromtxt(filename, delimiter='\t', dtype=str)
+    dti_matrix = dti_matrix[1:, :][:, 1:]
+    dti_matrix = dti_matrix.astype(np.int)
+    '''
     print('dti_matrix.shape', dti_matrix.shape)
-    print(dti_matrix[:, 0].sum())
+
     print(dti_matrix.sum())
+
+    print((dti_matrix.sum(axis=1)==0).sum())
+
+    raise Exception
 
     num_drugs, num_prots = dti_matrix.shape
 
-    '''
     sum_vec = dti_matrix.sum(axis=0)
 
     print('sumvec:', sum_vec.shape, sum_vec[:20])
@@ -137,8 +149,7 @@ def test_Yamanishi_AUC():
 
     dti_matrix = dti_matrix.flatten()
 
-    for i in range(200, num_prots+1, 1):
-        continue
+    for i in range(0, num_prots+1, 1):
         y_pred = np.zeros((num_drugs, num_prots))
         y_pred[:, idx[:i]] = 1
         y_pred = y_pred.flatten()
@@ -146,10 +157,10 @@ def test_Yamanishi_AUC():
         print(f'i: {i}, ROCAUC: {dti_utils.dti_auroc(y_true=dti_matrix, y_pred=y_pred)}, '
               f'microAUC: {dti_utils.micro_AUC_per_prot(y_true=dti_matrix, y_pred=y_pred, num_drugs=num_drugs)}, '
               f'{dti_utils.micro_AUC_per_drug(dti_matrix, y_pred,num_drugs)}')
-    '''
 
+    return
 
-    kf = KFold(n_splits=5, random_state=12, shuffle=True)
+    kf = KFold(n_splits=5, random_state=42, shuffle=True)
     X = np.zeros((num_drugs, 1))
 
     train_aucs = []

### 2021-01-29 paper/main.tex

@@
 scenario is motivated by the need to identify drugs that can target a
+
+\subsection{\name identifies drugs that target a protein}
+
+We developed \name as a computational model to predict drug--target
+interactions. Specifically, given a protein, \name will identify and
+rank drugs that likely target this protein. \name combines two 
+types of features: structural information for drugs and proteins that
+can be used to determine if they may physically interact, and
+
+This scenario is motivated by the need to identify drugs that can target a
 specific protein that may be associated with a disease or abnormal
@@
 Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
+
+To obtain a prediction for a given drug-protein pair\todo{should this
+  be about pairs, or rather something like ``given a protein, we aim
+  to find a drug...''? I.e., find a consistent way to separate
+  yourself language-wise from a naive prediction of pairs.}, we first
+build DL2vec representations\todo{In results: talk about what they are
+supposed to DO (``embeddings/representations of the phenotypes and
+functions...''), not how you implement them (you could use something
+else, for example).} for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
+Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.
+These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.
+Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs.
+The overall architecture can be seen in Figure~\ref{fig:Fullmodel}.
 

### 2021-01-30 paper/main.tex
minor updates

@@
 \subsection{Evaluation and metrics}
-As the number of drug-targets are sparsely given w.r.t.
-to the number of both drugs and proteins considered, the resulting training, validation and testing datasets are highly imbalanced.
-As there are only $~22.000$ links in the considered subset the ratio
+
+\todo[inline]{First part is not about evaluation but optimization
+  objectives, should go in the Model section:}
+As the number of drug-targets are sparsely given w.r.t.\todo{no
+  acronyms} to the number of both drugs and proteins considered, the
+resulting training, validation and testing datasets are highly
+imbalanced. As there are only $~22.000$\todo{exact number} links in
+the considered subset the ratio
 \begin{equation*}
@@
 \end{equation*}
+consequently needing compensation in the computed loss function and
+appropriate metrics for the evaluation.
 
-consequently needing compensation in the computed loss function and appropriate metrics for the evaluation.\\
-
-On that account, we weighted all positive drug-protein pair samples with this ratio by introducing the following loss function w.r.t. to binary cross-entropy:
+Therefore, we weighted all positive drug-protein pair samples with
+this ratio by introducing the following loss function w.r.t.\todo{no
+  acronym} to binary cross-entropy:
 \begin{equation}
@@
 \end{equation}
-for given prediction $x$ and target $y$.
-We average this loss among all drug-protein pairs in the training set, leading to a stable environment for the used optimizing scheme \textit{Adam} \citep{Adam2014}.
-We implemented a 5-fold cross validation among the proteins as justified in the results section.
-Furthermore, we  used early stopping to detect plateaus in the training process.
-\\
+for given prediction $x$ and target $y$. We average this loss among
+all drug-protein pairs in the training set, leading to a stable
+environment for the used optimizing scheme \textit{Adam}
+\citep{Adam2014}. We implemented a 5-fold cross validation among the
+proteins as justified in the results section. Furthermore, we used
+early stopping to detect plateaus in the training process.
 
-To assess each model, we computed the \textit{area under receiver operating characteristic}-score (AUROC), \textit{F1-score} (F1) and \textit{Matthews correlation coefficient} (MCC) for each considered hyperparameter configuration setting.\\
+\todo[inline]{This needs to be more elaborate: add the micro- vs
+  macro-averages here (formulas)}
+To assess each model, we computed the \textit{area under receiver
+  operating characteristic}-score (AUROC), \textit{F1-score} (F1) and
+\textit{Matthews correlation coefficient} (MCC) for each considered
+hyperparameter configuration setting.
 
-Avoiding validation overfitting, we eventually tested the best performing model on the Yamanishi dataset as presented and discussed.
-Furthermore, we compare our results on this very dataset with various other cutting edge approaches to DTI-prediction.\\
+\todo[inline]{Remove from here:}
+Avoiding validation overfitting, we eventually tested the best
+performing model on the Yamanishi dataset as presented and
+discussed. Furthermore, we compare our results on this very dataset
+with various other cutting edge approaches to DTI-prediction.
 
 \subsection{Model}
-In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each individually.
-Hence, we assemble rich molecular structure based features for drugs from \textit{SmilesTransformer}\citep{SmilesTransformer} and proteins from \textit{DeepGOPlus}\citep{DeepGoPlus}.
-\textit{SmilesTransformer} introduces an autoencoder, learning over the SMILES strings and thus the molecular organization of each drug in an unsupervised manner.
-On the other hand, we take advantage of the pretrained models of \textit{DeepGOPlus}, obtaining features from the proteins amino acid sequence, showing significant performance in the field of protein function prediction.
-Thus, both embeddings seem to suitably supplement the  following ontology based represenations.\\
 
-In the top-down section, we used \textit{DL2vec}\citep{DL2vec2020} to obtain ontology based representations.
-Hereby, DL2vec constructs a graph by introducing vertices and edges for each ontology class and axiom, respectively, followed by random walks starting from each entity.
-These walks are eventually learned on using a Word2vec \citep{Word2vec2013} model.
-Thus, we pick up rich, neighbourhood focused representations for each entity, which has shown great results for representing protein function and phenotypes.
-The overall structure of the ontology can be seen in Figure~\ref{fig:Onto}.\\
+\todo[inline]{Introduce the notion of top-down and bottom-up:}
+In order to build a method that incorporates both top-down and
+bottom-up features, we first created a model for each
+individually. Hence, we assemble rich molecular structure based
+features for drugs from
+\textit{SmilesTransformer} \citep{SmilesTransformer} and proteins from
+\textit{DeepGOPlus} \citep{DeepGoPlus}. \textit{SmilesTransformer}
+introduces an autoencoder, learning over the SMILES strings and thus
+the molecular organization of each drug in an unsupervised manner. On
+the other hand, we take advantage of the pretrained models of
+\textit{DeepGOPlus}, obtaining features from the proteins amino acid
+sequence, showing significant performance in the field of protein
+function prediction. Thus, both embeddings seem to suitably supplement
+the following ontology based representations.
+
+In the top-down section, we used \textit{DL2vec} \citep{DL2vec2020} to
+obtain ontology based representations. DL2vec constructs a graph by
+introducing vertices and edges for each ontology class and axiom,
+respectively, followed by random walks starting from each
+entity. These walks are eventually learned on using a Word2vec
+\citep{Word2vec2013} model. Thus, we pick up rich, neighbourhood
+focused representations for each entity, which has shown great results
+for representing protein function and phenotypes. The overall
+structure of the ontology can be seen in Figure~\ref{fig:Onto}.
 
@@
 
-As we wanted to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
-On the other hand we built a deep neural network for the molecular structure based features, also benefiting from the siamese network architecture.
-Therefore, the precomputed embeddings are run through a deep neural network transformation (DNNT) which also reduces the eventual representation size for drugs and proteins separately.
-An example structure for both types of features can be found in Figure~\ref{fig:SiameseNetwork}.\\
+As we wanted to learn from the similarity of drug side effects and
+protein phenotypes we opted for a deep siamese network approach, hence
+learning a high-dimensional embedding emphasizing this identity by
+forcing a maximal cosine similarity between these embeddings. On the
+other hand we built a deep neural network for the molecular structure
+based features, also benefiting from the siamese network
+architecture. Therefore, the precomputed embeddings are run through a
+deep neural network transformation (DNNT) which also reduces the
+eventual representation size for drugs and proteins separately. An
+example structure for both types of features can be found in
+Figure~\ref{fig:SiameseNetwork}.
 
@@
 	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
-	\caption{Siamese network applied to molecular and DL2vec features, utilizing deep neural network transformations (DNNT).
-The similarity function $\otimes$ yields the similarity between both transformed embeddings e.g.
-by computing the cosine similarity.}
+	\caption{Siamese network applied to molecular and DL2vec
+          features, utilizing deep neural network transformations
+          (DNNT). The similarity function $\otimes$ yields the
+          similarity between both transformed embeddings e.g. by
+          computing the cosine similarity.}
 	\label{fig:SiameseNetwork}
@@
 \subsubsection{Graph convolutional layers}
-These molecular and ontology based sub-models were added to a larger graph convolutional model, based on the protein-protein interaction (PPI) graph.
-The PPI dataset is represented by a graph $G=(V,E)$, where each protein is represented by a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$ symbolizes an interaction between two proteins.
-Additionally, we introduce mapping $x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node feature $x_v = x(v)$, where $d$ denotes the dimensionality of the node features.
-\\
+These molecular and ontology based sub-models were added to a larger
+graph convolutional model, based on the protein-protein interaction
+(PPI) graph. The PPI dataset is represented by a graph $G=(V,E)$,
+where each protein is represented by a vertex $v\in V$, and each edge
+$e\in E\subseteq V\times V$ symbolizes an interaction between two
+proteins. Additionally, we introduce mapping
+$x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node
+feature $x_v = x(v)$, where $d$ denotes the dimensionality of the node
+features.
 
-As described before, graph convolution has shown significant performance increase in a variety of tasks.
-While there are various methods out there we will only introduce the most basic one here.
-A graph convolutional layer w.r.t.
-\citet{GCNConv} hereby consists of a learnable weight matrix followed by an aggregation step, formalized by
-
+As described before, graph convolution has shown significant
+performance increase in a variety of tasks. While there are various
+methods out there we will only introduce the most basic one here. A
+graph convolutional layer \citet{GCNConv} consists of a
+learnable weight matrix followed by an aggregation step, formalized by
 \begin{equation}
@@
 \end{equation}
+where for a given graph $G=(V,E)$, $\hat{A} = A + I$ denotes the
+adjacency matrix with added self-loops for each vertex, $D$ is
+described by $\hat{D}_{ii} = \sum_{j=0} \hat{A}_{ij}$, a diagonal
+matrix displaying the degree of each node, and $\Theta$ denotes the
+learnable weight matrix. Added self-loops enforce that each node
+representation is directly dependent on its own preceding
+one. Notably, the number of graph convolutional layers stacked equals
+the radius of relevant nodes for each vertex within the graph.
 
-where for a given graph $G=(V,E)$, $\hat{A} = A + I$ denotes the adjacency matrix with added self-loops for each vertex, $D$ is described by $\hat{D}_{ii} = \sum_{j=0} \hat{A}_{ij}$, a diagonal matrix displaying the degree of each node, and $\Theta$ denotes the learnable weight matrix.
-Added self-loops enforce that each node representation is directly dependent on its own preceding one.
-Notably, the number of graph convolutional layers stacked equals the radius of relevant nodes for each vertex within the graph.
-\\
-
-The update rule for each individual node is denoted by
-
+The update rule for each individual node is
 \begin{equation}
@@
 \end{equation}
+where both $\hat{d}_i, \hat{d}_j$ are dependent on the edge weights
+$e_{ij}$ of the graph. With simple, single-valued edge weights such as
+$e_{ij}=1 \text{ }\forall (i,j)\in E$, all $\hat{d}_i$ reduce to
+$d_i$, i.e., the degree of each vertex $i$. We denote this type of
+graph convolutional neural layers with \textsc{GCNConv}.
 
-where both $\hat{d}_i, \hat{d}_j$ are dependent on the edge weights $e_{ij}$ of the graph.
-With simple, single valued edge weights such as $e_{ij}=1 \text{ }\forall (i,j)\in E$, all $\hat{d}_i$ reduce to $d_i$, i.e.
-the degree of each vertex $i$.
-We will denote this type of graph convolutional neural layers with \textsc{GCNConv}.
-\\
-
-While in this initial formulation the node-wise update step is defined by the sum over all neighbouring node representations, we are able to alter this formulation for a more sophisticated message passing scheme.
-As described we are able to rearrange the order of activation function $\sigma$, aggregation $\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this formulation as proposed by \citet{GENConv2020}:
-
+While in this initial formulation the node-wise update step is defined
+by the sum over all neighbouring node representations, we are able to
+alter this formulation to another message passing scheme.  We are able
+to rearrange the order of activation function $\sigma$, aggregation
+$\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this
+formulation as proposed by \citet{GENConv2020}:
 \begin{equation}
@@
 \end{equation}
-
-where we will generally only consider $\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$.
-We will denote this generalized layer type as \textsc{GENConv}, following the notation of \citet{PytorchGeometric}.
-While the reordering is merely import for numerical stability, the authors claim that this alteration of the original formulation eases the vanishing gradient problem for deeper convolutional networks.
-Additionally, the authors revise the aggregation function by generalizing it, allowing more sophisticated weighting functions such as learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming signals for each vertex, substituting the previously mentioned averaging step.
-Hence, while \textsc{GCNConv} suffers from both vanishing gradients and signal fading for large scale, highly connected graphs, each propagation step in \textsc{GENConv} emphasizes strong signals with values close to $0$ and $1$.
-Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
-Note additionally, that the resulting information hold no information on their own connectivity.\\
+where we will generally only consider
+$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We will denote
+this generalized layer type as \textsc{GENConv}, following the
+notation of \citet{PytorchGeometric}.  While the reordering is mainly
+import for numerical stability, this alteration also addresses the
+vanishing gradient problem for deeper convolutional networks
+\cite{GENConv2020}. Additionally, we can also generalize the
+aggregation function to allow different weighting functions such as
+learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming
+signals for each vertex, substituting the averaging step in
+\textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
+vanishing gradients and signal fading for large scale, highly
+connected graphs, each propagation step in \textsc{GENConv} emphasizes
+signals with values close to $0$ and $1$. The same convolutional
+filter and weight matrix are applied to and learned for all nodes
+simultaneously, and the resulting information\todo{Which information?
+  Specify} hold no information on their own connectivity.
 

### 2021-01-30 paper/main.tex
minor updates

@@
 
-We employ another mechanism to avoid redundancy and fading signals in stacked graph convolutional networks, which was introduced by \citet{DeepGCN2019} and refined in \citet{DeeperGCN2020}.
-The authors propose a residual network architecture and normalization scheme, solving these issues in a variety of graph predicting tasks.
-This structure is depicted in Figure~\ref{fig:ResGraphConvBlocks} resulting in reusable residual blocks, which can be stacked multiple times, thereby not losing focus of each node neighbourhood.
-\\
+We employ another mechanism to avoid redundancy and fading signals in
+stacked graph convolutional networks, which was introduced by
+\citet{DeepGCN2019} and refined in \citet{DeeperGCN2020}. The authors
+propose a residual network architecture and normalization scheme. This
+structure is depicted in Figure~\ref{fig:ResGraphConvBlocks} resulting
+in reusable residual blocks, which can be stacked multiple times,
+thereby not losing focus of each node neighbourhood.
 
 \subsubsection{Hyperparameter tuning}
-To find the best hyperparameter configuration for the proposed model we performed a grid search, to find the most expressive and non-redundant representation.
-Therefor, we pretrained the bottom-up and the top-down model separately and aimed at best performing models w.r.t.
-previously described metrics.
-Embedding sizes, depth of the neural network, optimizer, learning rate and layer types were found from an extensive, manual grid search.
+To find the best hyperparameter configuration for the proposed model
+we performed a grid search to find the most expressive and
+non-redundant representation. We pretrained the bottom-up and the
+top-down model separately and aimed at best performing models
+w.r.t. previously described metrics. We optimized embedding sizes,
+depth of the neural network, optimizer, learning rate and layer types
+using an extensive, manual grid search. \todo[inline]{Specify range of
+  values searched here; even better if you have the intermediate
+  results and can put them here (better: in the supplement).}
 

### 2021-01-30 paper/main.tex
minor updates

@@
 \subsection{Choosing a cross-validation splitting scheme}
-As in machine learning inference is derived from the underlying data, models and data are naturally and intrinsically linked.
-Thus, the more we understand about the pitfalls and biases within the data, the more we can try to bypass these difficulties.
-We will hereby abbreviate, e.g., a cross validation splitting scheme as \glqq split\grqq{} determining the train, validation and test subset of a given dataset.
-\\
-Especially within DTI prediction, one can observe elemental and basic skews \citep{Pahikkala2014}.
-(1.) Frequently, novel drugs are designed by altering non-functional components of a drug, leading to two and more very similar drugs, designed to target the same proteins (\textbf{CITATION?}).
-This bias can lead to \glqq hidden duplicates\grqq{} that can distribute among the train-test split, skewing the ultimate predictive performance.
-(2.) Additionally, within the realm of DTI prediction, some proteins, which we will denote as \textit{hub proteins}, have significantly more known interactions with drugs than others.
-In the considered subset of STITCH database, $5\%$ of the proteins had $40\%$ of the interactions, with similar biases for Yamanishi and Drugbank (\citet{Drugbank2007} and \citet{Drugbank2017}) datasets, making them easy and reliable predictors.
-(3.) The second distortion could be based on a publication bias for more \glqq valuable\grqq{} proteins due to their disease involvement, such as e.g., cancer, leading to more research and eventually more interacting drugs regarding those.
-Thus if we aim to find a drug for a given protein, the challenge is to develop a training and evaluation scheme, that does not simply overfit to those biases.\\
+
+% As in machine learning inference is derived from the underlying data,
+% models and data are naturally and intrinsically linked. Thus, the more
+% we understand about the pitfalls and biases within the data, the more
+% we can try to bypass these difficulties. We will hereby abbreviate,
+% e.g., a cross validation splitting scheme as \glqq split\grqq{}
+% determining the train, validation and test subset of a given dataset.
+
+Within DTI prediction, one can observe basic biases resulting from the
+underlying datasets \citep{Pahikkala2014}. First, novel drugs are
+often designed by altering non-functional components of a drug,
+leading to two and more very similar drugs designed to target the same
+proteins \cite{Overington2006}. This can lead to a bias when it leads
+to \glqq hidden duplicates\grqq{} that can distribute among the
+train--test split, resulting in a better (measured) predictive
+performance than would be expected when the model is applied to
+genuinely unknown interactions. Second, some proteins (which we will
+denote as \textit{hub proteins}) have significantly more known
+interactions with drugs than others. In the STITCH database, $5\%$
+of the proteins have $40\%$ of the interactions, and similar
+distributions are present in the Yamanishi and Drugbank datasets
+\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
+these proteins can often increase predictive performance while again
+not reflecting the actual performance when applied to a new protein
+(e.g., a protein for which no interactions are known). This bias in
+the number of drugs targeting certain proteins may be the result of
+study bias where more \glqq valuable\grqq{} proteins have more drugs
+designed to target them due to their involvement in more common
+diseases (or diseases for which drugs can be more profitably
+marketed).
+
+The aim of \name is to predict candidate drugs that target a given
+protein; the challenge is to develop a training and evaluation scheme
+that does not simply overfit to the inherent biases in training and
+testing data.
 
@@
 Our aim is further to combine two different types of features:
-structural information based on protein structure and molecular drug composition, and their respective observable characteristics. \\
+structural information based on protein structure and molecular drug composition, and their respective observable characteristics.
 

### 2021-01-30 paper/main.tex
minor updates

@@
 testing data.
+In general, when performing cross-validation for DTI prediction, the
+options are to split over drug--target pairs, split over drugs, and
+split over proteins.
 
-In general, when concerning cross-validation splitting schemes for DTI prediction, there are the following three options:
-\begin{enumerate}
-	\item Build split over drugs
-	\item Build split over drug-target pairs
-	\item Build split over proteins
-\end{enumerate}
 where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.

### 2021-01-30 paper/main.tex
minor updates

--- 2021-01-30 paper/main.tex
@@
 
+\todo[inline]{Likely discussion:}
 The aim of \name is to predict candidate drugs that target a given
@@
 testing data.
+In general, when performing cross-validation for DTI prediction, the
+options are to split over drug--target pairs, split over drugs, and
+split over proteins; the last two options ensure that at least parts
+of the interactions are not seen during training and evaluate either
+how well targets are predicted for unseen drugs or unseen proteins.
 
-In general, when concerning cross-validation splitting schemes for DTI prediction, there are the following three options:
-\begin{enumerate}
-	\item Build split over drugs
-	\item Build split over drug-target pairs
-	\item Build split over proteins
-\end{enumerate}
-where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
-Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
-\\
-The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.
-Thus, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
-Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions, while in the third scheme the number of interactions of the test proteins is not known.
-\\
-
-Considering the publications of the last
+\todo[inline]{Likely discussion:}
+The most common scheme for DTI prediction is the split over
+drug--target pairs \citep{Survey2018}, where likely all drugs and
+targets of the validation and testing phase have already occurred in
+the training phase, as part of other drug-target pairs. Thus, the
+first and second splitting scheme are exposed to the first dataset
+bias and are hence more likely vulnerable to transductive inference by
+just predicting recently seen structures, rather than implementing
+inductive inference and generalizing over the drug
+representations. Second, these two strategies are more susceptible to
+the second and third bias, as only in these cases the model may
+overfit on the number of existing interactions, while in the third
+scheme the number of interactions of the test proteins is not known.
 

### 2021-01-30 paper/main.tex
Update protein split part

--- 2021-01-30 paper/main.tex
@@
 \subsection{Evaluation and metrics}
+As the number of drug-targets are sparsely given w.r.t.
+to the number of both drugs and proteins considered, the resulting training, validation and testing datasets are highly imbalanced.
+As there are only $~22.000$ links in the considered subset the ratio
+\begin{equation*}
+	\frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
+\end{equation*}
 
-\todo[inline]{First part is not about evaluation but optimization
-  objectives, should go in the Model section:}
-As the number of drug-targets are sparsely given w.r.t.\todo{no
-  acronyms} to the number of both drugs and proteins considered, the
-resulting training, validation and testing datasets are highly
-imbalanced. As there are only $~22.000$\todo{exact number} links in
-the considered subset the ratio
-\begin{equation*}
-	\frac{\#dti\_links}{\#drugs \cdot \#proteins} \approx 360,
-\end{equation*}
-consequently needing compensation in the computed loss function and
-appropriate metrics for the evaluation.
+consequently needing compensation in the computed loss function and appropriate metrics for the evaluation.\\
 
-Therefore, we weighted all positive drug-protein pair samples with
-this ratio by introducing the following loss function w.r.t.\todo{no
-  acronym} to binary cross-entropy:
+On that account, we weighted all positive drug-protein pair samples with this ratio by introducing the following loss function w.r.t. to binary cross-entropy:
 \begin{equation}
@@
 \end{equation}
-for given prediction $x$ and target $y$. We average this loss among
-all drug-protein pairs in the training set, leading to a stable
-environment for the used optimizing scheme \textit{Adam}
-\citep{Adam2014}. We implemented a 5-fold cross validation among the
-proteins as justified in the results section. Furthermore, we used
-early stopping to detect plateaus in the training process.
+for given prediction $x$ and target $y$.
+We average this loss among all drug-protein pairs in the training set, leading to a stable environment for the used optimizing scheme \textit{Adam} \citep{Adam2014}.
+We implemented a 5-fold cross validation among the proteins as justified in the results section.
+Furthermore, we  used early stopping to detect plateaus in the training process.
+\\
 
-\todo[inline]{This needs to be more elaborate: add the micro- vs
-  macro-averages here (formulas)}
-To assess each model, we computed the \textit{area under receiver
-  operating characteristic}-score (AUROC), \textit{F1-score} (F1) and
-\textit{Matthews correlation coefficient} (MCC) for each considered
-hyperparameter configuration setting.
+To assess each model, we computed the \textit{area under receiver operating characteristic}-score (AUROC), \textit{F1-score} (F1) and \textit{Matthews correlation coefficient} (MCC) for each considered hyperparameter configuration setting.\\
 
-\todo[inline]{Remove from here:}
-Avoiding validation overfitting, we eventually tested the best
-performing model on the Yamanishi dataset as presented and
-discussed. Furthermore, we compare our results on this very dataset
-with various other cutting edge approaches to DTI-prediction.
+Avoiding validation overfitting, we eventually tested the best performing model on the Yamanishi dataset as presented and discussed.
+Furthermore, we compare our results on this very dataset with various other cutting edge approaches to DTI-prediction.\\
 
 \subsection{Model}
+In order to build a method that incorporates both top-down and bottom-up features, we first created a model for each individually.
+Hence, we assemble rich molecular structure based features for drugs from \textit{SmilesTransformer}\citep{SmilesTransformer} and proteins from \textit{DeepGOPlus}\citep{DeepGoPlus}.
+\textit{SmilesTransformer} introduces an autoencoder, learning over the SMILES strings and thus the molecular organization of each drug in an unsupervised manner.
+On the other hand, we take advantage of the pretrained models of \textit{DeepGOPlus}, obtaining features from the proteins amino acid sequence, showing significant performance in the field of protein function prediction.
+Thus, both embeddings seem to suitably supplement the  following ontology based represenations.\\
 
-\todo[inline]{Introduce the notion of top-down and bottom-up:}
-In order to build a method that incorporates both top-down and
-bottom-up features, we first created a model for each
-individually. Hence, we assemble rich molecular structure based
-features for drugs from
-\textit{SmilesTransformer} \citep{SmilesTransformer} and proteins from
-\textit{DeepGOPlus} \citep{DeepGoPlus}. \textit{SmilesTransformer}
-introduces an autoencoder, learning over the SMILES strings and thus
-the molecular organization of each drug in an unsupervised manner. On
-the other hand, we take advantage of the pretrained models of
-\textit{DeepGOPlus}, obtaining features from the proteins amino acid
-sequence, showing significant performance in the field of protein
-function prediction. Thus, both embeddings seem to suitably supplement
-the following ontology based representations.
-
-In the top-down section, we used \textit{DL2vec} \citep{DL2vec2020} to
-obtain ontology based representations. DL2vec constructs a graph by
-introducing vertices and edges for each ontology class and axiom,
-respectively, followed by random walks starting from each
-entity. These walks are eventually learned on using a Word2vec
-\citep{Word2vec2013} model. Thus, we pick up rich, neighbourhood
-focused representations for each entity, which has shown great results
-for representing protein function and phenotypes. The overall
-structure of the ontology can be seen in Figure~\ref{fig:Onto}.
+In the top-down section, we used \textit{DL2vec}\citep{DL2vec2020} to obtain ontology based representations.
+Hereby, DL2vec constructs a graph by introducing vertices and edges for each ontology class and axiom, respectively, followed by random walks starting from each entity.
+These walks are eventually learned on using a Word2vec \citep{Word2vec2013} model.
+Thus, we pick up rich, neighbourhood focused representations for each entity, which has shown great results for representing protein function and phenotypes.
+The overall structure of the ontology can be seen in Figure~\ref{fig:Onto}.\\
 
@@
 
-As we wanted to learn from the similarity of drug side effects and
-protein phenotypes we opted for a deep siamese network approach, hence
-learning a high-dimensional embedding emphasizing this identity by
-forcing a maximal cosine similarity between these embeddings. On the
-other hand we built a deep neural network for the molecular structure
-based features, also benefiting from the siamese network
-architecture. Therefore, the precomputed embeddings are run through a
-deep neural network transformation (DNNT) which also reduces the
-eventual representation size for drugs and proteins separately. An
-example structure for both types of features can be found in
-Figure~\ref{fig:SiameseNetwork}.
+As we wanted to learn from the similarity of drug side effects and protein phenotypes we opted for a deep siamese network approach, hence learning a high-dimensional embedding emphasizing this identity by forcing a maximal cosine similarity between these embeddings.
+On the other hand we built a deep neural network for the molecular structure based features, also benefiting from the siamese network architecture.
+Therefore, the precomputed embeddings are run through a deep neural network transformation (DNNT) which also reduces the eventual representation size for drugs and proteins separately.
+An example structure for both types of features can be found in Figure~\ref{fig:SiameseNetwork}.\\
 
@@
 	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
-	\caption{Siamese network applied to molecular and DL2vec
-          features, utilizing deep neural network transformations
-          (DNNT). The similarity function $\otimes$ yields the
-          similarity between both transformed embeddings e.g. by
-          computing the cosine similarity.}
+	\caption{Siamese network applied to molecular and DL2vec features, utilizing deep neural network transformations (DNNT).
+The similarity function $\otimes$ yields the similarity between both transformed embeddings e.g.
+by computing the cosine similarity.}
 	\label{fig:SiameseNetwork}
@@
 \subsubsection{Graph convolutional layers}
-These molecular and ontology based sub-models were added to a larger
-graph convolutional model, based on the protein-protein interaction
-(PPI) graph. The PPI dataset is represented by a graph $G=(V,E)$,
-where each protein is represented by a vertex $v\in V$, and each edge
-$e\in E\subseteq V\times V$ symbolizes an interaction between two
-proteins. Additionally, we introduce mapping
-$x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node
-feature $x_v = x(v)$, where $d$ denotes the dimensionality of the node
-features.
+These molecular and ontology based sub-models were added to a larger graph convolutional model, based on the protein-protein interaction (PPI) graph.
+The PPI dataset is represented by a graph $G=(V,E)$, where each protein is represented by a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$ symbolizes an interaction between two proteins.
+Additionally, we introduce mapping $x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node feature $x_v = x(v)$, where $d$ denotes the dimensionality of the node features.
+\\
 
-As described before, graph convolution has shown significant
-performance increase in a variety of tasks. While there are various
-methods out there we will only introduce the most basic one here. A
-graph convolutional layer \citet{GCNConv} consists of a
-learnable weight matrix followed by an aggregation step, formalized by
+As described before, graph convolution has shown significant performance increase in a variety of tasks.
+While there are various methods out there we will only introduce the most basic one here.
+A graph convolutional layer w.r.t.
+\citet{GCNConv} hereby consists of a learnable weight matrix followed by an aggregation step, formalized by
+
 \begin{equation}
@@
 \end{equation}
-where for a given graph $G=(V,E)$, $\hat{A} = A + I$ denotes the
-adjacency matrix with added self-loops for each vertex, $D$ is
-described by $\hat{D}_{ii} = \sum_{j=0} \hat{A}_{ij}$, a diagonal
-matrix displaying the degree of each node, and $\Theta$ denotes the
-learnable weight matrix. Added self-loops enforce that each node
-representation is directly dependent on its own preceding
-one. Notably, the number of graph convolutional layers stacked equals
-the radius of relevant nodes for each vertex within the graph.
 
-The update rule for each individual node is
+where for a given graph $G=(V,E)$, $\hat{A} = A + I$ denotes the adjacency matrix with added self-loops for each vertex, $D$ is described by $\hat{D}_{ii} = \sum_{j=0} \hat{A}_{ij}$, a diagonal matrix displaying the degree of each node, and $\Theta$ denotes the learnable weight matrix.
+Added self-loops enforce that each node representation is directly dependent on its own preceding one.
+Notably, the number of graph convolutional layers stacked equals the radius of relevant nodes for each vertex within the graph.
+\\
+
+The update rule for each individual node is denoted by
+
 \begin{equation}
@@
 \end{equation}
-where both $\hat{d}_i, \hat{d}_j$ are dependent on the edge weights
-$e_{ij}$ of the graph. With simple, single-valued edge weights such as
-$e_{ij}=1 \text{ }\forall (i,j)\in E$, all $\hat{d}_i$ reduce to
-$d_i$, i.e., the degree of each vertex $i$. We denote this type of
-graph convolutional neural layers with \textsc{GCNConv}.
 
-While in this initial formulation the node-wise update step is defined
-by the sum over all neighbouring node representations, we are able to
-alter this formulation to another message passing scheme.  We are able
-to rearrange the order of activation function $\sigma$, aggregation
-$\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this
-formulation as proposed by \citet{GENConv2020}:
+where both $\hat{d}_i, \hat{d}_j$ are dependent on the edge weights $e_{ij}$ of the graph.
+With simple, single valued edge weights such as $e_{ij}=1 \text{ }\forall (i,j)\in E$, all $\hat{d}_i$ reduce to $d_i$, i.e.
+the degree of each vertex $i$.
+We will denote this type of graph convolutional neural layers with \textsc{GCNConv}.
+\\
+
+While in this initial formulation the node-wise update step is defined by the sum over all neighbouring node representations, we are able to alter this formulation for a more sophisticated message passing scheme.
+As described we are able to rearrange the order of activation function $\sigma$, aggregation $\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this formulation as proposed by \citet{GENConv2020}:
+
 \begin{equation}
@@
 \end{equation}
-where we will generally only consider
-$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We will denote
-this generalized layer type as \textsc{GENConv}, following the
-notation of \citet{PytorchGeometric}.  While the reordering is mainly
-import for numerical stability, this alteration also addresses the
-vanishing gradient problem for deeper convolutional networks
-\cite{GENConv2020}. Additionally, we can also generalize the
-aggregation function to allow different weighting functions such as
-learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming
-signals for each vertex, substituting the averaging step in
-\textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
-vanishing gradients and signal fading for large scale, highly
-connected graphs, each propagation step in \textsc{GENConv} emphasizes
-signals with values close to $0$ and $1$. The same convolutional
-filter and weight matrix are applied to and learned for all nodes
-simultaneously, and the resulting information\todo{Which information?
-  Specify} hold no information on their own connectivity.
+
+where we will generally only consider $\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$.
+We will denote this generalized layer type as \textsc{GENConv}, following the notation of \citet{PytorchGeometric}.
+While the reordering is merely import for numerical stability, the authors claim that this alteration of the original formulation eases the vanishing gradient problem for deeper convolutional networks.
+Additionally, the authors revise the aggregation function by generalizing it, allowing more sophisticated weighting functions such as learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming signals for each vertex, substituting the previously mentioned averaging step.
+Hence, while \textsc{GCNConv} suffers from both vanishing gradients and signal fading for large scale, highly connected graphs, each propagation step in \textsc{GENConv} emphasizes strong signals with values close to $0$ and $1$.
+Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
+Note additionally, that the resulting information hold no information on their own connectivity.\\
 
@@
 
-We employ another mechanism to avoid redundancy and fading signals in
-stacked graph convolutional networks, which was introduced by
-\citet{DeepGCN2019} and refined in \citet{DeeperGCN2020}. The authors
-propose a residual network architecture and normalization scheme. This
-structure is depicted in Figure~\ref{fig:ResGraphConvBlocks} resulting
-in reusable residual blocks, which can be stacked multiple times,
-thereby not losing focus of each node neighbourhood.
+We employ another mechanism to avoid redundancy and fading signals in stacked graph convolutional networks, which was introduced by \citet{DeepGCN2019} and refined in \citet{DeeperGCN2020}.
+The authors propose a residual network architecture and normalization scheme, solving these issues in a variety of graph predicting tasks.
+This structure is depicted in Figure~\ref{fig:ResGraphConvBlocks} resulting in reusable residual blocks, which can be stacked multiple times, thereby not losing focus of each node neighbourhood.
+\\
 
 \subsubsection{Hyperparameter tuning}
-To find the best hyperparameter configuration for the proposed model
-we performed a grid search to find the most expressive and
-non-redundant representation. We pretrained the bottom-up and the
-top-down model separately and aimed at best performing models
-w.r.t. previously described metrics. We optimized embedding sizes,
-depth of the neural network, optimizer, learning rate and layer types
-using an extensive, manual grid search. \todo[inline]{Specify range of
-  values searched here; even better if you have the intermediate
-  results and can put them here (better: in the supplement).}
+To find the best hyperparameter configuration for the proposed model we performed a grid search, to find the most expressive and non-redundant representation.
+Therefor, we pretrained the bottom-up and the top-down model separately and aimed at best performing models w.r.t.
+previously described metrics.
+Embedding sizes, depth of the neural network, optimizer, learning rate and layer types were found from an extensive, manual grid search.
 
@@
 \subsection{Choosing a cross-validation splitting scheme}
-
-% As in machine learning inference is derived from the underlying data,
-% models and data are naturally and intrinsically linked. Thus, the more
-% we understand about the pitfalls and biases within the data, the more
-% we can try to bypass these difficulties. We will hereby abbreviate,
-% e.g., a cross validation splitting scheme as \glqq split\grqq{}
-% determining the train, validation and test subset of a given dataset.
-
-Within DTI prediction, one can observe basic biases resulting from the
-underlying datasets \citep{Pahikkala2014}. First, novel drugs are
-often designed by altering non-functional components of a drug,
-leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can lead to a bias when it leads
-to \glqq hidden duplicates\grqq{} that can distribute among the
-train--test split, resulting in a better (measured) predictive
-performance than would be expected when the model is applied to
-genuinely unknown interactions. Second, some proteins (which we will
-denote as \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$
-of the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank datasets
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins can often increase predictive performance while again
-not reflecting the actual performance when applied to a new protein
-(e.g., a protein for which no interactions are known). This bias in
-the number of drugs targeting certain proteins may be the result of
-study bias where more \glqq valuable\grqq{} proteins have more drugs
-designed to target them due to their involvement in more common
-diseases (or diseases for which drugs can be more profitably
-marketed).
-
-The aim of \name is to predict candidate drugs that target a given
-protein; the challenge is to develop a training and evaluation scheme
-that does not simply overfit to the inherent biases in training and
-testing data.
+As in machine learning inference is derived from the underlying data, models and data are naturally and intrinsically linked.
+Thus, the more we understand about the pitfalls and biases within the data, the more we can try to bypass these difficulties.
+We will hereby abbreviate, e.g., a cross validation splitting scheme as \glqq split\grqq{} determining the train, validation and test subset of a given dataset.
+\\
+Especially within DTI prediction, one can observe elemental and basic skews \citep{Pahikkala2014}.
+(1.) Frequently, novel drugs are designed by altering non-functional components of a drug, leading to two and more very similar drugs, designed to target the same proteins (\textbf{CITATION?}).
+This bias can lead to \glqq hidden duplicates\grqq{} that can distribute among the train-test split, skewing the ultimate predictive performance.
+(2.) Additionally, within the realm of DTI prediction, some proteins, which we will denote as \textit{hub proteins}, have significantly more known interactions with drugs than others.
+In the considered subset of STITCH database, $5\%$ of the proteins had $40\%$ of the interactions, with similar biases for Yamanishi and Drugbank (\citet{Drugbank2007} and \citet{Drugbank2017}) datasets, making them easy and reliable predictors.
+(3.) The second distortion could be based on a publication bias for more \glqq valuable\grqq{} proteins due to their disease involvement, such as e.g., cancer, leading to more research and eventually more interacting drugs regarding those.
+Thus if we aim to find a drug for a given protein, the challenge is to develop a training and evaluation scheme, that does not simply overfit to those biases.\\
 
@@
 The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.
-Thus, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
-Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions, while in the third scheme the number of interactions of the test proteins is not known.
+The second most prevalent arrangement is the split over drugs, while only close to none is aiming on a protein split.
+However, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
+Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions for a single protein, while in the third scheme the number of interactions of the test proteins is not known.
+
+Assuming a hypothetical, perfectly generalizing model built upon an unbiased dataset, this very model would yield similar performances for all three splitting schemes, not overfitting on the known structures.
+On the other hand, for a hypothetical entirely overfitting model, trained on a highly biased dataset, this model would show substantial deviations from the original performance over another split.
 \\
+We emphasize, that all real-world models are prone to some sort of overfitting, and unknown, deviant entities in both validation and testing set will likely lead to some sort of performance gap for relevant metrics.
+However, a large disparity may hint the biases stated above.
 
-Considering the publications of the last
+Hence, we will perform a cross-validation over proteins for the training and prediction phase of our model, despite predicting per protein is rather counter-intuitive as there only limited drug-targets \citep{Overington2006}, and thus novel drugs are more likely to arise than novel targets.
+Yet, we aim to find all interacting drugs for existing targets motivating our split choice even further.
 
-In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
-Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
-However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
+\subsection{MicroAUC}
+Continuing this line, we propose the well known metric \textit{MicroAUC} for a better evaluation for this purpose.
+For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
+With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
+\begin{equation*}
+	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
+\end{equation*}
 
-\todo[inline]{Relate these to the issues listed above:}
-\begin{enumerate}
-	\item Build split over drugs
-	\item Build split over drug-target pairs
-	\item Build split over proteins
-\end{enumerate} 
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results.
-This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}.
-The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
-Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive.
-As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes.
-\\
-
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
-However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier.
-Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE).
-When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
-Hence, when entirely new molecules are seen, the models perform much worse
-
-The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
-
-As mentioned in the introduction it is quite difficult to learn suitable features from proteins.
-In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features.
-\\
-Thus, building splitting over proteins is the most challenging of the three options. \\
+Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
+As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
+Furthermore,
 
@@
 Our aim is further to combine two different types of features:
-structural information based on protein structure and molecular drug composition, and their respective observable characteristics.
+structural information based on protein structure and molecular drug composition, and their respective observable characteristics. \\
 

### 2021-01-30 paper/main.tex

--- 2021-01-30 paper/main.tex
@@
 \begin{equation*}
-	\frac{\#dti\_links}{\#drugs \cdot \#proteins} \approx 360,
+	\frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
 \end{equation*}
@@
 \subsection{Choosing a cross-validation splitting scheme}
+As in machine learning inference is derived from the underlying data, models and data are naturally and intrinsically linked.
+Thus, the more we understand about the pitfalls and biases within the data, the more we can try to bypass these difficulties.
+We will hereby abbreviate, e.g., a cross validation splitting scheme as \glqq split\grqq{} determining the train, validation and test subset of a given dataset.
+\\
+Especially within DTI prediction, one can observe elemental and basic skews \citep{Pahikkala2014}.
+(1.) Frequently, novel drugs are designed by altering non-functional components of a drug, leading to two and more very similar drugs, designed to target the same proteins (\textbf{CITATION?}).
+This bias can lead to \glqq hidden duplicates\grqq{} that can distribute among the train-test split, skewing the ultimate predictive performance.
+(2.) Additionally, within the realm of DTI prediction, some proteins, which we will denote as \textit{hub proteins}, have significantly more known interactions with drugs than others.
+In the considered subset of STITCH database, $5\%$ of the proteins had $40\%$ of the interactions, with similar biases for Yamanishi and Drugbank (\citet{Drugbank2007} and \citet{Drugbank2017}) datasets, making them easy and reliable predictors.
+(3.) The second distortion could be based on a publication bias for more \glqq valuable\grqq{} proteins due to their disease involvement, such as e.g., cancer, leading to more research and eventually more interacting drugs regarding those.
+Thus if we aim to find a drug for a given protein, the challenge is to develop a training and evaluation scheme, that does not simply overfit to those biases.\\
 
-% As in machine learning inference is derived from the underlying data,
-% models and data are naturally and intrinsically linked. Thus, the more
-% we understand about the pitfalls and biases within the data, the more
-% we can try to bypass these difficulties. We will hereby abbreviate,
-% e.g., a cross validation splitting scheme as \glqq split\grqq{}
-% determining the train, validation and test subset of a given dataset.
-
-Within DTI prediction, one can observe basic biases resulting from the
-underlying datasets \citep{Pahikkala2014}. First, novel drugs are
-often designed by altering non-functional components of a drug,
-leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can lead to a bias when it leads
-to \glqq hidden duplicates\grqq{} that can distribute among the
-train--test split, resulting in a better (measured) predictive
-performance than would be expected when the model is applied to
-genuinely unknown interactions. Second, some proteins (which we will
-denote as \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$
-of the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank datasets
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins can often increase predictive performance while again
-not reflecting the actual performance when applied to a new protein
-(e.g., a protein for which no interactions are known). This bias in
-the number of drugs targeting certain proteins may be the result of
-study bias where more \glqq valuable\grqq{} proteins have more drugs
-designed to target them due to their involvement in more common
-diseases (or diseases for which drugs can be more profitably
-marketed).
-
-The aim of \name is to predict candidate drugs that target a given
-protein; the challenge is to develop a training and evaluation scheme
-that does not simply overfit to the inherent biases in training and
-testing data.
-In general, when performing cross-validation for DTI prediction, the
-options are to split over drug--target pairs, split over drugs, and
-split over proteins.
-
-where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
-Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
-\\
-The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.
-Thus, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
-Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions, while in the third scheme the number of interactions of the test proteins is not known.
-\\
-
-Considering the publications of the last
-
-In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
-Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
-However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:
-
-\todo[inline]{Relate these to the issues listed above:}
+In general, when concerning cross-validation splitting schemes for DTI prediction, there are the following three options:
 \begin{enumerate}
@@
 \end{enumerate}
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results.
-This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}.
-The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias.
-Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive.
-As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes.
+where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
+Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
 \\
+The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.
+The second most prevalent arrangement is the split over drugs, while only close to none is aiming on a protein split.
+However, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
+Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions for a single protein, while in the third scheme the number of interactions of the test proteins is not known.
 
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
-The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
-However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier.
-Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE).
-When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures.
-Hence, when entirely new molecules are seen, the models perform much worse
+Assuming a hypothetical, perfectly generalizing model built upon an unbiased dataset, this very model would yield similar performances for all three splitting schemes, not overfitting on the known structures.
+On the other hand, for a hypothetical entirely overfitting model, trained on a highly biased dataset, this model would show substantial deviations from the original performance over another split.
+\\
+We emphasize, that all real-world models are prone to some sort of overfitting, and unknown, deviant entities in both validation and testing set will likely lead to some sort of performance gap for relevant metrics.
+However, a large disparity may hint the biases stated above.
 
-The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
+Hence, we will perform a cross-validation over proteins for the training and prediction phase of our model, despite predicting per protein is rather counter-intuitive as there only limited drug-targets \citep{Overington2006}, and thus novel drugs are more likely to arise than novel targets.
+Yet, we aim to find all interacting drugs for existing targets motivating our split choice even further.
 
-As mentioned in the introduction it is quite difficult to learn suitable features from proteins.
-In general, attempts search for motifs in the protein sequences under usage of convolutional neural networks and filters, which is more suitable for tasks like protein function prediction, than for for drug-target interaction prediction, and lack a more in-depth hypothesis on the protein side, while investing in refined drug features.
-\\
-Thus, building splitting over proteins is the most challenging of the three options. \\
+\subsection{MicroAUC}
+Continuing this line, we propose the well known metric \textit{MicroAUC} for a better evaluation for this purpose.
+For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
+With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
+\begin{equation*}
+	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
+\end{equation*}
+
+Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
+As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
+Furthermore,
 

### 2021-01-30 paper/main.tex
minor updates

--- 2021-01-30 paper/main.tex
@@
 
-\subsection{Choosing a cross-validation splitting scheme}
+\subsection{\name: computational model to identify drugs that target a
+  protein}
 
@@
 
+We developed \name as a computational model to predict drug--target
+interactions. Specifically, given a protein, \name will identify and
+rank drugs that likely target this protein. \name combines two 
+types of features: structural information for drugs and proteins that
+can be used to determine if they may physically interact, and
+information about drug effects that may localize on an interaction
+network.
+
 \todo[inline]{Likely discussion:}
@@
 
-In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
+% In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism.
 Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
@@
 
-\todo[inline]{Relate these to the issues listed above:}
-\begin{enumerate}
-	\item Build split over drugs
-	\item Build split over drug-target pairs
-	\item Build split over proteins
-\end{enumerate} 
+% \todo[inline]{Relate these to the issues listed above:}
+% \begin{enumerate}
+% 	\item Build split over drugs
+% 	\item Build split over drug-target pairs
+% 	\item Build split over proteins
+% \end{enumerate} 
 In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
@@
 
-In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION).
+In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}% , CITATION).
 The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein.
@@
 
-The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
+% The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.
 
@@
 Thus, building splitting over proteins is the most challenging of the three options. \\
-
-\subsection{Baseline model}
-Our aim is to develop and evaluate a computational model to predict
-drug--target interactions. Specifically, given a protein, we are
-interested in identifying the drugs that may target the protein; this
-scenario is motivated by the need to identify drugs that can target a
-
-\subsection{\name identifies drugs that target a protein}
-
-We developed \name as a computational model to predict drug--target
-interactions. Specifically, given a protein, \name will identify and
-rank drugs that likely target this protein. \name combines two 
-types of features: structural information for drugs and proteins that
-can be used to determine if they may physically interact, and
-
-This scenario is motivated by the need to identify drugs that can target a
-specific protein that may be associated with a disease or abnormal
-phenotypes.
-Our aim is further to combine two different types of features:
-structural information based on protein structure and molecular drug composition, and their respective observable characteristics.
 
@@
 By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
+
+\subsection{Baseline model}
+Our aim is to develop and evaluate a computational model to predict
+drug--target interactions. Specifically, given a protein, we are
+interested in identifying the drugs that may target the protein; this
+scenario is motivated by the need to identify drugs that can target a
 

### 2021-01-30 paper/main.tex
updated microAUC part

--- 2021-01-30 paper/main.tex
@@
 
+1
 \subsubsection{Graph convolutional layers}
@@
 Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
+Note further that both are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are only plausible for protein split and drug split, respectively.
 As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
-Furthermore,
 
-\subsection{Baseline model}
 Our aim is to develop and evaluate a computational model to predict
@@
 
+\subsection{Baseline model}
 Before introducing the core of our eventual model, we propose a suitable naive baseline model in order to analyze and understand the existing approaches and this novel method for drug target interaction prediction.

### 2021-01-30 paper/main.tex
minor updates

--- 2021-01-30 paper/main.tex
@@
 interactions. Specifically, given a protein, \name will identify and
-rank drugs that likely target this protein. \name combines two 
+rank known drugs that likely target this protein. \name combines two 
 types of features: structural information for drugs and proteins that
@@
 network.
+As structural features, \name uses structural representations of drugs
+from the SMILES transformer \cite{} and representations of protein
+amino acid sequences from DeepGOPlus \cite{}.
+\name learns representations of drug effects and protein functions
+using the ontology-based machine learning method DL2Vec \cite{}.
+
+We use two Siamese neural networks to combine the molecular
+and phenotype representations of drugs and proteins,
+% and
+% ontology features, yielding the deep neural network transformers
+% (DNNT) for both methods, and both drugs and proteins,
+% individually. 
+and add these to a protein interaction network as node features.
+% neural network in which nodes representing interactions
+% between proteins as node features, 
+We then train this model in an end-to-end manner. % Notably, like in the
+% pretraining phase, we rely on siamese neural networks, enforcing
+% similar representations for fitting drug-protein pairs. 
+The overall architecture is depicted in Figure~\ref{fig:Fullmodel}.
+
+% For a given protein, we want to find all suitable interacting drugs, \textbf{among a set of known drugs}.
+%To obtain a prediction for a given drug-protein pair,
+
+Note that this approach can be easily generalized, profiting from
+other protein function and phenotype representation methods.
+
+\begin{figure}[!tpb]
+	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
+	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
+The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
+}
+	\label{fig:Fullmodel}
+\end{figure}
+
+Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model.
+Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
+By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
 
@@
 \subsection{High level description}
-
-For a given protein, we want to find all suitable interacting drugs, \textbf{among a set of known drugs}.
-To obtain a prediction for a given drug-protein pair, we first
-build representations embodying both phenotypes and functions for both drugs and proteins, while also preparing structural, molecular representations, e.g., utilizing DL2vec over PhenomeNET and the prebuilt features from SMILES transformer and DeepGOPlus.
-Note that this approach can be easily generalized, profiting from other protein function and phenotype representation methods.
-\\
-Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers (DNNT) for both methods, and both drugs and proteins, individually.
-These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.
-Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs.
-The overall architecture is depicted in Figure~\ref{fig:Fullmodel}.\\
-
-\begin{figure}[!tpb]
-	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
-	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
-The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
-}
-	\label{fig:Fullmodel}
-\end{figure}
-
-Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model.
-Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
-By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
 

### 2021-01-30 paper/main.tex

--- 2021-01-30 paper/main.tex
@@
 \subsection{Choosing a cross-validation splitting scheme}
-As in machine learning inference is derived from the underlying data, models and data are naturally and intrinsically linked.
-Thus, the more we understand about the pitfalls and biases within the data, the more we can try to bypass these difficulties.
-We will hereby abbreviate, e.g., a cross validation splitting scheme as \glqq split\grqq{} determining the train, validation and test subset of a given dataset.
-\\
-Especially within DTI prediction, one can observe elemental and basic skews \citep{Pahikkala2014}.
-(1.) Frequently, novel drugs are designed by altering non-functional components of a drug, leading to two and more very similar drugs, designed to target the same proteins (\textbf{CITATION?}).
-This bias can lead to \glqq hidden duplicates\grqq{} that can distribute among the train-test split, skewing the ultimate predictive performance.
-(2.) Additionally, within the realm of DTI prediction, some proteins, which we will denote as \textit{hub proteins}, have significantly more known interactions with drugs than others.
-In the considered subset of STITCH database, $5\%$ of the proteins had $40\%$ of the interactions, with similar biases for Yamanishi and Drugbank (\citet{Drugbank2007} and \citet{Drugbank2017}) datasets, making them easy and reliable predictors.
-(3.) The second distortion could be based on a publication bias for more \glqq valuable\grqq{} proteins due to their disease involvement, such as e.g., cancer, leading to more research and eventually more interacting drugs regarding those.
-Thus if we aim to find a drug for a given protein, the challenge is to develop a training and evaluation scheme, that does not simply overfit to those biases.\\
 
-In general, when concerning cross-validation splitting schemes for DTI prediction, there are the following three options:
+% As in machine learning inference is derived from the underlying data,
+% models and data are naturally and intrinsically linked. Thus, the more
+% we understand about the pitfalls and biases within the data, the more
+% we can try to bypass these difficulties. We will hereby abbreviate,
+% e.g., a cross validation splitting scheme as \glqq split\grqq{}
+% determining the train, validation and test subset of a given dataset.
+
+Within DTI prediction, one can observe basic biases resulting from the
+underlying datasets \citep{Pahikkala2014}. First, novel drugs are
+often designed by altering non-functional components of a drug,
+leading to two and more very similar drugs designed to target the same
+proteins \cite{Overington2006}. This can lead to a bias when it leads
+to \glqq hidden duplicates\grqq{} that can distribute among the
+train--test split, resulting in a better (measured) predictive
+performance than would be expected when the model is applied to
+genuinely unknown interactions. Second, some proteins (which we will
+denote as \textit{hub proteins}) have significantly more known
+interactions with drugs than others. In the STITCH database, $5\%$
+of the proteins have $40\%$ of the interactions, and similar
+distributions are present in the Yamanishi and Drugbank datasets
+\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
+these proteins can often increase predictive performance while again
+not reflecting the actual performance when applied to a new protein
+(e.g., a protein for which no interactions are known). This bias in
+the number of drugs targeting certain proteins may be the result of
+study bias where more \glqq valuable\grqq{} proteins have more drugs
+designed to target them due to their involvement in more common
+diseases (or diseases for which drugs can be more profitably
+marketed).
+
+\todo[inline]{Likely discussion:}
+The aim of \name is to predict candidate drugs that target a given
+protein; the challenge is to develop a training and evaluation scheme
+that does not simply overfit to the inherent biases in training and
+testing data.
+In general, when performing cross-validation for DTI prediction, the
+options are to split over 
 \begin{enumerate}
-	\item Build split over drugs
-	\item Build split over drug-target pairs
-	\item Build split over proteins
+	\item split over drugs,
+	\item split over drug--target pairs, or
+	\item split over proteins
 \end{enumerate}
 where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
+They that at least parts of the interactions are not seen during training and evaluate either how well targets are predicted for unseen drugs or unseen proteins.
 Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
 \\
+
+\todo[inline]{Likely discussion:}
 The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.

### 2021-01-30 paper/main.tex

@@
 
-\subsection{Choosing a cross-validation splitting scheme}
+\subsection{\name: computational model to identify drugs that target a
+	protein}
 
@@
 can be used to determine if they may physically interact, and
+information about drug effects that may localize on an interaction
+network.
 

### 2021-01-30 paper/main.tex

@@
 
+We developed \name as a computational model to predict drug--target
+interactions. Specifically, given a protein, \name will identify and
+rank known drugs that likely target this protein. \name combines two 
+types of features: structural information for drugs and proteins that
+can be used to determine if they may physically interact, and
+information about drug effects that may localize on an interaction
+network.
+As structural features, \name uses structural representations of drugs
+from the SMILES transformer \cite{SmilesTransformer} and representations of protein
+amino acid sequences from DeepGOPlus \cite{DeepGoPlus}.
+\name learns representations of drug effects and protein functions
+using the ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
+
+We use two Siamese neural networks to combine the molecular
+and phenotype representations of drugs and proteins,
+% and
+% ontology features, yielding the deep neural network transformers
+% (DNNT) for both methods, and both drugs and proteins,
+% individually. 
+and add these to a protein interaction network as node features.
+% neural network in which nodes representing interactions
+% between proteins as node features, 
+We then train this model in an end-to-end manner. % Notably, like in the
+% pretraining phase, we rely on siamese neural networks, enforcing
+% similar representations for fitting drug-protein pairs. 
+The overall architecture is depicted in Figure~\ref{fig:Fullmodel}.
+
+% For a given protein, we want to find all suitable interacting drugs, \textbf{among a set of known drugs}.
+%To obtain a prediction for a given drug-protein pair,
+
+Note that this approach can be easily generalized, profiting from
+other protein function and phenotype representation methods.
+
+\begin{figure}[!tpb]
+	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
+	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
+The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
+}
+	\label{fig:Fullmodel}
+\end{figure}
+
+Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model.
+Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
+By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
+
+\subsection{\name identifies drugs that target a protein}
+
 \todo[inline]{Likely discussion:}
@@
 where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
-They that at least parts of the interactions are not seen during training and evaluate either how well targets are predicted for unseen drugs or unseen proteins.
+They ensure that at least parts of the interactions are not seen during training and evaluate either how well targets are predicted for unseen drugs or unseen proteins.
 Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
@@
 However, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
-Second, these two strategies are more susceptible to the second and third bias, as only in these cases the model may overfit on the number of existing interactions for a single protein, while in the third scheme the number of interactions of the test proteins is not known.
+Second, these two strategies are more susceptible to the second bias, as only in these cases the model may overfit on the number of existing interactions for a single protein, while in the third scheme the number of interactions of the test proteins is entirely unknown during training process.
 
@@
 As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
-
-Our aim is to develop and evaluate a computational model to predict
-drug--target interactions. Specifically, given a protein, we are
-interested in identifying the drugs that may target the protein; this
-scenario is motivated by the need to identify drugs that can target a
-
-\subsection{\name identifies drugs that target a protein}
-
-We developed \name as a computational model to predict drug--target
-interactions. Specifically, given a protein, \name will identify and
-rank drugs that likely target this protein. \name combines two 
-types of features: structural information for drugs and proteins that
-can be used to determine if they may physically interact, and
-information about drug effects that may localize on an interaction
-network.
-
-This scenario is motivated by the need to identify drugs that can target a
-specific protein that may be associated with a disease or abnormal
-phenotypes.
-Our aim is further to combine two different types of features:
-structural information based on protein structure and molecular drug composition, and their respective observable characteristics.
 
@@
 Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
-
-To obtain a prediction for a given drug-protein pair\todo{should this
-  be about pairs, or rather something like ``given a protein, we aim
-  to find a drug...''? I.e., find a consistent way to separate
-  yourself language-wise from a naive prediction of pairs.}, we first
-build DL2vec representations\todo{In results: talk about what they are
-supposed to DO (``embeddings/representations of the phenotypes and
-functions...''), not how you implement them (you could use something
-else, for example).} for both drugs and proteins over PhenomeNET, while also preparing structural, molecular representations.
-Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers for both methods and both drugs and proteins, individually.
-These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.
-Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs.
-The overall architecture can be seen in Figure~\ref{fig:Fullmodel}.
-
-\subsection{High level description}
-
-For a given protein, we want to find all suitable interacting drugs, \textbf{among a set of known drugs}.
-To obtain a prediction for a given drug-protein pair, we first
-build representations embodying both phenotypes and functions for both drugs and proteins, while also preparing structural, molecular representations, e.g., utilizing DL2vec over PhenomeNET and the prebuilt features from SMILES transformer and DeepGOPlus.
-Note that this approach can be easily generalized, profiting from other protein function and phenotype representation methods.
-\\
-Hence, we pretrain the siamese networks for both molecular and ontology features, yielding the deep neural network transformers (DNNT) for both methods, and both drugs and proteins, individually.
-These embeddings are now added to the graph neural network as node features, where we start an end-to-end learning and training process.
-Notably, like in the pretraining phase, we rely on siamese neural networks, enforcing similar representations for fitting drug-protein pairs.
-The overall architecture is depicted in Figure~\ref{fig:Fullmodel}.\\
-
-\begin{figure}[!tpb]
-	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
-	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
-The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
-}
-	\label{fig:Fullmodel}
-\end{figure}
-
-Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model.
-Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
-By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
 

### 2021-01-30 paper/main.tex
Update model section and moved micro AUC to methods

@@
 
-\todo[inline]{First part is not about evaluation but optimization
-  objectives, should go in the Model section:}
-As the number of drug-targets are sparsely given w.r.t.\todo{no
-  acronyms} to the number of both drugs and proteins considered, the
-resulting training, validation and testing datasets are highly
-imbalanced. As there are only $~22.000$\todo{exact number} links in
-the considered subset the ratio
+To assess each model, we compute a variety of common metrics for binary classification.
+As the datasets are highly imbalanced, we chose the \textit{area under receiver operating characteristic}-score (AUROC) on training, validation and testing split.
+Therefor, we calculated the true positive rate (TPR), false positive rate (FPR) and precision score.
+We will further compute true positives (TP), false positives (FP), false negatives (FN) and finally true negatives (TN).
+We calculate the ROC by plotting the TPR against the FPR for various threshold settings for the model predictions.
+We eventually calculate the area under ROC curve utilizing trapezoidal approximations.
+We will refer to this measure as $MacroAUC$.
+
+In contrast we also calculate the $MicroAUC$ score.
+For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
+With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
 \begin{equation*}
-	\frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
+	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
 \end{equation*}
-consequently needing compensation in the computed loss function and
-appropriate metrics for the evaluation.
 
-Therefore, we weighted all positive drug-protein pair samples with
-this ratio by introducing the following loss function w.r.t.\todo{no
-  acronym} to binary cross-entropy:
-\begin{equation}
-	l(x,y) = - w \left[ y \cdot \log x + (1 - y) \cdot \log (1 - x) \right]
-\end{equation}
-for given prediction $x$ and target $y$. We average this loss among
-all drug-protein pairs in the training set, leading to a stable
-environment for the used optimizing scheme \textit{Adam}
-\citep{Adam2014}. We implemented a 5-fold cross validation among the
-proteins as justified in the results section. Furthermore, we used
-early stopping to detect plateaus in the training process.
-
-\todo[inline]{This needs to be more elaborate: add the micro- vs
-  macro-averages here (formulas)}
-To assess each model, we computed the \textit{area under receiver
-  operating characteristic}-score (AUROC), \textit{F1-score} (F1) and
-\textit{Matthews correlation coefficient} (MCC) for each considered
-hyperparameter configuration setting.
-
-\todo[inline]{Remove from here:}
-Avoiding validation overfitting, we eventually tested the best
-performing model on the Yamanishi dataset as presented and
-discussed. Furthermore, we compare our results on this very dataset
-with various other cutting edge approaches to DTI-prediction.
+Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
 
 \subsection{Model}
+Before introducing the models core components, we will clarify the previously mentioned notions of top-down and bottom-up.
+We will refer to features and methods, starting from the observable characteristics of either drug or protein inferring the likely molecular mechanisms as top-down.
+On the other hand, bottom-up refers to approaches deducing from even finer molecular properties, such as single functional groups or amino acid sequences for drugs and proteins, respectively.
 
-\todo[inline]{Introduce the notion of top-down and bottom-up:}
 In order to build a method that incorporates both top-down and
@@
 
-As we wanted to learn from the similarity of drug side effects and
+\subsubsection{Siamese neural networks and modular learnable feature transformation}
+As we want to learn from the similarity of drug side effects and
 protein phenotypes we opted for a deep siamese network approach, hence
 learning a high-dimensional embedding emphasizing this identity by
-forcing a maximal cosine similarity between these embeddings. On the
-other hand we built a deep neural network for the molecular structure
-based features, also benefiting from the siamese network
-architecture. Therefore, the precomputed embeddings are run through a
-deep neural network transformation (DNNT) which also reduces the
+forcing a similarity between these embeddings. On the other hand we built a deep neural network for the molecular structure based features, also benefiting from the siamese network
+architecture. Computing the similarity between two representations, allows for a variety of different methods.
+However, we decided for the cosine similarity measure being invariant to scaling.
+
+Therefore, the precomputed embeddings are run through a
+regular, neural learnable feature transformation (LFT) network, which also reduces the
 eventual representation size for drugs and proteins separately. An
@@
 
+While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we emphasize the versatility of this approach, as both ontology and molecular feature for both drugs and proteins are reduced to similar dimensionality.
+This allows for a high amount of modularity and different experimental setups by plugging different kinds of features into the model.
+Additionally, these pretrained features can be used for a variety of other tasks.
+Additionally, the ontology LFT can be reused for a variety of DL2vec based features with respect to other ontologies and hypotheses.
+We hereby followed the results of \textit{DL2vec}, indicating that utilizing the activation function $\sigma := \mathrm{LeakyReLU}$ leads to performance increase.
+
 \begin{figure}[!tpb]%figure1
@@
 	\caption{Siamese network applied to molecular and DL2vec
-          features, utilizing deep neural network transformations
-          (DNNT). The similarity function $\otimes$ yields the
+          features, utilizing deep learnable feature transformations
+          (LFT). The similarity function $\otimes$ yields the
           similarity between both transformed embeddings e.g. by
@@
 
-1
 \subsubsection{Graph convolutional layers}
@@
 \subsubsection{Hyperparameter tuning}
+As the number of drug-targets are sparsely given with respect to the number of both drugs and proteins considered, the resulting training, validation and testing datasets are highly
+imbalanced. As there are only $22.336$ links in
+the considered subset the ratio
+\begin{equation*}
+	\frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
+\end{equation*}
+consequently needing compensation in the computed loss function and
+appropriate metrics for the evaluation.
+
+Therefore, we weighted all positive drug-protein pair samples with
+this ratio by introducing the following loss function with respect to binary cross-entropy:
+\begin{equation}
+	l(x,y) = - w \left[ y \cdot \log x + (1 - y) \cdot \log (1 - x) \right]
+\end{equation}
+for given prediction $x$ and target $y$. We average this loss among
+all drug-protein pairs in the training set, leading to a stable
+environment for the used optimizing scheme \textit{Adam}
+\citep{Adam2014}. We implemented a 5-fold cross validation among the
+proteins as justified in the results section. Furthermore, we used
+early stopping to detect plateaus in the training process.
+
 To find the best hyperparameter configuration for the proposed model
@@
 	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
-	\caption{Full DTI prediction model based on the pretrained deep neural network transformers (DNNT) for either molecular structure or ontology based features.
+	\caption{Full DTI prediction model based on the pretrained learnable feature transformations (LFT) for either molecular structure or ontology based features.
 The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
@@
 
+\subsection{Experiments}
+In this section we will present the results of \name
+
 \subsection{MicroAUC}
 Continuing this line, we propose the well known metric \textit{MicroAUC} for a better evaluation for this purpose.
-For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
-With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
-\begin{equation*}
-	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
-\end{equation*}
 
-Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
 Note further that both are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are only plausible for protein split and drug split, respectively.

### 2021-01-30 paper/main.tex
Update experiments section

@@
 Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
+Additionally, the $MicroAUC$ score may not be defined as, in some datasets some targets or drugs, respectively, do not have any interactions, leading to an infeasible $TPR$ and an undefined $AUROC$ score for that entity.
+For those entities we impute the $MicroAUC$ linearly, by using the accuracy for this subset.
 
@@
 using the ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
+When combining ontology and molecular features with or without the graph model, we concatenate both protein features and both drugs features, before plugging them into the graph or the similarity computation.
 
@@
 %To obtain a prediction for a given drug-protein pair,
+If
 
@@
 
-\subsection{Experiments}
-In this section we will present the results of \name
-
 \subsection{MicroAUC}
-Continuing this line, we propose the well known metric \textit{MicroAUC} for a better evaluation for this purpose.
-
-Note further that both are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are only plausible for protein split and drug split, respectively.
+Continuing this line, we apply the well metric \textit{MicroAUC} for a better evaluation for this purpose.
+Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively.
 As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
 
+\subsection{Experiments}
+In this section we will present the results of \name on both STITCH and Yamanishi benchmark dataset, omitting any kind of comparison for now.
+We trained, validated and finally tested all considered models on STITCH dataset using a 5-fold cross-validation over a protein split, as reasoned and substantiated previously.
+We only evaluated the best performing model, with respect to \textit{AUROC} and $MicroAUC_p$, on the Yamanishi benchmark dataset.
+We will hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.
+As depicted in Figure \ref{tab:STITCHresults}, \textit{OntoPred} is showing significantly better performance on STITCH, while also only \textit{OntoPred} shows significant performance increase when adding \textit{GENConv} graph convolutional neural layers over the PPI graph.
+However, graph convolutional neural layers such as \textit{GCNConv} and \textit{GENConv}, especially when incorporated into the proposed \textit{ResGraphConv} blocks, add a lot of learnable parameters to the network, leading to more expressive power.
+This may lead to better performance, while deceiving, whether the respective features actually localize on the PPI graph.
+
+To eliminate these options we rebuild the graph model, whilst removing all graph convolutional neural layers in the residual blocks.
+This pruned network, with no information on the protein--protein interactions, reached very similarly results, as the original \textit{OntoPred} model.
+Secondly, when substituting the original \textit{GENConv} layers with \textit{GCNConv} or other related layers in the graph convolutional step of the ResGraphConv blocks, this model would achieve once again no significant performance gain in comparison to the plain \textit{OntoPred} one.
+This performance increase was only noticeable for \textit{GENConv} layers.
+The discrepancy of \textit{GENConv} and \textit{GCNConv} may be founded in numerical stability and fading signals, as described in the introduction of both methods.
+
+As this performance increase is quite significant, we \textbf{conclude that protein function does localize while molecular features do not (BETTER FORMULATION)}. ...
+
+\begin{figure}[!tpb]
+	\centerline{\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
+		\hline
+		STITCH results&\multicolumn{4}{c|}{PPI graph}\\
+		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
+		&AUROC&Micro $AUC_p$&AUROC&Micro $AUC_p$\\
+		\hline
+		MolPred&$0.69$&$0.65$&$0.69$&$0.67$\\
+		\hline
+		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$\\
+		\hline
+		MolPred + OntoPred & $0.89$ & $0.90$&$0.93$&$0.94$\\
+		\hline
+		\hline
+		Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
+		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
+		&AUROC&Micro AUC&AUROC&Micro AUC\\
+		\hline
+		MolPred + OntoPred & $0.83$ & $0.82$&$0.84$&$0.84$\\
+		\hline
+	\end{tabular}}
+	\caption{Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation. MolPred and OntoPred are the predictors for molecular and ontology based features, respectively. }
+	\label{tab:STITCHresults}
+\end{figure}
+
 \subsection{Baseline model}
-Before introducing the core of our eventual model, we propose a suitable naive baseline model in order to analyze and understand the existing approaches and this novel method for drug target interaction prediction.
+Before comparing \name to other methods, we propose a suitable naive baseline model in order to analyze and understand existing approaches and \name for drug target interaction prediction.
 For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, we construct an interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
@@
 \begin{equation*}
-	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=0}^{|D|}M_{ij}
+	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
 \end{equation*}
@@
 Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
+
+We evaluate this naive predictor on both STITCH and Yamanishi dataset, performing a 5-fold cross-validation over both drugs and drug--protein pairs.
+Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown.
+For each fold, we gradually increase $k$, eventually saving the best performance for each fold.
+The corresponding AUROC scores are summarized in the following table: \\
+
+\centerline{\begin{tabular}{|l|p{1.5cm}|p{1.5cm}|}
+	\hline
+	Dataset&\multicolumn{2}{c|}{Splitting scheme}\\
+	&DT pairs&Drugs\\
+	\hline
+	STITCH & ?? & ??\\
+	\hline
+	Yamanishi & $0.88$&$0.85$\\ 
+	\hline
+\end{tabular}}
+\vspace{0.5cm}
+This baseline predictor shows remarkable performance on both datasets, that
 
@@
 	\item using indications is like cheating, as not applicable for searching new drugs
-	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
+	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related[!tpb] drugs/proteins
 	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

### 2021-01-31 paper/main.tex
Results first part

@@
 		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
-		&AUROC&Micro AUC&AUROC&Micro AUC\\
+		&AUROC&Micro AUC&AUROC&Micro $AUC_p$\\
 		\hline
@@
 
-We evaluate this naive predictor on both STITCH and Yamanishi dataset, performing a 5-fold cross-validation over both drugs and drug--protein pairs.
+We evaluate this naive predictor on both STITCH and Yamanishi dataset, first running on the whole dataset followed by a 5-fold cross-validation (CV) over both drugs and drug--protein pairs.
 Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown.
@@
 
-\centerline{\begin{tabular}{|l|p{1.5cm}|p{1.5cm}|}
+\centerline{\begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
 	\hline
-	Dataset&\multicolumn{2}{c|}{Splitting scheme}\\
-	&DT pairs&Drugs\\
+	Dataset&\multicolumn{3}{c|}{Splitting scheme}\\
+	&Whole&DT pairs&Drugs\\
 	\hline
-	STITCH & ?? & ??\\
+	STITCH& $0.76$& $0.70$ & $0.73$\\
 	\hline
-	Yamanishi & $0.88$&$0.85$\\ 
+	Yamanishi& $0.88$& $0.84$&$0.85$\\ 
 	\hline
@@
 \vspace{0.5cm}
-This baseline predictor shows remarkable performance on both datasets, that
+The baseline predictor shows diverging performances on both datasets with remarkable results on the Yamanishi benchmark.
+As the baseline predictor is based on the second dataset bias i.e.
+the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
+
+\subsection{Comparison}
+
+In this section we will compare \name with the described baseline method and various other cutting-edge models.
+We therefor evaluate all models on their recommended splitting scheme choice, hyperparameters and fold amoung for cross-validation, measuring their respective AUROC score.
+We eventually assess each model, by performing a protein-wise cross-validation determining (macro) AUROC and $MicroAUC_p$.
+For this we allow stratification for the training process, but not for the validation and testing phase, as real world scenarios are rather imbalanced.
+This however, only impacts the not considered area under precision recall curve (AUPRC) and not the Macro AUROC score, while also supporting the expressiveness of $MicroAUC_p$ with more data points.
+
+For our comparison we chose the best performing methods for drug--target interaction prediction evaluating on the Yamanishi benchmark dataset.
+These state-of-the-art methods approaches include  DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed superior results in comparison to numerous works.
+Furthermore, we added DTINet \cite{DTINet2017}, which lays the foundation for a number of methods, such as NeoDTI \cite{NeoDTI2019}, with similar methodology.
+
+The results of this first analysis are summarized in the upper part of Figure \ref{tab:Results}.
+
+\begin{figure}[!tpb]%figure1
+
+	\centerline{\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+		\hline
+		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
+		\hline
+		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
+		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
+		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
+		\hline
+		\hline
+		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
+		\hline
+		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
+		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
+		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
+		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
+		\hline
+	\end{tabular}}
+	\caption{Test}
+	\label{tab:Results}
+\end{figure}
 

### 2021-02-03 paper/main.tex
.

@@
 
-Protein-protein interaction networks have shown great results in $\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context for molecular system biology.
-However, these contexts were never applied to the problem of drug-target-interaction prediction.
-Thus we formalized our hypotheses over these interaction graphs and will test them in the following chapters.
+Protein-protein interaction networks have shown great results in
+$\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context
+for molecular system biology. However, these contexts were never
+applied to the problem of drug-target-interaction prediction. Thus we
+formalized our hypotheses over these interaction graphs and will test
+them in the following chapters.
 
@@
 As we want to learn from the similarity of drug side effects and
-protein phenotypes we opted for a deep siamese network approach, hence
+protein phenotypes, we use a deep Siamese neural network.
+A Siamese neural network aims to learn a similarity between two
+embeddings.
+
 learning a high-dimensional embedding emphasizing this identity by
-forcing a similarity between these embeddings. On the other hand we built a deep neural network for the molecular structure based features, also benefiting from the siamese network
-architecture. Computing the similarity between two representations, allows for a variety of different methods.
-However, we decided for the cosine similarity measure being invariant to scaling.
+forcing a similarity between these embeddings. On the other hand we
+built a deep neural network for the molecular structure based
+features, also benefiting from the siamese network
+architecture. Computing the similarity between two representations,
+allows for a variety of different methods.  However, we decided for
+the cosine similarity measure being invariant to scaling.
 
-Therefore, the precomputed embeddings are run through a
-regular, neural learnable feature transformation (LFT) network, which also reduces the
-eventual representation size for drugs and proteins separately. An
-example structure for both types of features can be found in
-Figure~\ref{fig:SiameseNetwork}.
+Therefore, the precomputed embeddings are run through a regular,
+neural learnable feature transformation (LFT) network, which also
+reduces the eventual representation size for drugs and proteins
+separately. An example structure for both types of features can be
+found in Figure~\ref{fig:SiameseNetwork}.
 

### 2021-02-07 paper/main.tex
mehtods

@@
 \subsection{Datasets}
-The data for the different parts of this model were obtained from
-various sources. Starting with the protein-protein interactions, we
-fetched $\approx 11000$\todo{precise number} human proteins with over 170.000 links from
-STRING \citep{STRINGv10}. For the drug-target interactions themselves,
-we fetched 137.000 links from STITCH database \citep{STITCHv5}. As
-both STRING and STITCH provide probability scores for each
-association, we filtered them as advised by a threshold of 700, thus
-only obtaining likely interactions.
+We obtain a dataset consisting of $\approx 11000$\todo{precise number}
+human proteins with over 170.000\todo{exact number} links from STRING
+\citep{STRINGv10}. For the drug-target interactions, we
+fetched 137,000\todo{precise}} links from the STITCH database
+\citep{STITCHv5}. As both STRING and STITCH provide confidence scores
+for each association, we filtered them as advised by a threshold of
+$700$, therefore retaining only high-confidence interactions.
 
-For the ontology segment we utilized PhenomeNET \citep{PhenomeNET2011}, a collection of various ontologies such as Human Phenotype Ontology \citep{HPO2018}, Gene Ontology (\citet{GOoriginal2000} \citet{GOrecent2020}), Mammalian Phenotype Ontology \citep{MP2009} and numerous others.
-\todo{which ontologies to cite}
-Side effects and their links to drugs were obtained Side Effect Resource (SIDER)\citep{SIDER} and structured according MedDRA database \citep{MedDRA}.
-They were mapped to PhenomeNET with aid of \textit{Phenomebrowser.net}, which provides a SPARQL query endpoint for the mentioned resources.
-\\
+We utilize the PhenomeNET ontology \citep{PhenomeNET2011}, an ontology
+integrating ontologies such as the Human Phenotype Ontology
+\citep{HPO2018}, Gene Ontology \cite{GOoriginal2000, GOrecent2020},
+Mammalian Phenotype Ontology \citep{MP2009} and several others.
+We obtained side effects and their links to drugs from SIDER
+\citep{SIDER}; SIDER contains side effects encoded using identifiers
+from the MedDRA database \citep{MedDRA}. We mapped side effects to the
+PhenomeNET ontology using the \textit{Phenomebrowser.net}, which
+provides a SPARQL query endpoint for the mentioned resources.
 

### 2021-02-07 paper/main.tex
rearrange method

@@
 
-For comparative evaluation we used the gold standard dataset introduced by \citet{Yaminishi2008}, which includes both drug-target interaction pairs and side effects.\\
+We only use proteins in our analysis that have at least one link in
+either STITCH or STRING, and drugs with at least one side effect and
+one existing target. Therefore, the intersection between these
+resources yields 1,160 drugs and 6,680 human proteins for the training
+phase. We provide links to and methods for obtaining and processing
+the necessary data on Github.
 
-Eventually, we only considered proteins that had at least one link in either STITCH or STRING, and drugs with at least one side effect and one existing target.
-Thus, the intersection between these resources yielded 1160 drugs and 6680 human proteins for the training phase.
-We provide links to and methods for the necessary data in the provided Github repository.\\
-
-\subsection{Evaluation and metrics}
-
-To assess each model, we compute a variety of common metrics for binary classification.
-As the datasets are highly imbalanced, we chose the \textit{area under receiver operating characteristic}-score (AUROC) on training, validation and testing split.
-Therefor, we calculated the true positive rate (TPR), false positive rate (FPR) and precision score.
-We will further compute true positives (TP), false positives (FP), false negatives (FN) and finally true negatives (TN).
-We calculate the ROC by plotting the TPR against the FPR for various threshold settings for the model predictions.
-We eventually calculate the area under ROC curve utilizing trapezoidal approximations.
-We will refer to this measure as $MacroAUC$.
-
-In contrast we also calculate the $MicroAUC$ score.
-For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
-With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
-\begin{equation*}
-	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
-\end{equation*}
-
-Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
-Additionally, the $MicroAUC$ score may not be defined as, in some datasets some targets or drugs, respectively, do not have any interactions, leading to an infeasible $TPR$ and an undefined $AUROC$ score for that entity.
-For those entities we impute the $MicroAUC$ linearly, by using the accuracy for this subset.
+For comparative evaluation, we use the gold standard dataset
+introduced by \citet{Yaminishi2008} \todo[inline]{consisting of XXXX
+  interactions between XXX drugs and XXX protein targets}.
 
@@
   results and can put them here (better: in the supplement).}
+
+\subsection{Evaluation and metrics}
+
+To assess each model, we compute a variety of common metrics for binary classification.
+As the datasets are highly imbalanced, we chose the \textit{area under receiver operating characteristic}-score (AUROC) on training, validation and testing split.
+Therefor, we calculated the true positive rate (TPR), false positive rate (FPR) and precision score.
+We will further compute true positives (TP), false positives (FP), false negatives (FN) and finally true negatives (TN).
+We calculate the ROC by plotting the TPR against the FPR for various threshold settings for the model predictions.
+We eventually calculate the area under ROC curve utilizing trapezoidal approximations.
+We will refer to this measure as $MacroAUC$.
+
+In contrast we also calculate the $MicroAUC$ score.
+For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
+With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
+\begin{equation*}
+	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
+\end{equation*}
+
+Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
+Additionally, the $MicroAUC$ score may not be defined as, in some datasets some targets or drugs, respectively, do not have any interactions, leading to an infeasible $TPR$ and an undefined $AUROC$ score for that entity.
+For those entities we impute the $MicroAUC$ linearly, by using the accuracy for this subset.
 

### 2021-02-07 paper/main.tex
edit model sec

@@
 \subsection{Model}
-Before introducing the models core components, we will clarify the previously mentioned notions of top-down and bottom-up.
-We will refer to features and methods, starting from the observable characteristics of either drug or protein inferring the likely molecular mechanisms as top-down.
-On the other hand, bottom-up refers to approaches deducing from even finer molecular properties, such as single functional groups or amino acid sequences for drugs and proteins, respectively.
 
-In order to build a method that incorporates both top-down and
-bottom-up features, we first created a model for each
-individually. Hence, we assemble rich molecular structure based
-features for drugs from
-\textit{SmilesTransformer} \citep{SmilesTransformer} and proteins from
-\textit{DeepGOPlus} \citep{DeepGoPlus}. \textit{SmilesTransformer}
-introduces an autoencoder, learning over the SMILES strings and thus
-the molecular organization of each drug in an unsupervised manner. On
-the other hand, we take advantage of the pretrained models of
-\textit{DeepGOPlus}, obtaining features from the proteins amino acid
-sequence, showing significant performance in the field of protein
-function prediction. Thus, both embeddings seem to suitably supplement
-the following ontology based representations.
-
+Our model combines ``top-down'' and ``bottom-up'' information for
+drug--target identification. We consider an approach to be
+``top-down'' when observable characteristics of either a drug (such as
+a drug effect) or protein (such as a protein function, or phenotypes
+resulting from a loss of function) are used to provide information
+about a molecular mechanisms; we consider an approach ``bottom-up''
+when structural or other molecular information is used to determine a
+mechanism.  In order to build a method that incorporates both top-down
+and bottom-up features, we first create a model for each type of
+feature separately.  As features for the bottom-up model, we use
+features derived from molecular structures of drugs from the
+\textit{SmilesTransformer} \citep{SmilesTransformer} and molecular
+features for proteins from \textit{DeepGOPlus}
+\citep{DeepGoPlus}. \textit{SmilesTransformer} introduces an
+autoencoder, learning over the SMILES strings and therefore the
+molecular organization of each drug in an unsupervised manner. 
+\textit{DeepGOPlus} provides  features derived from protein amino acid
+sequences which are useful to predict protein function.
+% Thus, both embeddings seem to suitably supplement
+% the following ontology based representations.
 In the top-down section, we used \textit{DL2vec} \citep{DL2vec2020} to
@@
 \subsubsection{Siamese neural networks and modular learnable feature transformation}
+
 As we want to learn from the similarity of drug side effects and

### 2021-02-07 paper/main.tex
edit model sec

@@
 \citep{STRINGv10}. For the drug-target interactions, we
-fetched 137,000\todo{precise}} links from the STITCH database
+fetched 137,000\todo{precise} links from the STITCH database
 \citep{STITCHv5}. As both STRING and STITCH provide confidence scores
@@
 % the following ontology based representations.
-In the top-down section, we used \textit{DL2vec} \citep{DL2vec2020} to
-obtain ontology based representations. DL2vec constructs a graph by
-introducing vertices and edges for each ontology class and axiom,
-respectively, followed by random walks starting from each
-entity. These walks are eventually learned on using a Word2vec
-\citep{Word2vec2013} model. Thus, we pick up rich, neighbourhood
-focused representations for each entity, which has shown great results
-for representing protein function and phenotypes. The overall
-structure of the ontology can be seen in Figure~\ref{fig:Onto}.
 
+As phenotypes and functions are encoded through ontologies, we use
+DL2Vec \citep{DL2vec2020} to obtain ontology based representations for
+use as top-down features. DL2vec constructs a graph by introducing
+nodes for each ontology class and edges for ontology axioms, followed
+by random walks starting from each node in the graph. These walks are
+encoded using a Word2vec \citep{Word2vec2013} model. Therefore, DL2Vec
+generates representations that can encode drug effects or protein
+functions while preserving their semantic neighbourhood within that
+graph.
+
+% The overall structure of the ontology can be seen in
+% Figure~\ref{fig:Onto}.
+
+\todo[inline]{Consider supplement, or combine with other figures.}
 \begin{figure}[!tpb]%figure1

### 2021-02-07 paper/main.tex
siamese question

@@
 
+\todo[inline]{Looking at the code, I wonder if what is implemented is
+  actually a Siamese ANN; it does not look as if the weights are
+  shared between the two networks, so looks more like a twin or
+  half-twin network?}
+
 As we want to learn from the similarity of drug side effects and
-protein phenotypes, we use a deep Siamese neural network.
+protein phenotypes, we use a deep Siamese neural network with
+a contrastive loss using cosine similarity.
 A Siamese neural network aims to learn a similarity between two
@@
 
-learning a high-dimensional embedding emphasizing this identity by
-forcing a similarity between these embeddings. On the other hand we
-built a deep neural network for the molecular structure based
-features, also benefiting from the siamese network
-architecture. Computing the similarity between two representations,
-allows for a variety of different methods.  However, we decided for
-the cosine similarity measure being invariant to scaling.
+% learning a high-dimensional embedding emphasizing this identity by
+% forcing a similarity between these embeddings. On the other hand we
+% built a deep neural network for the molecular structure based
+% features, also benefiting from the siamese network
+% architecture. Computing the similarity between two representations,
+% allows for a variety of different methods.  
+% However, we decided for
+% the cosine similarity measure being invariant to scaling.
 

### 2021-02-07 paper/main.tex
siamese question

@@
   shared between the two networks, so looks more like a twin or
-  half-twin network?}
+  half-twin network? Siamese would be sim(f(e1),f(e2)) but the
+  implementation seems to be sim(f(e1), g(e2)).}
 

### 2021-02-07 paper/main.tex
done methods

@@
 \subsubsection{Graph convolutional layers}
-These molecular and ontology based sub-models were added to a larger
-graph convolutional model, based on the protein-protein interaction
-(PPI) graph. The PPI dataset is represented by a graph $G=(V,E)$,
-where each protein is represented by a vertex $v\in V$, and each edge
-$e\in E\subseteq V\times V$ symbolizes an interaction between two
-proteins. Additionally, we introduce mapping
+
+We included these molecular and ontology-based sub-models with a
+larger graph convolutional model. The graph underlying the graph
+convolutional neural network is based on the protein--protein
+interaction (PPI) graph. The PPI dataset is represented by a graph
+$G=(V,E)$, where each protein is represented by a vertex $v\in V$, and
+each edge $e\in E\subseteq V\times V$ symbolizes an interaction
+between two proteins. Additionally, we introduce a mapping
 $x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node
-feature $x_v = x(v)$, where $d$ denotes the dimensionality of the node
+feature $x_v := x(v)$, where $d$ denotes the dimensionality of the node
 features.
 
-As described before, graph convolution has shown significant
-performance increase in a variety of tasks. While there are various
-methods out there we will only introduce the most basic one here. A
-graph convolutional layer \citet{GCNConv} consists of a
-learnable weight matrix followed by an aggregation step, formalized by
+% As described before, graph convolution has shown significant
+% performance increase in a variety of tasks. While there are various
+% methods out there we will only introduce the most basic one here. 
+A graph convolutional layer \citet{GCNConv} consists of a learnable
+weight matrix followed by an aggregation step, formalized by
 \begin{equation}
@@
 learnable weight matrix. Added self-loops enforce that each node
-representation is directly dependent on its own preceding
-one. Notably, the number of graph convolutional layers stacked equals
-the radius of relevant nodes for each vertex within the graph.
+representation is directly dependent on its own preceding one. The
+number of graph convolutional layers stacked equals the radius of
+relevant nodes for each vertex within the graph.
 
@@
 this generalized layer type as \textsc{GENConv}, following the
-notation of \citet{PytorchGeometric}.  While the reordering is mainly
-import for numerical stability, this alteration also addresses the
-vanishing gradient problem for deeper convolutional networks
-\cite{GENConv2020}. Additionally, we can also generalize the
+notation of PyTorch Geometric \cite{PytorchGeometric}.  While the
+reordering is mainly import for numerical stability, this alteration
+also addresses the vanishing gradient problem for deeper convolutional
+networks \cite{GENConv2020}. Additionally, we can also generalize the
 aggregation function to allow different weighting functions such as
@@
 filter and weight matrix are applied to and learned for all nodes
-simultaneously, and the resulting information\todo{Which information?
-  Specify} hold no information on their own connectivity.
+simultaneously. % , and the resulting information\todo{Which information?
+  % Specify} hold no information on their own connectivity.
+We further employ another mechanism to avoid redundancy and fading
+signals in stacked graph convolutional networks, using residual
+connections and a normalization scheme \cite{DeepGCN2019,
+  DeeperGCN2020}.
+The residual blocks are reusable and can be stacked multiple times. %, thereby not losing focus of each node neighbourhood.
+\todo[inline]{Needs a clearer outline/description/figure of YOUR model
+structure.}
+The structure is depicted in Figure~\ref{fig:ResGraphConvBlocks}.
 
@@
 
-We employ another mechanism to avoid redundancy and fading signals in
-stacked graph convolutional networks, which was introduced by
-\citet{DeepGCN2019} and refined in \citet{DeeperGCN2020}. The authors
-propose a residual network architecture and normalization scheme. This
-structure is depicted in Figure~\ref{fig:ResGraphConvBlocks} resulting
-in reusable residual blocks, which can be stacked multiple times,
-thereby not losing focus of each node neighbourhood.
-
 \subsubsection{Hyperparameter tuning}
-As the number of drug-targets are sparsely given with respect to the number of both drugs and proteins considered, the resulting training, validation and testing datasets are highly
-imbalanced. As there are only $22.336$ links in
+As the number of drug-targets are sparse with respect to the number of
+both drugs and proteins considered, the training, validation and
+testing datasets are imbalanced. As there are only $22,336$ links in
 the considered subset the ratio
@@
 Therefore, we weighted all positive drug-protein pair samples with
-this ratio by introducing the following loss function with respect to binary cross-entropy:
+this ratio by introducing the following loss function with respect to
+binary cross-entropy:
 \begin{equation}
@@
 \end{equation}
-for given prediction $x$ and target $y$. We average this loss among
+for a given prediction $x$ and target $y$. We average this loss among
 all drug-protein pairs in the training set, leading to a stable
@@
 \citep{Adam2014}. We implemented a 5-fold cross validation among the
-proteins as justified in the results section. Furthermore, we used
-early stopping to detect plateaus in the training process.
+proteins. Furthermore, we used early stopping to detect plateaus in
+the training process.
 
@@
 top-down model separately and aimed at best performing models
-w.r.t. previously described metrics. We optimized embedding sizes,
+with respect to our evaluation metrics. We optimized embedding sizes,
 depth of the neural network, optimizer, learning rate and layer types
@@
 
-To assess each model, we compute a variety of common metrics for binary classification.
-As the datasets are highly imbalanced, we chose the \textit{area under receiver operating characteristic}-score (AUROC) on training, validation and testing split.
-Therefor, we calculated the true positive rate (TPR), false positive rate (FPR) and precision score.
-We will further compute true positives (TP), false positives (FP), false negatives (FN) and finally true negatives (TN).
-We calculate the ROC by plotting the TPR against the FPR for various threshold settings for the model predictions.
-We eventually calculate the area under ROC curve utilizing trapezoidal approximations.
-We will refer to this measure as $MacroAUC$.
+To assess each model, we compute a variety of common metrics for
+binary classification. As the datasets are highly imbalanced, we use
+the area under the receiver operating characteristic curve
+(AUROC) on training, validation and testing split. % We
+% calculated the true positive rate (TPR), false positive rate (FPR), and
+% precision score. We will further compute true positives (TP), false
+% positives (FP), false negatives (FN) and finally true negatives (TN).
 
+\todo[inline]{Needs a statement on how the PAIRS are ranked here and
+  TPR/FPR is calculated based on the ranks of the DT-PAIRS.}
+We calculate the AUROC by computing true positive rate at various
+false positive rate thresholds and using trapezoidal approximations to
+estimate the area under the curve. We refer to this measure as
+$MacroAUC$.
+
+\todo[inline]{For RH: check this again later:}
 In contrast we also calculate the $MicroAUC$ score.
@@
 
-Note that drugs and proteins can be interchanged in this formulation, which we will denote with $MicroAUC_p$ and $MicroAUC_d$, respectively.
-Additionally, the $MicroAUC$ score may not be defined as, in some datasets some targets or drugs, respectively, do not have any interactions, leading to an infeasible $TPR$ and an undefined $AUROC$ score for that entity.
-For those entities we impute the $MicroAUC$ linearly, by using the accuracy for this subset.
+Drugs and proteins can be interchanged in this formulation, and we
+refer to the different measures as a protein-centric microAUC
+($MicroAUC_p$) and a drug-centric microAUC
+($MicroAUC_d$). Additionally, the $MicroAUC$ score may not be defined
+as, in some datasets some targets or drugs, respectively, do not have
+any interactions, leading to an infeasible $TPR$ and an undefined
+$AUROC$ score for that entity.
+\todo[inline]{Don't understand this, can you add the formula:} For
+those entities we impute the $MicroAUC$ linearly, by using the
+accuracy for this subset.
 

### 2021-02-07 paper/main.tex
Results point 1

@@
 
-Within DTI prediction, one can observe basic biases resulting from the
+Within DTI prediction, there are potential biases resulting from the
 underlying datasets \citep{Pahikkala2014}. First, novel drugs are
@@
 leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can lead to a bias when it leads
-to \glqq hidden duplicates\grqq{} that can distribute among the
-train--test split, resulting in a better (measured) predictive
-performance than would be expected when the model is applied to
-genuinely unknown interactions. Second, some proteins (which we will
-denote as \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$
-of the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank datasets
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins can often increase predictive performance while again
-not reflecting the actual performance when applied to a new protein
-(e.g., a protein for which no interactions are known). This bias in
-the number of drugs targeting certain proteins may be the result of
-study bias where more \glqq valuable\grqq{} proteins have more drugs
-designed to target them due to their involvement in more common
+proteins \cite{Overington2006}. This can result in a bias when it
+leads to hidden duplicates that can distribute among the train/test
+split, resulting in a better (measured) predictive performance than
+would be expected when the model is applied to identify drugs that
+target a protein for which no drugs yet exist. Second, some proteins
+(which we will denote as \textit{hub proteins}) have significantly
+more known interactions with drugs than others. In the STITCH
+database, $5\%$ of the proteins have $40\%$ of the interactions, and
+similar distributions are present in the Yamanishi and Drugbank
+datasets \cite{Drugbank2007, Drugbank2017} datasets; preferentially
+predicting these proteins may increase predictive performance while
+again not reflecting the actual performance when applied to a new
+protein (e.g., a protein for which no interactions are known). These
+differences in the number of drugs targeting certain proteins may be
+the result of study bias where more ``valuable'' proteins have more
+drugs designed to target them due to their involvement in more common
 diseases (or diseases for which drugs can be more profitably
@@
 information about drug effects that may localize on an interaction
-network.
-As structural features, \name uses structural representations of drugs
-from the SMILES transformer \cite{SmilesTransformer} and representations of protein
-amino acid sequences from DeepGOPlus \cite{DeepGoPlus}.
-\name learns representations of drug effects and protein functions
-using the ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
-When combining ontology and molecular features with or without the graph model, we concatenate both protein features and both drugs features, before plugging them into the graph or the similarity computation.
+network.  As structural features, \name uses structural
+representations of drugs from the SMILES transformer
+\cite{SmilesTransformer} and representations of protein amino acid
+sequences from DeepGOPlus \cite{DeepGoPlus}.  \name learns
+representations of drug effects and protein functions using the
+ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
+
+When combining ontology and molecular features with or without the
+graph model, we concatenate both protein features and both drugs
+features, before plugging them into the graph or the similarity
+computation.
 

### 2021-02-07 paper/main.tex
Results point 1

@@
 
+\todo[inline]{Move to methods:}
 When combining ontology and molecular features with or without the
@@
 
-We use two Siamese neural networks to combine the molecular
-and phenotype representations of drugs and proteins,
-% and
-% ontology features, yielding the deep neural network transformers
-% (DNNT) for both methods, and both drugs and proteins,
-% individually. 
-and add these to a protein interaction network as node features.
-% neural network in which nodes representing interactions
-% between proteins as node features, 
-We then train this model in an end-to-end manner. % Notably, like in the
-% pretraining phase, we rely on siamese neural networks, enforcing
-% similar representations for fitting drug-protein pairs. 
-The overall architecture is depicted in Figure~\ref{fig:Fullmodel}.
+We use two Siamese neural networks to combine the molecular and
+phenotype representations of drugs and proteins, and add these to a
+protein interaction network as node labels.  We then train this model
+in an end-to-end manner using a graph neural network architecture.
+The overall workflow\todo{make sure it's a workflow} is depicted in
+Figure~\ref{fig:Fullmodel}.
 
-% For a given protein, we want to find all suitable interacting drugs, \textbf{among a set of known drugs}.
-%To obtain a prediction for a given drug-protein pair,
-If
-
-Note that this approach can be easily generalized, profiting from
-other protein function and phenotype representation methods.
+% Note that this approach can be easily generalized, profiting from
+% other protein function and phenotype representation methods.
 

### 2021-02-07 paper/main.tex
Results restructuring

@@
 
-Through the very nature of the graph convolutional neural network, we build the transformed representation for all proteins in every forwarding step of the model.
-Note particularly, that the same convolutional filter and weight matrix are applied to and learned for all nodes simultaneously.
-By construction, for a single drug we can compute and predict all its interactors in a single run of the model, leading to significantly less computing time.
+We first perform a cross-validation over proteins and validate our
+results. Cross-validation over proteins aims to evaluate how the model
+performs when tasked to identify drugs that may target a ``novel''
+protein, i.e., one not seen during training.
+\todo[inline]{Describe/add the CV results here.}
 
-\subsection{\name identifies drugs that target a protein}
+Second, we use the model trained on STITCH and evaluate the model on
+the Yamanishi benchmark dataset\todo{This needs another sentence
+  stating how you made sure there was no overlap in data.}.
 
-\todo[inline]{Likely discussion:}
-The aim of \name is to predict candidate drugs that target a given
-protein; the challenge is to develop a training and evaluation scheme
-that does not simply overfit to the inherent biases in training and
-testing data.
-In general, when performing cross-validation for DTI prediction, the
-options are to split over 
-\begin{enumerate}
-	\item split over drugs,
-	\item split over drug--target pairs, or
-	\item split over proteins
-\end{enumerate}
-where the first and third option concern splitting drugs and proteins, respectively, into train, validation and test sets, and arranging the corresponding drug-target interactions.
-They ensure that at least parts of the interactions are not seen during training and evaluate either how well targets are predicted for unseen drugs or unseen proteins.
-Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model.
-\\
+We trained, validated and finally tested all considered models on the
+STITCH dataset using a 5-fold cross-validation over a protein
+split. We only evaluated the best performing model (with respect to
+\textit{AUROC} and $MicroAUC_p$) on the Yamanishi benchmark dataset.
 
-\todo[inline]{Likely discussion:}
-The most common scheme for DTI prediction is the split over drug-target pairs \citep{Survey2018}, where likely all drugs and targets of the validation and testing phase have already occurred in the training phase, as part of other drug-target pairs.
-The second most prevalent arrangement is the split over drugs, while only close to none is aiming on a protein split.
-However, the first and second splitting scheme are exposed to the first dataset bias and are hence more likely vulnerable to transductive inference by just predicting recently seen structures, rather than implementing inductive inference and generalizing over the drug representations.
-Second, these two strategies are more susceptible to the second bias, as only in these cases the model may overfit on the number of existing interactions for a single protein, while in the third scheme the number of interactions of the test proteins is entirely unknown during training process.
-
-Assuming a hypothetical, perfectly generalizing model built upon an unbiased dataset, this very model would yield similar performances for all three splitting schemes, not overfitting on the known structures.
-On the other hand, for a hypothetical entirely overfitting model, trained on a highly biased dataset, this model would show substantial deviations from the original performance over another split.
-\\
-We emphasize, that all real-world models are prone to some sort of overfitting, and unknown, deviant entities in both validation and testing set will likely lead to some sort of performance gap for relevant metrics.
-However, a large disparity may hint the biases stated above.
-
-Hence, we will perform a cross-validation over proteins for the training and prediction phase of our model, despite predicting per protein is rather counter-intuitive as there only limited drug-targets \citep{Overington2006}, and thus novel drugs are more likely to arise than novel targets.
-Yet, we aim to find all interacting drugs for existing targets motivating our split choice even further.
-
-\subsection{MicroAUC}
-Continuing this line, we apply the well metric \textit{MicroAUC} for a better evaluation for this purpose.
-Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively.
-As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
-
-\subsection{Experiments}
-In this section we will present the results of \name on both STITCH and Yamanishi benchmark dataset, omitting any kind of comparison for now.
-We trained, validated and finally tested all considered models on STITCH dataset using a 5-fold cross-validation over a protein split, as reasoned and substantiated previously.
-We only evaluated the best performing model, with respect to \textit{AUROC} and $MicroAUC_p$, on the Yamanishi benchmark dataset.
 We will hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.
@@
 \end{figure}
+
+% despite predicting per protein is
+% rather counter-intuitive as there only limited drug-targets
+% \citep{Overington2006}, and thus novel drugs are more likely to arise
+% than novel targets. Yet, we aim to find all interacting drugs for
+% existing targets motivating our split choice even further.
+
+\todo[inline]{Methods:}
+Through the very nature of the graph convolutional neural network, we
+build the transformed representation for all proteins in every
+forwarding step of the model. Note particularly, that the same
+convolutional filter and weight matrix are applied to and learned for
+all nodes simultaneously. By construction, for a single drug we can
+compute and predict all its interactors in a single run of the model,
+leading to significantly less computing time.
+
+%\subsection{\name identifies drugs that target a protein}
+
+\todo[inline]{Likely discussion:}
+The aim of \name is to predict candidate drugs that target a given
+protein; the challenge is to develop a training and evaluation scheme
+that does not simply overfit to the inherent biases in training and
+testing data.
+In general, when performing cross-validation for DTI prediction, the
+options are to split over 
+\begin{enumerate}
+	\item split over drugs,
+	\item split over drug--target pairs, or
+	\item split over proteins
+\end{enumerate}
+where the first and third option concern splitting drugs and proteins,
+respectively, into train, validation and test sets, and arranging the
+corresponding drug-target interactions. They ensure that at least
+parts of the interactions are not seen during training and evaluate
+either how well targets are predicted for unseen drugs or unseen
+proteins. Hereby, different training and prediction schemes lead to
+divergent expressiveness of the resulting model.
+
+\todo[inline]{Likely discussion:} The most common scheme for DTI
+prediction is the split over drug-target pairs \citep{Survey2018},
+where likely all drugs and targets of the validation and testing phase
+have already occurred in the training phase, as part of other
+drug-target pairs. The second most prevalent arrangement is the split
+over drugs, while only close to none is aiming on a protein split.
+However, the first and second splitting scheme are exposed to the
+first dataset bias and are hence more likely vulnerable to
+transductive inference by just predicting recently seen structures,
+rather than implementing inductive inference and generalizing over the
+drug representations. Second, these two strategies are more
+susceptible to the second bias, as only in these cases the model may
+overfit on the number of existing interactions for a single protein,
+while in the third scheme the number of interactions of the test
+proteins is entirely unknown during training process.
+
+\todo[inline]{Likely discussion:}
+Assuming a hypothetical, perfectly generalizing model built upon an
+unbiased dataset, this very model would yield similar performances for
+all three splitting schemes, not overfitting on the known
+structures. On the other hand, for a hypothetical entirely overfitting
+model, trained on a highly biased dataset, this model would show
+substantial deviations from the original performance over another
+split.
+
+\todo[inline]{Likely discussion:}
+We emphasize, that all real-world models are prone to some sort of
+overfitting, and unknown, deviant entities in both validation and
+testing set will likely lead to some sort of performance gap for
+relevant metrics. However, a large disparity may hint the biases
+stated above.
+
+\subsection{MicroAUC}
+Continuing this line, we apply the well metric \textit{MicroAUC} for a better evaluation for this purpose.
+Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively.
+As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
+
+\subsection{Experiments}
 

### 2021-02-07 paper/main.tex
Results restructuring

@@
 STITCH dataset using a 5-fold cross-validation over a protein
-split. We only evaluated the best performing model (with respect to
-\textit{AUROC} and $MicroAUC_p$) on the Yamanishi benchmark dataset.
+split; we then selected the best-performing models (with respect to
+\textit{AUROC} and $MicroAUC_p$) and evaluate them on the Yamanishi
+benchmark dataset.
 

### 2021-02-07 paper/main.tex
Results restructuring

@@
 
-We first perform a cross-validation over proteins and validate our
-results. Cross-validation over proteins aims to evaluate how the model
-performs when tasked to identify drugs that may target a ``novel''
-protein, i.e., one not seen during training.
-\todo[inline]{Describe/add the CV results here.}
-
-Second, we use the model trained on STITCH and evaluate the model on
-the Yamanishi benchmark dataset\todo{This needs another sentence
-  stating how you made sure there was no overlap in data.}.
-
-We trained, validated and finally tested all considered models on the
-STITCH dataset using a 5-fold cross-validation over a protein
-split; we then selected the best-performing models (with respect to
-\textit{AUROC} and $MicroAUC_p$) and evaluate them on the Yamanishi
-benchmark dataset.
-
-We will hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.
-As depicted in Figure \ref{tab:STITCHresults}, \textit{OntoPred} is showing significantly better performance on STITCH, while also only \textit{OntoPred} shows significant performance increase when adding \textit{GENConv} graph convolutional neural layers over the PPI graph.
-However, graph convolutional neural layers such as \textit{GCNConv} and \textit{GENConv}, especially when incorporated into the proposed \textit{ResGraphConv} blocks, add a lot of learnable parameters to the network, leading to more expressive power.
-This may lead to better performance, while deceiving, whether the respective features actually localize on the PPI graph.
-
-To eliminate these options we rebuild the graph model, whilst removing all graph convolutional neural layers in the residual blocks.
-This pruned network, with no information on the protein--protein interactions, reached very similarly results, as the original \textit{OntoPred} model.
-Secondly, when substituting the original \textit{GENConv} layers with \textit{GCNConv} or other related layers in the graph convolutional step of the ResGraphConv blocks, this model would achieve once again no significant performance gain in comparison to the plain \textit{OntoPred} one.
-This performance increase was only noticeable for \textit{GENConv} layers.
-The discrepancy of \textit{GENConv} and \textit{GCNConv} may be founded in numerical stability and fading signals, as described in the introduction of both methods.
-
-As this performance increase is quite significant, we \textbf{conclude that protein function does localize while molecular features do not (BETTER FORMULATION)}. ...
-
 \begin{figure}[!tpb]
@@
 \end{figure}
+
+We first perform a cross-validation over proteins and validate our
+results. Cross-validation over proteins aims to evaluate how the model
+performs when tasked to identify drugs that may target a ``novel''
+protein, i.e., one not seen during training.
+\todo[inline]{Describe/add the CV results here.}
+
+Second, we use the model trained on STITCH and evaluate the model on
+the Yamanishi benchmark dataset\todo{This needs another sentence
+  stating how you made sure there was no overlap in data.}.
+
+We trained, validated and finally tested all considered models on the
+STITCH dataset using a 5-fold cross-validation over a protein
+split; we then selected the best-performing models (with respect to
+\textit{AUROC} and $MicroAUC_p$) and evaluate them on the Yamanishi
+benchmark dataset. To evaluate the influence of the different
+features separately, and determine whether they ``localize'' on the
+PPI graph (and therefore can be exploited successfully by the graph
+neural networks), we train and evaluate models with different types of
+features, and with and without inclusion of the PPI graph,
+separately.
+Table \ref{tab:STITCHresults} shows the results.
+
+\todo[inline]{Move to caption of table.}
+We will hereby denote the molecular feature based predictor with
+\textit{MolPred}, while abbreviating the ontology based, top-down
+predictor with \textit{OntoPred}.
+
+We find that the model using ontology-based features
+(\textit{OntoPred}) is showing significantly\todo{significance test
+  and p value? otherwise remove} better performance on STITCH compared
+to using only molecular features. We also observe that only the model
+using ontology-based features results in significantly\todo{p-val}
+increased performance when incorporating the PPI graph, demonstrating
+that these features localize on the PPI network whereas the molecular
+features do not.
+
+, while also only \textit{OntoPred} shows significant performance increase when adding \textit{GENConv} graph convolutional neural layers over the PPI graph.
+However, graph convolutional neural layers such as \textit{GCNConv} and \textit{GENConv}, especially when incorporated into the proposed \textit{ResGraphConv} blocks, add a lot of learnable parameters to the network, leading to more expressive power.
+This may lead to better performance, while deceiving, whether the respective features actually localize on the PPI graph.
+
+To eliminate these options we rebuild the graph model, whilst removing all graph convolutional neural layers in the residual blocks.
+This pruned network, with no information on the protein--protein interactions, reached very similarly results, as the original \textit{OntoPred} model.
+Secondly, when substituting the original \textit{GENConv} layers with \textit{GCNConv} or other related layers in the graph convolutional step of the ResGraphConv blocks, this model would achieve once again no significant performance gain in comparison to the plain \textit{OntoPred} one.
+This performance increase was only noticeable for \textit{GENConv} layers.
+The discrepancy of \textit{GENConv} and \textit{GCNConv} may be founded in numerical stability and fading signals, as described in the introduction of both methods.
+
+As this performance increase is quite significant, we \textbf{conclude that protein function does localize while molecular features do not (BETTER FORMULATION)}. ...
 

### 2021-02-07 paper/main.tex
Results restructuring

@@
 using ontology-based features results in significantly\todo{p-val}
-increased performance when incorporating the PPI graph, demonstrating
-that these features localize on the PPI network whereas the molecular
-features do not.
+increased performance when incorporating the PPI graph.  This increase
+can be observed with different graph neural network
+architectures\todo{data not in the table}. While the {\em GCNConv} and
+{\em GENConv} architecture already shows an improvement, the
+\textit{ResGraphConv} blocks add a large amount of additional
+learnable parameters to the network, leading to more expressive
+power. To test whether the improvement is due to the number of
+learnable parameters added or the result of better exploiting the
+information about PPIs, we experiment with a graph model in which all
+graph convolutional neural layers in the residual blocks are
+removed\todo{is this still a graph model?}. This pruned network, with
+no information on the protein--protein interactions, reached very
+similarly results as the original \textit{OntoPred} model\todo{data
+  not shown; and details missing from Methods} and showed no
+improvement.
 
-, while also only \textit{OntoPred} shows significant performance increase when adding \textit{GENConv} graph convolutional neural layers over the PPI graph.
-However, graph convolutional neural layers such as \textit{GCNConv} and \textit{GENConv}, especially when incorporated into the proposed \textit{ResGraphConv} blocks, add a lot of learnable parameters to the network, leading to more expressive power.
-This may lead to better performance, while deceiving, whether the respective features actually localize on the PPI graph.
-
-To eliminate these options we rebuild the graph model, whilst removing all graph convolutional neural layers in the residual blocks.
-This pruned network, with no information on the protein--protein interactions, reached very similarly results, as the original \textit{OntoPred} model.
-Secondly, when substituting the original \textit{GENConv} layers with \textit{GCNConv} or other related layers in the graph convolutional step of the ResGraphConv blocks, this model would achieve once again no significant performance gain in comparison to the plain \textit{OntoPred} one.
+\todo[inline]{Not clear, need to reformulate:}
+Furthermore, when substituting the original \textit{GENConv} layers with \textit{GCNConv} or other related layers in the graph convolutional step of the ResGraphConv blocks, this model would achieve once again no significant performance gain in comparison to the plain \textit{OntoPred} one.
 This performance increase was only noticeable for \textit{GENConv} layers.
@@
 
-As this performance increase is quite significant, we \textbf{conclude that protein function does localize while molecular features do not (BETTER FORMULATION)}. ...
+As the inclusion of graph information significantly\todo{p val}
+increases performance, we conclude that information about protein
+functions localizes on the graph while molecular features do not.
 

### 2021-02-08 paper/main.tex
microAUC

@@
 
+\todo[inline]{Maybe remove microAUC here, introduce in next section?}
 We trained, validated and finally tested all considered models on the
@@
 % existing targets motivating our split choice even further.
+
+\subsection{Protein-centric evaluation}
+The goal of \name is to find candidate drugs that target a specific
+protein; however, our evaluation does not evaluate this application
+but rather how \name would perform in finding plausible drug--target
+interactions among all possible interactions. To provide a better
+estimate on how \name performs for individual protein targets, we use
+micro-averages between proteins and compute the \textit{MicroAUC};
+micro-averages average the performance (true and false positive rates)
+per protein instead of across all drug--protein pairs.
+
+% Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively.
+As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
+
+% \subsection{Baseline model}
+Furthermore, we hypothesize that some of the predictions made by our
+model are the result of the biases in drug--target interaction
+datasets. We design a ``na\"ive'' baseline model that predicts the
+same list of proteins for each drugs based on the number of
+known drug--target interactions for a protein. Formally, given lists
+$D$ and $P$ of drugs and proteins and a set of known interactions
+$\mathcal{I} := \{(d_i, p_i) \}$, we construct an interaction matrix
+$M_{int}\in\{0,1\}^{|D|\times|P|}$ with
+\begin{equation*}
+	M_{ij} = \begin{cases}
+		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
+		0 & otherwise
+	\end{cases}
+\end{equation*}
+describing for all drug--protein pairs whether there is a known
+interaction or not. We now rank all proteins $p_j\in P$ descending by
+their number of drug interactors by summing over the columns of
+$M_{ij}$ and ranking these sums:
+\begin{equation*}
+	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
+\end{equation*}
+Our ``na\"ive'' predictor $P_k$  predicts all drugs to interact with
+the top $k$ targets
+% , denoted by $Top_k(P)$ 
+with respect to
+% $M_{ij}$ from
+the introduced ranking:
+\begin{equation*}
+	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
+		1 & \text{ if }p_j \in Top_{k}(P)\\
+		0 & otherwise
+	\end{cases}
+\end{equation*}
+with the only hyperparameter $k$.
+The prediction $P_k$ is not dependent on the drug $d_i$ and will
+predict the same ranked list of drugs for a given protein $p_j$;
+consequently, this na\"ive predictor does not predict any useful
+information about interactions between drugs and proteins.
+% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
+
+We evaluate this naive predictor on both STITCH and Yamanishi dataset, first running on the whole dataset followed by a 5-fold cross-validation (CV) over both drugs and drug--protein pairs.
+Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown.
+For each fold, we gradually increase $k$, eventually saving the best performance for each fold.
+The corresponding AUROC scores are summarized in the following table: \\
+
+\centerline{\begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
+	\hline
+	Dataset&\multicolumn{3}{c|}{Splitting scheme}\\
+	&Whole&DT pairs&Drugs\\
+	\hline
+	STITCH& $0.76$& $0.70$ & $0.73$\\
+	\hline
+	Yamanishi& $0.88$& $0.84$&$0.85$\\ 
+	\hline
+\end{tabular}}
+\vspace{0.5cm}
+The baseline predictor shows diverging performances on both datasets with remarkable results on the Yamanishi benchmark.
+As the baseline predictor is based on the second dataset bias i.e.
+the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
+
+\subsection{Comparison}
+
+In this section we will compare \name with the described baseline method and various other cutting-edge models.
+We therefor evaluate all models on their recommended splitting scheme choice, hyperparameters and fold amoung for cross-validation, measuring their respective AUROC score.
+We eventually assess each model, by performing a protein-wise cross-validation determining (macro) AUROC and $MicroAUC_p$.
+For this we allow stratification for the training process, but not for the validation and testing phase, as real world scenarios are rather imbalanced.
+This however, only impacts the not considered area under precision recall curve (AUPRC) and not the Macro AUROC score, while also supporting the expressiveness of $MicroAUC_p$ with more data points.
+
+For our comparison we chose the best performing methods for drug--target interaction prediction evaluating on the Yamanishi benchmark dataset.
+These state-of-the-art methods approaches include  DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed superior results in comparison to numerous works.
+Furthermore, we added DTINet \cite{DTINet2017}, which lays the foundation for a number of methods, such as NeoDTI \cite{NeoDTI2019}, with similar methodology.
+
+The results of this first analysis are summarized in the upper part of Figure \ref{tab:Results}.
+
+\begin{figure}[!tpb]%figure1
+
+	\centerline{\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+		\hline
+		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
+		\hline
+		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
+		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
+		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
+		\hline
+		\hline
+		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
+		\hline
+		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
+		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
+		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
+		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
+		\hline
+	\end{tabular}}
+	\caption{Test}
+	\label{tab:Results}
+\end{figure}
 
@@
 stated above.
-
-\subsection{MicroAUC}
-Continuing this line, we apply the well metric \textit{MicroAUC} for a better evaluation for this purpose.
-Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively.
-As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.
-
-\subsection{Experiments}
-
-\subsection{Baseline model}
-Before comparing \name to other methods, we propose a suitable naive baseline model in order to analyze and understand existing approaches and \name for drug target interaction prediction.
-For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, we construct an interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
-\begin{equation*}
-	M_{ij} = \begin{cases}
-		1 & \text{if } (d_i, p_j)\in Int\\
-		0 & otherwise
-	\end{cases}
-\end{equation*}
-describing for all drug--protein whether there is a known interaction or not. We now rank all proteins $p_j\in P$ descending by their number of drug interactors, characterized by 
-\begin{equation*}
-	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
-\end{equation*}
-by summing over the columns of $M_{ij}$ and ranking these sums.
-We now finish our baseline predictor $P_k$ by predicting all drugs to interact with the top $k$ targets, denoted by $TopK(P)$ w.r.t. $M_{ij}$ from the previously introduced ranking, formalized by
-
-\begin{equation*}
-	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
-		1 & \text{ if }p_j \in TopK(P)\\
-		0 & otherwise
-	\end{cases}
-\end{equation*}
-with the only hyperparameter $k$.
-Note particularly, that the consequent prediction of $P_k$ is not dependent on the considered drug $d_i$, and will thus predict all drugs similarly for a given protein $p_j$.
-Subsequently, this naive predictor is not yielding any valuable information on the individual interactions with drugs for a given protein.
-Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
-
-We evaluate this naive predictor on both STITCH and Yamanishi dataset, first running on the whole dataset followed by a 5-fold cross-validation (CV) over both drugs and drug--protein pairs.
-Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown.
-For each fold, we gradually increase $k$, eventually saving the best performance for each fold.
-The corresponding AUROC scores are summarized in the following table: \\
-
-\centerline{\begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-	\hline
-	Dataset&\multicolumn{3}{c|}{Splitting scheme}\\
-	&Whole&DT pairs&Drugs\\
-	\hline
-	STITCH& $0.76$& $0.70$ & $0.73$\\
-	\hline
-	Yamanishi& $0.88$& $0.84$&$0.85$\\ 
-	\hline
-\end{tabular}}
-\vspace{0.5cm}
-The baseline predictor shows diverging performances on both datasets with remarkable results on the Yamanishi benchmark.
-As the baseline predictor is based on the second dataset bias i.e.
-the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
-
-\subsection{Comparison}
-
-In this section we will compare \name with the described baseline method and various other cutting-edge models.
-We therefor evaluate all models on their recommended splitting scheme choice, hyperparameters and fold amoung for cross-validation, measuring their respective AUROC score.
-We eventually assess each model, by performing a protein-wise cross-validation determining (macro) AUROC and $MicroAUC_p$.
-For this we allow stratification for the training process, but not for the validation and testing phase, as real world scenarios are rather imbalanced.
-This however, only impacts the not considered area under precision recall curve (AUPRC) and not the Macro AUROC score, while also supporting the expressiveness of $MicroAUC_p$ with more data points.
-
-For our comparison we chose the best performing methods for drug--target interaction prediction evaluating on the Yamanishi benchmark dataset.
-These state-of-the-art methods approaches include  DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed superior results in comparison to numerous works.
-Furthermore, we added DTINet \cite{DTINet2017}, which lays the foundation for a number of methods, such as NeoDTI \cite{NeoDTI2019}, with similar methodology.
-
-The results of this first analysis are summarized in the upper part of Figure \ref{tab:Results}.
-
-\begin{figure}[!tpb]%figure1
-
-	\centerline{\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-		\hline
-		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
-		\hline
-		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
-		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
-		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
-		\hline
-		\hline
-		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
-		\hline
-		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
-		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
-		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
-		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
-		\hline
-	\end{tabular}}
-	\caption{Test}
-	\label{tab:Results}
-\end{figure}
 

### 2021-02-08 paper/main.tex
microAUC

@@
 
-We evaluate this naive predictor on both STITCH and Yamanishi dataset, first running on the whole dataset followed by a 5-fold cross-validation (CV) over both drugs and drug--protein pairs.
-Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown.
-For each fold, we gradually increase $k$, eventually saving the best performance for each fold.
-The corresponding AUROC scores are summarized in the following table: \\
+We apply this naive predictor on both the STITCH and Yamanishi
+datasets, using the full dataset as well as 5-fold cross-validation
+over drugs and drug--protein pairs.
+% Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown. 
+For each fold, we gradually increase $k$, eventually saving the best
+performance for each fold. Using the full dataset, drug--target split,
+and drug split, we obtain \todo[inline]{merge table results in
+  text}\todo{are these micro or macro AUC?}.
 
@@
 \end{tabular}}
-\vspace{0.5cm}
+
 The baseline predictor shows diverging performances on both datasets with remarkable results on the Yamanishi benchmark.

### 2021-02-08 paper/main.tex
microAUC

@@
 
-The baseline predictor shows diverging performances on both datasets with remarkable results on the Yamanishi benchmark.
-As the baseline predictor is based on the second dataset bias i.e.
+The baseline predictor shows significantly\todo{p-val} higher
+performance than a random classifier on both datasets, with better
+results on the Yamanishi benchmark.
+
+For comparison with the state of the art methods, we chose the best
+performing methods for drug--target interaction prediction that were
+previously evaluated on the Yamanishi benchmark dataset. These
+state-of-the-art methods approaches include  DTIGEMS+
+\cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed
+superior results in comparison to numerous works. Furthermore, we
+added DTINet \cite{DTINet2017} as method for comparison which has been
+used to develop a number of methods such as NeoDTI \cite{NeoDTI2019}
+with similar methodology.
+
+We evaluate all models on their recommended splitting scheme choice,
+hyperparameters and fold among for cross-validation, measuring their
+respective AUROC score. We eventually assess each model, by performing
+a protein-wise cross-validation determining (macro) AUROC and
+$MicroAUC_p$. For this we allow stratification for the training
+process, but not for the validation and testing phase, as real world
+scenarios are rather imbalanced. This however, only impacts the not
+considered area under precision recall curve (AUPRC) and not the Macro
+AUROC score, while also supporting the expressiveness of $MicroAUC_p$
+with more data points.
+
+%As the baseline predictor is based on the second dataset bias i.e.
 the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
@@
 \subsection{Comparison}
-
-In this section we will compare \name with the described baseline method and various other cutting-edge models.
-We therefor evaluate all models on their recommended splitting scheme choice, hyperparameters and fold amoung for cross-validation, measuring their respective AUROC score.
-We eventually assess each model, by performing a protein-wise cross-validation determining (macro) AUROC and $MicroAUC_p$.
-For this we allow stratification for the training process, but not for the validation and testing phase, as real world scenarios are rather imbalanced.
-This however, only impacts the not considered area under precision recall curve (AUPRC) and not the Macro AUROC score, while also supporting the expressiveness of $MicroAUC_p$ with more data points.
-
-For our comparison we chose the best performing methods for drug--target interaction prediction evaluating on the Yamanishi benchmark dataset.
-These state-of-the-art methods approaches include  DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed superior results in comparison to numerous works.
-Furthermore, we added DTINet \cite{DTINet2017}, which lays the foundation for a number of methods, such as NeoDTI \cite{NeoDTI2019}, with similar methodology.
 

### 2021-02-08 paper/main.tex
microAUC

@@
 We evaluate all models on their recommended splitting scheme choice,
-hyperparameters and fold among for cross-validation, measuring their
-respective AUROC score. We eventually assess each model, by performing
-a protein-wise cross-validation determining (macro) AUROC and
-$MicroAUC_p$. For this we allow stratification for the training
-process, but not for the validation and testing phase, as real world
-scenarios are rather imbalanced. This however, only impacts the not
-considered area under precision recall curve (AUPRC) and not the Macro
-AUROC score, while also supporting the expressiveness of $MicroAUC_p$
-with more data points.
+hyperparameters and folds in cross-validation, measuring their
+respective AUROC. We further evaluate each model by performing a
+protein-wise cross-validation determining (macro) AUROC and
+$MicroAUC_p$. For this evaluation we allow stratification for the
+training process but not for the validation and testing phase as
+real world applications of these models would have to deal with
+imbalanced data.
 
-%As the baseline predictor is based on the second dataset bias i.e.
-the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
-
-\subsection{Comparison}
-
-The results of this first analysis are summarized in the upper part of Figure \ref{tab:Results}.
+\todo[inline]{Merge the tables; add \name model to table to allow
+  DIRECT comparison:}
+The results of the analysis are summarized in the upper part of Figure
+\ref{tab:Results}.
 
@@
 \end{figure}
+
+\todo[inline]{Discussion:}
+The lack of stratification, only impacts the not
+considered area under precision recall curve (AUPRC) and not the Macro
+AUROC score, while also supporting the expressiveness of $MicroAUC_p$
+with more data points.
+
+%As the baseline predictor is based on the second dataset bias i.e.
+the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
+
+\subsection{Comparison}
 

### 2021-02-08 paper/main.tex
microAUC

@@
 
+We find that...\todo[inline]{summarize key findings from Table.}
+
+\todo[inline]{Replace/merge table:}
 \begin{figure}[!tpb]%figure1
@@
 
-\subsection{Comparison}
+\section{Discussion}
 
@@
 
-\section{Findings}
+%\section{Findings}
 
@@
 	\item using indications is like cheating, as not applicable for searching new drugs
-	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related[!tpb] drugs/proteins
+	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
 	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

### 2021-03-04 paper/main.tex
.

@@
 
+\todo[inline]{Introduction}
 Within DTI prediction, there are potential biases resulting from the
@@
 target a protein for which no drugs yet exist. Second, some proteins
-(which we will denote as \textit{hub proteins}) have significantly
-more known interactions with drugs than others. In the STITCH
-database, $5\%$ of the proteins have $40\%$ of the interactions, and
-similar distributions are present in the Yamanishi and Drugbank
-datasets \cite{Drugbank2007, Drugbank2017} datasets; preferentially
-predicting these proteins may increase predictive performance while
-again not reflecting the actual performance when applied to a new
-protein (e.g., a protein for which no interactions are known). These
-differences in the number of drugs targeting certain proteins may be
-the result of study bias where more ``valuable'' proteins have more
-drugs designed to target them due to their involvement in more common
-diseases (or diseases for which drugs can be more profitably
-marketed).
+(which we call \textit{hub proteins}) have significantly more known
+interactions with drugs than others. In the STITCH database, $5\%$ of
+the proteins have $40\%$ of the interactions, and similar
+distributions are present in the Yamanishi and Drugbank
+\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
+these proteins may increase predictive performance while not
+reflecting the actual performance when applied to a new protein (e.g.,
+a protein for which no interactions are known). These differences in
+the number of drugs targeting certain proteins may be the result of
+study bias where more ``valuable'' proteins have more drugs designed
+to target them due to their involvement in more common diseases (or
+diseases for which drugs can be more profitably marketed).
 
@@
 interactions. Specifically, given a protein, \name will identify and
-rank known drugs that likely target this protein. \name combines two
-types of features: structural information for drugs and proteins that
-can be used to determine if they may physically interact, and
-information about drug effects that may localize on an interaction
-network.  As structural features, \name uses structural
-representations of drugs from the SMILES transformer
+rank drugs that likely target this protein. \name combines two types
+of features: structural information for drugs and proteins that can be
+used to determine if the drug and protein physically interact, and
+information about phenotypic effects of drugs and changes in protein
+function that may ``localize'' on an interaction network (i.e.,
+neighboring nodes will share some of these features or be
+phenotypically similar).  As structural features, \name uses
+structural representations of drugs from the SMILES transformer
 \cite{SmilesTransformer} and representations of protein amino acid
@@
 representations of drug effects and protein functions using the
-ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
+ontology-based machine learning method DL2Vec \cite{DL2vec2020} and
+ontology-based annotations of drugs and proteins.
 
@@
 We use two Siamese neural networks to combine the molecular and
-phenotype representations of drugs and proteins, and add these to a
-protein interaction network as node labels.  We then train this model
-in an end-to-end manner using a graph neural network architecture.
-The overall workflow\todo{make sure it's a workflow} is depicted in
-Figure~\ref{fig:Fullmodel}.
+phenotype representations of drugs and proteins. Siamese neural
+networks are used to learn a kind of similarity between the molecular
+and phenotype representations. We add these similarities to a protein
+interaction network as node labels; for each drug, proteins receive
+two features (from the molecular and phenotype similarities of the
+drug to the protein).
+\todo[inline]{There is a missing sentence here explaining the graph
+  neural network, output, etc.}
+We then train this model in an end-to-end manner using a graph neural
+network architecture.  The overall workflow\todo{make sure it's a
+  workflow} is depicted in Figure~\ref{fig:Fullmodel}.
 

### 2021-03-04 paper/main.tex
.

@@
 	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
-	\caption{Full DTI prediction model based on the pretrained learnable feature transformations (LFT) for either molecular structure or ontology based features.
-The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
-}
+	\caption{Full DTI prediction model based on the pretrained
+          learnable feature transformations (LFT) for either molecular
+          structure or ontology based features. The transformed
+          protein representations are added to each corresponded
+          protein as node features for the graph convolutional
+          steps.}
 	\label{fig:Fullmodel}
@@
 
-We first perform a cross-validation over proteins and validate our
-results. Cross-validation over proteins aims to evaluate how the model
-performs when tasked to identify drugs that may target a ``novel''
-protein, i.e., one not seen during training.
-\todo[inline]{Describe/add the CV results here.}
+We evaluate our model's ability to identify drug--target interactions
+using different approaches and datasets.  First, we perform a
+cross-validation over proteins and validate our results. A
+cross-validation over proteins aims to evaluate how the model performs
+when tasked to identify drugs that may target a ``novel'' protein,
+i.e., one not seen during training, or a protein for which a drug that
+targets it should be predicted.  \todo[inline]{Describe/add the CV
+  results here.}
 
-Second, we use the model trained on STITCH and evaluate the model on
-the Yamanishi benchmark dataset\todo{This needs another sentence
-  stating how you made sure there was no overlap in data.}.
+Our model is trained on the STITCH database. We use the model trained
+on STITCH and evaluate the model on the Yamanishi benchmark
+dataset\todo{This needs another sentence stating how you made sure
+  there was no overlap in data.}.
 

### 2021-03-04 paper/main.tex
.

@@
 predict the same ranked list of drugs for a given protein $p_j$;
-consequently, this na\"ive predictor does not predict any useful
-information about interactions between drugs and proteins.
+consequently, this na\"ive predictor will not predict any useful
+information about interactions between drugs and proteins, but allows
+us to estimate the effect of potential biases in our evaluation
+dataset.
 % Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
@@
 % Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown. 
-For each fold, we gradually increase $k$, eventually saving the best
+For each fold, we gradually increase $k$ to determine the best
 performance for each fold. Using the full dataset, drug--target split,
 and drug split, we obtain \todo[inline]{merge table results in
-  text}\todo{are these micro or macro AUC?}.
+  text}\todo{are these micro or macro AUC? Add to caption!}.
 

### 2021-03-04 paper/main.tex
.

@@
 
-The baseline predictor shows significantly\todo{p-val} higher
+The baseline\todo{What is this? The naive one?} predictor shows significantly\todo{p-val} higher
 performance than a random classifier on both datasets, with better

### 2021-03-04 paper/main.tex
Updated full model all layers and some TODOs in methods

--- 2021-03-04 paper/main.tex
@@
 \subsection{Datasets}
-We obtain a dataset consisting of $\approx 11000$\todo{precise number}
-human proteins with over 170.000\todo{exact number} links from STRING
+We obtain a dataset consisting of 12884 human proteins with over 340627 links from STRING
 \citep{STRINGv10}. For the drug-target interactions, we
-fetched 137,000\todo{precise} links from the STITCH database
+fetched 229870 links from the STITCH database
 \citep{STITCHv5}. As both STRING and STITCH provide confidence scores
@@
 one existing target. Therefore, the intersection between these
-resources yields 1,160 drugs and 6,680 human proteins for the training
-phase. We provide links to and methods for obtaining and processing
+resources yields 1,428 drugs and 7,368 human proteins with 32212 interaction links for the training phase. We provide links to and methods for obtaining and processing
 the necessary data on Github.
@@
 For comparative evaluation, we use the gold standard dataset
-introduced by \citet{Yaminishi2008} \todo[inline]{consisting of XXXX
-  interactions between XXX drugs and XXX protein targets}.
+introduced by \citet{Yaminishi2008} consisting of 1923 interactions between 708 drugs and 1512 proteins.
 
@@
 
-\subsubsection{Siamese neural networks and modular learnable feature transformation}
+\subsubsection{Half-twin neural networks and modular learnable feature transformation}
 
 \todo[inline]{Looking at the code, I wonder if what is implemented is
-  actually a Siamese ANN; it does not look as if the weights are
+  actually a 
+   ANN; it does not look as if the weights are
   shared between the two networks, so looks more like a twin or
@@
 As we want to learn from the similarity of drug side effects and
-protein phenotypes, we use a deep Siamese neural network with
+protein phenotypes, we use a deep half-twin neural network with
 a contrastive loss using cosine similarity.
-A Siamese neural network aims to learn a similarity between two
+A half-twin neural network aims to learn a similarity between two
 embeddings.
@@
 separately. An example structure for both types of features can be
-found in Figure~\ref{fig:SiameseNetwork}.
+found in Figure~\ref{fig:HalfTwinNetwork}.
 
@@
 	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
-	\caption{Siamese network applied to molecular and DL2vec
+	\caption{Half-twin network applied to molecular and DL2vec
           features, utilizing deep learnable feature transformations
@@
           computing the cosine similarity.}
-	\label{fig:SiameseNetwork}
+	\label{fig:HalfTwinNetwork}
 \end{figure}
@@
 
+\begin{figure}[!tpb]%figure1
+	\centerline{\includegraphics[width=1\columnwidth]{figures/full_model_all_layers.png}}
+	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
+	\label{fig:FullModelAllLayers}
+\end{figure}
+
 \subsubsection{Hyperparameter tuning}
@@
 the considered subset the ratio
-\begin{equation*}
-	\frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
-\end{equation*}
+\begin{equation}
+	w:= \frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
+	\label{weight}
+\end{equation}
 consequently needing compensation in the computed loss function and
@@
 \end{equation}
-for a given prediction $x$ and target $y$. We average this loss among
+for a given prediction $x$ and target $y$, and positive weight $w$ defined by equation \eqref{weight}. We average this loss among
 all drug-protein pairs in the training set, leading to a stable
 environment for the used optimizing scheme \textit{Adam}
-\citep{Adam2014}. We implemented a 5-fold cross validation among the
-proteins. Furthermore, we used early stopping to detect plateaus in
-the training process.
+\citep{Adam2014}. We implemented a 5-fold cross validation split among the
+proteins. Furthermore, we used early stopping in
+the training process for optima detection.
 
@@
 depth of the neural network, optimizer, learning rate and layer types
-using an extensive, manual grid search. \todo[inline]{Specify range of
-  values searched here; even better if you have the intermediate
-  results and can put them here (better: in the supplement).}
+using an extensive, manual grid search, Starting from na\"ive, shallow feature transformations with an embedding size of 10, we scaled the network up to residual structures with up to 10 hidden layers leading to embeddings of size 4000, testing different network widths and learning rates for each configuration.
 
@@
 
-\todo[inline]{Needs a statement on how the PAIRS are ranked here and
-  TPR/FPR is calculated based on the ranks of the DT-PAIRS.}
 We calculate the AUROC by computing true positive rate at various
@@
 For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
-With respect to proteins, this can be formalized for given labels and prediction $l,y:D\times P \rightarrow \{0,1\}$  as
+With respect to proteins, this can be formalized for given labels $l:D\times P \rightarrow \{0,1\}$ and predictions $y:D\times P \rightarrow [0,1]$  as
 \begin{equation*}
@@
 
-Drugs and proteins can be interchanged in this formulation, and we
-refer to the different measures as a protein-centric microAUC
+Drugs and proteins can be interchanged in this formulation, while we
+refer to the different measures as protein-centric microAUC
 ($MicroAUC_p$) and a drug-centric microAUC
 ($MicroAUC_d$). Additionally, the $MicroAUC$ score may not be defined
-as, in some datasets some targets or drugs, respectively, do not have
-any interactions, leading to an infeasible $TPR$ and an undefined
-$AUROC$ score for that entity.
-\todo[inline]{Don't understand this, can you add the formula:} For
-those entities we impute the $MicroAUC$ linearly, by using the
-accuracy for this subset.
+as in some datasets some targets or drugs, respectively, do not have
+any interactions, leading to an infeasible $TPR=0$ for all thresholds and an undefined
+$AUROC$ score for that entity. For those entities we impute the $MicroAUC$ interpolating linearly, by using the accuracy for this subset.
 
@@
 
-\todo[inline]{Introduction}
 Within DTI prediction, there are potential biases resulting from the
@@
 target a protein for which no drugs yet exist. Second, some proteins
-(which we call \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$ of
-the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins may increase predictive performance while not
-reflecting the actual performance when applied to a new protein (e.g.,
-a protein for which no interactions are known). These differences in
-the number of drugs targeting certain proteins may be the result of
-study bias where more ``valuable'' proteins have more drugs designed
-to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed).
+(which we will denote as \textit{hub proteins}) have significantly
+more known interactions with drugs than others. In the STITCH
+database, $5\%$ of the proteins have $40\%$ of the interactions, and
+similar distributions are present in the Yamanishi and Drugbank
+datasets \cite{Drugbank2007, Drugbank2017} datasets; preferentially
+predicting these proteins may increase predictive performance while
+again not reflecting the actual performance when applied to a new
+protein (e.g., a protein for which no interactions are known). These
+differences in the number of drugs targeting certain proteins may be
+the result of study bias where more ``valuable'' proteins have more
+drugs designed to target them due to their involvement in more common
+diseases (or diseases for which drugs can be more profitably
+marketed).
 
@@
 interactions. Specifically, given a protein, \name will identify and
-rank drugs that likely target this protein. \name combines two types
-of features: structural information for drugs and proteins that can be
-used to determine if the drug and protein physically interact, and
-information about phenotypic effects of drugs and changes in protein
-function that may ``localize'' on an interaction network (i.e.,
-neighboring nodes will share some of these features or be
-phenotypically similar).  As structural features, \name uses
-structural representations of drugs from the SMILES transformer
+rank known drugs that likely target this protein. \name combines two
+types of features: structural information for drugs and proteins that
+can be used to determine if they may physically interact, and
+information about drug effects that may localize on an interaction
+network.  As structural features, \name uses structural
+representations of drugs from the SMILES transformer
 \cite{SmilesTransformer} and representations of protein amino acid
@@
 representations of drug effects and protein functions using the
-ontology-based machine learning method DL2Vec \cite{DL2vec2020} and
-ontology-based annotations of drugs and proteins.
+ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
 
@@
 
-We use two Siamese neural networks to combine the molecular and
-phenotype representations of drugs and proteins. Siamese neural
-networks are used to learn a kind of similarity between the molecular
-and phenotype representations. We add these similarities to a protein
-interaction network as node labels; for each drug, proteins receive
-two features (from the molecular and phenotype similarities of the
-drug to the protein).
-\todo[inline]{There is a missing sentence here explaining the graph
-  neural network, output, etc.}
-We then train this model in an end-to-end manner using a graph neural
-network architecture.  The overall workflow\todo{make sure it's a
-  workflow} is depicted in Figure~\ref{fig:Fullmodel}.
+We use two half-twin neural networks to combine the molecular and
+phenotype representations of drugs and proteins, and add these to a
+protein interaction network as node labels.  We then train this model
+in an end-to-end manner using a graph neural network architecture.
+The overall workflow\todo{make sure it's a workflow} is depicted in
+Figure~\ref{fig:Fullmodel}.
 
@@
 	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
-	\caption{Full DTI prediction model based on the pretrained
-          learnable feature transformations (LFT) for either molecular
-          structure or ontology based features. The transformed
-          protein representations are added to each corresponded
-          protein as node features for the graph convolutional
-          steps.}
+	\caption{Full DTI prediction model based on the pretrained learnable feature transformations (LFT) for either molecular structure or ontology based features.
+The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
+}
 	\label{fig:Fullmodel}
@@
 
-We evaluate our model's ability to identify drug--target interactions
-using different approaches and datasets.  First, we perform a
-cross-validation over proteins and validate our results. A
-cross-validation over proteins aims to evaluate how the model performs
-when tasked to identify drugs that may target a ``novel'' protein,
-i.e., one not seen during training, or a protein for which a drug that
-targets it should be predicted.  \todo[inline]{Describe/add the CV
-  results here.}
+We first perform a cross-validation over proteins and validate our
+results. Cross-validation over proteins aims to evaluate how the model
+performs when tasked to identify drugs that may target a ``novel''
+protein, i.e., one not seen during training.
+\todo[inline]{Describe/add the CV results here.}
 
-Our model is trained on the STITCH database. We use the model trained
-on STITCH and evaluate the model on the Yamanishi benchmark
-dataset\todo{This needs another sentence stating how you made sure
-  there was no overlap in data.}.
+Second, we use the model trained on STITCH and evaluate the model on
+the Yamanishi benchmark dataset\todo{This needs another sentence
+  stating how you made sure there was no overlap in data.}.
 

### 2021-03-04 paper/main.tex

@@
 
-\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
+\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}git
 \author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
@@
 
-The update rule for each individual node is
+The update rule for each individual node is given by a message passing scheme formalized by
 \begin{equation}
@@
 \end{figure}
+
+\subsubsection{title}
+Combining half-twin and graph convolutional neural networks, we map all protein representations to the respective node features, initializing the graph convolutional update steps.
+The resulting representations are used for a similarity prediction similar to the models presented in figure ~\ref{fig:HalfTwinNetwork}.
+When combining ontology and molecular features with or without the
+graph model, we concatenate both protein features and both drugs
+features, before plugging them into the graph for the similarity
+computation. A more in-depth image of the overall architecture, combining both feature types is depicted in figure ~\ref{fig:FullModelAllLayers}.
 
@@
 representations of drug effects and protein functions using the
-ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
+ontology-based machine learning method DL2Vec \cite{DL2vec2020} and
+ontology-based annotations of drugs and proteins.
 
-\todo[inline]{Move to methods:}
-When combining ontology and molecular features with or without the
-graph model, we concatenate both protein features and both drugs
-features, before plugging them into the graph or the similarity
-computation.
-
-We use two half-twin neural networks to combine the molecular and
-phenotype representations of drugs and proteins, and add these to a
-protein interaction network as node labels.  We then train this model
-in an end-to-end manner using a graph neural network architecture.
-The overall workflow\todo{make sure it's a workflow} is depicted in
-Figure~\ref{fig:Fullmodel}.
+We construct a PPI graph with proteins as nodes and protein-protein interactions as edges, mapping the protein features to each target as node features.
+\name then propagates information among the PPI network utilizing graph convolutional steps, calculating the similarity of drug and protein representation, and thus eventually predicting whether there is an interaction.
+The full workflow scheme is depicted in figure~\ref{fig:ModelWorkflow}
 
@@
 \begin{figure}[!tpb]
-	\centerline{\includegraphics[width=0.45\textwidth]{figures/full_model.png}}
-	\caption{Full DTI prediction model based on the pretrained learnable feature transformations (LFT) for either molecular structure or ontology based features.
-The transformed protein representations are added to each corresponded protein as node features for the graph convolutional steps.
-}
-	\label{fig:Fullmodel}
+	\centering
+	\includegraphics[width=0.9\columnwidth]{figures/model_workflow.png}
+	\caption{Full DTI prediction model based on the pretrained
+          learnable feature transformations (LFT) for either molecular
+          structure or ontology based features. The transformed
+          protein representations are added to each corresponded
+          protein as node features for the graph convolutional
+          steps.}
+	\label{fig:ModelWorkflow}
 \end{figure}
 
-\begin{figure}[!tpb]
-	\centerline{\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
+\begin{table}[!tpb]
+	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
 		\hline
@@
 		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
-		&AUROC&Micro $AUC_p$&AUROC&Micro $AUC_p$\\
+		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$\\
 		\hline
@@
 		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
-		&AUROC&Micro AUC&AUROC&Micro $AUC_p$\\
+		&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
 		\hline
@@
 		\hline
-	\end{tabular}}
-	\caption{Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation. MolPred and OntoPred are the predictors for molecular and ontology based features, respectively. }
+	\end{tabular} \medskip
+
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+		\hline
+		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
+		\hline
+		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
+		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
+		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
+		\hline
+		\hline
+		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
+		\hline
+		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
+		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
+		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
+		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
+		\hline
+	\end{tabular}\\
+
+	\caption{Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation.
+MolPred and OntoPred are the predictors for molecular and ontology based features, respectively.
+We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.}
 	\label{tab:STITCHresults}
-\end{figure}
+\end{table}
 
-We first perform a cross-validation over proteins and validate our
-results. Cross-validation over proteins aims to evaluate how the model
-performs when tasked to identify drugs that may target a ``novel''
-protein, i.e., one not seen during training.
+We evaluate our model's ability to identify drug--target interactions
+using different approaches and datasets. First, we perform a
+cross-validation over proteins and validate our results. A
+cross-validation over proteins aims to evaluate how the model performs
+when tasked to identify drugs that may target a ``novel'' protein,
+i.e., one not seen during training, or a protein for which a drug that
+targets it should be predicted.
+
 \todo[inline]{Describe/add the CV results here.}
-
-Second, we use the model trained on STITCH and evaluate the model on
-the Yamanishi benchmark dataset\todo{This needs another sentence
-  stating how you made sure there was no overlap in data.}.
-
-\todo[inline]{Maybe remove microAUC here, introduce in next section?}
+\begin{itemize}
+	\item 
+\end{itemize}
 We trained, validated and finally tested all considered models on the
@@
 split; we then selected the best-performing models (with respect to
-\textit{AUROC} and $MicroAUC_p$) and evaluate them on the Yamanishi
-benchmark dataset. To evaluate the influence of the different
+\textit{AUROC/MacroAUC} and $MicroAUC_p$) and evaluate them on in 5-fold protein split cross validation on the Yamanishi benchmark dataset to avoid validation overfitting and yield more realistic testing results.
+To evaluate the influence of the different
 features separately, and determine whether they ``localize'' on the
@@
 features, and with and without inclusion of the PPI graph,
-separately.
-Table \ref{tab:STITCHresults} shows the results.
-
-\todo[inline]{Move to caption of table.}
-We will hereby denote the molecular feature based predictor with
-\textit{MolPred}, while abbreviating the ontology based, top-down
-predictor with \textit{OntoPred}.
+separately, comparing the molecular (MolPred) and phenotype based (OntoPred) predictor and a combination of both, concatenating both features.
+Table \ref{tab:STITCHresults} shows the results of these experiments.
 
 We find that the model using ontology-based features
-(\textit{OntoPred}) is showing significantly\todo{significance test
-  and p value? otherwise remove} better performance on STITCH compared
+(\textit{OntoPred}) is showing better performance on STITCH compared
 to using only molecular features. We also observe that only the model
-using ontology-based features results in significantly\todo{p-val}
+using ontology-based features results in
 increased performance when incorporating the PPI graph. This increase
 can be observed with different graph neural network
-architectures\todo{data not in the table}. While the {\em GCNConv} and
-{\em GENConv} architecture already shows an improvement, the
+architectures and configurations. While the {\em GCNConv} and
+{\em GENConv} architecture already shows some minor improvement, the
 \textit{ResGraphConv} blocks add a large amount of additional
@@
 graph convolutional neural layers in the residual blocks are
-removed\todo{is this still a graph model?}. This pruned network, with
+removed, resulting in a non--graph model. This pruned network, with
 no information on the protein--protein interactions, reached very
-similarly results as the original \textit{OntoPred} model\todo{data
+similar results to the original \textit{OntoPred} model\todo{data
   not shown; and details missing from Methods} and showed no
@@
 
-\todo[inline]{Not clear, need to reformulate:}
-Furthermore, when substituting the original \textit{GENConv} layers with \textit{GCNConv} or other related layers in the graph convolutional step of the ResGraphConv blocks, this model would achieve once again no significant performance gain in comparison to the plain \textit{OntoPred} one.
-This performance increase was only noticeable for \textit{GENConv} layers.
-The discrepancy of \textit{GENConv} and \textit{GCNConv} may be founded in numerical stability and fading signals, as described in the introduction of both methods.
+Furthermore, these major improvements are only provided by the \textit{GENConv} graph convolution scheme as part of the ResGraphConv blocks, as \textit{GCNConv} and other related graph convolutional methods fail to achieve any gain in comparison to the plain \textit{OntoPred} performance even when combined with the residual blocks.
+The discrepancy of {\em GENConv} and other graph convolutional methods may be founded in numerical stability and fading signals, as described in the introduction of the {\em GENConv} method.
 
-As the inclusion of graph information significantly\todo{p val}
-increases performance, we conclude that information about protein
+As the inclusion of graph information increases performance, we conclude that information about protein
 functions localizes on the graph while molecular features do not.
@@
 over drugs and drug--protein pairs.
-% Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown. 
-For each fold, we gradually increase $k$, eventually saving the best
+Note, that this baseline predictor is not applicable for a protein split CV, as the number of interactions for each protein in the validation set is unknown. 
+For each fold, we gradually increase $k$ to determine the best
 performance for each fold. Using the full dataset, drug--target split,
-and drug split, we obtain \todo[inline]{merge table results in
-  text}\todo{are these micro or macro AUC?}.
+and drug split, we obtain the following MacroAUC/AUROC score results.
+For STITCH database we obtain a performance of $0.76$ on the whole dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the drug splitting scheme.
+While these results are fairly moderate, on Yamanishi dataset we yield MacroAUC scores of $0.88$, $0.84$ and $0.85$, for total dataset, DT pair split and drug split, respectively.
 
-\centerline{\begin{tabular}{|l|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-	\hline
-	Dataset&\multicolumn{3}{c|}{Splitting scheme}\\
-	&Whole&DT pairs&Drugs\\
-	\hline
-	STITCH& $0.76$& $0.70$ & $0.73$\\
-	\hline
-	Yamanishi& $0.88$& $0.84$&$0.85$\\ 
-	\hline
-\end{tabular}}
-
-The baseline predictor shows significantly\todo{p-val} higher
-performance than a random classifier on both datasets, with better
-results on the Yamanishi benchmark.
+The na\"ive predictor shows great performance on Yamanishi dataset, with substantial gain in comparison to a random predictor on both datasets.
+In the following we will denote this nai\"ve predictor as a baseline predictor correlating its performance to state of the art models and \name.
 
@@
 performing methods for drug--target interaction prediction that were
-previously evaluated on the Yamanishi benchmark dataset. These
-state-of-the-art methods approaches include  DTIGEMS+
+previously evaluated on the Yamanishi benchmark dataset. These methods include DTIGEMS+
 \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed
@@
 respective AUROC. We further evaluate each model by performing a
-protein-wise cross-validation determining (macro) AUROC and
+protein-wise cross-validation determining Macro AUC/AUROC score and
 $MicroAUC_p$. For this evaluation we allow stratification for the
 training process but not for the validation and testing phase as
-real world applications of these models would have to deal with
+real world applications of these models would have to deal with possibly
 imbalanced data.
 
-\todo[inline]{Merge the tables; add \name model to table to allow
+\todo[inline]{ add \name model to table to allow
   DIRECT comparison:}
@@
 
-	\centerline{\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-		\hline
-		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
-		\hline
-		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
-		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
-		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
-		\hline
-		\hline
-		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&AUROC score&AUROC score&Micro $AUC_p$\\
-		\hline
-		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
-		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
-		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
-		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
-		\hline
-	\end{tabular}}
+	\centerline{}
 	\caption{Test}

### 2021-03-09 paper/main.tex

@@
 
-\todo[inline]{Describe/add the CV results here.}
-\begin{itemize}
-	\item 
-\end{itemize}
 We trained, validated and finally tested all considered models on the
@@
 to using only molecular features. We also observe that only the model
-using ontology-based features results in
-increased performance when incorporating the PPI graph. This increase
-can be observed with different graph neural network
-architectures and configurations. While the {\em GCNConv} and
-{\em GENConv} architecture already shows some minor improvement, the
-\textit{ResGraphConv} blocks add a large amount of additional
-learnable parameters to the network, leading to more expressive
-power. To test whether the improvement is due to the number of
-learnable parameters added or the result of better exploiting the
+using ontology-based features results in increased performance when
+incorporating the PPI graph. This increase can be observed with
+different graph neural network architectures and configurations. While
+the {\em GCNConv} and {\em GENConv} architecture already shows some
+minor improvement, the \textit{ResGraphConv} blocks add a large amount
+of additional learnable parameters to the network, leading to more
+expressive power. To test whether the improvement is due to the number
+of learnable parameters added or the result of better exploiting the
 information about PPIs, we experiment with a graph model in which all
-graph convolutional neural layers in the residual blocks are
-removed, resulting in a non--graph model. This pruned network, with
-no information on the protein--protein interactions, reached very
-similar results to the original \textit{OntoPred} model\todo{data
-  not shown; and details missing from Methods} and showed no
+graph convolutional neural layers in the residual blocks are removed,
+resulting in a non--graph model. This pruned network, with no
+information on the protein--protein interactions, reached very similar
+results to the original \textit{OntoPred} model and showed no
 improvement.

### 2021-03-15 paper/main.tex
Update results section for DTA methods

@@
 
-\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}git
+\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
 \author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
@@
 		\hline
-		STITCH results&\multicolumn{4}{c|}{PPI graph}\\
+		(a) STITCH results&\multicolumn{4}{c|}{PPI graph}\\
 		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
@@
 		\hline
-		MolPred + OntoPred & $0.89$ & $0.90$&$0.93$&$0.94$\\
+		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$\\
 		\hline
 		\hline
-		Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
+		(b) Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
 		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
@@
 		\hline
-		MolPred + OntoPred & $0.83$ & $0.82$&$0.84$&$0.84$\\
+		\name (MolPred + OntoPred) & $0.83$ & $0.82$&$0.84$&$0.84$\\
 		\hline
@@
 		\hline
-		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		(c) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
 		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
 		\hline
-		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
-		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
-		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
+		Naive predictor&DP pairs&$0.85$& --&--\\
+		DTINet&DP pairs&$0.91$&$0.74$&$0.67$\\
+		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
+		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
+		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
 		\hline
 		\hline
-		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
+		(d) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
 		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
@@
 
-	\caption{Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation.
-MolPred and OntoPred are the predictors for molecular and ontology based features, respectively.
-We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.}
-	\label{tab:STITCHresults}
+	\caption{(a) + (b) Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation.
+We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.
+(c) + (d) Results for various state of the art (c) drug--target interaction prediction methods on Yamanishi dataset and (d) drug--target affinity prediction methods on BIOSNAP dataset, evaluated on their original and the protein cross-validation splitting scheme, approximately reproducing the results of MolTrans \textbf{CITATION}.}
+	\label{tab:Results}
 \end{table}
@@
 separately, comparing the molecular (MolPred) and phenotype based (OntoPred) predictor and a combination of both, concatenating both features.
-Table \ref{tab:STITCHresults} shows the results of these experiments.
+Table \ref{tab:Results} (a)+(b) shows the results of these experiments.
 
@@
 The na\"ive predictor shows great performance on Yamanishi dataset, with substantial gain in comparison to a random predictor on both datasets.
-In the following we will denote this nai\"ve predictor as a baseline predictor correlating its performance to state of the art models and \name.
+In the following, we will utilize this nai\"ve predictor as baseline predictor correlating its performance to state of the art models and \name.
 
@@
 
+The results of this analysis are summarized in part (c) of Table \ref{tab:Results}, calculating the performance of all considered methods over their original splitting scheme and over a protein split.
+We find that, first the majority of approaches evaluate over a drug--protein splitting scheme, while second, the highest performances on Yamanishi dataset are yielded over the drug--target pairs splitting scheme, with a noticeable gap to the methods performing their splits over proteins.
+Further, when evaluating the same methods over a protein split, we find a substantial performance divergence in comparison to their original splitting scheme.
+The authors of DTI-CDF evaluate their method on all three splitting schemes underlining this point.
+While \name provides comparable performance to DTI-CDF, it yields considerably better with respect to $MicroAUC_p$.
+
+As we were quite surprised about the MacroAUC discrepancy between the splitting schemes, we evaluated other cutting edge drug--target interaction and drug--target affinity prediction methods, training and evaluating on other datasets.
+Following the results of MolTrans (\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}), DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and MolTrans itself on BioSnap (\textbf{CITATION}) over a protein split, as shown in part (d) of Table \ref{tab:Results}.
+The authors of MolTrans evaluated all considered methods over the drug--target pair and the protein split, which we were able to reproduce with (d), showing once more a substantial divergence.
+However, we additionally computed the $MicroAUC_p$ score for all considered methods, leading to comparable scores as for the Yamanishi evaluated methods.
+
 \todo[inline]{ add \name model to table to allow
@@
 \todo[inline]{Replace/merge table:}
-\begin{figure}[!tpb]%figure1
-
-	\centerline{}
-	\caption{Test}
-	\label{tab:Results}
-\end{figure}
 
@@
 unbiased dataset, this very model would yield similar performances for
-all three splitting schemes, not overfitting on the known
+all three , not overfitting on the known
 structures. On the other hand, for a hypothetical entirely overfitting

### 2021-03-18 paper/main.tex
.

--- 2021-03-09 paper/main.tex
@@
 effects or its impact on diseases is seldom given for novel drugs.
+
+\todo[inline]{Introduction}
+Within DTI prediction, there are potential biases resulting from the
+underlying datasets \citep{Pahikkala2014}. First, novel drugs are
+often designed by altering non-functional components of a drug,
+leading to two and more very similar drugs designed to target the same
+proteins \cite{Overington2006}. This can result in a bias when it
+leads to hidden duplicates that can distribute among the train/test
+split, resulting in a better (measured) predictive performance than
+would be expected when the model is applied to identify drugs that
+target a protein for which no drugs yet exist. Second, some proteins
+(which we call \textit{hub proteins}) have significantly more known
+interactions with drugs than others. In the STITCH database, $5\%$ of
+the proteins have $40\%$ of the interactions, and similar
+distributions are present in the Yamanishi and Drugbank
+\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
+these proteins may increase predictive performance while not
+reflecting the actual performance when applied to a new protein (e.g.,
+a protein for which no interactions are known). These differences in
+the number of drugs targeting certain proteins may be the result of
+study bias where more ``valuable'' proteins have more drugs designed
+to target them due to their involvement in more common diseases (or
+diseases for which drugs can be more profitably marketed).
 
@@
 
-\todo[inline]{Introduction}
-Within DTI prediction, there are potential biases resulting from the
-underlying datasets \citep{Pahikkala2014}. First, novel drugs are
-often designed by altering non-functional components of a drug,
-leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can result in a bias when it
-leads to hidden duplicates that can distribute among the train/test
-split, resulting in a better (measured) predictive performance than
-would be expected when the model is applied to identify drugs that
-target a protein for which no drugs yet exist. Second, some proteins
-(which we call \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$ of
-the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins may increase predictive performance while not
-reflecting the actual performance when applied to a new protein (e.g.,
-a protein for which no interactions are known). These differences in
-the number of drugs targeting certain proteins may be the result of
-study bias where more ``valuable'' proteins have more drugs designed
-to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed).
-
 We developed \name as a computational model to predict drug--target
@@
 function that may ``localize'' on an interaction network (i.e.,
-neighboring nodes will share some of these features or be
+neighboring nodes will share some of these features or are
 phenotypically similar).  As structural features, \name uses
@@
 
-We construct a PPI graph with proteins as nodes and protein-protein interactions as edges, mapping the protein features to each target as node features.
-\name then propagates information among the PPI network utilizing graph convolutional steps, calculating the similarity of drug and protein representation, and thus eventually predicting whether there is an interaction.
-The full workflow scheme is depicted in figure~\ref{fig:ModelWorkflow}
+We construct a PPI graph with proteins as nodes and protein-protein
+interactions as edges, mapping the protein features to each target as
+node features. \name then propagates information among the PPI network
+utilizing graph convolutional steps, calculating the similarity of
+drug and protein representation, and thus eventually predicting
+whether there is an interaction. The full workflow scheme is depicted
+in figure~\ref{fig:ModelWorkflow}
 

### 2021-03-18 paper/main.tex
.

@@
 \todo[inline]{Looking at the code, I wonder if what is implemented is
-  actually a 
-   ANN; it does not look as if the weights are
-  shared between the two networks, so looks more like a twin or
-  half-twin network? Siamese would be sim(f(e1),f(e2)) but the
-  implementation seems to be sim(f(e1), g(e2)).}
+  actually a ANN; it does not look as if the weights are shared
+  between the two networks, so looks more like a twin or half-twin
+  network? Siamese would be sim(f(e1),f(e2)) but the implementation
+  seems to be sim(f(e1), g(e2)).}
 
 As we want to learn from the similarity of drug side effects and
-protein phenotypes, we use a deep half-twin neural network with
-a contrastive loss using cosine similarity.
-A half-twin neural network aims to learn a similarity between two
-embeddings.
-
-% learning a high-dimensional embedding emphasizing this identity by
-% forcing a similarity between these embeddings. On the other hand we
-% built a deep neural network for the molecular structure based
-% features, also benefiting from the siamese network
-% architecture. Computing the similarity between two representations,
-% allows for a variety of different methods.  
-% However, we decided for
-% the cosine similarity measure being invariant to scaling.
-
-Therefore, the precomputed embeddings are run through a regular,
-neural learnable feature transformation (LFT) network, which also
-reduces the eventual representation size for drugs and proteins
-separately. An example structure for both types of features can be
-found in Figure~\ref{fig:HalfTwinNetwork}.
-
-While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we emphasize the versatility of this approach, as both ontology and molecular feature for both drugs and proteins are reduced to similar dimensionality.
-This allows for a high amount of modularity and different experimental setups by plugging different kinds of features into the model.
-Additionally, these pretrained features can be used for a variety of other tasks.
-Additionally, the ontology LFT can be reused for a variety of DL2vec based features with respect to other ontologies and hypotheses.
-We hereby followed the results of \textit{DL2vec}, indicating that utilizing the activation function $\sigma := \mathrm{LeakyReLU}$ leads to performance increase.
+protein phenotypes, we use a deep half-twin neural network with a
+contrastive loss using cosine similarity.  A half-twin neural network
+aims to learn a similarity between two embeddings of the same
+dimension.  The precomputed embeddings are first processed by a
+learnable feature transformation (LFT) network, a fully connected
+neural network layer which takes as input an embedding and outputs a
+representation of a particular size. An example structure for both
+types of features can be found in Figure~\ref{fig:HalfTwinNetwork}. We
+use this LFT to reduce the eventual representation size for drugs and
+proteins separately.
+% While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we
+The use of an LFT enables flexible experimentation as both ontology
+and molecular feature for both drugs and proteins are reduced to
+similar dimensionality for varying sizes of inputs; this allows for a
+high amount of modularity across different experimental setups by
+adding different kinds of features into the model. Additionally, the
+generated features can be used for a variety of other tasks; for
+example, the ontology LFT can be reused for a variety of DL2vec-based
+features with respect to other ontologies and hypotheses. We follow
+the results of \textit{DL2vec} showing that the activation function
+$\sigma := \mathrm{LeakyReLU}$ leads to improved performance compared
+to other activation functions.
 

### 2021-03-20 paper/main.tex
.

@@
 \subsection{Problem Description}
-The issue of predicting drug-target interactions can be described
-quite briefly: For a given drug and a given protein we want to
-determine whether those interact or not.  We do not differentiate
-between types of interaction such as activation and inhibition, and do
-not predict the strength of the interaction.  If we additionally make
-the closed world assumption, i.e., assume that our knowledge is
-complete and all drug--protein pairs without a known interaction do
-not interact, we can formulate the problem as a binary classification
-task.
+\name aims to solve the following problem: for a given drug and a
+given protein we want to determine whether those interact or not.  We
+do not differentiate between types of interaction such as activation
+and inhibition, and do not predict the strength of the interaction.
+If we additionally assume that our knowledge is complete and all
+drug--protein pairs without a known interaction do not interact, we
+can formulate the problem as a binary classification task.
 
 \subsection{Datasets}
-We obtain a dataset consisting of 12884 human proteins with over 340627 links from STRING
-\citep{STRINGv10}. For the drug-target interactions, we
-fetched 229870 links from the STITCH database
+We obtain a dataset consisting of 12,884 human proteins with over
+340627 links from STRING \citep{STRINGv10}. For the drug-target
+interactions, we use 229,870 links from the STITCH database
 \citep{STITCHv5}. As both STRING and STITCH provide confidence scores
@@
 one existing target. Therefore, the intersection between these
-resources yields 1,428 drugs and 7,368 human proteins with 32212 interaction links for the training phase. We provide links to and methods for obtaining and processing
-the necessary data on Github.
+resources yields 1,428 drugs and 7,368 human proteins with 32,212
+interactions for the training phase. We provide links to and methods
+for obtaining and processing the necessary data on Github.
 
 For comparative evaluation, we use the gold standard dataset
-introduced by \citet{Yaminishi2008} consisting of 1923 interactions between 708 drugs and 1512 proteins.
+introduced by \citet{Yaminishi2008} consisting of 1,923 interactions
+between 708 drugs and 1,512 proteins.
 
@@
 
-\subsubsection{Half-twin neural networks and modular learnable feature transformation}
-
-\todo[inline]{Looking at the code, I wonder if what is implemented is
-  actually a ANN; it does not look as if the weights are shared
-  between the two networks, so looks more like a twin or half-twin
-  network? Siamese would be sim(f(e1),f(e2)) but the implementation
-  seems to be sim(f(e1), g(e2)).}
+\subsubsection{Half-twin neural networks and feature transformation}
 
@@
 aims to learn a similarity between two embeddings of the same
-dimension.  The precomputed embeddings are first processed by a
-learnable feature transformation (LFT) network, a fully connected
-neural network layer which takes as input an embedding and outputs a
-representation of a particular size. An example structure for both
-types of features can be found in Figure~\ref{fig:HalfTwinNetwork}. We
-use this LFT to reduce the eventual representation size for drugs and
+dimension.  The precomputed embeddings may have different
+dimensionality and we first process them using a learnable feature
+transformation (LFT) network, i.e., a fully connected neural network
+layer which takes as input an embedding and outputs a representation
+of a particular size. An example structure for both types of features
+can be found in Figure~\ref{fig:HalfTwinNetwork}. We use this LFT to
+reduce the representation size of the embeddings for drugs and
 proteins separately.
@@
 and molecular feature for both drugs and proteins are reduced to
-similar dimensionality for varying sizes of inputs; this allows for a
+the same dimensionality for varying sizes of inputs; this allows for a
 high amount of modularity across different experimental setups by
 adding different kinds of features into the model. Additionally, the
-generated features can be used for a variety of other tasks; for
-example, the ontology LFT can be reused for a variety of DL2vec-based
-features with respect to other ontologies and hypotheses. We follow
-the results of \textit{DL2vec} showing that the activation function
-$\sigma := \mathrm{LeakyReLU}$ leads to improved performance compared
-to other activation functions.
+generated features may be used for other tasks. % ; for
+% example, the ontology LFT can be reused for a variety of DL2vec-based
+% features with respect to other ontologies and hypotheses
+We follow the results of \textit{DL2vec} \cite{} and use as
+activation function $\sigma := \mathrm{LeakyReLU}$ which leads to
+improved performance compared to other activation functions.
 
@@
 % methods out there we will only introduce the most basic one here. 
-A graph convolutional layer \citet{GCNConv} consists of a learnable
+A graph convolutional layer \cite{GCNConv} consists of a learnable
 weight matrix followed by an aggregation step, formalized by
@@
 \subsubsection{title}
-Combining half-twin and graph convolutional neural networks, we map all protein representations to the respective node features, initializing the graph convolutional update steps.
-The resulting representations are used for a similarity prediction similar to the models presented in figure ~\ref{fig:HalfTwinNetwork}.
+Combining half-twin and graph convolutional neural networks, we map
+all protein representations to the respective node features,
+initializing the graph convolutional update steps. The resulting
+representations are used for a similarity prediction similar to the
+models presented in figure ~\ref{fig:HalfTwinNetwork}.
 When combining ontology and molecular features with or without the
@@
 features, before plugging them into the graph for the similarity
-computation. A more in-depth image of the overall architecture, combining both feature types is depicted in figure ~\ref{fig:FullModelAllLayers}.
+computation. A more in-depth image of the overall architecture,
+combining both feature types is depicted in figure
+~\ref{fig:FullModelAllLayers}.
 
@@
 \end{equation}
-for a given prediction $x$ and target $y$, and positive weight $w$ defined by equation \eqref{weight}. We average this loss among
-all drug-protein pairs in the training set, leading to a stable
-environment for the used optimizing scheme \textit{Adam}
-\citep{Adam2014}. We implemented a 5-fold cross validation split among the
-proteins. Furthermore, we used early stopping in
-the training process for optima detection.
+for a given prediction $x$ and target $y$, and positive weight $w$
+defined by equation \eqref{weight}. We average this loss among all
+drug-protein pairs in the training set, leading to a stable
+environment for the \textit{Adam} optimization algorithm
+\citep{Adam2014}. We implemented a 5-fold cross validation split among
+the proteins. Furthermore, we used early stopping in the training
+process.
 

### 2021-03-20 paper/main.tex
.

@@
 
-We included these molecular and ontology-based sub-models with a
-larger graph convolutional model. The graph underlying the graph
-convolutional neural network is based on the protein--protein
-interaction (PPI) graph. The PPI dataset is represented by a graph
-$G=(V,E)$, where each protein is represented by a vertex $v\in V$, and
-each edge $e\in E\subseteq V\times V$ symbolizes an interaction
-between two proteins. Additionally, we introduce a mapping
-$x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node
-feature $x_v := x(v)$, where $d$ denotes the dimensionality of the node
-features.
+We include these molecular and ontology-based sub-models within a
+graph neural network (GNN). The graph underlying the GNN is based on
+the protein--protein interaction (PPI) graph. The PPI dataset is
+represented by a graph $G=(V,E)$, where each protein is represented by
+a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$
+represents an interaction between two proteins. Additionally, we
+introduce a mapping $x:V\rightarrow\mathbb{R}^{d}$ projecting each
+vertex $v$ to its node feature $x_v := x(v)$, where $d$ denotes the
+dimensionality of the node features.
 
@@
 
-The update rule for each individual node is given by a message passing scheme formalized by
+The update rule for each node is given by a message passing scheme
+formalized by
 \begin{equation}

### 2021-03-20 paper/main.tex
.

@@
 
-While in this initial formulation the node-wise update step is defined
-by the sum over all neighbouring node representations, we are able to
-alter this formulation to another message passing scheme.  We are able
-to rearrange the order of activation function $\sigma$, aggregation
-$\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this
+While in this initial formulation of a GCNConv the node-wise update
+step is defined by the sum over all neighbouring node representations,
+we can alter this formulation to other message passing schemes.  We
+can rearrange the order of activation function $\sigma$, aggregation
+$\mathrm{AGG}$, and linear neural layer $\mathrm{MLP}$ with this
 formulation as proposed by \citet{GENConv2020}:
@@
 \end{equation}
-where we will generally only consider
-$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We will denote
-this generalized layer type as \textsc{GENConv}, following the
+where we only consider
+$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote
+this generalized layer type as \textsc{GENConv} following the
 notation of PyTorch Geometric \cite{PytorchGeometric}.  While the
@@
 \textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
-vanishing gradients and signal fading for large scale, highly
+vanishing gradients and signal fading for large scale and highly
 connected graphs, each propagation step in \textsc{GENConv} emphasizes
@@
 connections and a normalization scheme \cite{DeepGCN2019,
-  DeeperGCN2020}.
-The residual blocks are reusable and can be stacked multiple times. %, thereby not losing focus of each node neighbourhood.
-\todo[inline]{Needs a clearer outline/description/figure of YOUR model
-structure.}
-The structure is depicted in Figure~\ref{fig:ResGraphConvBlocks}.
+  DeeperGCN2020}.  The residual blocks are reusable and can be stacked
+multiple times. The structure of the GNN architecture we use is shown
+in Figure~\ref{fig:ResGraphConvBlocks}.
 

### 2021-03-20 paper/main.tex
.

@@
 
-\subsubsection{title}
+\subsubsection{Combined prediction model}
 Combining half-twin and graph convolutional neural networks, we map
-all protein representations to the respective node features,
+all protein representations to their respective node features,
 initializing the graph convolutional update steps. The resulting
-representations are used for a similarity prediction similar to the
-models presented in figure ~\ref{fig:HalfTwinNetwork}.
+representations are used for a similarity prediction similar
+to\todo{Similar, or identical to? Needs to describe the exact model
+  used} the models presented in figure ~\ref{fig:HalfTwinNetwork}.
 When combining ontology and molecular features with or without the
@@
 \end{equation}
-consequently needing compensation in the computed loss function and
-appropriate metrics for the evaluation.
+and consequently needs to be compensated in the computed loss function
+and appropriate metrics for the evaluation.
 
-Therefore, we weighted all positive drug-protein pair samples with
-this ratio by introducing the following loss function with respect to
+Therefore, we weight all positive drug-protein pair samples with this
+ratio by introducing the following loss function with respect to 
 binary cross-entropy:
@@
 
-To find the best hyperparameter configuration for the proposed model
+To find the best hyperparameter configuration for the proposed model,
 we performed a grid search to find the most expressive and
@@
 depth of the neural network, optimizer, learning rate and layer types
-using an extensive, manual grid search, Starting from na\"ive, shallow feature transformations with an embedding size of 10, we scaled the network up to residual structures with up to 10 hidden layers leading to embeddings of size 4000, testing different network widths and learning rates for each configuration.
+using an extensive, manual grid search. Starting from na\"ive, shallow
+feature transformations with an embedding size of $10$, we scaled the
+network up to residual structures with up to $10$ hidden layers
+leading to embeddings of size $4000$, testing different network widths
+and learning rates for each configuration.
 

### 2021-03-20 paper/main.tex
.

@@
 We calculate the AUROC by computing true positive rate at various
-false positive rate thresholds and using trapezoidal approximations to
+false positive rate thresholds and use trapezoidal approximations to
 estimate the area under the curve. We refer to this measure as
-$MacroAUC$.
+$\textrm{MacroAUC}$.
 
 \todo[inline]{For RH: check this again later:}
-In contrast we also calculate the $MicroAUC$ score.
-For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
-With respect to proteins, this can be formalized for given labels $l:D\times P \rightarrow \{0,1\}$ and predictions $y:D\times P \rightarrow [0,1]$  as
+We also calculate the $\textrm{MicroAUC}$ score. For given lists $D$
+and $P$ of drugs and proteins, respectively, and a set of known
+interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated
+as the average per entity (macro) \textit{AUROC} score. Specifically,
+for protein-centric score, this can be formalized as: given labels
+$l:D\times P \rightarrow \{0,1\}$ and predictions $y:D\times P
+\rightarrow [0,1]$, we define
 \begin{equation*}
-	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
+  \textrm{MicroAUC}_p := \underset{p\in P}{mean}\left(\left\{
+      \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\})
+    \right\}\right)
 \end{equation*}
 
-Drugs and proteins can be interchanged in this formulation, while we
+Drugs and proteins can be interchanged in this formulation, and we
 refer to the different measures as protein-centric microAUC
-($MicroAUC_p$) and a drug-centric microAUC
-($MicroAUC_d$). Additionally, the $MicroAUC$ score may not be defined
-as in some datasets some targets or drugs, respectively, do not have
-any interactions, leading to an infeasible $TPR=0$ for all thresholds and an undefined
-$AUROC$ score for that entity. For those entities we impute the $MicroAUC$ interpolating linearly, by using the accuracy for this subset.
+($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$). In some
+cases, the $MicroAUC$ score may not be defined because in some
+datasets some proteins or drugs have no interactions, leading to an
+infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
+for that entity. For those entities, we impute the $MicroAUC$
+interpolating linearly, by using the accuracy for this
+subset\todo{Last sentence not clear, can you give the formula instead?}.
 

### 2021-03-20 paper/main.tex
.

@@
 \subsection{Evaluation and metrics}
-
+\label{sec:evaluation}
 To assess each model, we compute a variety of common metrics for
@@
 
-We construct a PPI graph with proteins as nodes and protein-protein
+We construct a graph with proteins as nodes and protein-protein
 interactions as edges, mapping the protein features to each target as
 node features. \name then propagates information among the PPI network
-utilizing graph convolutional steps, calculating the similarity of
-drug and protein representation, and thus eventually predicting
-whether there is an interaction. The full workflow scheme is depicted
-in figure~\ref{fig:ModelWorkflow}
+utilizing graph convolutional steps, calculates the similarity of drug
+and protein representations, and predicts whether there is an
+interaction. The full workflow scheme is depicted in
+figure~\ref{fig:ModelWorkflow}
 
@@
 We trained, validated and finally tested all considered models on the
-STITCH dataset using a 5-fold cross-validation over a protein
-split; we then selected the best-performing models (with respect to
-\textit{AUROC/MacroAUC} and $MicroAUC_p$) and evaluate them on in 5-fold protein split cross validation on the Yamanishi benchmark dataset to avoid validation overfitting and yield more realistic testing results.
-To evaluate the influence of the different
-features separately, and determine whether they ``localize'' on the
-PPI graph (and therefore can be exploited successfully by the graph
-neural networks), we train and evaluate models with different types of
-features, and with and without inclusion of the PPI graph,
-separately, comparing the molecular (MolPred) and phenotype based (OntoPred) predictor and a combination of both, concatenating both features.
-Table \ref{tab:Results} (a)+(b) shows the results of these experiments.
+STITCH dataset using a 5-fold cross-validation over a protein split;
+we then selected the best-performing models (with respect to
+\textit{MacroAUC} and $MicroAUC_p$, see Section \ref{sec:evaluation})
+and evaluate them in a 5-fold protein-split cross-validation on the
+Yamanishi benchmark dataset to avoid validation overfitting and yield
+more realistic testing results. To evaluate the influence of the
+different features separately, and to determine whether they
+``localize'' on the PPI graph (and therefore can be exploited
+successfully by the graph neural networks), we train and evaluate
+models with different types of features, and with and without
+inclusion of the PPI graph, separately. We comparing the molecular
+(MolPred) and phenotype-based (OntoPred) prediction model, and a
+combination of both when concatenating both types of features.  Table
+\ref{tab:Results} (a)+(b) shows the results of these experiments.
 

### 2021-03-20 paper/main.tex
.

@@
 graph convolutional neural layers in the residual blocks are removed,
-resulting in a non--graph model. This pruned network, with no
+resulting in a model with similar parameters but without the ability
+to use graph-based information. This pruned network, with no
 information on the protein--protein interactions, reached very similar
@@
 
-Furthermore, these major improvements are only provided by the \textit{GENConv} graph convolution scheme as part of the ResGraphConv blocks, as \textit{GCNConv} and other related graph convolutional methods fail to achieve any gain in comparison to the plain \textit{OntoPred} performance even when combined with the residual blocks.
-The discrepancy of {\em GENConv} and other graph convolutional methods may be founded in numerical stability and fading signals, as described in the introduction of the {\em GENConv} method.
-
-As the inclusion of graph information increases performance, we conclude that information about protein
-functions localizes on the graph while molecular features do not.
+The improvements when including the graph are only provided by the
+\textit{GENConv} graph convolution scheme which includes the
+ResGraphConv blocks; \textit{GCNConv} and other graph
+convolutional methods fail to achieve any gain in comparison to the
+plain \textit{OntoPred} performance even when combined with the
+residual blocks. The discrepancy between {\em GENConv} and other graph
+convolutional methods may be the result of numerical instability and
+fading signals \cite{}.
+Nevertheless, our results demonstrate that the inclusion of graph
+information increases performance when ontology-based features are
+used, which allows us to conclude that information about protein
+functions localizes on the graph whereas molecular features do not.
 

### 2021-03-20 paper/main.tex
.

@@
 estimate on how \name performs for individual protein targets, we use
-micro-averages between proteins and compute the \textit{MicroAUC};
-micro-averages average the performance (true and false positive rates)
-per protein instead of across all drug--protein pairs.
+micro-averages between proteins and compute the \textit{MicroAUC} (see
+Section \ref{sec:evaluation}); micro-averages average the performance
+(true and false positive rates) per protein instead of across all
+drug--protein pairs, and can therefore better estimate how \name
+performs when tasked with finding a drug that targets a specific
+protein.
 
@@
 model are the result of the biases in drug--target interaction
-datasets. We design a ``na\"ive'' baseline model that predicts the
-same list of proteins for each drugs based on the number of
-known drug--target interactions for a protein. Formally, given lists
-$D$ and $P$ of drugs and proteins and a set of known interactions
-$\mathcal{I} := \{(d_i, p_i) \}$, we construct an interaction matrix
-$M_{int}\in\{0,1\}^{|D|\times|P|}$ with
+datasets where hub proteins may have a large number of interactions,
+or certain drugs interact with many proteins, and these may be easier
+to predict than others. We design a ``na\"ive'' baseline model that
+predicts the same list of proteins for each drug based only on the
+number of known drug--target interactions for a protein. Formally,
+given lists $D$ and $P$ of drugs and proteins and a set of known
+interactions $\mathcal{I} := \{(d_i, p_i) \}$, we construct an
+interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
 \begin{equation*}
@@
 Our ``na\"ive'' predictor $P_k$ predicts all drugs to interact with
-the top $k$ targets
-% , denoted by $Top_k(P)$ 
-with respect to
-% $M_{ij}$ from
-the introduced ranking:
+the top $k$ targets with respect to the introduced ranking:
 \begin{equation*}
@@
 with the only hyperparameter $k$.
+
 The prediction $P_k$ is not dependent on the drug $d_i$ and will
-predict the same ranked list of drugs for a given protein $p_j$;
-consequently, this na\"ive predictor will not predict any useful
-information about interactions between drugs and proteins, but allows
-us to estimate the effect of potential biases in our evaluation
-dataset.
+predict the same ranked list of drugs for all proteins; consequently,
+this na\"ive predictor will not predict any useful information about
+interactions between drugs and proteins, but allows us to estimate the
+effect of potential biases in our evaluation dataset.
 % Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
 
-We apply this naive predictor on both the STITCH and Yamanishi
-datasets, using the full dataset as well as 5-fold cross-validation
-over drugs and drug--protein pairs.
-Note, that this baseline predictor is not applicable for a protein split CV, as the number of interactions for each protein in the validation set is unknown. 
-For each fold, we gradually increase $k$ to determine the best
+Our na\"ive predictor is not applicable for a protein split
+cross-validation as the number of interactions for each protein in the
+validation set is unknown.  We apply this naive predictor on both the
+STITCH and Yamanishi datasets, using the full dataset as well as
+5-fold cross-validation over drugs as well as over drug--protein
+pairs.  For each fold, we gradually increase $k$ to determine the best
 performance for each fold. Using the full dataset, drug--target split,
-and drug split, we obtain the following MacroAUC/AUROC score results.
-For STITCH database we obtain a performance of $0.76$ on the whole dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the drug splitting scheme.
-While these results are fairly moderate, on Yamanishi dataset we yield MacroAUC scores of $0.88$, $0.84$ and $0.85$, for total dataset, DT pair split and drug split, respectively.
-
-The na\"ive predictor shows great performance on Yamanishi dataset, with substantial gain in comparison to a random predictor on both datasets.
-In the following, we will utilize this nai\"ve predictor as baseline predictor correlating its performance to state of the art models and \name.
+and drug split, we obtain the following MacroAUC results: for the
+STITCH database, we obtain a performance of $0.76$ on the whole
+dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the
+drug splitting scheme; on the Yamanishi dataset, we obtain MacroAUC
+scores of $0.88$, $0.84$ and $0.85$, for the total dataset,
+drug--target pair and drug split, respectively.  The na\"ive predictor
+shows higher performance on the Yamanishi dataset than on STITCH, and
+a substantial gain in comparison to an expected random predictor on
+both datasets. In the following, we utilize this nai\"ve predictor as
+baseline to compare its performance to state of the art models and
+\name.
 
@@
 performing methods for drug--target interaction prediction that were
-previously evaluated on the Yamanishi benchmark dataset. These methods include DTIGEMS+
-\cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed
-superior results in comparison to numerous works. Furthermore, we
-added DTINet \cite{DTINet2017} as method for comparison which has been
-used to develop a number of methods such as NeoDTI \cite{NeoDTI2019}
-with similar methodology.
+previously evaluated on the Yamanishi benchmark dataset. These methods
+include DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}
+which have showed superior results in comparison to numerous
+works. Furthermore, we added DTINet \cite{DTINet2017} as method for
+comparison which has been used to develop a number of methods such as
+NeoDTI \cite{NeoDTI2019} with similar methodology.
 
@@
 respective AUROC. We further evaluate each model by performing a
-protein-wise cross-validation determining Macro AUC/AUROC score and
-$MicroAUC_p$. For this evaluation we allow stratification for the
-training process but not for the validation and testing phase as
-real world applications of these models would have to deal with possibly
+protein-wise cross-validation determining the MacroAUC and
+$MicroAUC_p$. For this evaluation, we allow stratification for the
+training process but not for the validation and testing phase as real
+world applications of these models would have to deal with possibly
 imbalanced data.

### 2021-03-20 paper/main.tex
.

@@
 
-The results of this analysis are summarized in part (c) of Table \ref{tab:Results}, calculating the performance of all considered methods over their original splitting scheme and over a protein split.
-We find that, first the majority of approaches evaluate over a drug--protein splitting scheme, while second, the highest performances on Yamanishi dataset are yielded over the drug--target pairs splitting scheme, with a noticeable gap to the methods performing their splits over proteins.
-Further, when evaluating the same methods over a protein split, we find a substantial performance divergence in comparison to their original splitting scheme.
-The authors of DTI-CDF evaluate their method on all three splitting schemes underlining this point.
-While \name provides comparable performance to DTI-CDF, it yields considerably better with respect to $MicroAUC_p$.
+The results of our experiments are summarized in part (c) of Table
+\ref{tab:Results}; we calculated the performance of all compared
+methods over their original splitting scheme and over a protein
+split. First, we find that there is a large difference in performance
+when evaluating over a drug--target pair split compared to a
+protein split, with generally higher performance achieved when using
+the drug--target pair split.
+Second, when evaluating the same methods over a protein split, we find
+a substantial performance difference in comparison to the splitting
+scheme used in the original evaluation of each method. DTI-CDF was
+originally evaluated on all three splitting schemes underlining this
+point \cite{}. While \name provides comparable performance to DTI-CDF,
+it yields considerably better results with respect to $MicroAUC_p$.
 

### 2021-03-20 paper/main.tex
.

@@
 
-As we were quite surprised about the MacroAUC discrepancy between the splitting schemes, we evaluated other cutting edge drug--target interaction and drug--target affinity prediction methods, training and evaluating on other datasets.
-Following the results of MolTrans (\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}), DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and MolTrans itself on BioSnap (\textbf{CITATION}) over a protein split, as shown in part (d) of Table \ref{tab:Results}.
-The authors of MolTrans evaluated all considered methods over the drug--target pair and the protein split, which we were able to reproduce with (d), showing once more a substantial divergence.
-However, we additionally computed the $MicroAUC_p$ score for all considered methods, leading to comparable scores as for the Yamanishi evaluated methods.
+As the difference in performance with different splitting schemes is
+quite striking, we evaluated additional drug--target interaction and
+drug--target affinity prediction methods that were trained and
+evaluation on other datasets.  Following the results of MolTrans
+(\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}),
+DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and
+MolTrans itself on BioSnap\todo{What is that? Is it in methods?}
+(\textbf{CITATION}) over a protein split, as shown in part (d) of
+Table \ref{tab:Results}. The authors of MolTrans evaluated all
+considered methods over the drug--target pair and the protein split;
+we were able to reproduce the results (Table \ref{tab:Results} (d)),
+showing a substantial difference based on the splitting scheme. We
+additionally computed the $MicroAUC_p$ score for all considered
+methods, leading to comparable\todo{are the datasets comparable?}
+scores as for the results obtained on the Yamanishi dataset.
 

### 2021-03-20 paper/main.tex
.

@@
 
-\begin{table}[!tpb]
+\begin{table*}[!tpb]
+  \centering
 	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
@@
 	\label{tab:Results}
-\end{table}
+\end{table*}
 
@@
 
+\subsection{``Bottom-up'' and ``top-down'' prediction of drug-target interactions}
+
+What features localize on graph? Molecular interactions based on
+structure not likely -- only possible to find binding sites to other
+proteins.
+
+Phenotype-based features localize...
+
 \todo[inline]{Methods:}
@@
 
-%\subsection{\name identifies drugs that target a protein}
+using indications is like cheating, as not applicable for searching new drugs
+
+drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
+
+we built protein function and ontology based features based on DL2vec
+
+Ontology derived protein function focused features are highly predictive for dtis
+
+We built a versatile template for various features to test localization on the PPI graph
+
+normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
+
+protein functions localize on the PPI graph, while molecular features don't
+
+molecular features work for all (most?) drugs, top-down features only
+for those that have a side-effect profile
+
+\subsection{Evaluating drug--target interaction predictions}
+
+Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
+
+Running split over proteins is harder than, drug and drug protein pair split (see below table)
+
+this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
+
+Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
+
+microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
 
@@
 
-%\section{Findings}
+A naive predictor (ranking proteins) and predict each drug similarly
+achieves cutting edge performance ($87.5$ AUROC for whole dataset,
+$85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
+prot focused microAUC possible. $\rightarrow$ hub proteins
+
+Yamanishi Dataset is only partially suitable for comparing results, if
+everybody just derives a suitable subset (DTIGEMS)
+
+This also applies to drug target affinity prediction. We were hereby
+able to roughly reproduce the results from MolTrans (Bioinformatics)
+on BioSnap
+
+%\subsection{Features to use in predicting drug targets}
+
+%\subsection{\name identifies drugs that target a protein}
+
+\subsection{Conclusions}
+
+Evaluating DTI predictions is non-trivial due to numerous biases in
+the training and evaluation datasets.
+
+Combining bottom-up and top-down features improves state of the art.
 
@@
 \begin{itemize}
-	\item we built protein function and ontology based features based on DL2vec
-	\item Ontology derived protein function focused features are highly predictive for dtis
-	\item We built a versatile template for various features to test localization on the PPI graph
-	\item normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
-	\item protein functions localize on the PPI graph, while molecular features don't
 	\item all AUROC in \% AUROC score on STITCH
@@
 \begin{itemize}
-	\item Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
-	\item Running split over proteins is harder than, drug and drug protein pair split (see below table)
-	\item this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
-	\item using indications is like cheating, as not applicable for searching new drugs
-	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
 	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

### 2021-03-20 paper/main.tex
.

@@
 
-\subsection{Conclusions}
+\section{Conclusions}
+
+For the bottom-up approach we build a model that only relies on
+molecular features, which we will discuss in more detail in the
+following methods chapter. For the combination of both approaches we
+now attach the predictions to the protein-protein interaction graph as
+node features for future graph learning steps. In this graph we tried
+to find both patterns and regions for each drug that could be of
+interest through application of different graph convolutional layers,
+which in return represent the feature for each protein. Representing
+the drug we take the drug-drug interaction graph and the semantic
+similarity over side effects which we will explain in the following
+paragraphs.
+
+Combining bottom-up and top-down features improves state of the art.
 
@@
 
-Combining bottom-up and top-down features improves state of the art.
-
-\subsection{Deification of our method}
-\begin{itemize}
-	\item all AUROC in \% AUROC score on STITCH
-
-	\begin{tabular}{|l|r|r|}
-		\hline
-		DNNT model&Without graph model&With graph model\\
-		\hline
-		MolPred&$69$&$69$\\
-		PhenomeNETPred&$88$&$92$\\
-		\hline
-		MolPred + PhenomeNETPred & $89$ & $93$\\
-		\hline
-	\end{tabular}
-	\item microAUC for MolPred + PhenomeNETPred on graph is about $93+-$
-	\item and on yamanishi dataset
-
-	\begin{tabular}{|l|r|r|}
-		\hline
-		DNNT model&Without graph model&With graph model\\
-		\hline
-		PhenomeNETPred&$83$&$84$\\
-		\hline
-		MolPred + PhenomeNETPred & $83$ & $84.5$\\
-		\hline
-	\end{tabular}
-	\item MicroAUC is about $83$
-\end{itemize}
-
-\subsection{How to insult other methods}
-\begin{itemize}
-	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
-	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
-
-	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
-		\hline
-		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
-		\hline
-		DTINet&DP pairs&$91$&$84.1$&$67.2$\\
-		DTIGEMS+&DP pairs&$93$& $72.2$& $67.8$ \\
-		DTI-CDF&Proteins&$83$&$83$&$79$\\
-		\hline
-	\end{tabular}
-	\item A naive predictor (ranking proteins) and predict each drug similarly achieves cutting edge performance ($87.5$ AUROC for whole dataset, $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No prot focused microAUC possible.
+% \subsection{Deification of our method}
+% \begin{itemize}
+% 	\item all AUROC in \% AUROC score on STITCH
+
+% 	\begin{tabular}{|l|r|r|}
+% 		\hline
+% 		DNNT model&Without graph model&With graph model\\
+% 		\hline
+% 		MolPred&$69$&$69$\\
+% 		PhenomeNETPred&$88$&$92$\\
+% 		\hline
+% 		MolPred + PhenomeNETPred & $89$ & $93$\\
+% 		\hline
+% 	\end{tabular}
+% 	\item microAUC for MolPred + PhenomeNETPred on graph is about $93+-$
+% 	\item and on yamanishi dataset
+
+% 	\begin{tabular}{|l|r|r|}
+% 		\hline
+% 		DNNT model&Without graph model&With graph model\\
+% 		\hline
+% 		PhenomeNETPred&$83$&$84$\\
+% 		\hline
+% 		MolPred + PhenomeNETPred & $83$ & $84.5$\\
+% 		\hline
+% 	\end{tabular}
+% 	\item MicroAUC is about $83$
+% \end{itemize}
+
+% \subsection{How to insult other methods}
+% \begin{itemize}
+% 	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
+% 	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
+
+% 	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
+% 		\hline
+% 		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
+% 		\hline
+% 		DTINet&DP pairs&$91$&$84.1$&$67.2$\\
+% 		DTIGEMS+&DP pairs&$93$& $72.2$& $67.8$ \\
+% 		DTI-CDF&Proteins&$83$&$83$&$79$\\
+% 		\hline
+% 	\end{tabular}
+% 	\item A naive predictor (ranking proteins) and predict each drug similarly achieves cutting edge performance ($87.5$ AUROC for whole dataset, $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No prot focused microAUC possible.
 $\rightarrow$ hub proteins
-	\item Yamanishi Dataset is only partially suitable \todo{for
-            comparing results}, if everybody just derives a suitable subset (DTIGEMS)
-	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap
-
-\end{itemize}
-
-\begin{figure}[!tpb]%figure1
-	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
-		\hline
-		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
-		\hline
-		DeepDTI&Drugs&$87.6$&$75.9$&$70.1$\\
-		DeepDTA&DP pairs&87.6&$76.7$&$69.4$\\
-		DeepConv-DTI&DP pairs&88.3&$76.6$&$73.0$\\
-		MolTrans&DP pairs&89.5&$77.0$&$74.0$\\
-
-		\hline
-	\end{tabular}
-\end{figure}
-
-\subsection{Tested hypotheses}
-
-In this work we are testing the following hypotheses:
-\begin{enumerate}
-	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
-	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
-\end{enumerate}
-We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account.
+% 	\item Yamanishi Dataset is only partially suitable \todo{for
+%             comparing results}, if everybody just derives a suitable subset (DTIGEMS)
+% 	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap
+
+% \end{itemize}
+
+% \begin{figure}[!tpb]%figure1
+% 	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
+% 		\hline
+% 		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
+% 		\hline
+% 		DeepDTI&Drugs&$87.6$&$75.9$&$70.1$\\
+% 		DeepDTA&DP pairs&87.6&$76.7$&$69.4$\\
+% 		DeepConv-DTI&DP pairs&88.3&$76.6$&$73.0$\\
+% 		MolTrans&DP pairs&89.5&$77.0$&$74.0$\\
+
+% 		\hline
+% 	\end{tabular}
+% \end{figure}
+
+% \subsection{Tested hypotheses}
+
+% In this work we are testing the following hypotheses:
+% \begin{enumerate}
+% 	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
+% 	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
+% \end{enumerate}
+% We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account.
 Thus, we propose a novel approach to combine those mutual exclusive attempts, through the usage of interaction networks, similarity and molecular features.
@@
 
-For the bottom-up approach we build a model that only relies on molecular features, which we will discuss in more detail in the following methods chapter.
-For the combination of both approaches we now attach the predictions to the protein-protein interaction graph as node features for future graph learning steps.
-In this graph we tried to find both patterns and regions for each drug that could be of interest through application of different graph convolutional layers, which in return represent the feature for each protein.
-Representing the drug we take the drug-drug interaction graph and the semantic similarity over side effects which we will explain in the following paragraphs.
-
-\section{Methods}
-
-\subsection{Models}
-The used model consists of two separate models, that help to fuse together the two methods:
-\begin{enumerate}
-	\item The molecular predictor
-	\item The interaction network based predictor
-\end{enumerate}
-We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins.
+% \section{Methods}
+
+% \subsection{Models}
+% The used model consists of two separate models, that help to fuse together the two methods:
+% \begin{enumerate}
+% 	\item The molecular predictor
+% 	\item The interaction network based predictor
+% \end{enumerate}
+% We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins.
 Regarding proteins, we used the pretrained feature generator from \textit{DeepGoPlus} (\citep{DeepGoPlus}) that was originally designed for protein function prediction and is regarded as state of the art for this purpose.
@@
 \\
-The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE).
+% The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE).
 Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug.
@@
 \\
-The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node.
+% The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node.
 In contrast to learning over whole graphs we perform node classification within the graph.
@@
 \\
-The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
-For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy.
+% The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
+% For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy.
 Concerning this hierarchy, drug-drug similarity is computed by the Resnik similarity (\cite{Resnik1995}).
@@
 
-Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
+% Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
 This network finally yields our prediction.
@@
 
-\begin{figure}[!tpb]%figure1
-	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
-	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
-\end{figure}
-
-\begin{figure}[!tpb]%figure2
-\centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
-\caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.
+% \begin{figure}[!tpb]%figure1
+% 	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
+% 	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
+% \end{figure}
+
+% \begin{figure}[!tpb]%figure2
+% \centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
+% \caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.
 Protein and drug features are represented by blue and red, respectively.}\label{fig:02}
-\end{figure}
-
-\section{Discussion}
+% \end{figure}
+
+%\section{Discussion}
 

### 2021-03-20 paper/main.tex
.

@@
 The goal of \name is to find candidate drugs that target a specific
-protein; however, our evaluation does not evaluate this application
+protein; however, so far, we do not evaluate this application
 but rather how \name would perform in finding plausible drug--target
-interactions among all possible interactions. To provide a better
+interactions among all possible interactions. This evaluation does not
+correspond to the application of \name in finding drugs that target a
+specific protein.
+To provide a better
 estimate on how \name performs for individual protein targets, we use

### 2021-03-21 paper/main.tex
.

@@
 we then selected the best-performing models (with respect to
-\textit{MacroAUC} and $MicroAUC_p$, see Section \ref{sec:evaluation})
-and evaluate them in a 5-fold protein-split cross-validation on the
-Yamanishi benchmark dataset to avoid validation overfitting and yield
-more realistic testing results. To evaluate the influence of the
-different features separately, and to determine whether they
-``localize'' on the PPI graph (and therefore can be exploited
-successfully by the graph neural networks), we train and evaluate
-models with different types of features, and with and without
-inclusion of the PPI graph, separately. We comparing the molecular
-(MolPred) and phenotype-based (OntoPred) prediction model, and a
-combination of both when concatenating both types of features.  Table
-\ref{tab:Results} (a)+(b) shows the results of these experiments.
+\textit{MacroAUC} and $MicroAUC_p$\todo{which is it, macro or micro?},
+see Section \ref{sec:evaluation}) and evaluate them in a 5-fold
+protein-split cross-validation on the Yamanishi benchmark dataset to
+avoid validation overfitting and yield more realistic testing
+results. To evaluate the influence of the different features
+separately, and to determine whether they ``localize'' on the PPI
+graph (and therefore can be exploited successfully by the graph neural
+networks), we train and evaluate models with different types of
+features, and with and without inclusion of the PPI graph,
+separately. We comparing the molecular (MolPred) and phenotype-based
+(OntoPred) prediction model, and a combination of both where we
+concatenate both types of features.  Table
+\ref{tab:Results}\todo{check -- not currently in the table} shows the
+results of these experiments.
 
@@
 the {\em GCNConv} and {\em GENConv} architecture already shows some
-minor improvement, the \textit{ResGraphConv} blocks add a large amount
-of additional learnable parameters to the network, leading to more
-expressive power. To test whether the improvement is due to the number
-of learnable parameters added or the result of better exploiting the
-information about PPIs, we experiment with a graph model in which all
-graph convolutional neural layers in the residual blocks are removed,
-resulting in a model with similar parameters but without the ability
-to use graph-based information. This pruned network, with no
-information on the protein--protein interactions, reached very similar
-results to the original \textit{OntoPred} model and showed no
-improvement.
+minor improvement, the use of \textit{ResGraphConv} results in larger
+performance improvements.  {\em ResGraphConv} blocks add a large
+amount of additional learnable parameters to the network, leading to
+more expressive power. To test whether the observed improvement is due
+to the number of learnable parameters added or the result of better
+exploiting the information about PPIs, we experiment with a graph
+model in which all graph convolutional neural layers in the residual
+blocks are removed, resulting in a model with similar parameters but
+without the ability to use graph-based information. This pruned
+network, with no information on the protein--protein interactions,
+reached very similar results to the original \textit{OntoPred} model
+and showed no improvement.
 
@@
 \textit{GENConv} graph convolution scheme which includes the
-ResGraphConv blocks; \textit{GCNConv} and other graph
+{\em ResGraphConv} blocks; \textit{GCNConv} and other graph
 convolutional methods fail to achieve any gain in comparison to the
@@
 fading signals \cite{}.
-Nevertheless, our results demonstrate that the inclusion of graph
-information increases performance when ontology-based features are
-used, which allows us to conclude that information about protein
-functions localizes on the graph whereas molecular features do not.
+
+Our results demonstrate that the inclusion of graph information
+can increase performance when ontology-based features are used but not
+when molecular features are used alone. This observation allows us to
+conclude that information about protein functions localizes on the
+graph whereas molecular features do not.
 
@@
 The goal of \name is to find candidate drugs that target a specific
-protein; however, so far, we do not evaluate this application
-but rather how \name would perform in finding plausible drug--target
-interactions among all possible interactions. This evaluation does not
+protein; however, so far, we do not evaluate this application but
+rather how \name would perform in finding plausible drug--target
+interactions among all possible interactions (since we use the
+MacroAUC as our main evaluation measure). This evaluation does not
 correspond to the application of \name in finding drugs that target a
-specific protein.
-To provide a better
-estimate on how \name performs for individual protein targets, we use
-micro-averages between proteins and compute the \textit{MicroAUC} (see
-Section \ref{sec:evaluation}); micro-averages average the performance
-(true and false positive rates) per protein instead of across all
-drug--protein pairs, and can therefore better estimate how \name
+specific protein.  To provide a better estimate on how \name performs
+for individual protein targets, we use micro-averages between proteins
+and compute the \textit{MicroAUC} (see Section \ref{sec:evaluation});
+to determine {\em MicroAUC}, we average the performance (true and
+false positive rates) per protein instead of across all drug--protein
+pairs; the resulting measure can therefore better estimate how \name
 performs when tasked with finding a drug that targets a specific

### 2021-03-21 paper/main.tex
.

@@
 % \subsection{Baseline model}
-Furthermore, we hypothesize that some of the predictions made by our
-model are the result of the biases in drug--target interaction
-datasets where hub proteins may have a large number of interactions,
-or certain drugs interact with many proteins, and these may be easier
-to predict than others. We design a ``na\"ive'' baseline model that
+Furthermore, we hypothesize that it may be possible to exploit biases
+in drug--target interactions to achieve relatively high prediction
+performance without obtaining a biologically meaningful signal.
+For example, hub proteins may have a large number of interactions,
+or certain drugs interact with many proteins, and preferentially
+predicting these interactions may increase predictive performance even
+in the absence of any biological features.
+To test this hypothesis, we design a ``na\"ive'' baseline model that
 predicts the same list of proteins for each drug based only on the

### 2021-03-21 paper/main.tex
.

@@
 predict the same ranked list of drugs for all proteins; consequently,
-this na\"ive predictor will not predict any useful information about
-interactions between drugs and proteins, but allows us to estimate the
-effect of potential biases in our evaluation dataset.
+this na\"ive predictor does not rely on any biological features and
+will not predict any novel information about interactions between
+drugs and proteins; the na\"ive predictor only exploits imbalances in
+the evaluation set to make predictions that may perform
+well. % but allows us to estimate the effect of potential
+% biases in our evaluation dataset.
 % Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
 
-Our na\"ive predictor is not applicable for a protein split
-cross-validation as the number of interactions for each protein in the
-validation set is unknown.  We apply this naive predictor on both the
-STITCH and Yamanishi datasets, using the full dataset as well as
-5-fold cross-validation over drugs as well as over drug--protein
-pairs.  For each fold, we gradually increase $k$ to determine the best
-performance for each fold. Using the full dataset, drug--target split,
-and drug split, we obtain the following MacroAUC results: for the
-STITCH database, we obtain a performance of $0.76$ on the whole
+The way in which we formulated the na\"ive predictor, it is not
+applicable for a protein split cross-validation as the number of
+interactions for each protein in the validation set is unknown.  We
+apply this naive predictor on both the STITCH and Yamanishi datasets,
+using the full datasets as well as a 5-fold cross-validation over
+drugs and over drug--protein pairs to compare the prediction results
+directly to \name.  For each fold, we gradually increase $k$ to
+determine the best performance for each fold. Using the full dataset,
+drug--target split, and drug split, we obtain the following MacroAUC
+results\todo[inline]{Put in Results table, then remove from here}: for
+the STITCH database, we obtain a performance of $0.76$ on the whole
 dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the
@@
 protein-wise cross-validation determining the MacroAUC and
-$MicroAUC_p$. For this evaluation, we allow stratification for the
-training process but not for the validation and testing phase as real
-world applications of these models would have to deal with possibly
+$MicroAUC_p$. For this evaluation, we allow
+stratification\todo{Stratification may not be the right word here; do
+  you mean ``sub-sampling of negatives''?} for the training process
+but not for the validation and testing phase as real world
+applications of these models would have to deal with possibly
 imbalanced data.
@@
 The results of our experiments are summarized in part (c) of Table
-\ref{tab:Results}; we calculated the performance of all compared
-methods over their original splitting scheme and over a protein
-split. First, we find that there is a large difference in performance
-when evaluating over a drug--target pair split compared to a
-protein split, with generally higher performance achieved when using
-the drug--target pair split.
-Second, when evaluating the same methods over a protein split, we find
-a substantial performance difference in comparison to the splitting
+\ref{tab:Results}\todo{Check after restructure}; we calculated the
+performance of all compared methods over their original splitting
+scheme and over a protein split. We find that there is a large
+difference in performance when evaluating over a drug--target pair
+split compared to a protein split, with generally higher performance
+achieved when using the drug--target pair split.  Second, when
+evaluating the same methods over a protein split, we find a
+substantial performance difference in comparison to the splitting
 scheme used in the original evaluation of each method. DTI-CDF was
 originally evaluated on all three splitting schemes underlining this
-point \cite{}. While \name provides comparable performance to DTI-CDF,
-it yields considerably better results with respect to $MicroAUC_p$.
+point \cite{}. While \name provides comparable performance to the
+na\"ive predictor and DTI-CDF in terms of MacroAUC, it yields
+considerably better results with respect to $\textrm{MicroAUC}_p$. We
+also find that methods that are trained using a protein split
+generally result in higher $\textrm{MicroAUC}_p$ than methods trained
+using a drug--target pair split.
 
 As the difference in performance with different splitting schemes is
-quite striking, we evaluated additional drug--target interaction and
-drug--target affinity prediction methods that were trained and
+quite large, we further evaluated additional drug--target interaction
+and drug--target affinity prediction methods that were trained and
 evaluation on other datasets.  Following the results of MolTrans
-(\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}),
-DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and
-MolTrans itself on BioSnap\todo{What is that? Is it in methods?}
-(\textbf{CITATION}) over a protein split, as shown in part (d) of
-Table \ref{tab:Results}. The authors of MolTrans evaluated all
-considered methods over the drug--target pair and the protein split;
-we were able to reproduce the results (Table \ref{tab:Results} (d)),
-showing a substantial difference based on the splitting scheme. We
-additionally computed the $MicroAUC_p$ score for all considered
-methods, leading to comparable\todo{are the datasets comparable?}
-scores as for the results obtained on the Yamanishi dataset.
+\cite{}\todo{add}, we reevaluated DeepDTI \cite{}\todo{add}, DeepDTA
+\cite{}, DeepConv-DTI \cite{} and MolTrans itself on BioSnap\todo{What
+  is that? Is it described in methods?}  \cite{} over a protein split,
+as shown in part (d) of Table \ref{tab:Results}\todo{check table}. The
+authors of MolTrans evaluated all considered methods over the
+drug--target pair and the protein split; we were able to reproduce the
+results (Table \ref{tab:Results} (d))\todo{check table}, showing a
+substantial difference based on the splitting scheme. We additionally
+computed the $MicroAUC_p$ score for all considered methods, leading to
+comparable\todo{are the datasets comparable?} scores as for the
+results obtained on the Yamanishi dataset.
 

### 2021-03-21 paper/main.tex
.

@@
 computed the $MicroAUC_p$ score for all considered methods, leading to
-comparable\todo{are the datasets comparable?} scores as for the
-results obtained on the Yamanishi dataset.
+evaluation results that are comparable\todo{are the datasets
+  comparable?} to the ones obtained on the Yaminishi dataset.
 
@@
 
-\todo[inline]{Discussion:}
-The lack of stratification, only impacts the not
-considered area under precision recall curve (AUPRC) and not the Macro
-AUROC score, while also supporting the expressiveness of $MicroAUC_p$
-with more data points.
-
 %As the baseline predictor is based on the second dataset bias i.e.
@@
 \subsection{``Bottom-up'' and ``top-down'' prediction of drug-target interactions}
+
+There are many computational methods to predict drug--target
+interactions. They can broadly be grouped in two types; the first,
+which we refer to as ``bottom-up'' approaches, start from molecular
+information about a drug and protein and predict an interaction based
+on their molecular properties; the second, which we refer to as
+``top-down'' approaches, start from observable characteristics of an
+organism and infer drug--target interactions as the putative molecular
+mechanisms that explain these observations; many top-down approaches
+follow methods similar to those used in reverse genetics \cite{}.
+
+Another view on these two approaches is as direct and indirect ways to
+predict drug--target interactions. Molecular information can be used
+to directly determine whether two molecules have the ability to
+interact,
 
@@
 %\subsection{Features to use in predicting drug targets}
+
+\todo[inline]{Discussion:}
+The lack of stratification, only impacts the not
+considered area under precision recall curve (AUPRC) and not the Macro
+AUROC score, while also supporting the expressiveness of $MicroAUC_p$
+with more data points.
 

### 2021-03-21 paper/main.tex
.

@@
 predict drug--target interactions. Molecular information can be used
-to directly determine whether two molecules have the ability to
-interact,
-
-What features localize on graph? Molecular interactions based on
-structure not likely -- only possible to find binding sites to other
-proteins.
-
-Phenotype-based features localize...
-
-\todo[inline]{Methods:}
-Through the very nature of the graph convolutional neural network, we
-build the transformed representation for all proteins in every
-forwarding step of the model. Note particularly, that the same
-convolutional filter and weight matrix are applied to and learned for
-all nodes simultaneously. By construction, for a single drug we can
-compute and predict all its interactors in a single run of the model,
-leading to significantly less computing time.
-
-using indications is like cheating, as not applicable for searching new drugs
-
-drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
+to directly determine whether two molecules (such as a drug and
+protein) have the ability to interact, whereas information about
+phenotypic consequences of a drug (drug effects) or disruption of a
+protein function can be used to indirectly suggest candidate
+drug--target interactions. Molecular information will be specific to a
+drug--target pair and we would not expect this information to
+propagate through a protein--protein interaction network; the main
+information about drug--target interactions that could be obtained
+from interactions between proteins is information about binding sites
+between proteins that may also be used by a drug molecule (i.e.,
+information that protein $P_1$ binds to protein $P_2$ reveals
+information about the molecular structures of both $P_1$ and
+$P_2$). On the other hand, phenotypic consequences of changes in
+protein function, or drug effects, are often a result of aberrant
+pathway or network activity and involve more than one protein;
+consequently, we expect these features to benefit more from including
+information about protein--protein interactions. Our results confirm
+this hypothesis and demonstrate that molecular features do not benefit
+from including the interaction network whereas the indirect, top-down
+features benefit from the propagation over the interaction network.
+
+% \todo[inline]{Methods:}
+% Through the very nature of the graph convolutional neural network, we
+% build the transformed representation for all proteins in every
+% forwarding step of the model. Note particularly, that the same
+% convolutional filter and weight matrix are applied to and learned for
+% all nodes simultaneously. By construction, for a single drug we can
+% compute and predict all its interactors in a single run of the model,
+% leading to significantly less computing time.
+
+There are other types of indirect features that could be added to our
+model. A common feature that is added are drug indications which are
+highly predictive of drug--target interactions \cite{}; however, we do
+not include them in our model as including drug indications would
+allow our model to make many trivial predictions based only on
+remembering which targets are often used for which indication;
+including network information would likely benefit predictions based
+on drug indications because different drugs may target the same
+pathway through different mechanisms \cite{}.
 
@@
 normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
-
-protein functions localize on the PPI graph, while molecular features don't
 
@@
 \subsection{Evaluating drug--target interaction predictions}
+
+We spent a significant amount of time testing different evaluation
+schemes for predicting drug--target interactions. This was motivated
+by our finding that a na\"ive predictor can achieve a performance only
+slightly below state of the art drug--target prediction methods.
 

### 2021-03-21 paper/main.tex
.

@@
 We spent a significant amount of time testing different evaluation
-schemes for predicting drug--target interactions. This was motivated
-by our finding that a na\"ive predictor can achieve a performance only
-slightly below state of the art drug--target prediction methods.
+schemes for predicting drug--target interactions. We designed a
+na\"ive predictor and find that it can achieve a performance only
+slightly below state of the art drug--target prediction methods. We
+investigated this finding further and
+
+In the presence of imbalanced
 

### 2021-03-22 paper/main.tex
Updated introduction, added some TODO

--- 2021-03-21 paper/main.tex
@@
 
-\cite{Survey2018}\\
 In history, traditional remedies, that were known for their medicinal
@@
 could be utilized for feature generation, while rapid advances were
-made. Almost all state of the art models for drug-target-interaction
-prediction were based on the usage of neural networks with increasing
-size.\todo{Needs ref, or reformulate}
+made.
 
@@
 graphs. While convolutional filters are suitable for finding patterns
-among the the given graph, attention mechanisms are more relevant for
+among the given graph, attention mechanisms are more relevant for
 discovery of important regions within. Lately, graph learning
@@
 for DTI prediction.
+
+These approaches can be classified into top-down and bottom-up methods.
+Top-down or network approaches hereby start form the observable, phenotypical characteristics, such as side-effects, associated diseases and indications represented by knowledge graphs or ontologies, induced by a drug and infer targets based on the likely molecular mechanisms that result in these phenotypes.
+On the other hand, bottom-up approaches start from molecular features, such as molecular structure, molecular fingerprints, secondary structure and contact predictions and others, deriving predictions from suitable correlations between these features.
+
+However, both bottom--up and top--down approaches to drug--target interaction prediction contain some limitations that are not solvable within themselves.
+Generally, it is very challenging for bottom-up methods to accurately predict, whether there a structure is binding to a protein, as not only molecular structure, but also binding sites and molecular forces are crucial for accurate prediction.
+Thus, finding similarities for dissimilar drugs affecting the same targets is problematic but crucial.
+Furthermore, phenotypic features contain relevant information for DTI prediction \cite{Campillos2008} that are complementary to knowledge gained from molecular properties.
+While bottom-up representations possibly contain all information needed for accurate DTI prediction, we rely on indirect information for better generalization.
+Contrary, top-down approaches lack the ability to spot and cope with small differences within the molecular drug structure, thus both incapable to predict on novel entities and derive similarity from structural correlation.
+
+In order to design representations concerning both top-down and bottom-up features for proteins and drugs,respectively, we utilize protein--protein interaction (PPI) networks which have shown great results in protein function prediction \cite{Vazquez2003} and antiviral drug--target discovery \cite{Ackerman2019}.
+However, only few applied this context to the task of DTI prediction in humans.
+The method of graph convolutional neural networks has been applied successfully to PPI networks \cite{Zitnik2017}, granting valuable information for medical diagnostics, closely related to phenotypic information.
+
+Eventually, we further address potential biases and skews within DTI prediction datasets. \todo[inline]{Introduction}
+Within DTI prediction, there are potential biases resulting from the
+underlying datasets \citep{Pahikkala2014}. First, novel drugs are
+often designed by altering non-functional components of a drug,
+leading to two and more very similar drugs designed to target the same
+proteins \cite{Overington2006}. This can result in a bias when it
+leads to hidden duplicates that can distribute among the train/test
+split, resulting in a better (measured) predictive performance than
+would be expected when the model is applied to identify drugs that
+target a protein for which no drugs yet exist. Second, some proteins
+(which we call \textit{hub proteins}) have significantly more known
+interactions with drugs than others. In the STITCH database, $5\%$ of
+the proteins have $40\%$ of the interactions, and similar
+distributions are present in the Yamanishi and Drugbank
+\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
+these proteins may increase predictive performance while not
+reflecting the actual performance when applied to a new protein (e.g.,
+a protein for which no interactions are known). These differences in
+the number of drugs targeting certain proteins may be the result of
+study bias where more ``valuable'' proteins have more drugs designed
+to target them due to their involvement in more common diseases (or
+diseases for which drugs can be more profitably marketed).
+
+\cite{Survey2018}\\
+
+Almost all state of the art models for drug-target-interaction
+prediction were based on the usage of neural networks with increasing
+size.\todo{Needs ref, or reformulate}
 
@@
 
-\todo[inline]{Introduction}
-Within DTI prediction, there are potential biases resulting from the
-underlying datasets \citep{Pahikkala2014}. First, novel drugs are
-often designed by altering non-functional components of a drug,
-leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can result in a bias when it
-leads to hidden duplicates that can distribute among the train/test
-split, resulting in a better (measured) predictive performance than
-would be expected when the model is applied to identify drugs that
-target a protein for which no drugs yet exist. Second, some proteins
-(which we call \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$ of
-the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins may increase predictive performance while not
-reflecting the actual performance when applied to a new protein (e.g.,
-a protein for which no interactions are known). These differences in
-the number of drugs targeting certain proteins may be the result of
-study bias where more ``valuable'' proteins have more drugs designed
-to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed).
-
 In order to design such a feature for proteins and drugs,
@@
 \subsection{Problem Description}
-\name aims to solve the following problem: for a given drug and a
-given protein we want to determine whether those interact or not.  We
-do not differentiate between types of interaction such as activation
-and inhibition, and do not predict the strength of the interaction.
-If we additionally assume that our knowledge is complete and all
-drug--protein pairs without a known interaction do not interact, we
-can formulate the problem as a binary classification task.
+The issue of predicting drug-target interactions can be described
+quite briefly: For a given drug and a given protein we want to
+determine whether those interact or not.  We do not differentiate
+between types of interaction such as activation and inhibition, and do
+not predict the strength of the interaction.  If we additionally make
+the closed world assumption, i.e., assume that our knowledge is
+complete and all drug--protein pairs without a known interaction do
+not interact, we can formulate the problem as a binary classification
+task.
 
 \subsection{Datasets}
-We obtain a dataset consisting of 12,884 human proteins with over
-340627 links from STRING \citep{STRINGv10}. For the drug-target
-interactions, we use 229,870 links from the STITCH database
+We obtain a dataset consisting of 12884 human proteins with over 340627 links from STRING
+\citep{STRINGv10}. For the drug-target interactions, we
+fetched 229870 links from the STITCH database
 \citep{STITCHv5}. As both STRING and STITCH provide confidence scores
@@
 one existing target. Therefore, the intersection between these
-resources yields 1,428 drugs and 7,368 human proteins with 32,212
-interactions for the training phase. We provide links to and methods
-for obtaining and processing the necessary data on Github.
+resources yields 1,428 drugs and 7,368 human proteins with 32212 interaction links for the training phase. We provide links to and methods for obtaining and processing
+the necessary data on Github.
 
 For comparative evaluation, we use the gold standard dataset
-introduced by \citet{Yaminishi2008} consisting of 1,923 interactions
-between 708 drugs and 1,512 proteins.
+introduced by \citet{Yamanishi2008} consisting of 1923 interactions between 708 drugs and 1512 proteins.
 
@@
 
-\subsubsection{Half-twin neural networks and feature transformation}
+\subsubsection{Half-twin neural networks and modular learnable feature transformation}
+
+\todo[inline]{Looking at the code, I wonder if what is implemented is
+  actually a 
+   ANN; it does not look as if the weights are
+  shared between the two networks, so looks more like a twin or
+  half-twin network? Siamese would be sim(f(e1),f(e2)) but the
+  implementation seems to be sim(f(e1), g(e2)).}
 
 As we want to learn from the similarity of drug side effects and
-protein phenotypes, we use a deep half-twin neural network with a
-contrastive loss using cosine similarity.  A half-twin neural network
-aims to learn a similarity between two embeddings of the same
-dimension.  The precomputed embeddings may have different
-dimensionality and we first process them using a learnable feature
-transformation (LFT) network, i.e., a fully connected neural network
-layer which takes as input an embedding and outputs a representation
-of a particular size. An example structure for both types of features
-can be found in Figure~\ref{fig:HalfTwinNetwork}. We use this LFT to
-reduce the representation size of the embeddings for drugs and
-proteins separately.
-% While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we
-The use of an LFT enables flexible experimentation as both ontology
-and molecular feature for both drugs and proteins are reduced to
-the same dimensionality for varying sizes of inputs; this allows for a
-high amount of modularity across different experimental setups by
-adding different kinds of features into the model. Additionally, the
-generated features may be used for other tasks. % ; for
-% example, the ontology LFT can be reused for a variety of DL2vec-based
-% features with respect to other ontologies and hypotheses
-We follow the results of \textit{DL2vec} \cite{} and use as
-activation function $\sigma := \mathrm{LeakyReLU}$ which leads to
-improved performance compared to other activation functions.
+protein phenotypes, we use a deep half-twin neural network with
+a contrastive loss using cosine similarity.
+A half-twin neural network aims to learn a similarity between two
+embeddings.
+
+% learning a high-dimensional embedding emphasizing this identity by
+% forcing a similarity between these embeddings. On the other hand we
+% built a deep neural network for the molecular structure based
+% features, also benefiting from the siamese network
+% architecture. Computing the similarity between two representations,
+% allows for a variety of different methods.  
+% However, we decided for
+% the cosine similarity measure being invariant to scaling.
+
+Therefore, the precomputed embeddings are run through a regular,
+neural learnable feature transformation (LFT) network, which also
+reduces the eventual representation size for drugs and proteins
+separately. An example structure for both types of features can be
+found in Figure~\ref{fig:HalfTwinNetwork}.
+
+While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we emphasize the versatility of this approach, as both ontology and molecular feature for both drugs and proteins are reduced to similar dimensionality.
+This allows for a high amount of modularity and different experimental setups by plugging different kinds of features into the model.
+Additionally, these pretrained features can be used for a variety of other tasks.
+Additionally, the ontology LFT can be reused for a variety of DL2vec based features with respect to other ontologies and hypotheses.
+We hereby followed the results of \textit{DL2vec}, indicating that utilizing the activation function $\sigma := \mathrm{LeakyReLU}$ leads to performance increase.
 
@@
 
-We include these molecular and ontology-based sub-models within a
-graph neural network (GNN). The graph underlying the GNN is based on
-the protein--protein interaction (PPI) graph. The PPI dataset is
-represented by a graph $G=(V,E)$, where each protein is represented by
-a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$
-represents an interaction between two proteins. Additionally, we
-introduce a mapping $x:V\rightarrow\mathbb{R}^{d}$ projecting each
-vertex $v$ to its node feature $x_v := x(v)$, where $d$ denotes the
-dimensionality of the node features.
+We included these molecular and ontology-based sub-models with a
+larger graph convolutional model. The graph underlying the graph
+convolutional neural network is based on the protein--protein
+interaction (PPI) graph. The PPI dataset is represented by a graph
+$G=(V,E)$, where each protein is represented by a vertex $v\in V$, and
+each edge $e\in E\subseteq V\times V$ symbolizes an interaction
+between two proteins. Additionally, we introduce a mapping
+$x:V\rightarrow\mathbb{R}^{d}$ projecting each vertex $v$ to its node
+feature $x_v := x(v)$, where $d$ denotes the dimensionality of the node
+features.
 
@@
 % methods out there we will only introduce the most basic one here. 
-A graph convolutional layer \cite{GCNConv} consists of a learnable
+A graph convolutional layer \citet{GCNConv} consists of a learnable
 weight matrix followed by an aggregation step, formalized by
@@
 
-The update rule for each node is given by a message passing scheme
-formalized by
+The update rule for each individual node is given by a message passing scheme formalized by
 \begin{equation}
@@
 
-While in this initial formulation of a GCNConv the node-wise update
-step is defined by the sum over all neighbouring node representations,
-we can alter this formulation to other message passing schemes.  We
-can rearrange the order of activation function $\sigma$, aggregation
-$\mathrm{AGG}$, and linear neural layer $\mathrm{MLP}$ with this
+While in this initial formulation the node-wise update step is defined
+by the sum over all neighbouring node representations, we are able to
+alter this formulation to another message passing scheme.  We are able
+to rearrange the order of activation function $\sigma$, aggregation
+$\mathrm{AGG}$ and linear neural layer $\mathrm{MLP}$ with this
 formulation as proposed by \citet{GENConv2020}:
@@
 \end{equation}
-where we only consider
-$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote
-this generalized layer type as \textsc{GENConv} following the
+where we will generally only consider
+$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We will denote
+this generalized layer type as \textsc{GENConv}, following the
 notation of PyTorch Geometric \cite{PytorchGeometric}.  While the
@@
 \textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
-vanishing gradients and signal fading for large scale and highly
+vanishing gradients and signal fading for large scale, highly
 connected graphs, each propagation step in \textsc{GENConv} emphasizes
@@
 connections and a normalization scheme \cite{DeepGCN2019,
-  DeeperGCN2020}.  The residual blocks are reusable and can be stacked
-multiple times. The structure of the GNN architecture we use is shown
-in Figure~\ref{fig:ResGraphConvBlocks}.
+  DeeperGCN2020}.
+The residual blocks are reusable and can be stacked multiple times. %, thereby not losing focus of each node neighbourhood.
+\todo[inline]{Needs a clearer outline/description/figure of YOUR model
+structure.}
+The structure is depicted in Figure~\ref{fig:ResGraphConvBlocks}.
 
@@
 
-\subsubsection{Combined prediction model}
-Combining half-twin and graph convolutional neural networks, we map
-all protein representations to their respective node features,
-initializing the graph convolutional update steps. The resulting
-representations are used for a similarity prediction similar
-to\todo{Similar, or identical to? Needs to describe the exact model
-  used} the models presented in figure ~\ref{fig:HalfTwinNetwork}.
+\subsubsection{title}
+Combining half-twin and graph convolutional neural networks, we map all protein representations to the respective node features, initializing the graph convolutional update steps.
+The resulting representations are used for a similarity prediction similar to the models presented in figure ~\ref{fig:HalfTwinNetwork}.
 When combining ontology and molecular features with or without the
@@
 features, before plugging them into the graph for the similarity
-computation. A more in-depth image of the overall architecture,
-combining both feature types is depicted in figure
-~\ref{fig:FullModelAllLayers}.
+computation. A more in-depth image of the overall architecture, combining both feature types is depicted in figure ~\ref{fig:FullModelAllLayers}.
 
@@
 testing datasets are imbalanced. As there are only $22,336$ links in
-the considered subset the ratio
+the considered STITCH subset, the ratio
 \begin{equation}
@@
 \end{equation}
-and consequently needs to be compensated in the computed loss function
-and appropriate metrics for the evaluation.
-
-Therefore, we weight all positive drug-protein pair samples with this
-ratio by introducing the following loss function with respect to 
+consequently needs suitable compensation in the computed loss function and
+appropriate metrics for the evaluation.
+
+Therefore, we weighted all positive drug-protein pair samples with
+this ratio by introducing the following loss function with respect to
 binary cross-entropy:
@@
 \end{equation}
-for a given prediction $x$ and target $y$, and positive weight $w$
-defined by equation \eqref{weight}. We average this loss among all
-drug-protein pairs in the training set, leading to a stable
-environment for the \textit{Adam} optimization algorithm
-\citep{Adam2014}. We implemented a 5-fold cross validation split among
-the proteins. Furthermore, we used early stopping in the training
-process.
-
-To find the best hyperparameter configuration for the proposed model,
+for a given prediction $x$ and target $y$, and positive weight $w$ defined by equation \eqref{weight}. We average this loss among
+all drug-protein pairs in the training set, leading to a stable
+environment for the used optimizing scheme \textit{Adam}
+\citep{Adam2014}. We implemented a 5-fold cross validation split among the
+proteins. Furthermore, we used early stopping in
+the training process for optima detection.
+
+To find the best hyperparameter configuration for the proposed model
 we performed a grid search to find the most expressive and
@@
 depth of the neural network, optimizer, learning rate and layer types
-using an extensive, manual grid search. Starting from na\"ive, shallow
-feature transformations with an embedding size of $10$, we scaled the
-network up to residual structures with up to $10$ hidden layers
-leading to embeddings of size $4000$, testing different network widths
-and learning rates for each configuration.
+using an extensive, manual grid search, Starting from na\"ive, shallow feature transformations with an embedding size of 10, we scaled the network up to residual structures with up to 10 hidden layers leading to embeddings of size 4000, testing different network widths and learning rates for each configuration.
 
 \subsection{Evaluation and metrics}
-\label{sec:evaluation}
+
 To assess each model, we compute a variety of common metrics for
@@
 We calculate the AUROC by computing true positive rate at various
-false positive rate thresholds and use trapezoidal approximations to
+false positive rate thresholds and using trapezoidal approximations to
 estimate the area under the curve. We refer to this measure as
-$\textrm{MacroAUC}$.
+$MacroAUC$.
 
 \todo[inline]{For RH: check this again later:}
-We also calculate the $\textrm{MicroAUC}$ score. For given lists $D$
-and $P$ of drugs and proteins, respectively, and a set of known
-interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated
-as the average per entity (macro) \textit{AUROC} score. Specifically,
-for protein-centric score, this can be formalized as: given labels
-$l:D\times P \rightarrow \{0,1\}$ and predictions $y:D\times P
-\rightarrow [0,1]$, we define
+In contrast we also calculate the $MicroAUC$ score.
+For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated as the average per entity (macro) \textit{AUROC} score.
+With respect to proteins, this can be formalized for given labels $l:D\times P \rightarrow \{0,1\}$ and predictions $y:D\times P \rightarrow [0,1]$  as
 \begin{equation*}
-  \textrm{MicroAUC}_p := \underset{p\in P}{mean}\left(\left\{
-      \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\})
-    \right\}\right)
+	MicroAUC_p := \underset{p\in P}{mean}\left(\left\{ \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\}) \right\}\right)
 \end{equation*}
 
-Drugs and proteins can be interchanged in this formulation, and we
+Drugs and proteins can be interchanged in this formulation, while we
 refer to the different measures as protein-centric microAUC
-($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$). In some
-cases, the $MicroAUC$ score may not be defined because in some
-datasets some proteins or drugs have no interactions, leading to an
-infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
-for that entity. For those entities, we impute the $MicroAUC$
-interpolating linearly, by using the accuracy for this
-subset\todo{Last sentence not clear, can you give the formula instead?}.
+($MicroAUC_p$) and a drug-centric microAUC
+($MicroAUC_d$). Additionally, the $MicroAUC$ score may not be defined
+as in some datasets some targets or drugs, respectively, do not have
+any interactions, leading to an infeasible $TPR=0$ for all thresholds and an undefined
+$AUROC$ score for that entity. For those entities we impute the $MicroAUC$ interpolating linearly, by using the accuracy for this subset.
 
@@
 function that may ``localize'' on an interaction network (i.e.,
-neighboring nodes will share some of these features or are
+neighboring nodes will share some of these features or be
 phenotypically similar).  As structural features, \name uses
@@
 
-We construct a graph with proteins as nodes and protein-protein
-interactions as edges, mapping the protein features to each target as
-node features. \name then propagates information among the PPI network
-utilizing graph convolutional steps, calculates the similarity of drug
-and protein representations, and predicts whether there is an
-interaction. The full workflow scheme is depicted in
-figure~\ref{fig:ModelWorkflow}
+We construct a PPI graph with proteins as nodes and protein-protein interactions as edges, mapping the protein features to each target as node features.
+\name then propagates information among the PPI network utilizing graph convolutional steps, calculating the similarity of drug and protein representation, and thus eventually predicting whether there is an interaction.
+The full workflow scheme is depicted in figure~\ref{fig:ModelWorkflow}
 
@@
 
-\begin{table*}[!tpb]
-  \centering
+\begin{table}[!tpb]
 	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
@@
 	\label{tab:Results}
-\end{table*}
+\end{table}
 
@@
 We trained, validated and finally tested all considered models on the
-STITCH dataset using a 5-fold cross-validation over a protein split;
-we then selected the best-performing models (with respect to
-\textit{MacroAUC} and $MicroAUC_p$\todo{which is it, macro or micro?},
-see Section \ref{sec:evaluation}) and evaluate them in a 5-fold
-protein-split cross-validation on the Yamanishi benchmark dataset to
-avoid validation overfitting and yield more realistic testing
-results. To evaluate the influence of the different features
-separately, and to determine whether they ``localize'' on the PPI
-graph (and therefore can be exploited successfully by the graph neural
-networks), we train and evaluate models with different types of
+STITCH dataset using a 5-fold cross-validation over a protein
+split; we then selected the best-performing models (with respect to
+\textit{AUROC/MacroAUC} and $MicroAUC_p$) and evaluate them on in 5-fold protein split cross validation on the Yamanishi benchmark dataset to avoid validation overfitting and yield more realistic testing results.
+To evaluate the influence of the different
+features separately, and determine whether they ``localize'' on the
+PPI graph (and therefore can be exploited successfully by the graph
+neural networks), we train and evaluate models with different types of
 features, and with and without inclusion of the PPI graph,
-separately. We comparing the molecular (MolPred) and phenotype-based
-(OntoPred) prediction model, and a combination of both where we
-concatenate both types of features.  Table
-\ref{tab:Results}\todo{check -- not currently in the table} shows the
-results of these experiments.
+separately, comparing the molecular (MolPred) and phenotype based (OntoPred) predictor and a combination of both, concatenating both features.
+Table \ref{tab:Results} (a)+(b) shows the results of these experiments.
 
@@
 the {\em GCNConv} and {\em GENConv} architecture already shows some
-minor improvement, the use of \textit{ResGraphConv} results in larger
-performance improvements.  {\em ResGraphConv} blocks add a large
-amount of additional learnable parameters to the network, leading to
-more expressive power. To test whether the observed improvement is due
-to the number of learnable parameters added or the result of better
-exploiting the information about PPIs, we experiment with a graph
-model in which all graph convolutional neural layers in the residual
-blocks are removed, resulting in a model with similar parameters but
-without the ability to use graph-based information. This pruned
-network, with no information on the protein--protein interactions,
-reached very similar results to the original \textit{OntoPred} model
-and showed no improvement.
-
-The improvements when including the graph are only provided by the
-\textit{GENConv} graph convolution scheme which includes the
-{\em ResGraphConv} blocks; \textit{GCNConv} and other graph
-convolutional methods fail to achieve any gain in comparison to the
-plain \textit{OntoPred} performance even when combined with the
-residual blocks. The discrepancy between {\em GENConv} and other graph
-convolutional methods may be the result of numerical instability and
-fading signals \cite{}.
-
-Our results demonstrate that the inclusion of graph information
-can increase performance when ontology-based features are used but not
-when molecular features are used alone. This observation allows us to
-conclude that information about protein functions localizes on the
-graph whereas molecular features do not.
+minor improvement, the \textit{ResGraphConv} blocks add a large amount
+of additional learnable parameters to the network, leading to more
+expressive power. To test whether the improvement is due to the number
+of learnable parameters added or the result of better exploiting the
+information about PPIs, we experiment with a graph model in which all
+graph convolutional neural layers in the residual blocks are removed,
+resulting in a non--graph model. This pruned network, with no
+information on the protein--protein interactions, reached very similar
+results to the original \textit{OntoPred} model and showed no
+improvement.
+
+Furthermore, these major improvements are only provided by the \textit{GENConv} graph convolution scheme as part of the ResGraphConv blocks, as \textit{GCNConv} and other related graph convolutional methods fail to achieve any gain in comparison to the plain \textit{OntoPred} performance even when combined with the residual blocks.
+The discrepancy of {\em GENConv} and other graph convolutional methods may be founded in numerical stability and fading signals, as described in the introduction of the {\em GENConv} method.
+
+As the inclusion of graph information increases performance, we conclude that information about protein
+functions localizes on the graph while molecular features do not.
 
@@
 The goal of \name is to find candidate drugs that target a specific
-protein; however, so far, we do not evaluate this application but
-rather how \name would perform in finding plausible drug--target
-interactions among all possible interactions (since we use the
-MacroAUC as our main evaluation measure). This evaluation does not
-correspond to the application of \name in finding drugs that target a
-specific protein.  To provide a better estimate on how \name performs
-for individual protein targets, we use micro-averages between proteins
-and compute the \textit{MicroAUC} (see Section \ref{sec:evaluation});
-to determine {\em MicroAUC}, we average the performance (true and
-false positive rates) per protein instead of across all drug--protein
-pairs; the resulting measure can therefore better estimate how \name
-performs when tasked with finding a drug that targets a specific
-protein.
+protein; however, our evaluation does not evaluate this application
+but rather how \name would perform in finding plausible drug--target
+interactions among all possible interactions. To provide a better
+estimate on how \name performs for individual protein targets, we use
+micro-averages between proteins and compute the \textit{MicroAUC};
+micro-averages average the performance (true and false positive rates)
+per protein instead of across all drug--protein pairs.
 
@@
 % \subsection{Baseline model}
-Furthermore, we hypothesize that it may be possible to exploit biases
-in drug--target interactions to achieve relatively high prediction
-performance without obtaining a biologically meaningful signal.
-For example, hub proteins may have a large number of interactions,
-or certain drugs interact with many proteins, and preferentially
-predicting these interactions may increase predictive performance even
-in the absence of any biological features.
-To test this hypothesis, we design a ``na\"ive'' baseline model that
-predicts the same list of proteins for each drug based only on the
-number of known drug--target interactions for a protein. Formally,
-given lists $D$ and $P$ of drugs and proteins and a set of known
-interactions $\mathcal{I} := \{(d_i, p_i) \}$, we construct an
-interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
+Furthermore, we hypothesize that some of the predictions made by our
+model are the result of the biases in drug--target interaction
+datasets. We design a ``na\"ive'' baseline model that predicts the
+same list of proteins for each drugs based on the number of
+known drug--target interactions for a protein. Formally, given lists
+$D$ and $P$ of drugs and proteins and a set of known interactions
+$\mathcal{I} := \{(d_i, p_i) \}$, we construct an interaction matrix
+$M_{int}\in\{0,1\}^{|D|\times|P|}$ with
 \begin{equation*}
@@
 Our ``na\"ive'' predictor $P_k$  predicts all drugs to interact with
-the top $k$ targets with respect to the introduced ranking:
+the top $k$ targets
+% , denoted by $Top_k(P)$ 
+with respect to
+% $M_{ij}$ from
+the introduced ranking:
 \begin{equation*}
@@
 with the only hyperparameter $k$.
-
 The prediction $P_k$ is not dependent on the drug $d_i$ and will
-predict the same ranked list of drugs for all proteins; consequently,
-this na\"ive predictor does not rely on any biological features and
-will not predict any novel information about interactions between
-drugs and proteins; the na\"ive predictor only exploits imbalances in
-the evaluation set to make predictions that may perform
-well. % but allows us to estimate the effect of potential
-% biases in our evaluation dataset.
+predict the same ranked list of drugs for a given protein $p_j$;
+consequently, this na\"ive predictor will not predict any useful
+information about interactions between drugs and proteins, but allows
+us to estimate the effect of potential biases in our evaluation
+dataset.
 % Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
 
-The way in which we formulated the na\"ive predictor, it is not
-applicable for a protein split cross-validation as the number of
-interactions for each protein in the validation set is unknown.  We
-apply this naive predictor on both the STITCH and Yamanishi datasets,
-using the full datasets as well as a 5-fold cross-validation over
-drugs and over drug--protein pairs to compare the prediction results
-directly to \name.  For each fold, we gradually increase $k$ to
-determine the best performance for each fold. Using the full dataset,
-drug--target split, and drug split, we obtain the following MacroAUC
-results\todo[inline]{Put in Results table, then remove from here}: for
-the STITCH database, we obtain a performance of $0.76$ on the whole
-dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the
-drug splitting scheme; on the Yamanishi dataset, we obtain MacroAUC
-scores of $0.88$, $0.84$ and $0.85$, for the total dataset,
-drug--target pair and drug split, respectively.  The na\"ive predictor
-shows higher performance on the Yamanishi dataset than on STITCH, and
-a substantial gain in comparison to an expected random predictor on
-both datasets. In the following, we utilize this nai\"ve predictor as
-baseline to compare its performance to state of the art models and
-\name.
+We apply this naive predictor on both the STITCH and Yamanishi
+datasets, using the full dataset as well as 5-fold cross-validation
+over drugs and drug--protein pairs.
+Note, that this baseline predictor is not applicable for a protein split CV, as the number of interactions for each protein in the validation set is unknown. 
+For each fold, we gradually increase $k$ to determine the best
+performance for each fold. Using the full dataset, drug--target split,
+and drug split, we obtain the following MacroAUC/AUROC score results.
+For STITCH database we obtain a performance of $0.76$ on the whole dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the drug splitting scheme.
+While these results are fairly moderate, on Yamanishi dataset we yield MacroAUC scores of $0.88$, $0.84$ and $0.85$, for total dataset, DT pair split and drug split, respectively.
+
+The na\"ive predictor shows great performance on Yamanishi dataset, with substantial gain in comparison to a random predictor on both datasets.
+In the following, we will utilize this nai\"ve predictor as baseline predictor correlating its performance to state of the art models and \name.
 
@@
 performing methods for drug--target interaction prediction that were
-previously evaluated on the Yamanishi benchmark dataset. These methods
-include DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}
-which have showed superior results in comparison to numerous
-works. Furthermore, we added DTINet \cite{DTINet2017} as method for
-comparison which has been used to develop a number of methods such as
-NeoDTI \cite{NeoDTI2019} with similar methodology.
+previously evaluated on the Yamanishi benchmark dataset. These methods include DTIGEMS+
+\cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}, which have showed
+superior results in comparison to numerous works. Furthermore, we
+added DTINet \cite{DTINet2017} as method for comparison which has been
+used to develop a number of methods such as NeoDTI \cite{NeoDTI2019}
+with similar methodology.
 
@@
 respective AUROC. We further evaluate each model by performing a
-protein-wise cross-validation determining the MacroAUC and
-$MicroAUC_p$. For this evaluation, we allow
-stratification\todo{Stratification may not be the right word here; do
-  you mean ``sub-sampling of negatives''?} for the training process
-but not for the validation and testing phase as real world
-applications of these models would have to deal with possibly
+protein-wise cross-validation determining Macro AUC/AUROC score and
+$MicroAUC_p$. For this evaluation we allow stratification for the
+training process but not for the validation and testing phase as
+real world applications of these models would have to deal with possibly
 imbalanced data.
 
-The results of our experiments are summarized in part (c) of Table
-\ref{tab:Results}\todo{Check after restructure}; we calculated the
-performance of all compared methods over their original splitting
-scheme and over a protein split. We find that there is a large
-difference in performance when evaluating over a drug--target pair
-split compared to a protein split, with generally higher performance
-achieved when using the drug--target pair split.  Second, when
-evaluating the same methods over a protein split, we find a
-substantial performance difference in comparison to the splitting
-scheme used in the original evaluation of each method. DTI-CDF was
-originally evaluated on all three splitting schemes underlining this
-point \cite{}. While \name provides comparable performance to the
-na\"ive predictor and DTI-CDF in terms of MacroAUC, it yields
-considerably better results with respect to $\textrm{MicroAUC}_p$. We
-also find that methods that are trained using a protein split
-generally result in higher $\textrm{MicroAUC}_p$ than methods trained
-using a drug--target pair split.
-
-As the difference in performance with different splitting schemes is
-quite large, we further evaluated additional drug--target interaction
-and drug--target affinity prediction methods that were trained and
-evaluation on other datasets.  Following the results of MolTrans
-\cite{}\todo{add}, we reevaluated DeepDTI \cite{}\todo{add}, DeepDTA
-\cite{}, DeepConv-DTI \cite{} and MolTrans itself on BioSnap\todo{What
-  is that? Is it described in methods?}  \cite{} over a protein split,
-as shown in part (d) of Table \ref{tab:Results}\todo{check table}. The
-authors of MolTrans evaluated all considered methods over the
-drug--target pair and the protein split; we were able to reproduce the
-results (Table \ref{tab:Results} (d))\todo{check table}, showing a
-substantial difference based on the splitting scheme. We additionally
-computed the $MicroAUC_p$ score for all considered methods, leading to
-evaluation results that are comparable\todo{are the datasets
-  comparable?} to the ones obtained on the Yaminishi dataset.
+The results of this analysis are summarized in part (c) of Table \ref{tab:Results}, calculating the performance of all considered methods over their original splitting scheme and over a protein split.
+We find that, first the majority of approaches evaluate over a drug--protein splitting scheme, while second, the highest performances on Yamanishi dataset are yielded over the drug--target pairs splitting scheme, with a noticeable gap to the methods performing their splits over proteins.
+Further, when evaluating the same methods over a protein split, we find a substantial performance divergence in comparison to their original splitting scheme.
+The authors of DTI-CDF evaluate their method on all three splitting schemes underlining this point.
+While \name provides comparable performance to DTI-CDF, it yields considerably better with respect to $MicroAUC_p$.
+
+As we were quite surprised about the MacroAUC discrepancy between the splitting schemes, we evaluated other cutting edge drug--target interaction and drug--target affinity prediction methods, training and evaluating on other datasets.
+Following the results of MolTrans (\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}), DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and MolTrans itself on BioSnap (\textbf{CITATION}) over a protein split, as shown in part (d) of Table \ref{tab:Results}.
+The authors of MolTrans evaluated all considered methods over the drug--target pair and the protein split, which we were able to reproduce with (d), showing once more a substantial divergence.
+However, we additionally computed the $MicroAUC_p$ score for all considered methods, leading to comparable scores as for the Yamanishi evaluated methods.
 
@@
 
+\todo[inline]{Discussion:}
+The lack of stratification, only impacts the not
+considered area under precision recall curve (AUPRC) and not the Macro
+AUROC score, while also supporting the expressiveness of $MicroAUC_p$
+with more data points.
+
 %As the baseline predictor is based on the second dataset bias i.e.
@@
 \section{Discussion}
-
-\subsection{``Bottom-up'' and ``top-down'' prediction of drug-target interactions}
-
-There are many computational methods to predict drug--target
-interactions. They can broadly be grouped in two types; the first,
-which we refer to as ``bottom-up'' approaches, start from molecular
-information about a drug and protein and predict an interaction based
-on their molecular properties; the second, which we refer to as
-``top-down'' approaches, start from observable characteristics of an
-organism and infer drug--target interactions as the putative molecular
-mechanisms that explain these observations; many top-down approaches
-follow methods similar to those used in reverse genetics \cite{}.
-
-Another view on these two approaches is as direct and indirect ways to
-predict drug--target interactions. Molecular information can be used
-to directly determine whether two molecules have the ability to
-interact,
-
-What features localize on graph? Molecular interactions based on
-structure not likely -- only possible to find binding sites to other
-proteins.
-
-Phenotype-based features localize...
 
@@
 
-using indications is like cheating, as not applicable for searching new drugs
-
-drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
-
-we built protein function and ontology based features based on DL2vec
-
-Ontology derived protein function focused features are highly predictive for dtis
-
-We built a versatile template for various features to test localization on the PPI graph
-
-normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
-
-protein functions localize on the PPI graph, while molecular features don't
-
-molecular features work for all (most?) drugs, top-down features only
-for those that have a side-effect profile
-
-\subsection{Evaluating drug--target interaction predictions}
-
-Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
-
-Running split over proteins is harder than, drug and drug protein pair split (see below table)
-
-this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
-
-Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
-
-microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
+%\subsection{\name identifies drugs that target a protein}
 
@@
 
-A naive predictor (ranking proteins) and predict each drug similarly
-achieves cutting edge performance ($87.5$ AUROC for whole dataset,
-$85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
-prot focused microAUC possible. $\rightarrow$ hub proteins
-
-Yamanishi Dataset is only partially suitable for comparing results, if
-everybody just derives a suitable subset (DTIGEMS)
-
-This also applies to drug target affinity prediction. We were hereby
-able to roughly reproduce the results from MolTrans (Bioinformatics)
-on BioSnap
-
-%\subsection{Features to use in predicting drug targets}
-
-\todo[inline]{Discussion:}
-The lack of stratification, only impacts the not
-considered area under precision recall curve (AUPRC) and not the Macro
-AUROC score, while also supporting the expressiveness of $MicroAUC_p$
-with more data points.
-
-%\subsection{\name identifies drugs that target a protein}
-
-\section{Conclusions}
-
-For the bottom-up approach we build a model that only relies on
-molecular features, which we will discuss in more detail in the
-following methods chapter. For the combination of both approaches we
-now attach the predictions to the protein-protein interaction graph as
-node features for future graph learning steps. In this graph we tried
-to find both patterns and regions for each drug that could be of
-interest through application of different graph convolutional layers,
-which in return represent the feature for each protein. Representing
-the drug we take the drug-drug interaction graph and the semantic
-similarity over side effects which we will explain in the following
-paragraphs.
-
-Combining bottom-up and top-down features improves state of the art.
-
-Evaluating DTI predictions is non-trivial due to numerous biases in
-the training and evaluation datasets.
-
-% \subsection{Deification of our method}
-% \begin{itemize}
-% 	\item all AUROC in \% AUROC score on STITCH
-
-% 	\begin{tabular}{|l|r|r|}
-% 		\hline
-% 		DNNT model&Without graph model&With graph model\\
-% 		\hline
-% 		MolPred&$69$&$69$\\
-% 		PhenomeNETPred&$88$&$92$\\
-% 		\hline
-% 		MolPred + PhenomeNETPred & $89$ & $93$\\
-% 		\hline
-% 	\end{tabular}
-% 	\item microAUC for MolPred + PhenomeNETPred on graph is about $93+-$
-% 	\item and on yamanishi dataset
-
-% 	\begin{tabular}{|l|r|r|}
-% 		\hline
-% 		DNNT model&Without graph model&With graph model\\
-% 		\hline
-% 		PhenomeNETPred&$83$&$84$\\
-% 		\hline
-% 		MolPred + PhenomeNETPred & $83$ & $84.5$\\
-% 		\hline
-% 	\end{tabular}
-% 	\item MicroAUC is about $83$
-% \end{itemize}
-
-% \subsection{How to insult other methods}
-% \begin{itemize}
-% 	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
-% 	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
-
-% 	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
-% 		\hline
-% 		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
-% 		\hline
-% 		DTINet&DP pairs&$91$&$84.1$&$67.2$\\
-% 		DTIGEMS+&DP pairs&$93$& $72.2$& $67.8$ \\
-% 		DTI-CDF&Proteins&$83$&$83$&$79$\\
-% 		\hline
-% 	\end{tabular}
-% 	\item A naive predictor (ranking proteins) and predict each drug similarly achieves cutting edge performance ($87.5$ AUROC for whole dataset, $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No prot focused microAUC possible.
+%\section{Findings}
+
+\subsection{Deification of our method}
+\begin{itemize}
+	\item we built protein function and ontology based features based on DL2vec
+	\item Ontology derived protein function focused features are highly predictive for dtis
+	\item We built a versatile template for various features to test localization on the PPI graph
+	\item normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
+	\item protein functions localize on the PPI graph, while molecular features don't
+	\item all AUROC in \% AUROC score on STITCH
+
+	\begin{tabular}{|l|r|r|}
+		\hline
+		DNNT model&Without graph model&With graph model\\
+		\hline
+		MolPred&$69$&$69$\\
+		PhenomeNETPred&$88$&$92$\\
+		\hline
+		MolPred + PhenomeNETPred & $89$ & $93$\\
+		\hline
+	\end{tabular}
+	\item microAUC for MolPred + PhenomeNETPred on graph is about $93+-$
+	\item and on yamanishi dataset
+
+	\begin{tabular}{|l|r|r|}
+		\hline
+		DNNT model&Without graph model&With graph model\\
+		\hline
+		PhenomeNETPred&$83$&$84$\\
+		\hline
+		MolPred + PhenomeNETPred & $83$ & $84.5$\\
+		\hline
+	\end{tabular}
+	\item MicroAUC is about $83$
+\end{itemize}
+
+\subsection{How to insult other methods}
+\begin{itemize}
+	\item Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
+	\item Running split over proteins is harder than, drug and drug protein pair split (see below table)
+	\item this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
+	\item using indications is like cheating, as not applicable for searching new drugs
+	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
+	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
+	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
+
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
+		\hline
+		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
+		\hline
+		DTINet&DP pairs&$91$&$84.1$&$67.2$\\
+		DTIGEMS+&DP pairs&$93$& $72.2$& $67.8$ \\
+		DTI-CDF&Proteins&$83$&$83$&$79$\\
+		\hline
+	\end{tabular}
+	\item A naive predictor (ranking proteins) and predict each drug similarly achieves cutting edge performance ($87.5$ AUROC for whole dataset, $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No prot focused microAUC possible.
 $\rightarrow$ hub proteins
-% 	\item Yamanishi Dataset is only partially suitable \todo{for
-%             comparing results}, if everybody just derives a suitable subset (DTIGEMS)
-% 	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap
-
-% \end{itemize}
-
-% \begin{figure}[!tpb]%figure1
-% 	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
-% 		\hline
-% 		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
-% 		\hline
-% 		DeepDTI&Drugs&$87.6$&$75.9$&$70.1$\\
-% 		DeepDTA&DP pairs&87.6&$76.7$&$69.4$\\
-% 		DeepConv-DTI&DP pairs&88.3&$76.6$&$73.0$\\
-% 		MolTrans&DP pairs&89.5&$77.0$&$74.0$\\
-
-% 		\hline
-% 	\end{tabular}
-% \end{figure}
-
-% \subsection{Tested hypotheses}
-
-% In this work we are testing the following hypotheses:
-% \begin{enumerate}
-% 	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
-% 	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
-% \end{enumerate}
-% We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account.
+	\item Yamanishi Dataset is only partially suitable \todo{for
+            comparing results}, if everybody just derives a suitable subset (DTIGEMS)
+	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap
+
+\end{itemize}
+
+\begin{figure}[!tpb]%figure1
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
+		\hline
+		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
+		\hline
+		DeepDTI&Drugs&$87.6$&$75.9$&$70.1$\\
+		DeepDTA&DP pairs&87.6&$76.7$&$69.4$\\
+		DeepConv-DTI&DP pairs&88.3&$76.6$&$73.0$\\
+		MolTrans&DP pairs&89.5&$77.0$&$74.0$\\
+
+		\hline
+	\end{tabular}
+\end{figure}
+
+\subsection{Tested hypotheses}
+
+In this work we are testing the following hypotheses:
+\begin{enumerate}
+	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
+	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
+\end{enumerate}
+We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account.
 Thus, we propose a novel approach to combine those mutual exclusive attempts, through the usage of interaction networks, similarity and molecular features.
@@
 
-% \section{Methods}
-
-% \subsection{Models}
-% The used model consists of two separate models, that help to fuse together the two methods:
-% \begin{enumerate}
-% 	\item The molecular predictor
-% 	\item The interaction network based predictor
-% \end{enumerate}
-% We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins.
+For the bottom-up approach we build a model that only relies on molecular features, which we will discuss in more detail in the following methods chapter.
+For the combination of both approaches we now attach the predictions to the protein-protein interaction graph as node features for future graph learning steps.
+In this graph we tried to find both patterns and regions for each drug that could be of interest through application of different graph convolutional layers, which in return represent the feature for each protein.
+Representing the drug we take the drug-drug interaction graph and the semantic similarity over side effects which we will explain in the following paragraphs.
+
+\section{Methods}
+
+\subsection{Models}
+The used model consists of two separate models, that help to fuse together the two methods:
+\begin{enumerate}
+	\item The molecular predictor
+	\item The interaction network based predictor
+\end{enumerate}
+We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins.
 Regarding proteins, we used the pretrained feature generator from \textit{DeepGoPlus} (\citep{DeepGoPlus}) that was originally designed for protein function prediction and is regarded as state of the art for this purpose.
@@
 \\
-% The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE).
+The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE).
 Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug.
@@
 \\
-% The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node.
+The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node.
 In contrast to learning over whole graphs we perform node classification within the graph.
@@
 \\
-% The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
-% For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy.
+The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
+For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy.
 Concerning this hierarchy, drug-drug similarity is computed by the Resnik similarity (\cite{Resnik1995}).
@@
 
-% Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
+Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
 This network finally yields our prediction.
@@
 
-% \begin{figure}[!tpb]%figure1
-% 	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
-% 	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
-% \end{figure}
-
-% \begin{figure}[!tpb]%figure2
-% \centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
-% \caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.
+\begin{figure}[!tpb]%figure1
+	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
+	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
+\end{figure}
+
+\begin{figure}[!tpb]%figure2
+\centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
+\caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.
 Protein and drug features are represented by blue and red, respectively.}\label{fig:02}
-% \end{figure}
-
-%\section{Discussion}
+\end{figure}
+
+\section{Discussion}
 

### 2021-03-22 paper/main.tex

--- 2021-03-21 paper/main.tex
@@
 \end{equation}
-and consequently needs to be compensated in the computed loss function
-and appropriate metrics for the evaluation.
+consequently needs suitable compensation in the computed loss function and
+appropriate metrics for the evaluation.
 

### 2021-03-22 paper/main.tex
Updated intro but trailing transition, added citations, updated results table, Updated overall model methods section

@@
 to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed).
-
-\cite{Survey2018}\\
-
-Almost all state of the art models for drug-target-interaction
-prediction were based on the usage of neural networks with increasing
-size.\todo{Needs ref, or reformulate}
-
-Approaches on this rather sophisticated\todo{Try to avoid}
-problem\todo{Need to state the problem clearly here or at end of prev
-  paragraph} can divided into top-down or network approaches
-(\textbf{CITATION}), and bottom-up or molecular approaches
-(\textbf{CITATION}). Top-down approaches take advantage of other data
-such as diseases (CITATION), side effects, knowledge graphs or
-ontologies, in order to learn representations for both compound and
-protein. \todo[inline]{We call these approaches ``top-down'' because
-  they start with the observable characteristics induced by a drug and
-  infer the targets based on the likely molecular mechanisms that
-  result in these phenotypes.} On the other hand, bottom-up approaches
-attempt to learn from chemical properties of proteins or drugs to
-infer candidate drug--target interactions. For drugs, molecular
-structure (CITATION GraphDTA), molecular fingerprints, similarity to
-other drugs (See Bioinf Survey), and other molecular features may be
-used. On the protein side, secondary structure prediction (CITATION),
-contact prediction (CITATION), or simply\todo{Don't use ``simply''.}
-convolution over the amino acid sequences can be used to obtain a
-feature representation for a given proteins. However, both bottom-up
-and top-down approaches to drug--target interaction prediction
-\todo[inline]{replace: ``contain and share some problems'' with
-  something like ``have some limitations''} that are not solvable within
-themselves.
-\todo[inline]{Following is not sufficiently precise; here, you need to
-clearly state the challenges faced by both approaches, ideally with
-references.}
-Thus, bottom-up approaches share the lack of ability to
-generalize, which we will show in later sections, and usually focus on
-engineering sophisticated features for the drugs, while neglecting to
-formulate meaningful features on the protein side. Top-down approaches
-lack the ability to spot small differences to cope with small
-differences within the drug structure and rely heavily on given data
-for the considered drug-target pair. The latter is not suitable for
-predictions on novel or unseen compounds, as e.g., data on side
-effects or its impact on diseases is seldom given for novel drugs.
-
-\todo[inline]{Introduction}
-Within DTI prediction, there are potential biases resulting from the
-underlying datasets \citep{Pahikkala2014}. First, novel drugs are
-often designed by altering non-functional components of a drug,
-leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can result in a bias when it
-leads to hidden duplicates that can distribute among the train/test
-split, resulting in a better (measured) predictive performance than
-would be expected when the model is applied to identify drugs that
-target a protein for which no drugs yet exist. Second, some proteins
-(which we call \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$ of
-the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
-these proteins may increase predictive performance while not
-reflecting the actual performance when applied to a new protein (e.g.,
-a protein for which no interactions are known). These differences in
-the number of drugs targeting certain proteins may be the result of
-study bias where more ``valuable'' proteins have more drugs designed
-to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed).
-
-In order to design such a feature for proteins and drugs,
-respectively, we make use of the interaction networks for both
-proteins and compounds. Drug-drug interaction networks were introduced
-and standardized by \citet{Boyce2015} and have been used for clinical
-decision support \cite{Scheife2015}. Drug-drug interaction networks
-may give a hint on common targeted pathways. As an additional compound
-feature we will use semantic side effect similarity, which we will
-discuss later on\todo{Generally, try to avoid pointers to ``later''.}.
-
-Protein-protein interaction networks have shown great results in
-$\dots$(\cite{Vazquez2003}, \cite{Ackerman2019}) in granting context
-for molecular system biology. However, these contexts were never
-applied to the problem of drug-target-interaction prediction. Thus we
-formalized our hypotheses over these interaction graphs and will test
-them in the following chapters.
+diseases for which drugs can be more profitably marketed). This may affect common \cite{Survey2018} cross-validation splitting schemes, that allow exploitation of these biases within DTI prediction.
+
+\todo[inline]{Somehow smooth this transition}
 
@@
 initializing the graph convolutional update steps. The resulting
-representations are used for a similarity prediction similar
-to\todo{Similar, or identical to? Needs to describe the exact model
-  used} the models presented in figure ~\ref{fig:HalfTwinNetwork}.
+representations are used for a similarity prediction as presented in figure ~\ref{fig:HalfTwinNetwork}. 
 When combining ontology and molecular features with or without the
 graph model, we concatenate both protein features and both drugs
-features, before plugging them into the graph for the similarity
-computation. A more in-depth image of the overall architecture,
-combining both feature types is depicted in figure
-~\ref{fig:FullModelAllLayers}.
+features, before plugging them into the graph model for the similarity
+computation. A more in-depth image of the overall and final architecture,
+combining both feature types, is depicted in figure
+~\ref{fig:FullModelAllLayers}. Here the original representations are transformed by LFTs and then run through a stack (here with height 3) of residual graph convolutional blocks.
 
@@
 \begin{equation*}
-  \textrm{MicroAUC}_p := \underset{p\in P}{mean}\left(\left\{
+  \textrm{MicroAUC}'_p(l,y) := \underset{p\in P}{mean}\left(\left\{
       \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\})
@@
 
-Drugs and proteins can be interchanged in this formulation, and we
-refer to the different measures as protein-centric microAUC
-($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$). In some
-cases, the $MicroAUC$ score may not be defined because in some
+In some cases, the $MicroAUC'$ score may not be defined as in some
 datasets some proteins or drugs have no interactions, leading to an
@@
 interpolating linearly, by using the accuracy for this
-subset\todo{Last sentence not clear, can you give the formula instead?}.
+subset:
+
+\begin{equation*}
+	\textrm{MicroAUC}_p(l,y) := 
+	\begin{cases}
+		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in D}l(d_i,p)\neq0\\
+		Accuracy(l,y)&otherwise
+	\end{cases}
+\end{equation*}
+
+Note that drugs and proteins can be interchanged in this formulation, and we
+refer to the different measures as protein-centric microAUC
+($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$).
 
@@
   \centering
-	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
-		\hline
-		(a) STITCH results&\multicolumn{4}{c|}{PPI graph}\\
-		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
-		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$\\
-		\hline
-		MolPred&$0.69$&$0.65$&$0.69$&$0.67$\\
-		\hline
-		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$\\
-		\hline
-		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$\\
-		\hline
-		\hline
-		(b) Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
-		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
-		&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
-		\hline
-		\name (MolPred + OntoPred) & $0.83$ & $0.82$&$0.84$&$0.84$\\
+	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
+		\hline
+		&\multicolumn{4}{c|}{(a) STITCH results}& \multicolumn{4}{c|}{(b) Yamanishi results}\\
+		\hline
+		\name results&\multicolumn{4}{c|}{PPI graph}&\multicolumn{4}{c|}{PPI graph}\\
+		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
+		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
+		\hline
+		MolPred&$0.69$&$0.65$&$0.69$&$0.67$&$0.66$&$0.67$&$0.66$&$0.64$\\
+		\hline
+		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$&$0.80$&$0.79$&$0.83$&$0.82$\\
+		\hline
+		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
 		\hline
@@
 STITCH dataset using a 5-fold cross-validation over a protein split;
-we then selected the best-performing models (with respect to
-\textit{MacroAUC} and $MicroAUC_p$\todo{which is it, macro or micro?},
+we then selected the best-performing models (with respect to $MicroAUC_p$,
 see Section \ref{sec:evaluation}) and evaluate them in a 5-fold
@@
 $MicroAUC_p$. For this evaluation, we allow
-stratification\todo{Stratification may not be the right word here; do
-  you mean ``sub-sampling of negatives''?} for the training process
+sub-sampling of negatives for the training process
 but not for the validation and testing phase as real world
@@
 evaluation on other datasets.  Following the results of MolTrans
-\cite{}\todo{add}, we reevaluated DeepDTI \cite{}\todo{add}, DeepDTA
-\cite{}, DeepConv-DTI \cite{} and MolTrans itself on BioSnap\todo{What
+\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017}, DeepDTA
+\cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019} and MolTrans itself on BioSnap\todo{What
   is that? Is it described in methods?}  \cite{} over a protein split,

### 2021-03-23 paper/main.tex
Added BioSnap integration to preprocessing and predictor

@@
 		\hline
+		Naive predictor&DP pairs&$0.79$&-- &--\\
 		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
@@
 
+\begin{itemize}
+	\item \name comparison to other approaches on BioSnap prone to bias, as $ BioSnap: drugs 
+	\times prots = 5500 \times 3000 $, while only 
+	$\approx$ 1000 drugs have Sider Annotation, and 
+\end{itemize}
+
 \subsection{Evaluating drug--target interaction predictions}

### 2021-03-23 src/DTI_data_preparation.py

@@ -368,12 +368,11 @@ def get_protein_to_EnsemblProtein_id():
     return protein_to_Ensembl_protein_id_mapping
 
 def get_yamanishi_data(original_drug_list, original_protein_list):
+    path = '../data/Yamanishi_data/'
+
     # yamanishi_drug_mapping = DDI_utils.get_Yamanishi_db_to_PubChem_mapping_dict()
     yamanishi_drug_mapping = DDI_utils.get_STITCH_db_Pubchem_mapping_dict()
-
     yamanishi_protein_mapping = PPI_utils.get_protein_Yamanishi_to_STITCH_mapping()
-
-    path = '../data/Yamanishi_data/'
 
     drug_list = []
     with open(file=path + 'drug.txt', mode='r') as f:
@@ -407,6 +406,68 @@ def get_yamanishi_data(original_drug_list, original_protein_list):
 
     return drug_list[drug_indices], protein_list[protein_indices], dti_matrix[drug_indices,:][:, protein_indices]
 
+def get_BioSnap_data(original_drug_list, original_protein_list):
+    path = '../data/BioSnap_data/'
+
+    # fetch DrugBank to Pubchem and UniProt to STRING mapping
+    drug_mapping = DDI_utils.get_STITCH_db_Pubchem_mapping_dict()
+    protein_mapping = {}
+    with open(file=path+'UniProt_to_STRING_mapping.tab', mode='r') as f:
+        for line in f:
+            uniprot_id, STRING_id = line.strip().split('\t')
+            protein_mapping[uniprot_id] = STRING_id
+
+    # build drug list
+    drug_list = []
+    with open(file=path+'BioSnap_DB_drug_list', mode='r') as f:
+        for line in f:
+            drug_list.append(line.strip())
+    print('drug_list', len(drug_list))
+
+    # build protein list
+    protein_list = []
+    with open(file=path + 'BioSnap_gene_list', mode='r') as f:
+        for line in f:
+            protein_list.append(line.strip())
+    print('protein_list', len(protein_list))
+
+    train_file = 'train.csv'
+    val_file = 'val.csv'
+    test_file = 'test.csv'
+
+    # parse all BioSnap data as preprocessed in MolTrans
+    full_list = []
+    for file in [train_file, val_file, test_file]:
+        print(f'Parsing {path + file}')
+        with open(file=path + file, mode='r') as f:
+            f.readline()
+            for i, line in enumerate(f):
+                full_list.append(line.strip().split(','))
[…]

### 2021-03-23 src/HPO_predictor.py

@@ -26,7 +26,7 @@ class HPODTIDataBuilder:
         self.config = config
 
         print("Loading data ...")
-        if config.yamanishi_test:
+        if config.yamanishi_test or config.biosnap_test:
             self.drug_list = np.array(PhenomeNET_DL2vec_utils.get_PhenomeNET_drug_list())
             if config.include_indications:
                 indication_drug_list = set(PhenomeNET_DL2vec_utils.get_UMLS_drug_list())
@@ -43,13 +43,16 @@ class HPODTIDataBuilder:
         dti_graph = DTI_data_preparation.get_human_DTI_graph(mode=config.mode)
         PPI_graph = PPI_utils.get_PPI_graph(min_score=700)
 
-        if config.yamanishi_test:
+        if config.yamanishi_test or config.biosnap_test:
             self.protein_list = np.array(list(set(PPI_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
             print("Loading Yamanishi data ...")
-            self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
+            if config.yamanishi_test:
+                self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
+            elif config.biosnap_test:
+                self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
+
             print(self.drug_list.shape, self.y_dti_data.shape, self.protein_list.shape)
         else:
-            self.protein_list = np.array(list(set(PPI_graph.nodes()) & set(dti_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
             if config.include_indications:
                 print("Drug indications can only be selected with yamanishi test for now.")
                 raise ValueError
@@ -71,7 +74,7 @@ class HPODTIDataBuilder:
     def build_data(self, config):
 
         # DTI data
-        if not config.yamanishi_test:
+        if not config.yamanishi_test and not config.biosnap_test:
             print("Loading DTI links ...")
             y_dti_data = DTI_data_preparation.get_DTIs(drug_list=self.drug_list, protein_list=self.protein_list,
                                                        mode=config.mode)
@@ -428,6 +431,7 @@ if __name__ == '__main__':
[…]

### 2021-03-23 src/PhenomeNET_DL2vec_utils.py

@@ -48,7 +48,7 @@ def get_UMLS_drug_list():
         return pkl.load(f)
 
 
-def write_PhenomeNET_files(mode='all'):
+def write_PhenomeNET_files(mode='all', include_indications=False):
     path_prefix = "../data/PhenomeNET_data/"
     onto_prefix = "<http://purl.obolibrary.org/obo/{entity}>"
 

### 2021-03-23 src/dti_utils.py

@@ -51,11 +51,11 @@ def micro_AUC_per_prot_DT_pairs(y_true, y_pred, num_drugs, indices):
 
     # collect all interactors for each protein w.r.t. indices
     for index in indices:
-        print(index, len(prot_wise_index_list), index%num_drugs, num_prots)
+        print(index, len(prot_wise_index_list))
         prot_wise_index_list[index%num_drugs].append(index)
 
     for prot_index in range(num_prots):
-        if prot_wise_index_list:
+        if prot_wise_index_list[prot_index]:
             pair_indices = np.array(prot_wise_index_list[prot_index])
             prot_y_true = y_true.flatten()[pair_indices]
             prot_y_pred = y_pred.flatten()[pair_indices]

### 2021-03-23 src/results_evaluation.py

@@ -236,7 +236,7 @@ def test_Yamanishi_AUC():
         print(results.mean())
 
 def parse_BioSnap():
-    path = 'data/BioSnap_data/'
+    path = '../data/BioSnap_data/'
 
     train_file = 'train.csv'
     val_file = 'val.csv'
@@ -311,6 +311,9 @@ def parse_BioSnap():
               f'{dti_utils.micro_AUC_per_drug(dti_matrix, y_pred, num_drugs)}')
 
 def naive_predictor_pair_split(dti_matrix):
+    # naive predictor on DT pair data
+    # note that the specific micro_AUC computation needed, is extremely slow.
+
     kf = KFold(n_splits=5, random_state=42, shuffle=False)
 
     num_drugs, num_prots = dti_matrix.shape

### 2021-03-23 src/torch_dti_predictor.py

@@ -229,6 +229,7 @@ if __name__ == '__main__':
     parser.add_argument("--include_mol_features", action='store_true')
 
     parser.add_argument("--yamanishi_test", action='store_true')
+    parser.add_argument("--biosnap_test", action='store_true')
     parser.add_argument("--include_indications", action='store_true')
 
 

### 2021-03-23 src/torch_dti_utils.py

@@ -24,13 +24,12 @@ class QuickProtFuncDTINetworkData:
         self.config = config
 
         print("Loading data ...")
-        if config.yamanishi_test:
+        if config.yamanishi_test or config.biosnap_test:
             self.drug_list = np.array(PhenomeNET_DL2vec_utils.get_PhenomeNET_drug_list())
         else:
             self.drug_list = np.array(DTI_data_preparation.get_drug_list(config.mode))
         print(len(self.drug_list), "drugs present")
 
-        print(len(self.drug_list), "drugs present")
         # self.protein_list = np.array(DTI_data_preparation.get_human_prot_func_proteins())[:config.num_proteins]
         # get protein lists for each ontology
         uberon_protein_list = PhenomeNET_DL2vec_utils.get_PhenomeNET_protein_list(mode='uberon')
@@ -40,7 +39,6 @@ class QuickProtFuncDTINetworkData:
         dti_graph = DTI_data_preparation.get_human_DTI_graph(mode=config.mode)
         self.PPI_graph = PPI_utils.get_PPI_graph(min_score=config.PPI_min_score)
         self.protein_list = np.array(list(set(self.PPI_graph.nodes()) & set(dti_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
-
 
         if config.include_mol_features:
             print("Building molecular features")
@@ -62,11 +60,13 @@ class QuickProtFuncDTINetworkData:
             self.protein_mol_encodings = torch.Tensor([protein_to_feature_dict[protein] for protein in self.protein_list])
 
             print(len(self.protein_list), "proteins present with mol_pred_intersection.\n")
-        if config.yamanishi_test:
+        if config.yamanishi_test or config.biosnap_test:
             self.protein_list = np.array(list(set(self.PPI_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
             print("Loading Yamanishi data ...")
-            self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
-
+            if config.yamanishi_test:
+                self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
+            elif config.biosnap_test:
[…]

### 2021-03-28 paper/main.tex
.

@@
 \usepackage{natbib}
-\bibliographystyle{natbib.bst}
+\bibliographystyle{natbib}
 
@@
 \subtitle{Subject Section}
-\newcommand{\name}{Voodoo }
-
-\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
+\newcommand{\name}{Voodoo}
+
+\title[\name{}]{\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
 \author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
@@
 \subsection{Problem Description}
-\name aims to solve the following problem: for a given drug and a
+\name{} aims to solve the following problem: for a given drug and a
 given protein we want to determine whether those interact or not.  We
@@
 
-\subsection{\name: computational model to identify drugs that target a
+\subsection{\name{}: computational model to identify drugs that target a
   protein}
@@
 
-We developed \name as a computational model to predict drug--target
-interactions. Specifically, given a protein, \name will identify and
-rank drugs that likely target this protein. \name combines two types
+We developed \name{} as a computational model to predict drug--target
+interactions. Specifically, given a protein, \name{} will identify and
+rank drugs that likely target this protein. \name{} combines two types
 of features: structural information for drugs and proteins that can be
@@
 neighboring nodes will share some of these features or are
-phenotypically similar).  As structural features, \name uses
+phenotypically similar).  As structural features, \name{} uses
 structural representations of drugs from the SMILES transformer
 \cite{SmilesTransformer} and representations of protein amino acid
-sequences from DeepGOPlus \cite{DeepGoPlus}.  \name learns
+sequences from DeepGOPlus \cite{DeepGoPlus}.  \name{} learns
 representations of drug effects and protein functions using the
@@
 interactions as edges, mapping the protein features to each target as
-node features. \name then propagates information among the PPI network
+node features. \name{} then propagates information among the PPI network
 utilizing graph convolutional steps, calculates the similarity of drug
@@
 		\hline
-		\name results&\multicolumn{4}{c|}{PPI graph}&\multicolumn{4}{c|}{PPI graph}\\
+		\name{} results&\multicolumn{4}{c|}{PPI graph}&\multicolumn{4}{c|}{PPI graph}\\
 		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
@@
 		\hline
-		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
+		\name{} (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
 		\hline
@@
 		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
-		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
+		\name{}&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
 		\hline
@@
 		\hline
-	\end{tabular}\\
-
-	\caption{(a) + (b) Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation.
+	\end{tabular}
+
+	\caption{(a) + (b) Results for \name{} on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation.
 We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.
@@
 \subsection{Protein-centric evaluation}
-The goal of \name is to find candidate drugs that target a specific
+The goal of \name{} is to find candidate drugs that target a specific
 protein; however, so far, we do not evaluate this application but
-rather how \name would perform in finding plausible drug--target
+rather how \name{} would perform in finding plausible drug--target
 interactions among all possible interactions (since we use the
 MacroAUC as our main evaluation measure). This evaluation does not
-correspond to the application of \name in finding drugs that target a
-specific protein.  To provide a better estimate on how \name performs
+correspond to the application of \name{} in finding drugs that target a
+specific protein.  To provide a better estimate on how \name{} performs
 for individual protein targets, we use micro-averages between proteins
@@
 false positive rates) per protein instead of across all drug--protein
-pairs; the resulting measure can therefore better estimate how \name
+pairs; the resulting measure can therefore better estimate how \name{}
 performs when tasked with finding a drug that targets a specific
@@
 drugs and over drug--protein pairs to compare the prediction results
-directly to \name.  For each fold, we gradually increase $k$ to
+directly to \name{}.  For each fold, we gradually increase $k$ to
 determine the best performance for each fold. Using the full dataset,
@@
 baseline to compare its performance to state of the art models and
-\name.
+\name{}.
 
@@
 originally evaluated on all three splitting schemes underlining this
-point \cite{}. While \name provides comparable performance to the
+point \cite{}. While \name{} provides comparable performance to the
 na\"ive predictor and DTI-CDF in terms of MacroAUC, it yields
@@
 
-\todo[inline]{ add \name model to table to allow
+\todo[inline]{ add \name{} model to table to allow
   DIRECT comparison:}
@@
 \begin{itemize}
-	\item \name comparison to other approaches on BioSnap prone to bias, as $ BioSnap: drugs 
+	\item \name{} comparison to other approaches on BioSnap prone to bias, as $ BioSnap: drugs 
 	\times prots = 5500 \times 3000 $, while only 

### 2021-03-29 paper/main.tex
Update table, naive predictor on BioSnap

--- 2021-03-23 paper/main.tex
@@
 introduced by \citet{Yaminishi2008} consisting of 1,923 interactions
-between 708 drugs and 1,512 proteins.
+between 708 drugs and 1,512 proteins, and BioSnap dataset \cite{BioSnap2018}, which consists of 5,017 drug nodes, 2,324 gene nodes and 15,138 edges.
 
@@
 
-	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
 		\hline
-		(c) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
+		&&\multicolumn{3}{c|}{(c) Yamanishi results}&\multicolumn{3}{c|}{(d) BioSnap results}\\
+		Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}&Original scheme&\multicolumn{2}{c|}{Protein split}\\
+
+		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$&Macro AUC&Macro AUC&Micro $AUC_p$\\
 		\hline
-		Naive predictor&DP pairs&$0.85$& --&--\\
-		DTINet&DP pairs&$0.91$&$0.74$&$0.67$\\
-		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
-		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
-		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
+		Naive predictor&DP pairs&$0.85$& --&--&$0.79$&-- &--\\
+		DTINet&DP pairs&$0.91$&$0.74$&$0.67$ & --&--&--\\
+		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ & --&--&--\\
+		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$& --&--&--\\
 		\hline
+		DeepDTI&Drugs& --&--&--&$0.88$&$0.76$&$0.70$\\
+		DeepDTA&DP pairs& --&--&--&$0.88$&$0.77$&$0.69$\\
+		DeepConv-DTI&DP pairs& --&--&--&$0.88$&$0.76$&$0.73$\\
+		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
+		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$& & & \\
 		\hline
-		(d) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
-		\hline
-		Naive predictor&DP pairs&$0.79$&-- &--\\
-		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
-		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
-		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
-		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
-		\hline
-	\end{tabular}\\
+	\end{tabular}\medskip
 

### 2021-03-29 src/results_evaluation.py

@@ -236,7 +236,7 @@ def test_Yamanishi_AUC():
         print(results.mean())
 
 def parse_BioSnap():
-    path = '../data/BioSnap_data/'
+    path = 'data/BioSnap_data/'
 
     train_file = 'train.csv'
     val_file = 'val.csv'
@@ -283,6 +283,8 @@ def parse_BioSnap():
         dti_matrix[drug_index, gene_index] = label
 
     print(dti_matrix.sum(), dti_matrix.shape)
+
+    raise Exception
 
     print(f'Number of genes with at least one interactor {(dti_matrix.sum(axis=0)>0).sum()}')
 

### 2021-03-29 paper/main.tex
Fix typo in graph convolution biosnap data preparation

--- 2021-03-28 paper/main.tex
@@
 \subsection{Evaluation and metrics}
+
+\subsubsection{Splitting schemes}
+
+As DTI prediction is dependent on both drugs and proteins, there are multiple ways of determining training, validation and testing sets of pairs to evaluate each model.
+For cross-validation, we can perform the split over DTI pairs, over drugs and over proteins, respectively, where only in the latter two schemes unseen entities are shown to the model in the validation and testing phase.
+Additionally, these schemes have diverging expressiveness with respect to DTI prediction, as different information is known and leaked across training and testing phase.
+
+\subsubsection{Metrics}
 \label{sec:evaluation}
@@
 
-Note that drugs and proteins can be interchanged in this formulation, and we
+Note that drugs and proteins can be interchanged in this formulation. We
 refer to the different measures as protein-centric microAUC
@@
 
-	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+
 		\hline
-		(c) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
+		&&\multicolumn{3}{c|}{(c) Yamanishi results}&\multicolumn{3}{c|}{(d) BioSnap results}\\
+		Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}&Original scheme&\multicolumn{2}{c|}{Protein split}\\
+
+		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$&Macro AUC&Macro AUC&Micro $AUC_p$\\
 		\hline
-		Naive predictor&DP pairs&$0.85$& --&--\\
-		DTINet&DP pairs&$0.91$&$0.74$&$0.67$\\
-		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
-		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
-		\name{}&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
+		Naive predictor&DP pairs&$0.85$& --&--&$0.79$&-- &--\\
+		DTINet&DP pairs&$0.91$&$0.74$&$0.67$ & --&--&--\\
+		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ & --&--&--\\
+		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$& --&--&--\\
 		\hline
+		DeepDTI&Drugs& --&--&--&$0.88$&$0.76$&$0.70$\\
+		DeepDTA&DP pairs& --&--&--&$0.88$&$0.77$&$0.69$\\
+		DeepConv-DTI&DP pairs& --&--&--&$0.88$&$0.76$&$0.73$\\
+		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
+		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$& & & \\
 		\hline
-		(d) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
-		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
-		\hline
-		Naive predictor&DP pairs&$0.79$&-- &--\\
-		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
-		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
-		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
-		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
-		\hline
-	\end{tabular}
+	\end{tabular}\medskip
 
@@
 concatenate both types of features.  Table
-\ref{tab:Results}\todo{check -- not currently in the table} shows the
+\ref{tab:Results} shows the
 results of these experiments.

### 2021-03-29 src/HPO_predictor.py

@@ -353,21 +353,9 @@ def siamese_drug_protein_network(config):
                           dti_utils.dti_f1_score(test_labels, test_predictions),
                           dti_utils.dti_mcc(test_labels, test_predictions))#@TODO, file=f)
 
-                    metrics_func_list = [metrics.accuracy_score, dti_utils.dti_auroc, metrics.average_precision_score, dti_utils.dti_f1_score,
-                                         metrics.matthews_corrcoef]
-                    ret = [list_fun(test_labels, test_predictions) for list_fun in metrics_func_list]
-
-                    test_loss = metrics.log_loss(test_labels, test_predictions, eps=0.000001)
-                    if test_loss < best_loss:
-                        best_loss = test_loss
-                        best_epoch = epoch
-
-                        print('rmse improved at epoch ', best_epoch, '; best_test_loss, best_test_ci:', best_test_loss,
-                              best_test_ci, model_st)
-                    else:
-                        print(test_loss, 'No improvement since epoch ', best_epoch, ';', model_st)
 
                     test_AUROC = dti_utils.dti_auroc(test_labels, test_predictions)
+
                     if test_AUROC > best_AUROC:
                         state_dict_path = '../models/HPO_models/hpo_pred_fold_'+str(fold)+'_model'
                         torch.save(model.state_dict(), state_dict_path)

### 2021-03-29 src/torch_dti_utils.py

@@ -118,7 +118,7 @@ class QuickProtFuncDTINetworkData:
         # self.edge_attr = torch.ones((self.edge_list.size(1),1), dtype=torch.float)
 
         # DTI data
-        if not config.yamanishi_test and not config.biosnapt_test:
+        if not config.yamanishi_test and not config.biosnap_test:
             print("Loading DTI links ...")
             y_dti_data = DTI_data_preparation.get_DTIs(drug_list=self.drug_list, protein_list=self.protein_list, mode=config.mode)
             self.y_dti_data = y_dti_data.reshape((len(self.drug_list), len(self.protein_list)))

### 2021-03-29 src/HPO_predictor.py
Fix typo in graph convolution biosnap data preparation 2

@@ -232,7 +232,7 @@ class HPOPredNet(nn.Module):
         self.model2 = nn.Sequential(
             nn.Linear(600, 256),
             # nn.Dropout(0.5),
-            nn.BatchNorm1d(256, affine=True),
+            # nn.BatchNorm1d(256, affine=True),
             nn.LeakyReLU(0.2, inplace=True),
             nn.Linear(256, 200),
             # nn.BatchNorm1d(200),

### 2021-04-08 src/DTI_data_preparation.py
Update BioSnap data and results in paper

@@ -211,7 +211,7 @@ def get_DTIs(drug_list,
              mode=''):
     DTI_graph = get_human_DTI_graph(mode=mode)
 
-    y_data = np.zeros(len(drug_list)*len(protein_list))
+    y_data = np.zeros((len(drug_list), len(protein_list)))
 
     for i in range(len(protein_list)):
         protein = protein_list[i]
@@ -407,7 +407,7 @@ def get_yamanishi_data(original_drug_list, original_protein_list):
     return drug_list[drug_indices], protein_list[protein_indices], dti_matrix[drug_indices,:][:, protein_indices]
 
 def get_BioSnap_data(original_drug_list, original_protein_list):
-    path = '../data/BioSnap_data/'
+    path = 'data/BioSnap_data/'
 
     drug_list = []
     protein_list = []
@@ -420,6 +420,8 @@ def get_BioSnap_data(original_drug_list, original_protein_list):
             dti_pairs.append((drug, protein))
 
     drug_list = list(set(drug_list))
+    print(drug_list[:10])
+    raise Exception
     protein_list = list(set(protein_list))
 
     with open(file=path+'BioSnap_DB_drug_list', mode='w') as f:
@@ -455,8 +457,9 @@ def get_BioSnap_data(original_drug_list, original_protein_list):
     drug_list = list(map(lambda d: drug_mapping.get(d, None), drug_list))
     protein_list = list(map(lambda p: protein_mapping.get(p, None), protein_list))
 
-    drug_indices = list(set([drug_list.index(drug) for drug in drug_list if drug in original_drug_list]))
-    protein_indices = list(set([protein_list.index(protein) for protein in protein_list if protein in original_protein_list]))
+    drug_indices = list(set([i for i, drug in enumerate(drug_list) if drug in original_drug_list]))
+    protein_indices = list(set([i for i, protein in enumerate(protein_list) if protein in original_protein_list]))
+
 
     drug_list = np.array(drug_list)
     protein_list = np.array(protein_list)

### 2021-04-08 src/HPO_predictor.py

@@ -45,8 +45,8 @@ class HPODTIDataBuilder:
 
         if config.yamanishi_test or config.biosnap_test:
             self.protein_list = np.array(list(set(PPI_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
-            print("Loading Yamanishi data ...")
             if config.yamanishi_test:
+                print("Loading Yamanishi data ...")
                 self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
             elif config.biosnap_test:
                 self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_BioSnap_data(self.drug_list, self.protein_list)

### 2021-04-09 paper/main.tex
Update BioSnap data and results in paper
.

@@
 \subtitle{Subject Section}
-\newcommand{\name}{Voodoo}
+\newcommand{\name}{DTI-Voodoo}
 
@@
 
-As DTI prediction is dependent on both drugs and proteins, there are multiple ways of determining training, validation and testing sets of pairs to evaluate each model.
-For cross-validation, we can perform the split over DTI pairs, over drugs and over proteins, respectively, where only in the latter two schemes unseen entities are shown to the model in the validation and testing phase.
-Additionally, these schemes have diverging expressiveness with respect to DTI prediction, as different information is known and leaked across training and testing phase.
+As DTI prediction is dependent on both drugs and proteins, there are
+multiple ways of determining training, validation and testing sets of
+pairs to evaluate each model. For cross-validation, we can perform the
+split over DTI pairs, over drugs and over proteins, respectively,
+where only in the latter two schemes unseen entities are guaranteed to
+be shown to the model in the validation and testing
+phase. Additionally, these schemes may have different expressiveness
+and use different information in DTI prediction, as different
+information is known in the training and testing phase.
 
@@
 		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
-		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$& & & \\
+		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
 		\hline

### 2021-04-09 paper/main.tex
.

@@
 pairs to evaluate each model. For cross-validation, we can perform the
-split over DTI pairs, over drugs and over proteins, respectively,
-where only in the latter two schemes unseen entities are guaranteed to
+split over DTI pairs, over drugs and over proteins, respectively; when
+splitting over drugs or proteins, the entities (drugs or proteins) are
+separated and all their associations included in the split.
+Only when splitting by protein or drug are unseen entities guaranteed to
 be shown to the model in the validation and testing
-phase. Additionally, these schemes may have different expressiveness
-and use different information in DTI prediction, as different
-information is known in the training and testing phase.
+phase. The models resulting from the different splitting schemes may
+have different expressiveness and exploit different information in DTI
+prediction, as different information is known in the training and
+testing phase.
 

### 2021-04-09 paper/main.tex
.

@@
 the area under the receiver operating characteristic curve
-(AUROC) on training, validation and testing split. % We
-% calculated the true positive rate (TPR), false positive rate (FPR), and
-% precision score. We will further compute true positives (TP), false
-% positives (FP), false negatives (FN) and finally true negatives (TN).
+(AUROC) on training, validation and testing split.
+% We calculated the true positive rate (TPR), false positive rate
+% (FPR), and precision score. We will further compute true positives
+% (TP), false positives (FP), false negatives (FN) and finally true
+% negatives (TN).
 
@@
 
-\todo[inline]{For RH: check this again later:}
 We also calculate the $\textrm{MicroAUC}$ score. For given lists $D$
@@
 interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated
-as the average per entity (macro) \textit{AUROC} score. Specifically,
-for protein-centric score, this can be formalized as: given labels
-$l:D\times P \rightarrow \{0,1\}$ and predictions $y:D\times P
-\rightarrow [0,1]$, we define
+as the average per entity \textit{AUROC} score. For example, the
+protein-centric score can be formalized as: given labels
+$l:D\times P \rightarrow \{0,1\}$ and predictions
+$y:D\times P \rightarrow [0,1]$, we define
 \begin{equation*}
@@
 
-In some cases, the $MicroAUC'$ score may not be defined as in some
+In some cases, the $MicroAUC$ score may not be defined as in some
 datasets some proteins or drugs have no interactions, leading to an
@@
 subset:
-
 \begin{equation*}
@@
 	\begin{cases}
-		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in D}l(d_i,p)\neq0\\
+		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
+                  D}l(d_i,p)\neq 0\\
 		Accuracy(l,y)&otherwise
@@
 \end{equation*}
-
-Note that drugs and proteins can be interchanged in this formulation. We
-refer to the different measures as protein-centric microAUC
-($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$).
+Drugs and proteins can be interchanged in this formulation. We refer
+to the different measures as protein-centric microAUC ($MicroAUC_p$)
+and a drug-centric microAUC ($MicroAUC_d$).
 

### 2021-04-09 paper/main.tex
.

@@
 
+\begin{table*}
+
+	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+
+		\hline
+		&&\multicolumn{3}{c|}{Yamanishi results}&\multicolumn{3}{c|}{BioSnap results}\\
+		Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}&Original scheme&\multicolumn{2}{c|}{Protein split}\\
+
+		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$&Macro AUC&Macro AUC&Micro $AUC_p$\\
+		\hline
+		Naive predictor&DP pairs&$0.85$& --&--&$0.79$&-- &--\\
+		DTINet&DP pairs&$0.91$&$0.74$&$0.67$ & --&--&--\\
+		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ & --&--&--\\
+		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$& --&--&--\\
+		\hline
+		DeepDTI&Drugs& --&--&--&$0.88$&$0.76$&$0.70$\\
+		DeepDTA&DP pairs& --&--&--&$0.88$&$0.77$&$0.69$\\
+		DeepConv-DTI&DP pairs& --&--&--&$0.88$&$0.76$&$0.73$\\
+		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
+		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
+		\hline
+	\end{tabular}
+        \caption{\label{tab:comparison} Comparison of \name{} with
+          state of the art drug--target interaction prediction methods
+          (Yamanishi dataset) and drug--target affinity prediction
+          methods (BioSnap dataset); we evaluated the original and the
+          protein-based split in a cross-validation}.
+\end{table*}
+
+We evaluate our model's ability to identify drug--target interactions
+using different approaches and datasets. First, we perform a
+cross-validation over proteins and validate our results. A
+cross-validation over proteins aims to evaluate how the model performs
+when tasked to identify drugs that may target a ``novel'' protein,
+i.e., one not seen during training, or a protein for which a drug that
+targets it should be predicted.
+
 \begin{table*}[!tpb]
@@
     \hline
-	\end{tabular} \medskip
-
-	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-
-		\hline
-		&&\multicolumn{3}{c|}{(c) Yamanishi results}&\multicolumn{3}{c|}{(d) BioSnap results}\\
-		Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}&Original scheme&\multicolumn{2}{c|}{Protein split}\\
-
-		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$&Macro AUC&Macro AUC&Micro $AUC_p$\\
-		\hline
-		Naive predictor&DP pairs&$0.85$& --&--&$0.79$&-- &--\\
-		DTINet&DP pairs&$0.91$&$0.74$&$0.67$ & --&--&--\\
-		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ & --&--&--\\
-		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$& --&--&--\\
-		\hline
-		DeepDTI&Drugs& --&--&--&$0.88$&$0.76$&$0.70$\\
-		DeepDTA&DP pairs& --&--&--&$0.88$&$0.77$&$0.69$\\
-		DeepConv-DTI&DP pairs& --&--&--&$0.88$&$0.76$&$0.73$\\
-		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
-		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
-		\hline
-	\end{tabular}\medskip
-
-	\caption{(a) + (b) Results for \name{} on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation.
-We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.
-(c) + (d) Results for various state of the art (c) drug--target interaction prediction methods on Yamanishi dataset and (d) drug--target affinity prediction methods on BIOSNAP dataset, evaluated on their original and the protein cross-validation splitting scheme, approximately reproducing the results of MolTrans \textbf{CITATION}.}
-	\label{tab:Results}
+  \end{tabular}
+  \caption{\label{tab:Results}Results for \name{} on the STITCH
+    and Yamanishi datasets evaluated with 5-fold cross-validation. We 
+    call the model using only molecular features
+    \textit{MolPred} and the model using only ontology-based
+    features \textit{OntoPred}. \name{} combines both types of
+    features.}
 \end{table*}
-
-We evaluate our model's ability to identify drug--target interactions
-using different approaches and datasets. First, we perform a
-cross-validation over proteins and validate our results. A
-cross-validation over proteins aims to evaluate how the model performs
-when tasked to identify drugs that may target a ``novel'' protein,
-i.e., one not seen during training, or a protein for which a drug that
-targets it should be predicted.
 
@@
 STITCH dataset using a 5-fold cross-validation over a protein split;
-we then selected the best-performing models (with respect to $MicroAUC_p$,
-see Section \ref{sec:evaluation}) and evaluate them in a 5-fold
-protein-split cross-validation on the Yamanishi benchmark dataset to
-avoid validation overfitting and yield more realistic testing
-results. To evaluate the influence of the different features
+we then selected the best-performing models (with respect to
+$MicroAUC_p$, see Section \ref{sec:evaluation}) and evaluate them in a
+5-fold protein-split cross-validation on the Yamanishi benchmark
+dataset to avoid validation overfitting and yield more realistic
+testing results. To evaluate the influence of the different features
 separately, and to determine whether they ``localize'' on the PPI
@@
 (OntoPred) prediction model, and a combination of both where we
-concatenate both types of features.  Table
-\ref{tab:Results} shows the
+concatenate both types of features.  Table \ref{tab:Results} shows the
 results of these experiments.

### 2021-04-09 paper/main.tex
.

@@
 
-\begin{table*}
-
-	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
-
-		\hline
-		&&\multicolumn{3}{c|}{Yamanishi results}&\multicolumn{3}{c|}{BioSnap results}\\
-		Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}&Original scheme&\multicolumn{2}{c|}{Protein split}\\
-
-		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$&Macro AUC&Macro AUC&Micro $AUC_p$\\
-		\hline
-		Naive predictor&DP pairs&$0.85$& --&--&$0.79$&-- &--\\
-		DTINet&DP pairs&$0.91$&$0.74$&$0.67$ & --&--&--\\
-		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ & --&--&--\\
-		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$& --&--&--\\
-		\hline
-		DeepDTI&Drugs& --&--&--&$0.88$&$0.76$&$0.70$\\
-		DeepDTA&DP pairs& --&--&--&$0.88$&$0.77$&$0.69$\\
-		DeepConv-DTI&DP pairs& --&--&--&$0.88$&$0.76$&$0.73$\\
-		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
-		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
-		\hline
-	\end{tabular}
-        \caption{\label{tab:comparison} Comparison of \name{} with
-          state of the art drug--target interaction prediction methods
-          (Yamanishi dataset) and drug--target affinity prediction
-          methods (BioSnap dataset); we evaluated the original and the
-          protein-based split in a cross-validation}.
-\end{table*}
-
 We evaluate our model's ability to identify drug--target interactions
@@
 
-\begin{table*}[!tpb]
+\begin{table}[ht]
   \centering
-  \begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
+  \begin{tabular}{|p{1.1cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|}
     \hline
@@
     \hline
-    \name{} (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
+    \name{} & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
     \hline
@@
     features.}
-\end{table*}
+\end{table}
 
@@
 incorporating the PPI graph. This increase can be observed with
-different graph neural network architectures and configurations. While
-the {\em GCNConv} and {\em GENConv} architecture already shows some
-minor improvement, the use of \textit{ResGraphConv} results in larger
+different graph neural network architectures and configurations
+(Supplementary Table X\todo{Add to Supplement!}). While the {\em
+  GCNConv} and {\em GENConv} architecture already shows some minor
+improvement, the use of \textit{ResGraphConv} results in larger
 performance improvements.  {\em ResGraphConv} blocks add a large
@@
 reached very similar results to the original \textit{OntoPred} model
-and showed no improvement.
+and showed no improvement (Supplementary Table X\todo{Add!}).
 
@@
 convolutional methods may be the result of numerical instability and
-fading signals \cite{}.
+fading signals \cite{}\todo{Add citation to original paper}.
 
@@
 % \subsection{Baseline model}
-Furthermore, we hypothesize that it may be possible to exploit biases
-in drug--target interactions to achieve relatively high prediction
-performance without obtaining a biologically meaningful signal.
-For example, hub proteins may have a large number of interactions,
-or certain drugs interact with many proteins, and preferentially
-predicting these interactions may increase predictive performance even
-in the absence of any biological features.
-To test this hypothesis, we design a ``na\"ive'' baseline model that
-predicts the same list of proteins for each drug based only on the
-number of known drug--target interactions for a protein. Formally,
-given lists $D$ and $P$ of drugs and proteins and a set of known
-interactions $\mathcal{I} := \{(d_i, p_i) \}$, we construct an
-interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with
+Furthermore, we hypothesize that it may be possible for a machine
+learning model to exploit biases in drug--target interaction data to
+achieve relatively high prediction performance without obtaining a
+biologically meaningful signal.  For example, hub proteins may have a
+large number of interactions, or certain drugs interact with many
+proteins, and preferentially predicting these interactions may
+increase predictive performance even in the absence of any biological
+features.  To test this hypothesis, we design a ``na\"ive'' baseline
+model that predicts the same list of proteins for each drug based only
+on the number of known drug--target interactions for a
+protein. Specifically, given lists $D$ and $P$ of drugs and proteins
+and a set of known interactions $\mathcal{I} := \{(d_i, p_i) \}$, we
+construct an interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$
+with
 \begin{equation*}
@@
 drugs and proteins; the na\"ive predictor only exploits imbalances in
-the evaluation set to make predictions that may perform
-well. % but allows us to estimate the effect of potential
-% biases in our evaluation dataset.
-% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
-
-The way in which we formulated the na\"ive predictor, it is not
-applicable for a protein split cross-validation as the number of
-interactions for each protein in the validation set is unknown.  We
-apply this naive predictor on both the STITCH and Yamanishi datasets,
-using the full datasets as well as a 5-fold cross-validation over
-drugs and over drug--protein pairs to compare the prediction results
-directly to \name{}.  For each fold, we gradually increase $k$ to
-determine the best performance for each fold. Using the full dataset,
-drug--target split, and drug split, we obtain the following MacroAUC
-results\todo[inline]{Put in Results table, then remove from here}: for
-the STITCH database, we obtain a performance of $0.76$ on the whole
-dataset, $0.70$ for the drug--target pairs and $0.73$ in case of the
-drug splitting scheme; on the Yamanishi dataset, we obtain MacroAUC
-scores of $0.88$, $0.84$ and $0.85$, for the total dataset,
-drug--target pair and drug split, respectively.  The na\"ive predictor
-shows higher performance on the Yamanishi dataset than on STITCH, and
-a substantial gain in comparison to an expected random predictor on
-both datasets. In the following, we utilize this nai\"ve predictor as
-baseline to compare its performance to state of the art models and
-\name{}.
+the evaluation set to make predictions that may perform well.  The way
+in which we formulated the na\"ive predictor, it is not applicable for
+a protein split cross-validation as the number of interactions for
+each protein in the validation set is unknown.
+
+We apply this naive predictor on both the STITCH and Yamanishi
+datasets, using the full datasets as well as a 5-fold cross-validation
+over drugs and over drug--protein pairs to compare the prediction
+results directly to \name{}.  For each fold, we gradually increase $k$
+to determine the best performance for each fold. Using the full
+dataset, drug--target split, and drug split, we obtain the following
+MacroAUC results: for the STITCH database, we obtain a performance of
+$0.76$ on the whole dataset, $0.70$ for the drug--target pairs and
+$0.73$ in case of the drug splitting scheme; on the Yamanishi dataset,
+we obtain MacroAUC scores of $0.88$, $0.84$ and $0.85$, for the total
+dataset, drug--target pair and drug split, respectively\todo{In the
+  Table is says 0.85 on Yaminishi for DP pair split, here it says 0.84
+  -- which is correct? Need fix}.  The na\"ive
+predictor shows higher performance on the Yamanishi dataset than on
+STITCH, and a substantial gain in comparison to an expected random
+predictor on both datasets. In the following, we utilize this nai\"ve
+predictor as baseline to compare its performance to state of the art
+models and \name{}.
 
@@
 
-The results of our experiments are summarized in part (c) of Table
-\ref{tab:Results}\todo{Check after restructure}; we calculated the
-performance of all compared methods over their original splitting
-scheme and over a protein split. We find that there is a large
-difference in performance when evaluating over a drug--target pair
-split compared to a protein split, with generally higher performance
-achieved when using the drug--target pair split.  Second, when
-evaluating the same methods over a protein split, we find a
-substantial performance difference in comparison to the splitting
-scheme used in the original evaluation of each method. DTI-CDF was
-originally evaluated on all three splitting schemes underlining this
-point \cite{}. While \name{} provides comparable performance to the
-na\"ive predictor and DTI-CDF in terms of MacroAUC, it yields
-considerably better results with respect to $\textrm{MicroAUC}_p$. We
-also find that methods that are trained using a protein split
-generally result in higher $\textrm{MicroAUC}_p$ than methods trained
-using a drug--target pair split.
+\begin{table}[ht]
+  \centering
+  \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+    \hline
+    Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}\\
+
+    &Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
+    \hline
+    Naive predictor&DP pairs&$0.85$& --&--\\
+    DTINet&DP pairs&$0.91$&$0.74$&$0.67$ \\
+    DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
+    DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
+    \name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
+    \hline
+  \end{tabular}
+  \caption{\label{tab:comparison1} Comparison of \name{} with
+    state of the art drug--target interaction prediction methods on
+    the Yamanishi dataset; we evaluated the original and the
+    protein-based split in a cross-validation}.
+\end{table}
+
+The results of our experiments are summarized in Table
+\ref{tab:comparison1}; we calculated the performance of all compared
+methods over their original splitting scheme and over a protein
+split. We find that there is a large difference in performance when
+evaluating over a drug--target pair split compared to a protein split,
+with generally higher performance achieved when using the drug--target
+pair split.  Second, when evaluating the same methods over a protein
+split, we find a substantial performance difference in comparison to
+the splitting scheme used in the original evaluation of each
+method. DTI-CDF was originally evaluated on all three splitting
+schemes underlining this point \cite{}. While \name{} provides
+comparable performance to the na\"ive predictor and DTI-CDF in terms
+of MacroAUC, it yields considerably better results with respect to
+$\textrm{MicroAUC}_p$. We also find that methods that are trained
+using a protein split generally result in higher $\textrm{MicroAUC}_p$
+than methods trained using a drug--target pair split, indicating that
+they may generalize better to unseen protein targets whereas methods
+trained on a drug--target split potentially exploit hidden biases and
+therefore generalize less well.
+
+\begin{table}[ht]
+  \centering
+  \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+    \hline
+    Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}\\
+
+    &Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
+    \hline
+    Naive predictor&DP pairs&$0.79$&-- &--\\
+    DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
+    DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
+    DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
+    MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
+    \name&Proteins&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
+    \hline
+  \end{tabular}
+  \caption{\label{tab:comparison2} Comparison of \name{} with
+    state of the art drug--target interaction prediction methods
+    on the BioSnap dataset; we evaluated the original and the
+    protein-based split in a cross-validation}.
+\end{table}
 
@@
 evaluation on other datasets.  Following the results of MolTrans
-\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017}, DeepDTA
-\cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019} and MolTrans itself on BioSnap\todo{What
-  is that? Is it described in methods?}  \cite{} over a protein split,
-as shown in part (d) of Table \ref{tab:Results}\todo{check table}. The
-authors of MolTrans evaluated all considered methods over the
-drug--target pair and the protein split; we were able to reproduce the
-results (Table \ref{tab:Results} (d))\todo{check table}, showing a
-substantial difference based on the splitting scheme. We additionally
-computed the $MicroAUC_p$ score for all considered methods, leading to
-evaluation results that are comparable\todo{are the datasets
-  comparable?} to the ones obtained on the Yaminishi dataset.
-
-\todo[inline]{ add \name{} model to table to allow
-  DIRECT comparison:}
-The results of the analysis are summarized in the upper part of Figure
-\ref{tab:Results}.
-
-We find that...\todo[inline]{summarize key findings from Table.}
-
-\todo[inline]{Replace/merge table:}
-
-%As the baseline predictor is based on the second dataset bias i.e.
-the existence of hub proteins with significantly more interactions within the dataset, this may hint on a more severe skew within the Yamanishi benchmark.
+\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
+DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
+MolTrans itself on the BioSnap dataset \cite{} and compared it to our
+``na\"ive'' predictor as well as \name{} (see Table
+\ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
+pair and the protein split; we were able to reproduce the MolTrans
+results (Table \ref{tab:comparison2}), showing a substantial
+difference based on the splitting scheme. We additionally computed the
+$MicroAUC_p$ score for all considered methods, leading to similar 
+results as observed on the Yaminishi dataset.
+
+In all our experiments, \name{} has the highest Micro $AUC_p$,
+demonstrating that \name{} can improve over other methods in the task
+of identifying drugs that target a specific protein. \name{}
+
+(i.e., higher MacroAUC
+when evaluated on a drug--target split compared to a protein split;
+large differences in macro- and micro measures)
 

### 2021-04-09 paper/main.tex
.

@@
 demonstrating that \name{} can improve over other methods in the task
-of identifying drugs that target a specific protein. \name{}
-
-(i.e., higher MacroAUC
-when evaluated on a drug--target split compared to a protein split;
-large differences in macro- and micro measures)
+of identifying drugs that target a specific protein. Several methods
+achieve a higher Macro AUC than \name{}, in particular when evaluated
+using a drug--target pair splitting scheme; our results with the
+``na\"ive'' prediction method show that it may be possible that models
+trained on a drug--target split utilize certain biases in the dataset
+without necessarily producing novel biological insights.
 

### 2021-04-09 paper/main.tex
.

@@
 organism and infer drug--target interactions as the putative molecular
-mechanisms that explain these observations; many top-down approaches
-follow methods similar to those used in reverse genetics \cite{}.
+mechanisms that explain these observations.
 
@@
 consequently, we expect these features to benefit more from including
-information about protein--protein interactions. Our results confirm
-this hypothesis and demonstrate that molecular features do not benefit
-from including the interaction network whereas the indirect, top-down
-features benefit from the propagation over the interaction network.
+information about protein--protein interactions. Our results (Table
+\ref{tab:Results}) confirm this hypothesis and demonstrate that
+molecular features do not benefit from including the interaction
+network whereas the indirect, top-down features benefit from the
+propagation over the interaction network.
 
@@
 There are other types of indirect features that could be added to our
-model. A common feature that is added are drug indications which are
-highly predictive of drug--target interactions \cite{}; however, we do
+model. A common feature that may be added are drug indications which are
+predictive of drug--target interactions \cite{}; however, we do
 not include them in our model as including drug indications would
@@
 on drug indications because different drugs may target the same
-pathway through different mechanisms \cite{}.
+pathway through different mechanisms.
 

### 2021-04-09 paper/main.tex
.

@@
 
+Combining bottom-up and top-down approaches in a single model can
+follow different strategies. Graph neural networks are a natural way
+to represent interactions between proteins, and interaction networks
+are used widely to determine indirect effects of molecular changes
+\cite{}.
+
+Graph neural networks \cite{} have the ability to
+represent interaction networks
+
 we built protein function and ontology based features based on DL2vec

### 2021-04-09 paper/main.tex
.

--- 2021-04-09 paper/main.tex
@@
 
-we built protein function and ontology based features based on DL2vec
-
-Ontology derived protein function focused features are highly predictive for dtis
-
-We built a versatile template for various features to test localization on the PPI graph
-
-normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
-
-molecular features work for all (most?) drugs, top-down features only
-for those that have a side-effect profile
-
-\begin{itemize}
-	\item \name{} comparison to other approaches on BioSnap prone to bias, as $ BioSnap: drugs 
-	\times prots = 5500 \times 3000 $, while only 
-	$\approx$ 1000 drugs have Sider Annotation, and 
-\end{itemize}
-
-\subsection{Evaluating drug--target interaction predictions}
+Combining bottom-up and top-down approaches in a single model can
+follow different strategies. Interaction networks are used widely to
+determine indirect effects of molecular changes \cite{} and predict
+drug--target interactions \cite{}. Our work relies on graph neural
+networks as a way to combine qualitative information about
+interactions with additional features (molecular interaction,
+phenotypic and functional features); even if only some of these
+features benefit from the information the graph provides, graph neural
+networks will allow further extension of our model with additional
+features in the future.
+
+% Graph neural networks \cite{} have the ability to
+% represent interaction networks
+
+% we built protein function and ontology based features based on DL2vec
+
+% Ontology derived protein function focused features are highly predictive for dtis
+
+% We built a versatile template for various features to test localization on the PPI graph
+
+% normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
+
+% molecular features work for all (most?) drugs, top-down features only
+% for those that have a side-effect profile
+
+% \begin{itemize}
+% 	\item \name{} comparison to other approaches on BioSnap prone to bias, as $ BioSnap: drugs 
+% 	\times prots = 5500 \times 3000 $, while only 
+% 	$\approx$ 1000 drugs have Sider Annotation, and 
+% \end{itemize}
+
+\subsection{The challenge of evaluating drug--target interaction
+  predictions}
 

### 2021-04-10 paper/main.tex
.

@@
 
-We spent a significant amount of time testing different evaluation
-schemes for predicting drug--target interactions. We designed a
+One major component of our experiments was to determine how the
+information that is available to a machine learning model during
+training affects the performance of the model. Similarly to previous
+work, we find significant differences in predictive performance across
+different splitting schemes.
+
+The most common scheme for drug--target interaction prediction is the
+split over drug-target pairs \cite{Survey2018} where it may happen
+that most drugs and targets that are including in the model's
+validation and testing phase have also been included in the training
+phase (as part of other drug--target pairs).
+
+The second most prevalent arrangement is the split
+over drugs, while only close to none is aiming on a protein split.
+However, the first and second splitting scheme are exposed to the
+first dataset bias and are hence more likely vulnerable to
+transductive inference by just predicting recently seen structures,
+rather than implementing inductive inference and generalizing over the
+drug representations. Second, these two strategies are more
+susceptible to the second bias, as only in these cases the model may
+overfit on the number of existing interactions for a single protein,
+while in the third scheme the number of interactions of the test
+proteins is entirely unknown during training process.
+
+We designed a
 na\"ive predictor and find that it can achieve a performance only
@@
 divergent expressiveness of the resulting model.
-
-\todo[inline]{Likely discussion:} The most common scheme for DTI
-prediction is the split over drug-target pairs \citep{Survey2018},
-where likely all drugs and targets of the validation and testing phase
-have already occurred in the training phase, as part of other
-drug-target pairs. The second most prevalent arrangement is the split
-over drugs, while only close to none is aiming on a protein split.
-However, the first and second splitting scheme are exposed to the
-first dataset bias and are hence more likely vulnerable to
-transductive inference by just predicting recently seen structures,
-rather than implementing inductive inference and generalizing over the
-drug representations. Second, these two strategies are more
-susceptible to the second bias, as only in these cases the model may
-overfit on the number of existing interactions for a single protein,
-while in the third scheme the number of interactions of the test
-proteins is entirely unknown during training process.
 

### 2021-04-10 paper/main.tex
.

@@
 training affects the performance of the model. Similarly to previous
-work, we find significant differences in predictive performance across
-different splitting schemes.
+work \cite{}, we find significant differences in predictive
+performance across different splitting schemes.
 
@@
 validation and testing phase have also been included in the training
-phase (as part of other drug--target pairs).
+phase (as part of other drug--target pairs). This scheme is prone to a
+number of biases. If the number of interactions for a drug or protein
+are imbalanced, i.e., some drugs or proteins have many more
+interactions than others, these will be seen more often during
+training and they will likely also have more interactions in the
+testing and validation sets. Preferentially predicting
 
@@
 microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
+
+Other biases: evaluate on balanced data, not all
 
@@
 
-\todo[inline]{Likely discussion:}
-Assuming a hypothetical, perfectly generalizing model built upon an
-unbiased dataset, this very model would yield similar performances for
-all three , not overfitting on the known
-structures. On the other hand, for a hypothetical entirely overfitting
-model, trained on a highly biased dataset, this model would show
-substantial deviations from the original performance over another
-split.
-
-\todo[inline]{Likely discussion:}
-We emphasize, that all real-world models are prone to some sort of
-overfitting, and unknown, deviant entities in both validation and
-testing set will likely lead to some sort of performance gap for
-relevant metrics. However, a large disparity may hint the biases
-stated above.
-
 A naive predictor (ranking proteins) and predict each drug similarly
@@
 %\subsection{\name identifies drugs that target a protein}
+
+% \todo[inline]{Likely discussion:}
+% Assuming a hypothetical, perfectly generalizing model built upon an
+% unbiased dataset, this very model would yield similar performances for
+% all three , not overfitting on the known
+% structures. On the other hand, for a hypothetical entirely overfitting
+% model, trained on a highly biased dataset, this model would show
+% substantial deviations from the original performance over another
+% split.
+
+% \todo[inline]{Likely discussion:}
+% We emphasize, that all real-world models are prone to some sort of
+% overfitting, and unknown, deviant entities in both validation and
+% testing set will likely lead to some sort of performance gap for
+% relevant metrics. However, a large disparity may hint the biases
+% stated above.
 

### 2021-04-10 paper/main.tex
.

@@
 training and they will likely also have more interactions in the
-testing and validation sets. Preferentially predicting
-
-The second most prevalent arrangement is the split
-over drugs, while only close to none is aiming on a protein split.
-However, the first and second splitting scheme are exposed to the
-first dataset bias and are hence more likely vulnerable to
-transductive inference by just predicting recently seen structures,
-rather than implementing inductive inference and generalizing over the
-drug representations. Second, these two strategies are more
-susceptible to the second bias, as only in these cases the model may
-overfit on the number of existing interactions for a single protein,
-while in the third scheme the number of interactions of the test
-proteins is entirely unknown during training process.
+testing and validation sets; because some entities have more
+interactions, i.e., they are more likely to interact, any model that
+preferentially predicts these as interaction partners will improve its
+predictive performance. While this accurately captures the
+distribution, predicting based on biases in the number of interaction
+partners is not desirable when applying the model to novel
+entities. We have demonstrated that even a ``na\"ive'' classifier that
+makes predictions only by exploiting the imbalanced number of
+interaction partners can achieve performance close to state of the art
+methods (when measuring Macro AUC). When training a machine learning
+model on such imbalanced data, it will eventually overfit to this
+imbalance.  Splitting by entity (protein or drug) can reduce the
+impact of this spurious correlation but not reduce it entirely,
+because similar entities will still exhibit similar interaction
+patterns. In our experiments, we observed the impact of splitting
+training and evaluation sets by protein as a decrease in overall
+performance (Macro AUC), providing some evidence that models trained
+using this splitting scheme are less sensitive to overfitting to this
+type of bias.
+
+The way in which training and evaluation data is generated is related
+to how the model is evaluated. An evaluation based on Macro AUC
+evaluated the application scenario where a set of drugs and proteins
+are given, and out of all possible pairs, the more likely interactions
+need to be identified. However, this does not correspond to most
+scenarios for drug repurposing where a drug that targets a {\em
+  specific} protein (e.g., a protein involved in a disease) needs to
+be identified. We use an evaluation measure based on micro-averaged
+per protein (Micro AUC$_p$) to evaluate this application.
+
+Finally, a potential source of differences in evaluation is how
+negatives are treated during evaluation (and training).
+Random sampling of unknown...
+
+% The second most prevalent arrangement is the split
+% over drugs, while only close to none is aiming on a protein split.
+% However, the first and second splitting scheme are exposed to the
+% first dataset bias and are hence more likely vulnerable to
+% transductive inference by just predicting recently seen structures,
+% rather than implementing inductive inference and generalizing over the
+% drug representations. Second, these two strategies are more
+% susceptible to the second bias, as only in these cases the model may
+% overfit on the number of existing interactions for a single protein,
+% while in the third scheme the number of interactions of the test
+% proteins is entirely unknown during training process.
 

### 2021-04-10 paper/main.tex
.

@@
   specific} protein (e.g., a protein involved in a disease) needs to
-be identified. We use an evaluation measure based on micro-averaged
-per protein (Micro AUC$_p$) to evaluate this application.
-
-Finally, a potential source of differences in evaluation is how
-negatives are treated during evaluation (and training).
-Random sampling of unknown...
+be identified. We use an evaluation measure based on micro-averages
+per protein (Micro AUC$_p$) to evaluate this scenario, and we often
+find substantial differences in predictive performance when evaluating
+with Macro AUC and Micro AUC$_p$; generally, models that are trained
+using a split over drug--target pairs perform worse in Micro AUC$_p$
+than models that use a protein-based split, further providing evidence
+that a drug--target split results in overfitting to dataset biases.
+
+Finally, a potential source of differences in model performance is how
+negatives are identified and treated during evaluation (and
+training). There are few large sets of validated negative drug--target
+interactions; consequently, many models (including \name{}) use all
+unknown interactions as negatives. As there are many more negative
+than positive interactions, negatives are then sub-sampled during
+training resulting in a training set that is balanced between
+positives and negatives (or a certain ratio is preserved). While this
+is a reasonable strategy to deal with imbalanced data, it may be
+problematic when the same sub-sampling is applied on the model's
+evaluation set because it over-simplifies the evaluation process. The
+performance differences is not usually visible when using ROC curves
+but results in unrealistically high precision and therefore high area
+under a precision-recall curve.
+
+In summary, drug--target interaction prediction is not a single
+computational problem in bioinformatics but a set of related
+problems. Let $P$ be a set of proteins $P$ and $D$ a set of drugs; one
+task can be to identify arbitrary pairs $(p,d)$ with $p \in P$ and
+$d \in D$ that interact, another to identify a set of interacting
+drugs for each $p \in P$, and yet another to identify a set of
+proteins for each drug $d \in D$. The first task may be useful when no
+particular drug or protein is considered; the second task when
+searching for a drug that targets a specific (disease-associated)
+protein; and the third when aiming to find new applications for a
+given drug. The first task would best be evaluated using a Macro AUC,
+the second and third using a Micro AUC$_p$ and Micro AUC$_d$.
 
@@
 
-We designed a
-na\"ive predictor and find that it can achieve a performance only
-slightly below state of the art drug--target prediction methods. We
-investigated this finding further and
-
-In the presence of imbalanced
-
-Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
-
-Running split over proteins is harder than, drug and drug protein pair split (see below table)
-
-this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
-
-Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
-
-microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
+% We designed a
+% na\"ive predictor and find that it can achieve a performance only
+% slightly below state of the art drug--target prediction methods. We
+% investigated this finding further and
+
+% In the presence of imbalanced
+
+% Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
+
+% Running split over proteins is harder than, drug and drug protein pair split (see below table)
+
+% this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
+
+% Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
+
+% microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
 

### 2021-04-10 paper/main.tex
.

@@
 
-Other biases: evaluate on balanced data, not all
-
-\todo[inline]{Likely discussion:}
-The aim of \name is to predict candidate drugs that target a given
-protein; the challenge is to develop a training and evaluation scheme
-that does not simply overfit to the inherent biases in training and
-testing data.
-In general, when performing cross-validation for DTI prediction, the
-options are to split over 
-\begin{enumerate}
-	\item split over drugs,
-	\item split over drug--target pairs, or
-	\item split over proteins
-\end{enumerate}
-where the first and third option concern splitting drugs and proteins,
-respectively, into train, validation and test sets, and arranging the
-corresponding drug-target interactions. They ensure that at least
-parts of the interactions are not seen during training and evaluate
-either how well targets are predicted for unseen drugs or unseen
-proteins. Hereby, different training and prediction schemes lead to
-divergent expressiveness of the resulting model.
-
-A naive predictor (ranking proteins) and predict each drug similarly
-achieves cutting edge performance ($87.5$ AUROC for whole dataset,
-$85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
-prot focused microAUC possible. $\rightarrow$ hub proteins
-
-Yamanishi Dataset is only partially suitable for comparing results, if
-everybody just derives a suitable subset (DTIGEMS)
-
-This also applies to drug target affinity prediction. We were hereby
-able to roughly reproduce the results from MolTrans (Bioinformatics)
-on BioSnap
-
-%\subsection{Features to use in predicting drug targets}
-
-\todo[inline]{Discussion:}
-The lack of stratification, only impacts the not
-considered area under precision recall curve (AUPRC) and not the Macro
-AUROC score, while also supporting the expressiveness of $MicroAUC_p$
-with more data points.
+% \todo[inline]{Likely discussion:}
+% The aim of \name is to predict candidate drugs that target a given
+% protein; the challenge is to develop a training and evaluation scheme
+% that does not simply overfit to the inherent biases in training and
+% testing data.
+% In general, when performing cross-validation for DTI prediction, the
+% options are to split over 
+% \begin{enumerate}
+% 	\item split over drugs,
+% 	\item split over drug--target pairs, or
+% 	\item split over proteins
+% \end{enumerate}
+% where the first and third option concern splitting drugs and proteins,
+% respectively, into train, validation and test sets, and arranging the
+% corresponding drug-target interactions. They ensure that at least
+% parts of the interactions are not seen during training and evaluate
+% either how well targets are predicted for unseen drugs or unseen
+% proteins. Hereby, different training and prediction schemes lead to
+% divergent expressiveness of the resulting model.
+
+% A naive predictor (ranking proteins) and predict each drug similarly
+% achieves cutting edge performance ($87.5$ AUROC for whole dataset,
+% $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
+% prot focused microAUC possible. $\rightarrow$ hub proteins
+
+% Yamanishi Dataset is only partially suitable for comparing results, if
+% everybody just derives a suitable subset (DTIGEMS)
+
+% This also applies to drug target affinity prediction. We were hereby
+% able to roughly reproduce the results from MolTrans (Bioinformatics)
+% on BioSnap
+
+% %\subsection{Features to use in predicting drug targets}
+
+%\todo[inline]{Discussion:}
+% The lack of stratification, only impacts the not
+% considered area under precision recall curve (AUPRC) and not the Macro
+% AUROC score, while also supporting the expressiveness of $MicroAUC_p$
+% with more data points.
 
@@
 
-For the bottom-up approach we build a model that only relies on
-molecular features, which we will discuss in more detail in the
-following methods chapter. For the combination of both approaches we
-now attach the predictions to the protein-protein interaction graph as
-node features for future graph learning steps. In this graph we tried
-to find both patterns and regions for each drug that could be of
-interest through application of different graph convolutional layers,
-which in return represent the feature for each protein. Representing
-the drug we take the drug-drug interaction graph and the semantic
-similarity over side effects which we will explain in the following
-paragraphs.
-
-Combining bottom-up and top-down features improves state of the art.
-
-Evaluating DTI predictions is non-trivial due to numerous biases in
-the training and evaluation datasets.
+We developed \name{} as a machine learning model that combines
+molecular features and functional information with an interaction
+network using graph neural networks to predict drugs that may target
+specific proteins. In this task, \name{} improves over several state
+of the art methods. We demonstrated that functional and phenotypic
+information localizes on the interaction network whereas molecular
+information does not. Moreover, we showed that drug--target
+interaction prediction datasets have some inherent biases that affect
+the performance of models. This led us to conclude that drug--target
+interaction prediction is not a single computational problem but a set
+of multiple problems. Experimental evaluation of drug--target
+interaction prediction methods must be carefully designed to reflect
+the problem the model aims to solve, and the interpretation of
+performance results should be aligned with the specific problem.
 

### 2021-04-10 paper/main.tex
.

@@
 
-In history, traditional remedies, that were known for their medicinal
-properties lead to drugs by extraction of the functional
-ingredients. Alternatively, characteristics and features of potential
-drugs were detected by accident like in the case of penicillin. More
-recently, biological drug targets can be found \textit{in silico}
-through discovery of suitable computational predictors.
-
-The challenge of accurately predicting drug-target-interactions (DTI)
-has shown its importance in the fields of drug repurposing and
-repositioning, and in the exploration of novel drugs and their
-interaction partners. Knowledge about those links between compounds
-and their target proteins help in an array of medical and
-pharmaceutical studies. Additionally, those associations can be
-utilized to identify disease specific targets, leading to desirable
-therapeutic effects.
-
-With the rapidly growing field of machine learning approaches and
-their application to bioscientifical problems in the realm of
-bioinformatics, different kinds of data, such as long DNA sequences
-could be utilized for feature generation, while rapid advances were
-made.
+% In history, traditional remedies, that were known for their medicinal
+% properties lead to drugs by extraction of the functional
+% ingredients. Alternatively, characteristics and features of potential
+% drugs were detected by accident like in the case of penicillin. More
+% recently, biological drug targets can be found \textit{in silico}
+% through discovery of suitable computational predictors.
+
+% Knowledge about those links between compounds and their
+% target proteins help in an array of medical and pharmaceutical
+% studies. Additionally, those associations can be utilized to identify
+% disease specific targets, leading to desirable therapeutic effects.
+
+% With the rapidly growing field of machine learning approaches and
+% their application to bioscientifical problems in the realm of
+% bioinformatics, different kinds of data, such as long DNA sequences
+% could be utilized for feature generation, while rapid advances were
+% made. 
+Identifying drug--target interactions (DTIs) is important in drug
+repurposing and repositioning, and in the exploration of novel drug
+candidates and their interaction partners. Computational methods are
+widely applied to predict DTIs and many computational methods have
+been developed.  These methods can be broadly classified into
+``top-down'' and ``bottom-up'' approaches. Top-down approaches start
+form observable characteristics resulting from a drug--target
+interaction, such as side-effects or the diseases treated by a drug,
+and infer likely molecular mechanisms (i.e., the interaction) using
+these observations.
+% and indications represented by knowledge graphs or ontologies, induced by a drug and infer targets based on the likely molecular mechanisms that result in these phenotypes
+Bottom-up approaches start from molecular features such as molecular
+structure or fingerprints associated with drug and protein, and
+predict interactions from this information.
+
+Both bottom-up and top-down approaches to drug--target interaction
+prediction have some advantages and limitations. Generally, bottom-up
+methods face the challenge to predict whether a structure binds to a
+protein given their molecular properties; whether two entities
+interact depends not only on the molecular structure of the entities
+(where binding sites and molecular forces need to be determined for
+accurate prediction) but also properties such as where a protein is
+expressed or in which celltypes and anatomical structures. Top-down
+methods use information for DTI prediction such as side-effect
+similarity \cite{Campillos2008} that is largely complementary to
+knowledge gained from molecular properties. While methods that rely on
+molecular information are directly predicting whether two molecular
+could interact, top-down methods rely on more indirect means and 
+infer a DTI from observable effects resulting from the interaction.
+
+Both approaches may be used in conjunction with network inference
+\cite{}.
+
+In order to design representations concerning both top-down and bottom-up features for proteins and drugs,respectively, we utilize protein--protein interaction (PPI) networks which have shown great results in protein function prediction \cite{Vazquez2003} and antiviral drug--target discovery \cite{Ackerman2019}.
+However, only few applied this context to the task of DTI prediction in humans.
+The method of graph convolutional neural networks has been applied successfully to PPI networks \cite{Zitnik2017}, granting valuable information for medical diagnostics, closely related to phenotypic information.
 
@@
 for DTI prediction.
-
-These approaches can be classified into top-down and bottom-up methods.
-Top-down or network approaches hereby start form the observable, phenotypical characteristics, such as side-effects, associated diseases and indications represented by knowledge graphs or ontologies, induced by a drug and infer targets based on the likely molecular mechanisms that result in these phenotypes.
-On the other hand, bottom-up approaches start from molecular features, such as molecular structure, molecular fingerprints, secondary structure and contact predictions and others, deriving predictions from suitable correlations between these features.
-
-However, both bottom--up and top--down approaches to drug--target interaction prediction contain some limitations that are not solvable within themselves.
-Generally, it is very challenging for bottom-up methods to accurately predict, whether there a structure is binding to a protein, as not only molecular structure, but also binding sites and molecular forces are crucial for accurate prediction.
-Thus, finding similarities for dissimilar drugs affecting the same targets is problematic but crucial.
-Furthermore, phenotypic features contain relevant information for DTI prediction \cite{Campillos2008} that are complementary to knowledge gained from molecular properties.
-While bottom-up representations possibly contain all information needed for accurate DTI prediction, we rely on indirect information for better generalization.
-Contrary, top-down approaches lack the ability to spot and cope with small differences within the molecular drug structure, thus both incapable to predict on novel entities and derive similarity from structural correlation.
-
-In order to design representations concerning both top-down and bottom-up features for proteins and drugs,respectively, we utilize protein--protein interaction (PPI) networks which have shown great results in protein function prediction \cite{Vazquez2003} and antiviral drug--target discovery \cite{Ackerman2019}.
-However, only few applied this context to the task of DTI prediction in humans.
-The method of graph convolutional neural networks has been applied successfully to PPI networks \cite{Zitnik2017}, granting valuable information for medical diagnostics, closely related to phenotypic information.
 

### 2021-04-10 paper/main.tex
.

@@
 Both approaches may be used in conjunction with network inference
-\cite{}.
+\cite{}. Biological networks used for DTI prediction include
+protein--protein interaction networks \cite{} and networks including
+several other types of biological relations, including similarity
+between represented entities \cite{PREDICT}. Network-based DTI
+prediction methods use the guilt-by-association principle \cite{} and
+assume that a protein is a likely target for a drug if many of the
+protein's neighbors in the interaction network are targets of the
+drug \cite{}.
 

### 2021-04-10 paper/main.tex
.

--- 2021-04-10 paper/main.tex
@@
 Both approaches may be used in conjunction with network inference
-\cite{}.
+\cite{}. Biological networks used for DTI prediction include
+protein--protein interaction networks \cite{} and networks including
+several other types of biological relations, including similarity
+between represented entities \cite{PREDICT}. Network-based DTI
+prediction methods use the guilt-by-association principle
+\cite{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002444}
+and assume that a protein is a likely target for a drug if many of the
+protein's neighbors in the interaction network are targets of the drug
+\cite{https://academic.oup.com/bioinformatics/article/32/7/1057/1743938,
+  others}. Network-based methods have been applied successfully
+to DTI prediction. However, if DTIs are taken as direct physical
+interactions between a drug and protein, it remains an unresolved
+question whether the network-based guilt-by-association hypothesis is
+true, or whether an interaction of a drug and protein dys-regulates
+several of the protein's interaction partners, and therefore resulting
+in downstream effects that are not direct interactions but only
+downstream consequences of an interaction.
 

### 2021-04-10 paper/main.tex
.

@@
 
-In order to design representations concerning both top-down and bottom-up features for proteins and drugs,respectively, we utilize protein--protein interaction (PPI) networks which have shown great results in protein function prediction \cite{Vazquez2003} and antiviral drug--target discovery \cite{Ackerman2019}.
-However, only few applied this context to the task of DTI prediction in humans.
-The method of graph convolutional neural networks has been applied successfully to PPI networks \cite{Zitnik2017}, granting valuable information for medical diagnostics, closely related to phenotypic information.
-
-Only recently, the technique of graph learning was introduced by
-\citet{GCNConv} through graph convolution algorithms, and improved and
-altered under usage of different kernels \cite{ChebConv, ARMAConv},
-attention mechanisms \cite{GATConv}, random walks \cite{APPNPConv},
-and mixtures of both \cite{SAGEConv}. While based on diverse systems,
-they can be relevant for testing distinct hypothesis for given
-graphs. While convolutional filters are suitable for finding patterns
-among the given graph, attention mechanisms are more relevant for
-discovery of important regions within. Lately, graph learning
-approaches found application for computing compound representations
-for DTI prediction.
-
-Eventually, we further address potential biases and skews within DTI prediction datasets. \todo[inline]{Introduction}
-Within DTI prediction, there are potential biases resulting from the
-underlying datasets \citep{Pahikkala2014}. First, novel drugs are
-often designed by altering non-functional components of a drug,
-leading to two and more very similar drugs designed to target the same
-proteins \cite{Overington2006}. This can result in a bias when it
-leads to hidden duplicates that can distribute among the train/test
-split, resulting in a better (measured) predictive performance than
-would be expected when the model is applied to identify drugs that
-target a protein for which no drugs yet exist. Second, some proteins
-(which we call \textit{hub proteins}) have significantly more known
-interactions with drugs than others. In the STITCH database, $5\%$ of
-the proteins have $40\%$ of the interactions, and similar
-distributions are present in the Yamanishi and Drugbank
-\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
+Progress in machine learning using graph neural networks can allow us
+to test this hypothesis and combine both bottom-up and top-down
+features with a network in a single machine learning model. 
+In particular, Graph Convolutional Networks \cite{GCNConv} and
+their variants using different types of kernels \cite{ChebConv,
+  ARMAConv}, including attention mechanisms \cite{GATConv}, and
+different forms of exploring node neighborhoods \cite{APPNPConv,
+  SAGEConv} can combine different types of features and graph-based
+information. % While based on diverse systems,
+% they can be relevant for testing distinct hypothesis for given
+% graphs. While convolutional filters are suitable for finding patterns
+% among the given graph, attention mechanisms are more relevant for
+% discovery of important regions within. Lately, graph learning
+% approaches found application for computing compound representations
+% for DTI prediction.
+They have previously been applied for a number of tasks, including
+prediction of protein functions \cite{Zitnik2017}, \todo{add 2 more,
+  and then one on DTI prediction}.
+
+One challenge in DTI prediction are potential biases resulting from
+the underlying datasets \citep{Pahikkala2014} which may affect model
+evaluation and comparison. First, novel drugs are often designed by
+altering non-functional components of a drug, leading to two and more
+very similar drugs designed to target the same proteins
+\cite{Overington2006}. This can result in a bias when it leads to
+hidden duplicates or highly similar compounds that are distributed
+among training and evaluation dataset, resulting in a better
+(measured) predictive performance than would be expected when the
+model is applied to identify drugs that target a protein for which no
+drugs yet exist. Second, some proteins (which we call \textit{hub
+  proteins}) have significantly more known interactions with drugs
+than others. In the STITCH database, $5\%$ of the proteins have $40\%$
+of the interactions, and similar distributions are present in other
+datasets \cite{Drugbank2007, Drugbank2017}; preferentially predicting
 these proteins may increase predictive performance while not
-reflecting the actual performance when applied to a new protein (e.g.,
+reflecting the actual performance when applied to a new protein (i.e.,
 a protein for which no interactions are known). These differences in
@@
 to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed). This may affect common \cite{Survey2018} cross-validation splitting schemes, that allow exploitation of these biases within DTI prediction.
+diseases for which drugs can be more profitably marketed). This may
+affect common evaluation schemes where it is possible
+to exploit these biases within DTI prediction \cite{Survey2018}.
+
+% In order to design representations concerning both top-down and bottom-up features for proteins and drugs,respectively, we utilize protein--protein interaction (PPI) networks which have shown great results in protein function prediction \cite{Vazquez2003} and antiviral drug--target discovery \cite{Ackerman2019}.
+% However, only few applied this context to the task of DTI prediction in humans.
 

### 2021-04-10 paper/main.tex
.

@@
 to exploit these biases within DTI prediction \cite{Survey2018}.
+
+We developed \name{} as a method for predicting DTIs. We use an
+ontology-based machine learning method \cite{} to encode phenotypic
+consequences of DTIs and deep learning methods to encode molecular
+features. We combine both using a protein interaction network which we
+exploit using a graph neural network. We use this model to test
+whether molecular or phenotype features benefit from the network
+information and find that only phenotype features localize on the
+graph whereas molecular features do not. We further evaluate and
+compare \name{} against several DTI prediction methods and
+demonstrate a substantial improvement of \name{} over the state of the
+art in predicting drugs that target a protein. We also identify and
+characterize several biases in both training and evaluating DTI
+prediction methods, and make recommendations on how to avoid
+them. \name is available as Free Software at \url{}\todo{add}.
 

### 2021-04-11 paper/main.tex
.

@@
 and inhibition, and do not predict the strength of the interaction.
-If we additionally assume that our knowledge is complete and all
-drug--protein pairs without a known interaction do not interact, we
-can formulate the problem as a binary classification task.
+We treat all drug--protein pairs without a known interaction as
+negatives and therefore formulate the problem as a binary
+classification task.
 

### 2021-04-11 paper/main.tex
.

@@
 \subsection{Datasets}
-We obtain a dataset consisting of 12,884 human proteins with over
-340627 links from STRING \citep{STRINGv10}. For the drug-target
+We obtain a dataset consisting of 12,884 human proteins with 
+340,627 links from STRING \citep{STRINGv10}. For the drug-target
 interactions, we use 229,870 links from the STITCH database
@@
 \citep{HPO2018}, Gene Ontology \cite{GOoriginal2000, GOrecent2020},
-Mammalian Phenotype Ontology \citep{MP2009} and several others.
-We obtained side effects and their links to drugs from SIDER
+Mammalian Phenotype Ontology \citep{MP2009} and several others.  We
+obtained side effects and their links to drugs from SIDER
 \citep{SIDER}; SIDER contains side effects encoded using identifiers
@@
 PhenomeNET ontology using the \textit{Phenomebrowser.net}, which
-provides a SPARQL query endpoint for the mentioned resources.
+provides a SPARQL \cite{} query endpoint for the mentioned resources.
 
@@
 introduced by \citet{Yaminishi2008} consisting of 1,923 interactions
-between 708 drugs and 1,512 proteins, and BioSnap dataset \cite{BioSnap2018}, which consists of 5,017 drug nodes, 2,324 gene nodes and 15,138 edges.
+between 708 drugs and 1,512 proteins, and the BioSnap dataset
+\cite{BioSnap2018} which consists of 5,017 drug nodes, 2,324 gene
+nodes and 15,138 edges.
 
@@
 features for proteins from \textit{DeepGOPlus}
-\citep{DeepGoPlus}. \textit{SmilesTransformer} introduces an
-autoencoder, learning over the SMILES strings and therefore the
-molecular organization of each drug in an unsupervised manner. 
+\citep{DeepGoPlus}. \textit{SmilesTransformer} is an autoencoder
+trained over the SMILES strings, and therefore captures (some aspects
+of) the molecular organization of each drug in an unsupervised manner.
 \textit{DeepGOPlus} provides features derived from protein amino acid

### 2021-04-11 paper/main.tex
.

@@
 dimension.  The precomputed embeddings may have different
-dimensionality and we first process them using a learnable feature
-transformation (LFT) network, i.e., a fully connected neural network
+dimensionality and we first process them using a fully connected neural network
 layer which takes as input an embedding and outputs a representation
-of a particular size. An example structure for both types of features
+of a particular size.
+We use this as a trainable feature transformation.
+
+An example structure for both types of features
 can be found in Figure~\ref{fig:HalfTwinNetwork}. We use this LFT to

### 2021-04-11 paper/main.tex
.

@@
 dimension.  The precomputed embeddings may have different
-dimensionality and we first process them using a fully connected neural network
-layer which takes as input an embedding and outputs a representation
-of a particular size.
-We use this as a trainable feature transformation.
-
-An example structure for both types of features
-can be found in Figure~\ref{fig:HalfTwinNetwork}. We use this LFT to
-reduce the representation size of the embeddings for drugs and
-proteins separately.
+dimensionality and we first process them using a fully connected
+neural network layer which takes as input an embedding and outputs a
+representation of a particular size, i.e., we use this layer as a
+trainable feature transformation and apply it to reduce the
+representation size of the embeddings for drugs and proteins
+separately.  An example structure for both types of features can be
+found in Figure~\ref{fig:HalfTwinNetwork}.
 % While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we
-The use of an LFT enables flexible experimentation as both ontology
-and molecular feature for both drugs and proteins are reduced to
-the same dimensionality for varying sizes of inputs; this allows for a
-high amount of modularity across different experimental setups by
-adding different kinds of features into the model. Additionally, the
-generated features may be used for other tasks. % ; for
+The use of this trainable feature transformation layer enables
+flexible experimentation as both ontology and molecular feature for
+both drugs and proteins are reduced to the same dimensionality for
+varying sizes of inputs; this allows for a high amount of modularity
+across different experimental setups by adding different kinds of
+features into the model. Additionally, the generated features may be
+used for other tasks. % ; for
 % example, the ontology LFT can be reused for a variety of DL2vec-based

### 2021-04-11 paper/main.tex
.

@@
 \bibliographystyle{natbib}
+\usepackage{url}
 
@@
 
-\todo[inline]{Somehow smooth this transition}
+%\todo[inline]{Somehow smooth this transition}
 
@@
 
-\todo[inline]{Consider supplement, or combine with other figures.}
-\begin{figure}[!tpb]%figure1
-	\centerline{\includegraphics[width=0.5\textwidth]{figures/drug_protein_ontology_network.png}}
-	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}
-	\label{fig:Onto}
-\end{figure}
+% \todo[inline]{Consider supplement, or combine with other figures.}
+% \begin{figure}[!tpb]%figure1
+% 	\centerline{\includegraphics[width=0.5\textwidth]{figures/drug_protein_ontology_network.png}}
+% 	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}
+% 	\label{fig:Onto}
+% \end{figure}
 
@@
 
-\begin{figure}[!tpb]%figure1
-	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
-	\caption{Half-twin network applied to molecular and DL2vec
-          features, utilizing deep learnable feature transformations
-          (LFT). The similarity function $\otimes$ yields the
-          similarity between both transformed embeddings e.g. by
-          computing the cosine similarity.}
-	\label{fig:HalfTwinNetwork}
-\end{figure}
+% \todo[inline]{Combine in some larger figure}
+% \begin{figure}[!tpb]%figure1
+% 	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
+% 	\caption{Half-twin network applied to molecular and DL2vec
+%           features, utilizing deep learnable feature transformations
+%           (LFT). The similarity function $\otimes$ yields the
+%           similarity between both transformed embeddings e.g. by
+%           computing the cosine similarity.}
+% 	\label{fig:HalfTwinNetwork}
+% \end{figure}
 
@@
 We include these molecular and ontology-based sub-models within a
-graph neural network (GNN). The graph underlying the GNN is based on
-the protein--protein interaction (PPI) graph. The PPI dataset is
-represented by a graph $G=(V,E)$, where each protein is represented by
-a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$
+graph neural network (GNN) \cite{}. The graph underlying the GNN is
+based on the protein--protein interaction (PPI) graph. The PPI dataset
+is represented by a graph $G=(V,E)$, where each protein is represented
+by a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$
 represents an interaction between two proteins. Additionally, we
@@
 where we only consider
-$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote
-this generalized layer type as \textsc{GENConv} following the
-notation of PyTorch Geometric \cite{PytorchGeometric}.  While the
-reordering is mainly import for numerical stability, this alteration
-also addresses the vanishing gradient problem for deeper convolutional
-networks \cite{GENConv2020}. Additionally, we can also generalize the
+$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote this
+generalized layer type as \textsc{GENConv} following the notation of
+PyTorch Geometric \cite{PytorchGeometric}.  While the reordering is
+mainly import for numerical stability, this alteration also addresses
+the vanishing gradient problem for deeper convolutional networks
+\cite{GENConv2020}. Additionally, we can also generalize the
 aggregation function to allow different weighting functions such as
@@
   DeeperGCN2020}.  The residual blocks are reusable and can be stacked
-multiple times. The structure of the GNN architecture we use is shown
-in Figure~\ref{fig:ResGraphConvBlocks}.
-
-\begin{figure}[!tpb]%figure1
-	\centerline{\includegraphics[width=0.5\columnwidth]{figures/ResGraphConvBlocks.png}}
-	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
-	\label{fig:ResGraphConvBlocks}
-\end{figure}
+multiple times. % The structure of the GNN architecture we use is shown
+% in Figure~\ref{fig:ResGraphConvBlocks}.
+
+% \begin{figure}[!tpb]%figure1
+% 	\centerline{\includegraphics[width=0.5\columnwidth]{figures/ResGraphConvBlocks.png}}
+% 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
+% 	\label{fig:ResGraphConvBlocks}
+% \end{figure}
 
@@
 initializing the graph convolutional update steps. The resulting
-representations are used for a similarity prediction as presented in figure ~\ref{fig:HalfTwinNetwork}. 
+representations are used for a similarity prediction.
+% as presented in
+% figure ~\ref{fig:HalfTwinNetwork}
 When combining ontology and molecular features with or without the
@@
 features, before plugging them into the graph model for the similarity
-computation. A more in-depth image of the overall and final architecture,
-combining both feature types, is depicted in figure
-~\ref{fig:FullModelAllLayers}. Here the original representations are transformed by LFTs and then run through a stack (here with height 3) of residual graph convolutional blocks.
+computation. An overview of the model architecture, combining both
+feature types, is shown in Figure ~\ref{fig:FullModelAllLayers}. Here
+the original representations are transformed by a dense layer and then
+used as input of a stack (with height 3) of residual graph
+convolutional blocks.
 

### 2021-04-11 paper/main.tex
.

@@
 binary classification. As the datasets are highly imbalanced, we use
-the area under the receiver operating characteristic curve
-(AUROC) on training, validation and testing split.
+the area under the receiver operating characteristic curve (AUROC) on
+training, validation and testing split.
 % We calculated the true positive rate (TPR), false positive rate
@@
 subset:
+\todo[inline]{Is the accuracy usually close to $1$? I am still not so
+  sure this imputation with accuracy is well justified -- is there a
+  precedent/paper to cite somewhere?}
 \begin{equation*}
@@
 
-\section{Conclusion}
-
-\vspace*{-10pt}
-
 \section*{Acknowledgements}
-
-\vspace*{-12pt}
+We acknowledge the use of computational resources from the 
+KAUST Supercomputing Core Laboratory.
 
@@
 
-This work has been supported by the... Text Text  Text Text.\vspace*{-12pt}
+This work was supported by funding from King Abdullah University of
+Science and Technology (KAUST) Office of Sponsored Research (OSR)
+under Award No. URF/1/3790-01-01 and URF/1/4355-01-01.
 

### 2021-04-11 paper/main.tex
.

@@
 
-\title[\name{}]{\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
-\author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
+\title[\name{}]{\name{}: Combining Bottom-up and top-down approaches
+  through graph learning over interaction networks for
+  drug-target-interaction prediction} \author[Sample
+\textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert
+  Hoehndorf\,$^{\text{\sfb 2}}$}
 \address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
-$^{\text{\sf 2}}$Department, Institution, City, Post Code,
-Country.}
+  $^{\text{\sf 2}}$Computational Bioscience Research Center, Computer,
+  Electrical and Mathematical Sciences \& Engineering Division, King
+  Abdullah University of Science and Technology, 4700 King Abdullah
+  University of Science and Technology, Thuwal 23955, Saudi Arabia.}
 

### 2021-04-11 paper/main.tex
.

@@
 
-\begin{figure}[!tpb]%figure1
-	\centerline{\includegraphics[width=1\columnwidth]{figures/full_model_all_layers.png}}
-	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
-	\label{fig:FullModelAllLayers}
-\end{figure}
+% \begin{figure}[!tpb]%figure1
+% 	\centerline{\includegraphics[width=1\columnwidth]{figures/full_model_all_layers.png}}
+% 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
+% 	\label{fig:FullModelAllLayers}
+% \end{figure}
 
@@
 computation. An overview of the model architecture, combining both
-feature types, is shown in Figure ~\ref{fig:FullModelAllLayers}. Here
+feature types, is shown in Supplementary Figure
+\ref{fig:FullModelAllLayers}\todo{Move to supplement}. Here
 the original representations are transformed by a dense layer and then
@@
 non-redundant representation. We pretrained the bottom-up and the
-top-down model separately and aimed at best performing models
-with respect to our evaluation metrics. We optimized embedding sizes,
-depth of the neural network, optimizer, learning rate and layer types
-using an extensive, manual grid search. Starting from na\"ive, shallow
-feature transformations with an embedding size of $10$, we scaled the
-network up to residual structures with up to $10$ hidden layers
-leading to embeddings of size $4000$, testing different network widths
-and learning rates for each configuration.
+top-down model separately and aimed at best performing models with
+respect to our evaluation metrics. We optimized embedding sizes, depth
+of the neural network, optimizer, learning rate and layer types using
+an extensive, manual grid search. Starting from shallow feature
+transformations with an embedding size of $10$, we scaled the network
+up to residual structures with up to $10$ hidden layers leading to
+embeddings of size $4000$, testing different network widths and
+learning rates for each configuration.
 
@@
 we then selected the best-performing models (with respect to
-$MicroAUC_p$, see Section \ref{sec:evaluation}) and evaluate them in a
+$MicroAUC_p$, see Section \ref{sec:evaluation}) and evaluate\todo{Only
+evaluate or retrain (from scratch)? Should say here.} them in a
 5-fold protein-split cross-validation on the Yamanishi benchmark

### 2021-04-11 paper/main.tex
.

@@
 but results in unrealistically high precision and therefore high area
-under a precision-recall curve.
+under a precision-recall curve\todo{I note that we do not report AUPR
+  -- why not?}.
 

### 2021-04-13 paper/main.tex
Update citations and fix some TODOs

--- 2021-04-11 paper/main.tex
@@
 similarity \cite{Campillos2008} that is largely complementary to
-knowledge gained from molecular properties. While methods that rely on
+knowledge gained from molecular properties. While methods tha3t rely on
 molecular information are directly predicting whether two molecular
@@
 several other types of biological relations, including similarity
-between represented entities \cite{PREDICT}. Network-based DTI
+between represented entities \cite{Gottlieb2011}. Network-based DTI
 prediction methods use the guilt-by-association principle
-\cite{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002444}
+\cite{Oliver2000}
 and assume that a protein is a likely target for a drug if many of the
 protein's neighbors in the interaction network are targets of the drug
-\cite{https://academic.oup.com/bioinformatics/article/32/7/1057/1743938,
-  others}. Network-based methods have been applied successfully
+\cite{Gillis2012}. Network-based methods have been applied successfully
 to DTI prediction. However, if DTIs are taken as direct physical
@@
 They have previously been applied for a number of tasks, including
-prediction of protein functions \cite{Zitnik2017}, \todo{add 2 more,
-  and then one on DTI prediction}.
+prediction of protein functions \cite{Zitnik2017}, cancer drug response \cite{Liu2020} and drug--target affinity prediction \cite{GraphDTA2020}.
 
@@
 We developed \name{} as a method for predicting DTIs. We use an
-ontology-based machine learning method \cite{} to encode phenotypic
+ontology-based machine learning method \cite{DL2vec2020} to encode phenotypic
 consequences of DTIs and deep learning methods to encode molecular
@@
 prediction methods, and make recommendations on how to avoid
-them. \name is available as Free Software at \url{}\todo{add}.
+them. \name{} is available as Free Software at \url{}\todo{add}.
 
@@
 PhenomeNET ontology using the \textit{Phenomebrowser.net}, which
-provides a SPARQL \cite{} query endpoint for the mentioned resources.
+provides a SPARQL query endpoint for the mentioned resources.
 
@@
 For comparative evaluation, we use the gold standard dataset
-introduced by \citet{Yaminishi2008} consisting of 1,923 interactions
+introduced by \citet{Yamanishi2008} consisting of 1,923 interactions
 between 708 drugs and 1,512 proteins, and the BioSnap dataset
@@
 we then selected the best-performing models (with respect to
-$MicroAUC_p$, see Section \ref{sec:evaluation}) and evaluate\todo{Only
-evaluate or retrain (from scratch)? Should say here.} them in a
+$MicroAUC_p$, see Section \ref{sec:evaluation}), and retrain them from scratch in a
 5-fold protein-split cross-validation on the Yamanishi benchmark
@@
 convolutional methods may be the result of numerical instability and
-fading signals \cite{}\todo{Add citation to original paper}.
+fading signals \cite{DeeperGCN2020}.
 
@@
 we obtain MacroAUC scores of $0.88$, $0.84$ and $0.85$, for the total
-dataset, drug--target pair and drug split, respectively\todo{In the
-  Table is says 0.85 on Yaminishi for DP pair split, here it says 0.84
-  -- which is correct? Need fix}.  The na\"ive
+dataset, drug--target pair and drug split, respectively.  The na\"ive
 predictor shows higher performance on the Yamanishi dataset than on
@@
     \hline
-    Naive predictor&DP pairs&$0.85$& --&--\\
+    Naive predictor&Drugs&$0.85$& --&--\\
     DTINet&DP pairs&$0.91$&$0.74$&$0.67$ \\

### 2021-04-13 src/DTI_predictor.py
Add explanation for AUPRC, some citations

@@ -1,546 +1,1 @@
-import numpy as np
-import networkx as nx
-import pandas as pd
-
-import DTI_data_preparation
-
-from sklearn.model_selection import KFold
-from sklearn import metrics
-
-import matplotlib.pyplot as plt
-
-import tensorflow as tf
-
-from tensorflow.keras import models, layers, optimizers, losses, regularizers
-# import tensorflow.keras.layers as layers
-import tensorflow.keras.backend as K
-# import tensorflow.keras.regularizers as regularizers
-# import tensorflow.keras.optimizers as optimizers
-# import tensorflow.keras.losses as losses
-# from tensorflow.keras.utils import plot_model
-
-
-from keras_dgl.layers.graph_cnn_layer import GraphCNN
-from keras_dgl.layers.graph_attention_cnn_layer import GraphAttentionCNN
-from keras_dgl.layers.multi_graph_cnn_layer import MultiGraphCNN
-from keras_dgl.layers.multi_graph_attention_cnn_layer import MultiGraphAttentionCNN
-from keras_dgl.utils import *
-
-import stellargraph as sg
-from stellargraph import globalvar
-from stellargraph.mapper import GraphSAGENodeGenerator, \
-    Attri2VecNodeGenerator, Attri2VecLinkGenerator, \
-    HinSAGENodeGenerator, \
-    FullBatchNodeGenerator
-from stellargraph.layer import GraphSAGE, Attri2Vec, HinSAGE, GCN, link_classification, GAT
-from stellargraph.layer.ppnp import PPNP
-from stellargraph.layer.appnp import APPNP
-from stellargraph.data import UnsupervisedSampler
-
-import dti_utils
-
-
-
-def missing_target_predictor(results_filename='../results/results_log',
-                             nb_epochs=3,
-                             batch_size=500,
-                             plot=False,
-                             num_samples=[100,20],
-                             embedding_layer_sizes=[32,64],
-                             embedding_output_size=32,
-                             embedding_method='graphsage',
-                             supervised=True):
-    results_table_filename = "../results/results_table"
-
-    print("Loading data ...")
-    drug_list = np.array(DTI_data_preparation.get_drug_list())
-    print("Get protein list ...")
-    protein_list = np.array(DTI_data_preparation.get_human_proteins())[:5000]
-    print("Finished.\n")
-
-    print("Scaling data ...")
[…]

### 2021-04-15 paper/main.tex
Add explanation for AUPRC, some citations
Finished abstract, updated README

@@
 
-\abstract{\textbf{Motivation:} Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
-Text Text Text Text Text Text
-Text Text Text Text Text.\\
-\textbf{Results:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text\\
-\textbf{Availability:} Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text
-Text Text Text Text Text Text Text Text Text Text Text Text Text Text  Text\\
+\abstract{\textbf{Motivation:} \textit{In silico} drug--target interaction (DTI) prediction is crucial for drug discovery and down-stream tasks, such as drug indication and drug--disease prediction.
+However, DTI datasets are prone to certain biases, skewing the effective predictive expressiveness over those datasets.
+Additionally, only few approaches combine phenotypical and molecular features for DTI prediction.\\
+\textbf{Results:} We present \name{} to combine molecular features and functional information with protein--protein interaction networks under usage of graph convolutional neural networks, showing that phenotypic information localizes on the interaction network.
+We hereby focus on protein-centric evaluation, predicting drugs that may target specific targets.
+We further show that common drug--target interaction datasets contain intrinsic biases, majorly affecting predictive performance of compared models.
+Additionally, we propose an modified evaluation scheme and metric to circumvent such skews and to better reflect the problem DTI prediction methods aim to solve.
+\\
+\textbf{Availability:} Source code and data used in \name{} are available at \url{https://github.com/THinnerichs/DTI-VOODOO}\\
 \textbf{Contact:} \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}\\
@@
 prediction methods, and make recommendations on how to avoid
-them. \name{} is available as Free Software at \url{}\todo{add}.
+them. \name{} is available as Free Software at \url{https://github.com/THinnerichs/DTI-VOODOO}.
 
@@
 subset:
-\todo[inline]{Is the accuracy usually close to $1$? I am still not so
-  sure this imputation with accuracy is well justified -- is there a
-  precedent/paper to cite somewhere?}
+
 \begin{equation*}
@@
 and a drug-centric microAUC ($MicroAUC_d$).
+
+Further, we choose AUROC over area under precision recall curve (AUPRC) as the latter is prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases, undermining cross-approach evaluation.\todo[inline]{More Information?
+Right place?}
 

### 2021-04-16 paper/main.tex
Add suplement tables and figures, captions still open, Added TODO for citation

@@
 \usepackage[english]{babel}
-\copyrightyear{2015} \pubyear{2015}
+\copyrightyear{2021} \pubyear{2021}
 \usepackage{natbib}
@@
 Both approaches may be used in conjunction with network inference
-\cite{}. Biological networks used for DTI prediction include
-protein--protein interaction networks \cite{} and networks including
+\todo[inline]{CITE}\cite{}. Biological networks used for DTI prediction include
+protein--protein interaction networks \todo[inline]{CITE}\cite{} and networks including
 several other types of biological relations, including similarity
@@
 % features with respect to other ontologies and hypotheses
-We follow the results of \textit{DL2vec} \cite{} and use as
+We follow the results of \textit{DL2vec} \cite{DL2vec2020} and use as
 activation function $\sigma := \mathrm{LeakyReLU}$ which leads to
@@
 We include these molecular and ontology-based sub-models within a
-graph neural network (GNN) \cite{}. The graph underlying the GNN is
+graph neural network (GNN) \cite{GCNConv}. The graph underlying the GNN is
 based on the protein--protein interaction (PPI) graph. The PPI dataset
@@
 feature types, is shown in Supplementary Figure
-\ref{fig:FullModelAllLayers}\todo{Move to supplement}. Here
+\ref{fig:FullModelAllLayers}. Here
 the original representations are transformed by a dense layer and then
@@
 infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
-for that entity. For those entities, we impute the $MicroAUC$
-interpolating linearly, by using the accuracy for this
+for that entity. As this is quite common for DTI datasets, we do not omit but impute the $MicroAUC$
+interpolating linearly for those entities, by using the accuracy for this
 subset:
@@
 \end{equation*}
-Drugs and proteins can be interchanged in this formulation. We refer
+Note that drugs and proteins can be interchanged in this formulation. We refer
 to the different measures as protein-centric microAUC ($MicroAUC_p$)
@@
 method. DTI-CDF was originally evaluated on all three splitting
-schemes underlining this point \cite{}. While \name{} provides
+schemes underlining this point \cite{DTI-CDF2019}. While \name{} provides
 comparable performance to the na\"ive predictor and DTI-CDF in terms
@@
 DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
-MolTrans itself on the BioSnap dataset \cite{} and compared it to our
+MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared it to our
 ``na\"ive'' predictor as well as \name{} (see Table
@@
 model. A common feature that may be added are drug indications which are
-predictive of drug--target interactions \cite{}; however, we do
+predictive of drug--target interactions \todo[inline]{CITE}\cite{}; however, we do
 not include them in our model as including drug indications would
@@
 follow different strategies. Interaction networks are used widely to
-determine indirect effects of molecular changes \cite{} and predict
-drug--target interactions \cite{}. Our work relies on graph neural
+determine indirect effects of molecular changes \todo[inline]{CITE}\cite{} and predict
+drug--target interactions \todo[inline]{CITE}\cite{}. Our work relies on graph neural
 networks as a way to combine qualitative information about
@@
 training affects the performance of the model. Similarly to previous
-work \cite{}, we find significant differences in predictive
+work \cite{MolTrans2020}\cite{DeepConvDTI2019}, we find significant differences in predictive
 performance across different splitting schemes.
@@
 
-% \subsection{Tested hypotheses}
-
-% In this work we are testing the following hypotheses:
-% \begin{enumerate}
-% 	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
-% 	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
-% \end{enumerate}
-% We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account.
-Thus, we propose a novel approach to combine those mutual exclusive attempts, through the usage of interaction networks, similarity and molecular features.
-Additionally, we test the latter by building a simple molecular DTI predictor and enhance it under usage of the interaction networks.\\
-
-% \section{Methods}
-
-% \subsection{Models}
-% The used model consists of two separate models, that help to fuse together the two methods:
-% \begin{enumerate}
-% 	\item The molecular predictor
-% 	\item The interaction network based predictor
-% \end{enumerate}
-% We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins.
-Regarding proteins, we used the pretrained feature generator from \textit{DeepGoPlus} (\citep{DeepGoPlus}) that was originally designed for protein function prediction and is regarded as state of the art for this purpose.
-For drugs we used a pretrained fingerprint model from \textit{SMILES transformer} (\cite{SmilesTransformer}), that provides a simple and fast method to compute fingerprints through autoencoder models.
-The encodings from these two models were funneled into a simple deep neural network (see Figure~\ref{fig:01}) with few fully connected.
-\\
-% The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE).
-Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug.
-Thus, given a compound-target pair, the nodes of the PPI graph now hold bottom-up features, which can now be processed by the graph learning algorithms.
-\\
-% The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node.
-In contrast to learning over whole graphs we perform node classification within the graph.
-These layers are either graph convolutional layers, that learn a certain kernel over the graph, or attention based.
-Different layers of both and other types such as were tested.
-\\
-% The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
-% For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy.
-Concerning this hierarchy, drug-drug similarity is computed by the Resnik similarity (\cite{Resnik1995}).
-For the given compound we take the corresponding row of this symmetric similarity matrix.
-\\
-
-% Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}.
-This network finally yields our prediction.
-We hereby perform splits over both drugs and proteins, in order to test and show the discrepancy and increasing difficulty.\\ Implementation was done in PyTorch (\cite{Pytorch}) and is available on Github under \href{github.com/thinnerichs/KAUST-dti-metabol}{github.com/thinnerichs/KAUST-dti-metabol}.
-Graph learning methods were build with help of PyTorch-Geometric (\cite{PytorchGeometric}), a geometric deep learning extension library for PyTorch, that recently got a lot of attention in the machine learning community.
-This library gives the potential to use many state of the art graph learning mechanisms, such as plain but effective graph convolution (\citet{GCNConv}), Chebychev kernels (\cite{ChebConv}), ARMA kernels (\cite{ARMAConv}), translation-invariant operators (\cite{FeaStConv}), attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of the latter two (\cite{SAGEConv}).
-The performance of these various layer types were tested for this particular problem, as discussed in the results section.
-
-% \begin{figure}[!tpb]%figure1
-% 	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
-% 	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
-% \end{figure}
-
-% \begin{figure}[!tpb]%figure2
-% \centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
-% \caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins.
-Protein and drug features are represented by blue and red, respectively.}\label{fig:02}
-% \end{figure}
-
-%\section{Discussion}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-04-16 paper/supplement/supplement.tex
Add suplement tables and figures, captions still open, Added TODO for citation

\documentclass[]{article}

\usepackage[a4paper, margin=25mm]{geometry}
\usepackage{graphicx}

\newcommand{\name}{DTI-Voodoo}

%opening
\title{\name{}: Supplement data}
\author{Tilman Hinnerichs and Robert Hoehndorf}
\date{}

\begin{document}

\maketitle

\begin{figure}[ht]%figure1
 	\centerline{\includegraphics[width=0.9\textwidth]{../figures/drug_protein_ontology_network.png}}
 	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}
 	\label{fig:Onto}
\end{figure}

\begin{figure}[ht]%figure1
 	\centerline{\includegraphics[width=0.75\textwidth]{../figures/siamese_network.png}}
 	\caption{Half-twin network applied to molecular and DL2vec
        features, utilizing deep learnable feature transformations
        (LFT). The similarity function $\otimes$ yields the
        similarity between both transformed embeddings e.g. by
        computing the cosine similarity.}
 	\label{fig:HalfTwinNetwork}
\end{figure}

\begin{figure}[ht]%figure1
 	\centerline{\includegraphics[width=0.3\textwidth]{../figures/ResGraphConvBlocks.png}}
 	\caption{Residual architecture built by \textit{Li et al. (2019, 2020b)} enabling deeper graph convolutional models}
 	\label{fig:ResGraphConvBlocks}
\end{figure}

\begin{figure}[ht]%figure1
	\centerline{\includegraphics[width=1\columnwidth]{../figures/full_model_all_layers.png}}
	\caption{Residual architecture built by \textit{Li et al. (2019, 2020b)} enabling deeper graph convolutional models}
	\label{fig:FullModelAllLayers}
\end{figure}

\begin{table}[ht]
	\centering

	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
		\hline
		Approach&\multicolumn{5}{c|}{Graph convolution method}\\
		&without&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
		\hline
		MolPred & $0.69$ & $0.67$& $0.68$& $0.65$& $0.69$\\
		\hline
		OntoPred &$0.88$ &$0.88$&$0.90$&$0.86$&$0.92$\\
		\hline
		\name& $0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
		\hline
	\end{tabular}
	\caption{\label{tab:Results}Results for \name{} on the STITCH
		and Yamanishi datasets evaluated with 5-fold cross-validation. We 
		call the model using only molecular features
		\textit{MolPred} and the model using only ontology-based
		features \textit{OntoPred}. \name{} combines both types of
		features.}
\end{table}

\begin{table}[ht]
	\centering
	\begin{tabular}{|p{1.4cm}|p{0.6cm}|p{0.7cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.7cm}|}
		\hline
		&\multicolumn{6}{c|}{(a) STITCH results}\\
		\hline
		\name{} results&\multicolumn{6}{c|}{PPI graph}\\
		&\multicolumn{3}{c|}{without}&\multicolumn{3}{c|}{with}\\
		&Macro AUC&Micro $AUC'_p$&Micro $AUC_p$&Macro AUC&Micro $AUC'_p$&Micro $AUC_p$\\
		\hline
		MolPred&$0.69$&$0.66$&$0.65$&$0.69$&$0.68$&$0.67$\\
		\hline
		OntoPred&$0.88$&$0.89$&$0.87$&$0.92$&$0.94$&$0.93$\\
		\hline
		\name{} & $0.89$& $0.91$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$&$\mathbf{0.94}$\\
		\hline
	\end{tabular}
	\caption{\label{tab:Results}Results for \name{} on the STITCH
		and Yamanishi datasets evaluated with 5-fold cross-validation. We 
		call the model using only molecular features
		\textit{MolPred} and the model using only ontology-based
		features \textit{OntoPred}. \name{} combines both types of
		features.}
\end{table}

\end{document}

### 2021-04-18 paper/main.tex
.

@@
 similarity \cite{Campillos2008} that is largely complementary to
-knowledge gained from molecular properties. While methods tha3t rely on
+knowledge gained from molecular properties. While methods that rely on
 molecular information are directly predicting whether two molecular
@@
 Both approaches may be used in conjunction with network inference
-\todo[inline]{CITE}\cite{}. Biological networks used for DTI prediction include
-protein--protein interaction networks \todo[inline]{CITE}\cite{} and networks including
-several other types of biological relations, including similarity
-between represented entities \cite{Gottlieb2011}. Network-based DTI
-prediction methods use the guilt-by-association principle
-\cite{Oliver2000}
-and assume that a protein is a likely target for a drug if many of the
-protein's neighbors in the interaction network are targets of the drug
-\cite{Gillis2012}. Network-based methods have been applied successfully
-to DTI prediction. However, if DTIs are taken as direct physical
-interactions between a drug and protein, it remains an unresolved
-question whether the network-based guilt-by-association hypothesis is
-true, or whether an interaction of a drug and protein dys-regulates
-several of the protein's interaction partners, and therefore resulting
-in downstream effects that are not direct interactions but only
-downstream consequences of an interaction.
+\cite{Chen2015}. Biological networks used for DTI prediction include
+protein--protein interaction networks \cite{Feng2017, Lee2018} and
+networks including several other types of biological relations,
+including similarity between represented entities
+\cite{Gottlieb2011}. Network-based DTI prediction methods use the
+guilt-by-association principle \cite{Oliver2000} and assume that a
+protein is a likely target for a drug if many of the protein's
+neighbors in the interaction network are targets of the drug
+\cite{Gillis2012}. Network-based methods have been applied
+successfully to DTI prediction. However, if DTIs are taken as direct
+physical interactions between a drug and protein, it remains an
+unresolved question whether the network-based guilt-by-association
+hypothesis is true, or whether an interaction of a drug and protein
+dys-regulates several of the protein's interaction partners, and
+therefore resulting in effects that are not direct interactions but
+only downstream consequences of an interaction.
 

### 2021-04-18 paper/main.tex
.

@@
 networks including several other types of biological relations,
-including similarity between represented entities
-\cite{Gottlieb2011}. Network-based DTI prediction methods use the
+including similarity between represented entities \cite{Ding2013,
+  Gottlieb2011}. Network-based DTI prediction methods use the
 guilt-by-association principle \cite{Oliver2000} and assume that a
@@
 They have previously been applied for a number of tasks, including
-prediction of protein functions \cite{Zitnik2017}, cancer drug response \cite{Liu2020} and drug--target affinity prediction \cite{GraphDTA2020}.
+prediction of protein functions \cite{Zitnik2017}, cancer drug
+response \cite{Liu2020} and drug--target affinity prediction
+\cite{GraphDTA2020}.
 
@@
 infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
-for that entity. As this is quite common for DTI datasets, we do not omit but impute the $MicroAUC$
-interpolating linearly for those entities, by using the accuracy for this
-subset:
-
+for that entity. As this is quite common for DTI datasets, we do not
+omit but impute the $MicroAUC$ interpolating linearly for those
+entities using the accuracy for this subset:
 \begin{equation*}
@@
 
-Further, we choose AUROC over area under precision recall curve (AUPRC) as the latter is prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases, undermining cross-approach evaluation.\todo[inline]{More Information?
-Right place?}
+Further, we choose the AUROC as primary metric to compare different
+methods, over area under precision recall curve (AUPRC); AUPRC is
+prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases,
+undermining cross-approach evaluation.
 

### 2021-04-18 paper/main.tex
.

@@
 We developed \name{} as a method for predicting DTIs. We use an
-ontology-based machine learning method \cite{DL2vec2020} to encode phenotypic
-consequences of DTIs and deep learning methods to encode molecular
-features. We combine both using a protein interaction network which we
-exploit using a graph neural network. We use this model to test
-whether molecular or phenotype features benefit from the network
-information and find that only phenotype features localize on the
-graph whereas molecular features do not. We further evaluate and
-compare \name{} against several DTI prediction methods and
-demonstrate a substantial improvement of \name{} over the state of the
-art in predicting drugs that target a protein. We also identify and
+ontology-based machine learning method \cite{DL2vec2020} to encode
+phenotypic consequences of DTIs and deep learning methods to encode
+molecular features. We combine both using a protein interaction
+network which we exploit using a graph neural network. We use this
+model to test whether molecular or phenotype features benefit from the
+network information and find that only phenotype features localize on
+the graph whereas molecular features do not. We further evaluate and
+compare \name{} against several DTI prediction methods and demonstrate
+a substantial improvement of \name{} over the state of the art in
+predicting drugs that target a protein. We also identify and
 characterize several biases in both training and evaluating DTI
 prediction methods, and make recommendations on how to avoid
-them. \name{} is available as Free Software at \url{https://github.com/THinnerichs/DTI-VOODOO}.
+them. \name{} is available as Free Software at
+\url{https://github.com/THinnerichs/DTI-VOODOO}.
 
@@
 Further, we choose the AUROC as primary metric to compare different
-methods, over area under precision recall curve (AUPRC); AUPRC is
-prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases,
-undermining cross-approach evaluation.
+methods over other measures such as the area under precision recall
+curve (AUPRC); AUPRC is sensitive to data imbalances \cite{Jeni2013}
+and therefore more challenging to apply to comparing different DTI
+prediction methods.
 

### 2021-04-18 paper/main.tex
.

@@
 There are other types of indirect features that could be added to our
-model. A common feature that may be added are drug indications which are
-predictive of drug--target interactions \todo[inline]{CITE}\cite{}; however, we do
-not include them in our model as including drug indications would
-allow our model to make many trivial predictions based only on
-remembering which targets are often used for which indication;
-including network information would likely benefit predictions based
-on drug indications because different drugs may target the same
-pathway through different mechanisms.
+model. A common feature that may be added are drug indications which
+are predictive of drug--target interactions \cite{Gottlieb2011};
+however, we do not include them in our model as including drug
+indications would allow our model to make many trivial predictions
+based only on remembering which targets are often used for which
+indication; including network information would likely benefit
+predictions based on drug indications because different drugs may
+target the same pathway through different mechanisms.
 
@@
 follow different strategies. Interaction networks are used widely to
-determine indirect effects of molecular changes \todo[inline]{CITE}\cite{} and predict
-drug--target interactions \todo[inline]{CITE}\cite{}. Our work relies on graph neural
-networks as a way to combine qualitative information about
-interactions with additional features (molecular interaction,
+determine indirect effects of molecular changes and predict
+drug--target interactions \cite{LotfiShahreza2017}. Our work relies on
+graph neural networks as a way to combine qualitative information
+about interactions with additional features (molecular interaction,
 phenotypic and functional features); even if only some of these
@@
 training affects the performance of the model. Similarly to previous
-work \cite{MolTrans2020}\cite{DeepConvDTI2019}, we find significant differences in predictive
-performance across different splitting schemes.
+work \cite{MolTrans2020, DeepConvDTI2019}, we find significant
+differences in predictive performance across different splitting
+schemes.
 

### 2021-04-18 paper/main.tex
.

@@
 
-\title[\name{}]{\name{}: Combining Bottom-up and top-down approaches
-  through graph learning over interaction networks for
-  drug-target-interaction prediction} \author[Sample
-\textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert
+\title[\name{}]{\name{}: machine learning over interaction networks and
+  ontology-based background knowledge predicts drug--target
+  interactions}
+\author[Hinnerichs \& Hoehndorf]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert
   Hoehndorf\,$^{\text{\sfb 2}}$}
@@
 
-\abstract{\textbf{Motivation:} \textit{In silico} drug--target interaction (DTI) prediction is crucial for drug discovery and down-stream tasks, such as drug indication and drug--disease prediction.
-However, DTI datasets are prone to certain biases, skewing the effective predictive expressiveness over those datasets.
-Additionally, only few approaches combine phenotypical and molecular features for DTI prediction.\\
-\textbf{Results:} We present \name{} to combine molecular features and functional information with protein--protein interaction networks under usage of graph convolutional neural networks, showing that phenotypic information localizes on the interaction network.
-We hereby focus on protein-centric evaluation, predicting drugs that may target specific targets.
-We further show that common drug--target interaction datasets contain intrinsic biases, majorly affecting predictive performance of compared models.
-Additionally, we propose an modified evaluation scheme and metric to circumvent such skews and to better reflect the problem DTI prediction methods aim to solve.
+\abstract{\textbf{Motivation:} \textit{In silico} drug--target
+  interaction (DTI) prediction is crucial for drug discovery and
+  down-stream tasks, such as drug indication and drug--disease
+  prediction. However, DTI datasets are prone to certain biases,
+  skewing the effective predictive expressiveness over those
+  datasets. Additionally, only few approaches combine phenotypical and
+  molecular features for DTI prediction. \\
+  \textbf{Results:} We present \name{} to combine molecular features
+  and functional information with protein--protein interaction
+  networks under usage of graph convolutional neural networks, showing
+  that phenotypic information localizes on the interaction network. We
+  hereby focus on protein-centric evaluation, predicting drugs that
+  may target specific targets. We further show that common
+  drug--target interaction datasets contain intrinsic biases, majorly
+  affecting predictive performance of compared models. Additionally,
+  we propose an modified evaluation scheme and metric to circumvent
+  such skews and to better reflect the problem DTI prediction methods
+  aim to solve. \\
+  \textbf{Availability:} Source code and data used in \name{} are
+  available at \url{https://github.com/THinnerichs/DTI-VOODOO} \\
+  \textbf{Contact:}
+  \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}
   \\
-\textbf{Availability:} Source code and data used in \name{} are available at \url{https://github.com/THinnerichs/DTI-VOODOO}\\
-\textbf{Contact:} \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}\\
-\textbf{Supplementary information:}10264703 Supplementary data are available at \textit{Bioinformatics}
-online.}
+  \textbf{Supplementary information:}Supplementary data are
+  available at \textit{Bioinformatics} online.}
 

### 2021-04-18 paper/main.tex
.

@@
 \abstract{\textbf{Motivation:} \textit{In silico} drug--target
-  interaction (DTI) prediction is crucial for drug discovery and
-  down-stream tasks, such as drug indication and drug--disease
-  prediction. However, DTI datasets are prone to certain biases,
+  interaction (DTI) prediction is important for drug discovery and
+  drug repurposing.  Approaches to predict DTIs can proceed
+  indirectly, top-down, using phenotypic effects of drugs to identify
+  potential drug targets, or they can be direct, bottom-up and use
+  molecular information to directly predict binding potentials.
+  Both approaches can be combined with information about interaction
+  networks,
+
+  However, DTI datasets are prone to certain biases,
   skewing the effective predictive expressiveness over those
-  datasets. Additionally, only few approaches combine phenotypical and
-  molecular features for DTI prediction. \\
+  datasets.
   \textbf{Results:} We present \name{} to combine molecular features

### 2021-04-18 paper/main.tex
.

@@
   potential drug targets, or they can be direct, bottom-up and use
-  molecular information to directly predict binding potentials.
-  Both approaches can be combined with information about interaction
-  networks,
-
-  However, DTI datasets are prone to certain biases,
-  skewing the effective predictive expressiveness over those
-  datasets.
-  \textbf{Results:} We present \name{} to combine molecular features
-  and functional information with protein--protein interaction
-  networks under usage of graph convolutional neural networks, showing
-  that phenotypic information localizes on the interaction network. We
-  hereby focus on protein-centric evaluation, predicting drugs that
-  may target specific targets. We further show that common
-  drug--target interaction datasets contain intrinsic biases, majorly
-  affecting predictive performance of compared models. Additionally,
-  we propose an modified evaluation scheme and metric to circumvent
-  such skews and to better reflect the problem DTI prediction methods
-  aim to solve. \\
-  \textbf{Availability:} Source code and data used in \name{} are
-  available at \url{https://github.com/THinnerichs/DTI-VOODOO} \\
+  molecular information to directly predict binding potentials.  Both
+  approaches can be combined with information about interaction
+  networks. \\
+  \textbf{Results:} We developed \name{} as a computational method
+  that combines molecular features and ontology-encoded phenotypic
+  effects of drugs with protein--protein interaction networks, and
+  uses a graph neural networks to predict DTIs.  We demonstrate that
+  drug effect features can exploit information in the interaction
+  network whereas molecular features do not.  \name{} is designed to
+  predict candidate drugs for a given protein; we use this formulation
+  to show that common DTI datasets contain intrinsic biases with major
+  affects on performance evaluation and comparison of DTI prediction
+  methods. Using a modified evaluation scheme, we demonstrate that
+  \name{} improves substantially over state of the art DTI prediction
+  methods. \\
+  \textbf{Availability:} \name{} source code and data necessary to
+  reproduce results are freely available at
+  \url{https://github.com/THinnerichs/DTI-VOODOO}. \\
   \textbf{Contact:}
@@
   \\
-  \textbf{Supplementary information:}Supplementary data are
-  available at \textit{Bioinformatics} online.}
+  \textbf{Supplementary information:}Supplementary data are available
+  at \textit{Bioinformatics} online.}
 

### 2021-04-20 paper/main.tex
Update some download scripts, update BioSnap Subset text, remove trash files and presentations

@@
 diseases for which drugs can be more profitably marketed). This may
-affect common evaluation schemes where it is possible
-to exploit these biases within DTI prediction \cite{Survey2018}.
+affect common evaluation schemes\cite{Survey2018} where it is possible
+to exploit these biases within DTI prediction \cite{}. \citet{vanLaarhoven2014} showed that another bias can be heavily capitalized over the dataset of \citet{Yamanishi2008}.
 
@@
 
-We only use proteins in our analysis that have at least one link in
-either STITCH or STRING, and drugs with at least one side effect and
-one existing target. Therefore, the intersection between these
-resources yields 1,428 drugs and 7,368 human proteins with 32,212
-interactions for the training phase. We provide links to and methods
-for obtaining and processing the necessary data on Github.
-
 For comparative evaluation, we use the gold standard dataset
@@
 nodes and 15,138 edges.
+
+We only use proteins in our analysis that have at least one link in STRING or one association in PhenomeNET, and drugs with at least one side effect. Therefore, the intersection between these
+resources yields 1,428 drugs and 7,368 human proteins with 32,212
+interactions for STITCH, 1,837 interactions between 680 drugs and 1,458 proteins for Yamanishi, and 6,498 links between 949 drugs and 2,221 proteins for BioSnap dataset.
+We provide links to and methods
+for obtaining and processing the necessary data on Github.
 
@@
 feature types, is shown in Supplementary Figure
-\ref{fig:FullModelAllLayers}. Here
+4. Here
 the original representations are transformed by a dense layer and then
@@
 different graph neural network architectures and configurations
-(Supplementary Table X\todo{Add to Supplement!}). While the {\em
+(Supplementary Table 1). While the {\em
   GCNConv} and {\em GENConv} architecture already shows some minor
@@
 reached very similar results to the original \textit{OntoPred} model
-and showed no improvement (Supplementary Table X\todo{Add!}).
+and showed no improvement (Supplementary Table 1).
 
@@
     DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
-    MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
+    MolTrans&DP pairs&$\mathbf{0.90}$&$0.77$&$0.74$\\
     \name&Proteins&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
@@
     on the BioSnap dataset; we evaluated the original and the
-    protein-based split in a cross-validation}.
+    protein-based split in a cross-validation.}.
 \end{table}

### 2021-04-20 paper/supplement/supplement.tex
Update some download scripts, update BioSnap Subset text, remove trash files and presentations

@@
 
-	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
+	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{1.2cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
 		\hline
-		Approach&\multicolumn{5}{c|}{Graph convolution method}\\
-		&without&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
+		Approach&\multicolumn{6}{c|}{Graph convolution method}\\
+		&without&without/ empty RGC&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
 		\hline
-		MolPred & $0.69$ & $0.67$& $0.68$& $0.65$& $0.69$\\
+		MolPred & $0.69$ & $0.67$& $0.67$& $0.68$& $0.65$& $0.69$\\
 		\hline
-		OntoPred &$0.88$ &$0.88$&$0.90$&$0.86$&$0.92$\\
+		OntoPred &$0.88$ &$0.87$ &$0.88$&$0.90$&$0.86$&$0.92$\\
 		\hline
-		\name& $0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
+		\name& $0.89$&$0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
 		\hline
 	\end{tabular}
-	\caption{\label{tab:Results}Results for \name{} on the STITCH
-		and Yamanishi datasets evaluated with 5-fold cross-validation. We 
-		call the model using only molecular features
-		\textit{MolPred} and the model using only ontology-based
-		features \textit{OntoPred}. \name{} combines both types of
-		features.}
+	\caption{\label{tab:Results}Results for various neural graph convolutional methods enhancing MolPred, OntoPred and \name{} over the STITCH dataset.
+We evaluated all methods over different stacking heights without and with residual graph convolution blocks (RGC, ResGraphConv) as depicted in Supplementary figure \ref{fig:ResGraphConvBlocks}.
+``Empty RGC`` denotes a RGC block with a plain neural layer substituting the GraphConv layer.}
 \end{table}
@@
 	\centering
-	\begin{tabular}{|p{1.4cm}|p{0.6cm}|p{0.7cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.7cm}|}
+	\begin{tabular}{|p{2.0cm}|p{0.8cm}|p{0.8cm}|p{0.8cm}|p{0.8cm}|p{0.8cm}|p{0.8cm}|}
 		\hline
@@
 	\end{tabular}
-	\caption{\label{tab:Results}Results for \name{} on the STITCH
-		and Yamanishi datasets evaluated with 5-fold cross-validation. We 
-		call the model using only molecular features
-		\textit{MolPred} and the model using only ontology-based
-		features \textit{OntoPred}. \name{} combines both types of
-		features.}
+	\caption{Extended results table for MolPred, OntoPred and \name{} over STITCH with added $MicroAUC'_p$ for comparison of both metrics.}
 \end{table}

### 2021-04-20 src/DTI_predictor.py

@@ -1,1 +1,0 @@
-import numpy as np
import networkx as nx
import pandas as pd

import DTI_data_preparation

from sklearn.model_selection import KFold
from sklearn import metrics

import matplotlib.pyplot as plt

import tensorflow as tf

from tensorflow.keras import models, layers, optimizers, losses, regularizers
# import tensorflow.keras.layers as layers
import tensorflow.keras.backend as K
# import tensorflow.keras.regularizers as regularizers
# import tensorflow.keras.optimizers as optimizers
# import tensorflow.keras.losses as losses
# from tensorflow.keras.utils import plot_model


from keras_dgl.layers.graph_cnn_layer import GraphCNN
from keras_dgl.layers.graph_attention_cnn_layer import GraphAttentionCNN
from keras_dgl.layers.multi_graph_cnn_layer import MultiGraphCNN
from keras_dgl.layers.multi_graph_attention_cnn_layer import MultiGraphAttentionCNN
from keras_dgl.utils import *

import stellargraph as sg
from stellargraph import globalvar
from stellargraph.mapper import GraphSAGENodeGenerator, \
    Attri2VecNodeGenerator, Attri2VecLinkGenerator, \
    HinSAGENodeGenerator, \
    FullBatchNodeGenerator
from stellargraph.layer import GraphSAGE, Attri2Vec, HinSAGE, GCN, link_classification, GAT
from stellargraph.layer.ppnp import PPNP
from stellargraph.layer.appnp import APPNP
from stellargraph.data import UnsupervisedSampler

import dti_utils



def missing_target_predictor(results_filename='../results/results_log',
                             nb_epochs=3,
                             batch_size=500,
                             plot=False,
                             num_samples=[100,20],
                             embedding_layer_sizes=[32,64],
                             embedding_output_size=32,
                             embedding_method='graphsage',
                             supervised=True):
    results_table_filename = "../results/results_table"

    print("Loading data ...")
    drug_list = np.array(DTI_data_preparation.get_drug_list())
    print("Get protein list ...")
    protein_list = np.array(DTI_data_preparation.get_human_proteins())[:5000]
    print("Finished.\n")

    print("Scaling data ...")
    side_effect_features = np.tile(DTI_data_preparation.get_side_effect_similarity_feature_list(drug_list), (len(protein_list),1))
    side_effect_features = side_effect_features.reshape((len(protein_list), len(drug_list), 1430))
[…]

### 2021-04-21 paper/main.tex
Submit revisioned paper

@@
 ``top-down'' and ``bottom-up'' approaches. Top-down approaches start
-form observable characteristics resulting from a drug--target
+from observable characteristics resulting from a drug--target
 interaction, such as side-effects or the diseases treated by a drug,
@@
 Both bottom-up and top-down approaches to drug--target interaction
-prediction have some advantages and limitations. Generally, bottom-up
+prediction bear some advantages and limitations. Generally, bottom-up
 methods face the challenge to predict whether a structure binds to a
@@
 (where binding sites and molecular forces need to be determined for
-accurate prediction) but also properties such as where a protein is
+accurate prediction) but also properties namely where a protein is
 expressed or in which celltypes and anatomical structures. Top-down
@@
 molecular information are directly predicting whether two molecular
-could interact, top-down methods rely on more indirect means and 
+could interact, top-down methods base on more indirect means and 
 infer a DTI from observable effects resulting from the interaction.
@@
 hypothesis is true, or whether an interaction of a drug and protein
-dys-regulates several of the protein's interaction partners, and
+dysregulates several of the protein's interaction partners, and
 therefore resulting in effects that are not direct interactions but
@@
 In particular, Graph Convolutional Networks \cite{GCNConv} and
-their variants using different types of kernels \cite{ChebConv,
+their variants operate on different types of kernels \cite{ChebConv,
   ARMAConv}, including attention mechanisms \cite{GATConv}, and
@@
 
-One challenge in DTI prediction are potential biases resulting from
+Potential biases resulting from
 the underlying datasets \citep{Pahikkala2014} which may affect model
-evaluation and comparison. First, novel drugs are often designed by
+evaluation and comparison pose a challenge for DTI prediction. Firstly, novel drugs are often developed by
 altering non-functional components of a drug, leading to two and more
@@
 among training and evaluation dataset, resulting in a better
-(measured) predictive performance than would be expected when the
+(measured) predictive performance than it would be expected when the
 model is applied to identify drugs that target a protein for which no
-drugs yet exist. Second, some proteins (which we call \textit{hub
+drugs yet exist. Secondly, some proteins (which we call \textit{hub
   proteins}) have significantly more known interactions with drugs
@@
 to target them due to their involvement in more common diseases (or
-diseases for which drugs can be more profitably marketed). This may
+diseases for which drugs can be more profitably marketed). This might
 affect common evaluation schemes\cite{Survey2018} where it is possible
@@
 molecular features. We combine both using a protein interaction
-network which we exploit using a graph neural network. We use this
+network which we exploit with the aid of a graph neural network. We use this
 model to test whether molecular or phenotype features benefit from the
@@
 encoded using a Word2vec \citep{Word2vec2013} model. Therefore, DL2Vec
-generates representations that can encode drug effects or protein
-functions while preserving their semantic neighbourhood within that
+generates representations that enable to encode drug effects or protein
+functions while preserving their semantic neighborhood within that
 graph.
@@
 contrastive loss using cosine similarity.  A half-twin neural network
-aims to learn a similarity between two embeddings of the same
-dimension.  The precomputed embeddings may have different
-dimensionality and we first process them using a fully connected
+aims to learn a similarity between two embeddings of variable but same
+dimension. As the original feature space may have varying dimensionality, we first process them using a fully connected
 neural network layer which takes as input an embedding and outputs a
@@
 While in this initial formulation of a GCNConv the node-wise update
-step is defined by the sum over all neighbouring node representations,
+step is defined by the sum over all neighboring node representations,
 we can alter this formulation to other message passing schemes.  We
@@
 PyTorch Geometric \cite{PytorchGeometric}.  While the reordering is
-mainly import for numerical stability, this alteration also addresses
+mainly important for numerical stability, this alteration also addresses
 the vanishing gradient problem for deeper convolutional networks
@@
 \subsubsection{Hyperparameter tuning}
-As the number of drug-targets are sparse with respect to the number of
+As the number of drug-targets are sparse with respect to the amount of
 both drugs and proteins considered, the training, validation and
@@
 interaction. The full workflow scheme is depicted in
-figure~\ref{fig:ModelWorkflow}
+Figure~\ref{fig:ModelWorkflow}
 
@@
 Another view on these two approaches is as direct and indirect ways to
-predict drug--target interactions. Molecular information can be used
+predict drug--target interactions. On one hand, molecular information can be used
 to directly determine whether two molecules (such as a drug and
@@
 protein function can be used to indirectly suggest candidate
-drug--target interactions. Molecular information will be specific to a
+drug--target interactions. Molecular features will be specific to a
 drug--target pair and we would not expect this information to
@@
 model. A common feature that may be added are drug indications which
-are predictive of drug--target interactions \cite{Gottlieb2011};
-however, we do not include them in our model as including drug
+are predictive of drug--target interactions \cite{Gottlieb2011}.
+However, we do not include them in our model as including drug
 indications would allow our model to make many trivial predictions

### 2021-04-21 paper/main.tex
.

@@
 the underlying datasets \citep{Pahikkala2014} which may affect model
-evaluation and comparison pose a challenge for DTI prediction. Firstly, novel drugs are often developed by
+evaluation and comparison pose a challenge for DTI
+prediction. Firstly, novel drugs are often developed by
 altering non-functional components of a drug, leading to two and more
@@
 diseases for which drugs can be more profitably marketed). This might
-affect common evaluation schemes\cite{Survey2018} where it is possible
-to exploit these biases within DTI prediction \cite{}. \citet{vanLaarhoven2014} showed that another bias can be heavily capitalized over the dataset of \citet{Yamanishi2008}.
+affect common evaluation schemes \cite{Survey2018} where it is possible
+to exploit these biases within DTI prediction \cite{}.
+\citet{vanLaarhoven2014} showed that several bias can be exploited on
+the dataset of \citet{Yamanishi2008}.
 
@@
 we then selected the best-performing models (with respect to
-$MicroAUC_p$, see Section \ref{sec:evaluation}), and retrain them from scratch in a
-5-fold protein-split cross-validation on the Yamanishi benchmark
-dataset to avoid validation overfitting and yield more realistic
-testing results. To evaluate the influence of the different features
-separately, and to determine whether they ``localize'' on the PPI
-graph (and therefore can be exploited successfully by the graph neural
-networks), we train and evaluate models with different types of
+$MicroAUC_p$, see Section \ref{sec:evaluation}), and retrain them from
+scratch in a 5-fold protein-split cross-validation on the Yamanishi
+benchmark dataset to avoid validation overfitting and yield more
+realistic testing results. To evaluate the influence of the different
+features separately, and to determine whether they ``localize'' on the
+PPI graph (and therefore can be exploited successfully by the graph
+neural networks), we train and evaluate models with different types of
 features, and with and without inclusion of the PPI graph,

### 2021-04-21 paper/cover_letter/cover_letter.tex
Update download scripts

\documentclass[]{letter}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{TO\_ADDRESS}
\address{ADDRESS}

Dear Editor,

Please find our manuscript entitled "DTI-Voodoo: machine learning over interaction networks and
ontology-based background knowledge predicts drug--target
interactions" for your consideration. We therefore developed a novel method for drug--target interaction prediction (DTI) utilizing both ontology-encoded phenotypical and molecular features for both drugs and proteins. We further combine these methods with interaction networks, aiming for a more protein-centric prediction and evaluation.

We designed DTI-Voodoo under usage of graph convolutional neural network, exploiting the ``guilt-by-association`` principle and incorporate both top-down and bottom-up features into protein-protein interaction networks. Furthermore, we show that common DTI datasets share immanent biases, majorly influencing predictive DTI model performance evaluations. Thus, we propose a modified evaluation scheme, showing that DTI-Voodoo advances significantly over other state of the art DTI prediction approaches over several datasets.

The importance of DTI prediction for other challenges such as drug repurposing and repositioning makes DTI-Voodoo highly influential for a variety of researchers. Eventually, DTI-Voodoo's versatile methodology may be used over alternative protein-based representations to improve performances within other predictive tasks in the realm of bioinformatics and computational biology.

Therefore, we believe that
DTI-Voodoo will be well suited for the broad readership of
Bioinformatics.

Please do not hesitate to contact us should you require any further
information.

With best wishes,
Tilman Hinnerichs and Robert Hoehndorf.\\

\signature{}

%enclosure listing
%\encl{}

\end{letter}
\end{document}

### 2021-04-21 paper/main.tex
Update download scripts

--- 2021-04-21 paper/main.tex
@@
   effects of drugs with protein--protein interaction networks, and
-  uses a graph neural networks to predict DTIs.  We demonstrate that
+  uses a graph convolutional neural network to predict DTIs.  We demonstrate that
   drug effect features can exploit information in the interaction
@@
 % features with respect to other ontologies and hypotheses
-We follow the results of \textit{DL2vec} \cite{DL2vec2020} and use as
-activation function $\sigma := \mathrm{LeakyReLU}$ which leads to
+We follow the results of \textit{DL2vec} \cite{DL2vec2020} and use $\sigma := \mathrm{LeakyReLU}$ as
+activation function  which leads to
 improved performance compared to other activation functions.

### 2021-04-21 data/STRING_data/download.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+URL=https://stringdb-static.org/download/protein.links.full.v11.0/9606.protein.links.full.v11.0.txt.gz
+curl -O "$URL" && gunzip -f "${URL##*/}"
+wait
+
+URL=https://stringdb-static.org/download/protein.aliases.v11.0/9606.protein.aliases.v11.0.txt.gz
+curl -O "$URL" && gunzip -f "${URL##*/}"
+wait
+
+URL=https://stringdb-static.org/download/protein.sequences.v11.0/9606.protein.sequences.v11.0.fa.gz
+curl -O "$URL" && gunzip -f "${URL##*/}"
+wait
+

### 2021-04-22 paper/main.tex
Finalize download, preprocessing and training script

@@
 affect common evaluation schemes \cite{Survey2018} where it is possible
-to exploit these biases within DTI prediction \cite{}.
+to exploit these biases within DTI prediction \cite{}\todo[inline]{cite or delete}.
 \citet{vanLaarhoven2014} showed that several bias can be exploited on
@@
 separately.  An example structure for both types of features can be
-found in Figure~\ref{fig:HalfTwinNetwork}.
+found in Supplementary Figure 2.
 % While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we
@@
 $MicroAUC_p$ score for all considered methods, leading to similar 
-results as observed on the Yaminishi dataset.
+results as observed on the Yaminishi dataset. As \name{} considers less drugs 
+than alternative approaches, we perform a two-sided T-test yielding $p<0.0001$ for 
+$\alpha<0.01$ showing the significance of \name's performance gain over other state-of-the-art methods.\todo[inline]{Formulation correct? Confidence interval?}
 

### 2021-04-22 src/DTI_data_preparation.py

@@ -466,78 +466,28 @@ def get_BioSnap_data(original_drug_list, original_protein_list):
 
     return drug_list[drug_indices], protein_list[protein_indices], dti_matrix[drug_indices, :][:, protein_indices]
 
-def test():
-    # print("DTI", len(get_human_proteins()))
-    # print("PPI", len(PPI_utils.get_human_protein_list()))
-
-    dti_graph = get_human_DTI_graph()
-    print(len(dti_graph.nodes()))
-    print(len(dti_graph.edges()))
-
-    ppi_graph = PPI_utils.get_PPI_graph()
-    print(len(set(dti_graph.nodes()) & set(ppi_graph.nodes())))
-
 
 if __name__ == '__main__':
+    DDI_utils.write_chemical_stereo_to_normal_mapping()
+    DDI_utils.write_STITCH_db_Pubchem_mapping_dict()
 
-    parser = argparse.ArgumentParser()
+    PPI_utils.prune_protein_protein_db(min_score=700, mode='')
+    PPI_utils.write_PPI_graph(700)
 
-    parser.add_argument("--mode", type=str, default='')
+    PhenomeNET_DL2vec_utils.write_UMLS_NET_files()
+    PhenomeNET_DL2vec_utils.write_PhenomeNET_files('drug')
+    PhenomeNET_DL2vec_utils.write_PhenomeNET_files('GO')
+    PhenomeNET_DL2vec_utils.write_PhenomeNET_files('uberon')
+    PhenomeNET_DL2vec_utils.write_PhenomeNET_files('MP')
 
-    config = parser.parse_args()
+    write_human_DTI_graph(700, mode='database')
 
-    mode=config.mode
-    # write_human_DTI_graph(300, mode='experimental')
-    # write_human_DTI_graph(500, mode='experimental')
-    # write_human_DTI_graph(700, mode=mode)
-
-    # dti_graph = get_human_DTI_graph()
-    # print("Nodes", len(dti_graph.nodes()))
-    # print("Edges", len(dti_graph.edges()))
-
-    # drug_list = get_drug_list()
-    # print('Drugs:', len(drug_list))
-
-    write_human_protein_list(mode=mode)
-
-
-    # write_truncated_drug_to_SMILES_dict()
-    # get_truncated_drug_to_SMILES_dict()
-
-    # protein_list = get_human_proteins()
-    # PPI_utils.write_protein_fasta(protein_list=protein_list)
+    print('\n----------')
[…]

### 2021-04-22 src/PhenomeNET_DL2vec_utils.py

@@ -213,13 +213,13 @@ def write_PhenomeNET_files(mode='all', yamanishi_test=False):
             for protein in protein_list:
                 f.write(protein+'\n')
 
-def output_example_DL2vec_command(workers=48, embedsize=200):
+def output_example_DL2vec_command(prefix='', workers=48, embedsize=200):
     print('Running DL2vec embedding generator ...')
     path = '../../data/PhenomeNET_data/'
     onto = path + 'phenomenet.owl'
-    asso = path + 'association_file'
-    outfile = path + 'embedding_model'
-    ents = path + 'entity_list'
+    asso = path + prefix+'association_file'
+    outfile = path + prefix+'embedding_model'
+    ents = path + prefix +'entity_list'
     command = 'python runDL2vec.py -embedsize {embedsize} -ontology {onto} -associations {asso} -outfile {outfile} -entity_list {ents} -num_workers {num_workers}'.format(
         embedsize=embedsize,
         onto=onto,

### 2021-04-22 src/molecular_predictor.py

@@ -117,7 +117,8 @@ def write_encoded_proteins():
     in_file = protein_dir+'data/PPI_graph_protein_seqs.fasta'
     out_file = protein_dir+'results/output'
 
-    subprocess.call('source' + protein_dir+f'predict.sh {in_file} {out_file}')
+    print('Execute this command in',protein_dir)
+    print(f'source {protein_dir}predict.sh {in_file} {out_file}')
 
 def molecular_predictor(config):
     model_st = 'molecular_predictor'
@@ -529,10 +530,7 @@ if __name__=='__main__':
         write_encoded_drugs(drug_list, mode='rnn')
         print('Successfully computed molecular drug embeddings. Aborting rest of script.')
 
-        raise Exception
-
-    # Run classifier
-    if config.model == 'protein':
+    elif config.model == 'protein':
         molecular_predictor(config)
     elif config.model == 'drug':
         drug_split_molecular_predictor(config)

### 2021-04-22 paper/main.tex
Add 20% of drugs discussion todo to discussion

@@
 
+\todo[inline]{Why is having only having 20\% of drugs reasonable -> significance test in results}
 One major component of our experiments was to determine how the

### 2021-04-22 paper/main.tex
.

@@
 
-Potential biases resulting from
-the underlying datasets \citep{Pahikkala2014} which may affect model
-evaluation and comparison pose a challenge for DTI
-prediction. Firstly, novel drugs are often developed by
-altering non-functional components of a drug, leading to two and more
-very similar drugs designed to target the same proteins
+Potential biases resulting from the underlying datasets
+\citep{Pahikkala2014} which may affect model evaluation and comparison
+pose a challenge for DTI prediction. Firstly, novel drugs are often
+developed by altering non-functional components of a drug, leading to
+two and more very similar drugs designed to target the same proteins
 \cite{Overington2006}. This can result in a bias when it leads to
@@
 diseases for which drugs can be more profitably marketed). This might
-affect common evaluation schemes \cite{Survey2018} where it is possible
-to exploit these biases within DTI prediction \cite{}\todo[inline]{cite or delete}.
+affect common evaluation schemes where it is possible to exploit these
+biases within DTI prediction \cite{Survey2018}.
 \citet{vanLaarhoven2014} showed that several bias can be exploited on
@@
 $MicroAUC_p$ score for all considered methods, leading to similar 
-results as observed on the Yaminishi dataset. As \name{} considers less drugs 
-than alternative approaches, we perform a two-sided T-test yielding $p<0.0001$ for 
-$\alpha<0.01$ showing the significance of \name's performance gain over other state-of-the-art methods.\todo[inline]{Formulation correct? Confidence interval?}
+results as observed on the Yaminishi dataset. % As \name{} considers less drugs 
+% than alternative approaches, we perform a two-sided T-test yielding $p<0.0001$ for 
+% $\alpha<0.01$ showing the significance of \name's performance gain over other state-of-the-art methods.\todo[inline]{Formulation correct? Confidence interval?}
 

### 2021-04-24 paper/main.tex
Update significance test part in results

--- 2021-04-22 paper/main.tex
@@
 $MicroAUC_p$ score for all considered methods, leading to similar 
-results as observed on the Yaminishi dataset. As \name{} considers less drugs 
-than alternative approaches, we perform a two-sided T-test yielding $p<0.0001$ for 
-$\alpha<0.01$ showing the significance of \name's performance gain over other state-of-the-art methods.\todo[inline]{Formulation correct? Confidence interval?}
-
-In all our experiments, \name{} has the highest Micro $AUC_p$,
-demonstrating that \name{} can improve over other methods in the task
+results as observed on the Yaminishi dataset.
+As \name{} considers less drugs than alternative approaches, we execute a one-tailed T-test with $\alpha=0.01$ over all considered methods.
+Considering MacroAUCs, \name{} improves significantly over MolTrans ($p<0.0001$), DeepDTI ($p<0.0001$), DeepDTA ($p<0.0001$), and DeepConv-DTI ($p<0.0001$) with respect to BioSnap dataset, and significantly over DTINet ($p<0.0001$), and DTIGEMS+ ($p<0.0001$) with respect to Yamanishi dataset.
+However, DTI-CDF performs significantly better than \name{} ($\alpha=0.05, p=0.02$) when considering MacroAUCs.
+As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF.
+We thus run another one-tailed T-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferron correction.
+We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$).
+
+In all our experiments, \name{} has the highest $MicroAUC_p$,
+demonstrating that \name{} can improve significantly over other methods in the task
 of identifying drugs that target a specific protein. Several methods

### 2021-04-24 src/DTI_data_preparation.py

@@ -420,8 +420,6 @@ def get_BioSnap_data(original_drug_list, original_protein_list):
             dti_pairs.append((drug, protein))
 
     drug_list = list(set(drug_list))
-    print(drug_list[:10])
-    raise Exception
     protein_list = list(set(protein_list))
 
     with open(file=path+'BioSnap_DB_drug_list', mode='w') as f:

### 2021-04-24 paper/main.tex

--- 2021-04-22 paper/main.tex
@@
 $MicroAUC_p$ score for all considered methods, leading to similar 
-results as observed on the Yaminishi dataset. % As \name{} considers less drugs 
-% than alternative approaches, we perform a two-sided T-test yielding $p<0.0001$ for 
-% $\alpha<0.01$ showing the significance of \name's performance gain over other state-of-the-art methods.\todo[inline]{Formulation correct? Confidence interval?}
-
-In all our experiments, \name{} has the highest Micro $AUC_p$,
-demonstrating that \name{} can improve over other methods in the task
+results as observed on the Yaminishi dataset.
+As \name{} considers less drugs than alternative approaches, we execute a one-tailed T-test with $\alpha=0.01$ over all considered methods.
+Considering Macro AUCs, \name{} improves significantly over MolTrans ($p<0.0001$), DeepDTI ($p<0.0001$), DeepDTA ($p<0.0001$), and DeepConv-DTI ($p<0.0001$) with respect to BioSnap dataset, and significantly over DTINet ($p<0.0001$) and DTIGEMS+ ($p<0.0001$) with respect to Yamanishi dataset.
+However, DTI-CDF performs significantly better than \name{} ($\alpha=0.05, p=0.02$) when considering Macro AUC.
+As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF.
+We further run another one-tailed T-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction.
+We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$) regarding Yamanishi dataset.
+
+In all our experiments, \name{} has the highest $MicroAUC_p$,
+demonstrating that \name{} can improve significantly over other methods in the task
 of identifying drugs that target a specific protein. Several methods

### 2021-04-25 paper/main.tex
.

@@
 
-We only use proteins in our analysis that have at least one link in STRING or one association in PhenomeNET, and drugs with at least one side effect. Therefore, the intersection between these
-resources yields 1,428 drugs and 7,368 human proteins with 32,212
-interactions for STITCH, 1,837 interactions between 680 drugs and 1,458 proteins for Yamanishi, and 6,498 links between 949 drugs and 2,221 proteins for BioSnap dataset.
-We provide links to and methods
-for obtaining and processing the necessary data on Github.
+We only use proteins in our analysis that have at least one link in
+STRING or one association in PhenomeNET, and drugs with at least one
+side effect. Therefore, the intersection between these resources
+yields 1,428 drugs and 7,368 human proteins with 32,212 interactions
+for STITCH, 1,837 interactions between 680 drugs and 1,458 proteins
+for Yamanishi, and 6,498 links between 949 drugs and 2,221 proteins
+for BioSnap dataset. We provide links to and methods for obtaining and
+processing the necessary data in our Github repository.
 
@@
 $MicroAUC_p$ score for all considered methods, leading to similar 
-results as observed on the Yaminishi dataset.
-As \name{} considers less drugs than alternative approaches, we execute a one-tailed T-test with $\alpha=0.01$ over all considered methods.
-Considering Macro AUCs, \name{} improves significantly over MolTrans ($p<0.0001$), DeepDTI ($p<0.0001$), DeepDTA ($p<0.0001$), and DeepConv-DTI ($p<0.0001$) with respect to BioSnap dataset, and significantly over DTINet ($p<0.0001$) and DTIGEMS+ ($p<0.0001$) with respect to Yamanishi dataset.
-However, DTI-CDF performs significantly better than \name{} ($\alpha=0.05, p=0.02$) when considering Macro AUC.
+results as observed on the Yaminishi dataset. As \name{} considers
+less drugs than alternative approaches, we execute a one-tailed T-test
+with $\alpha=0.01$ over all considered methods. Considering Macro
+AUCs, \name{} improves significantly ($p<0.0001$, one-sided t-test)
+over MolTrans, DeepDTI, DeepDTA, and DeepConv-DTI on the BioSnap
+dataset, and significantly ($p<0.0001$, one-sided t-test) over DTINet
+and DTIGEMS+ on the Yamanishi dataset. However, DTI-CDF performs
+significantly better than \name{} when considering Macro AUC ($p=0.02$).
 As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF.

### 2021-04-25 paper/main.tex
.

@@
 dataset, and significantly ($p<0.0001$, one-sided t-test) over DTINet
-and DTIGEMS+ on the Yamanishi dataset. However, DTI-CDF performs
-significantly better than \name{} when considering Macro AUC ($p=0.02$).
-As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF.
-We further run another one-tailed T-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction.
+and DTIGEMS+ on the Yamanishi dataset' DTI-CDF performs
+significantly better than \name{} ($p=0.02$, one-sided t-test). % As
+                                % we compare \name{} to 4 different
+                                % methods over BioSnap and 3 over
+                                % Yamanishi dataset, we add a
+                                % Bonferroni correction by multiplying
+                                % the $p$-values by 4 and 3,
+                                % respectively, leading to $p<0.0002$
+                                % ($\alpha=0.01$) for all previous
+                                % comparisons except DTI-CDF.
+We further run another one-tailed t-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction.
 We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$) regarding Yamanishi dataset.

### 2021-04-25 paper/main.tex
Update significance test results reporting

--- 2021-04-24 paper/main.tex
@@
 results as observed on the Yaminishi dataset.
-As \name{} considers less drugs than alternative approaches, we execute a one-tailed T-test with $\alpha=0.01$ over all considered methods.
-Considering Macro AUCs, \name{} improves significantly over MolTrans ($p<0.0001$), DeepDTI ($p<0.0001$), DeepDTA ($p<0.0001$), and DeepConv-DTI ($p<0.0001$) with respect to BioSnap dataset, and significantly over DTINet ($p<0.0001$) and DTIGEMS+ ($p<0.0001$) with respect to Yamanishi dataset.
-However, DTI-CDF performs significantly better than \name{} ($\alpha=0.05, p=0.02$) when considering Macro AUC.
-As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<0.0002$ ($\alpha=0.01$) for all previous comparisons except DTI-CDF.
-We further run another one-tailed T-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction.
-We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$) regarding Yamanishi dataset.
+As \name{} considers less drugs than alternative approaches, we execute a one-tailed T-test comparing \name{} to all examined methods.
+Considering Macro AUCs, \name{} improves significantly over MolTrans ($p<10^{-4}$), DeepDTI ($p<10^{-4}$), DeepDTA ($p<10^{-4}$), and DeepConv-DTI ($p<10^{-4}$) with respect to BioSnap dataset, and significantly over DTINet ($p<10^{-4}$) and DTIGEMS+ ($p<10^{-4}$) with respect to Yamanishi dataset.
+However, DTI-CDF performs significantly better than \name{} ($p=0.02$) when considering Macro AUC.
+As we compare \name{} to 4 different methods over BioSnap and 3 over Yamanishi dataset, we add a Bonferroni correction by multiplying the $p$-values by 4 and 3, respectively, leading to $p<10^{-3}$ for all previous comparisons except DTI-CDF.
+We further run another one-tailed T-test concerning $MicroAUC_p$ including Bonferroni correction.
+We find that \name{} advances significantly over MolTrans ($p<10^{-4}$), DeepDTI ($p<10^{-4}$), DeepDTA ($p<10^{-4}$), and DeepConv-DTI ($p<10^{-4}$) regarding BioSnap dataset, and over DTINet ($p<10^{-4}$), DTIGEMS+ ($p<10^{-4}$) and DTI-CDF ($p<10^{-4}$) regarding Yamanishi dataset.
 

### 2021-04-26 paper/main.tex

--- 2021-04-25 paper/main.tex
@@
 PhenomeNET ontology using the \textit{Phenomebrowser.net}, which
-provides a SPARQL query endpoint for the mentioned resources.
+provides a SPARQL query endpoint for the mentioned resources. The overall structure is depicted in Supplementary Figure 1.
 
@@
 used as input of a stack (with height 3) of residual graph
-convolutional blocks.
+convolutional blocks. The full eventual model of \name{} is shown in Supplementary Figure 4.
 
@@
     features \textit{OntoPred}. \name{} combines both types of
-    features.}
+    features. For a comparison of MicroAUC$_p$ and MicroAUC$'_p$ (See Section \ref{sec:evaluation} for more information) see Supplementary Table 2. }
 \end{table}
@@
 and drug--target affinity prediction methods that were trained and
-evaluation on other datasets.  Following the results of MolTrans
+evaluated on other datasets. Following the results of MolTrans
 \cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
@@
 with $\alpha=0.01$ over all considered methods. Considering Macro
-AUCs, \name{} improves significantly ($p<0.0001$, one-sided t-test)
-over MolTrans, DeepDTI, DeepDTA, and DeepConv-DTI on the BioSnap
-dataset, and significantly ($p<0.0001$, one-sided t-test) over DTINet
-and DTIGEMS+ on the Yamanishi dataset' DTI-CDF performs
+AUCs, \name{} improves significantly ($p<10^{-4}$, one-sided t-test)
+over MolTrans ($t=36.6$), DeepDTI ($t=40.5$), DeepDTA ($t=36.6$), and DeepConv-DTI ($t=40.5$) on the BioSnap
+dataset, and significantly ($p<10^{-4}$, one-sided t-test) over DTINet ($t=20.8$)
+and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset' DTI-CDF performs
 significantly better than \name{} ($p=0.02$, one-sided t-test). % As
@@
 We further run another one-tailed t-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction.
-We find that \name{} advances significantly over MolTrans ($p<0.00001$), DeepDTI ($p<0.00001$), DeepDTA ($p<0.00001$), and DeepConv-DTI ($p<0.00001$) regarding BioSnap dataset, and over DTINet ($p<0.00001$), DTIGEMS+ ($p<0.00001$) and DTI-CDF ($p<0.00001$) regarding Yamanishi dataset.
+We find that \name{} advances significantly over MolTrans ($p<10^{-4}, t=21.2$), DeepDTI ($p<10^{-4}, t=31.0$), DeepDTA ($p<10^{-4}, t=33.4$), and DeepConv-DTI ($p<10^{-4}, t=23.7$) regarding BioSnap dataset, and over DTINet ($p<10^{-4}, t=29.6$), DTIGEMS+ ($p<10^{-4}, t=28.0$) and DTI-CDF ($p<10^{-4}, t=9.5$) regarding Yamanishi dataset.
 

### 2021-04-27 paper/main.tex
.

@@
 PhenomeNET ontology using the \textit{Phenomebrowser.net}, which
-provides a SPARQL query endpoint for the mentioned resources. The overall structure is depicted in Supplementary Figure 1.
+provides a SPARQL query endpoint for the mentioned resources. The
+overall structure is shown in Supplementary Figure 1.
 
@@
 aims to learn a similarity between two embeddings of variable but same
-dimension. As the original feature space may have varying dimensionality, we first process them using a fully connected
-neural network layer which takes as input an embedding and outputs a
+dimension. As the original feature space may have varying
+dimensionality, we first process them using a fully connected neural
+network layer which takes as input an embedding and outputs a
 representation of a particular size, i.e., we use this layer as a
@@
 % features with respect to other ontologies and hypotheses
-We follow the results of \textit{DL2vec} \cite{DL2vec2020} and use $\sigma := \mathrm{LeakyReLU}$ as
-activation function  which leads to
+We follow the results of \textit{DL2vec} \cite{DL2vec2020} and use
+$\sigma := \mathrm{LeakyReLU}$ as activation function which leads to
 improved performance compared to other activation functions.
@@
   DeeperGCN2020}.  The residual blocks are reusable and can be stacked
-multiple times. % The structure of the GNN architecture we use is shown
+multiple times.
+% The structure of the GNN architecture we use is shown
 % in Figure~\ref{fig:ResGraphConvBlocks}.
@@
 features, before plugging them into the graph model for the similarity
-computation. An overview of the model architecture, combining both
-feature types, is shown in Supplementary Figure
-4. Here
-the original representations are transformed by a dense layer and then
-used as input of a stack (with height 3) of residual graph
-convolutional blocks. The full eventual model of \name{} is shown in Supplementary Figure 4.
+computation. % An overview of the model architecture, combining both
+% feature types, is shown in Supplementary Figure
+% 4. 
+Here the original representations are transformed by a dense layer and
+then used as input of a stack (with height 3) of residual graph
+convolutional blocks. The full model of \name{} is shown in
+Supplementary Figure 4.
 

### 2021-04-27 paper/main.tex
.

--- 2021-04-26 paper/main.tex
@@
 connections and a normalization scheme \cite{DeepGCN2019,
-  DeeperGCN2020}.  The residual blocks are reusable and can be stacked
-multiple times. % The structure of the GNN architecture we use is shown
+  DeeperGCN2020} as shown in Supplemtary Figure 3.  The residual
+blocks are reusable and can be stacked multiple times.
+% The structure of the GNN architecture we use is shown
 % in Figure~\ref{fig:ResGraphConvBlocks}.
@@
 computation.  An overview of the model architecture, combining both
-feature types, is shown in Supplementary Figure
-4. Here
-the original representations are transformed by a dense layer and then
-used as input of a stack (with height 3) of residual graph
-convolutional blocks. The full eventual model of \name{} is shown in Supplementary Figure 4.
+feature types, is shown in Supplementary Figure 4.  Here the original
+representations are transformed by a dense layer and then used as
+input of a stack (with height 3) of residual graph convolutional
+blocks. %The full model of \name{} is shown in Supplementary Figure 4.
 

### 2021-04-27 paper/main.tex
.

@@
 connections and a normalization scheme \cite{DeepGCN2019,
-  DeeperGCN2020} as shown in Supplemtary Figure 3.  The residual
+  DeeperGCN2020} as shown in Supplementary Figure 3.  The residual
 blocks are reusable and can be stacked multiple times.
@@
 \end{equation*}
-Note that drugs and proteins can be interchanged in this formulation. We refer
+Drugs and proteins can be interchanged in this formulation. We refer
 to the different measures as protein-centric microAUC ($MicroAUC_p$)
-and a drug-centric microAUC ($MicroAUC_d$).
-
-Further, we choose the AUROC as primary metric to compare different
-methods over other measures such as the area under precision recall
-curve (AUPRC); AUPRC is sensitive to data imbalances \cite{Jeni2013}
-and therefore more challenging to apply to comparing different DTI
-prediction methods.
+and a drug-centric microAUC ($MicroAUC_d$).  We further compare
+MicroAUC$_p$ with imputation and MicroAUC$'_p$ without imputing
+undefined $MicroAUC$ values (but omitting them) in Supplementary Table
+1.
+
+Further, we choose the AUROC over other measures such as the area
+under precision recall curve (AUPRC) as primary metric to compare
+different methods; AUPRC is sensitive to data imbalances
+\cite{Jeni2013} and therefore more challenging to apply to comparing
+different DTI prediction methods.
 
@@
 interactions as edges, mapping the protein features to each target as
-node features. \name{} then propagates information among the PPI network
-utilizing graph convolutional steps, calculates the similarity of drug
-and protein representations, and predicts whether there is an
+node features. \name{} then propagates information among the PPI
+network utilizing graph convolutional steps, calculates the similarity
+of drug and protein representations, and predicts whether there is an
 interaction. The full workflow scheme is depicted in
@@
     features \textit{OntoPred}. \name{} combines both types of
-    features. For a comparison of MicroAUC$_p$ and MicroAUC$'_p$ (See Section \ref{sec:evaluation} for more information) see Supplementary Table 2. }
+    features.}
 \end{table}
@@
 different graph neural network architectures and configurations
-(Supplementary Table 1). While the {\em
+(Supplementary Table 2). While the {\em
   GCNConv} and {\em GENConv} architecture already shows some minor
@@
 reached very similar results to the original \textit{OntoPred} model
-and showed no improvement (Supplementary Table 1).
+and showed no improvement (Supplementary Table 2).
 

### 2021-04-27 paper/supplement/supplement.tex
.

@@
 
+\clearpage
+
 \begin{figure}[ht]%figure1
@@
 
+\clearpage
 \begin{figure}[ht]%figure1
@@
 
-\begin{table}[ht]
-	\centering
-
-	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{1.2cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
-		\hline
-		Approach&\multicolumn{6}{c|}{Graph convolution method}\\
-		&without&without/ empty RGC&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
-		\hline
-		MolPred & $0.69$ & $0.67$& $0.67$& $0.68$& $0.65$& $0.69$\\
-		\hline
-		OntoPred &$0.88$ &$0.87$ &$0.88$&$0.90$&$0.86$&$0.92$\\
-		\hline
-		\name& $0.89$&$0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
-		\hline
-	\end{tabular}
-	\caption{\label{tab:Results}Results for various neural graph convolutional methods enhancing MolPred, OntoPred and \name{} over the STITCH dataset.
-We evaluated all methods over different stacking heights without and with residual graph convolution blocks (RGC, ResGraphConv) as depicted in Supplementary figure \ref{fig:ResGraphConvBlocks}.
-``Empty RGC`` denotes a RGC block with a plain neural layer substituting the GraphConv layer.}
-\end{table}
+\clearpage
 
@@
 	\end{tabular}
-	\caption{Extended results table for MolPred, OntoPred and \name{} over STITCH with added $MicroAUC'_p$ for comparison of both metrics.}
+	\caption{Extended results table for MolPred, OntoPred and
+          \name{} over STITCH with added $MicroAUC'_p$ for comparison
+          of both metrics.}
+\end{table}
+
+\clearpage
+\begin{table}[ht]
+	\centering
+
+	\begin{tabular}{|p{2.0cm}|p{1.1cm}|p{1.2cm}|p{0.8cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|}
+		\hline
+		Approach&\multicolumn{6}{c|}{Graph convolution method}\\
+		&without&without/ empty RGC&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
+		\hline
+		MolPred & $0.69$ & $0.67$& $0.67$& $0.68$& $0.65$& $0.69$\\
+		\hline
+		OntoPred &$0.88$ &$0.87$ &$0.88$&$0.90$&$0.86$&$0.92$\\
+		\hline
+		\name& $0.89$&$0.89$& $0.89$& $0.91$& $0.86$& $\mathbf{0.93}$\\
+		\hline
+	\end{tabular}
+	\caption{\label{tab:Results}Results for various neural graph
+          convolutional methods enhancing MolPred, OntoPred and
+          \name{} over the STITCH dataset. We evaluated all methods
+          over different stacking heights without and with residual
+          graph convolution blocks (RGC, ResGraphConv, see
+          Supplementary figure \ref{fig:ResGraphConvBlocks}). ``Empty
+          RGC'' denotes a RGC block with a plain neural layer
+          substituting the GraphConv layer.}
 \end{table}

### 2021-04-27 paper/main.tex
.

@@
 DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
-MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared it to our
-``na\"ive'' predictor as well as \name{} (see Table
+MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared
+it to our ``na\"ive'' predictor as well as \name{} (see Table
 \ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
@@
 AUCs, \name{} improves significantly ($p<10^{-4}$, one-sided t-test)
-over MolTrans ($t=36.6$), DeepDTI ($t=40.5$), DeepDTA ($t=36.6$), and DeepConv-DTI ($t=40.5$) on the BioSnap
-dataset, and significantly ($p<10^{-4}$, one-sided t-test) over DTINet ($t=20.8$)
-and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset' DTI-CDF performs
-significantly better than \name{} ($p=0.02$, one-sided t-test). % As
+over MolTrans ($t=36.6$), DeepDTI ($t=40.5$), DeepDTA ($t=36.6$), and
+DeepConv-DTI ($t=40.5$) on the BioSnap dataset, and significantly
+($p<10^{-4}$, one-sided t-test) over DTINet ($t=20.8$) and DTIGEMS+
+($t=18.0$) on the Yamanishi dataset; DTI-CDF performs significantly
+better than \name{} ($p=0.02$, one-sided t-test). % As
                                 % we compare \name{} to 4 different
@@
                                 % comparisons except DTI-CDF.
-We further run another one-tailed t-test concerning $MicroAUC_p$ with $\alpha=0.01$ including Bonferroni correction.
-We find that \name{} advances significantly over MolTrans ($p<10^{-4}, t=21.2$), DeepDTI ($p<10^{-4}, t=31.0$), DeepDTA ($p<10^{-4}, t=33.4$), and DeepConv-DTI ($p<10^{-4}, t=23.7$) regarding BioSnap dataset, and over DTINet ($p<10^{-4}, t=29.6$), DTIGEMS+ ($p<10^{-4}, t=28.0$) and DTI-CDF ($p<10^{-4}, t=9.5$) regarding Yamanishi dataset.
-
-In all our experiments, \name{} has the highest $MicroAUC_p$,
-demonstrating that \name{} can improve significantly over other methods in the task
-of identifying drugs that target a specific protein. Several methods
-achieve a higher Macro AUC than \name{}, in particular when evaluated
-using a drug--target pair splitting scheme; our results with the
-``na\"ive'' prediction method show that it may be possible that models
-trained on a drug--target split utilize certain biases in the dataset
-without necessarily producing novel biological insights.
+We further perform another one-tailed t-test to compare $MicroAUC_p$
+performance. We find that \name{} improved significantly ($p <
+10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
+($t=33.4$), and DeepConv-DTI ($t=23.7$) on the BioSnap dataset, and
+over DTINet ($t=29.6$), DTIGEMS+ ($t=28.0$), and DTI-CDF ($t=9.5$) on
+the Yamanishi dataset.
+
+In all our experiments, \name{} improves over all other methods with
+respect to $MicroAUC_p$, demonstrating that \name{} can identify drugs
+that target a specific protein more accurately than other
+methods. Several methods achieve a higher MacroAUC than \name{}, in
+particular when evaluated using a drug--target pair splitting scheme;
+our results with the ``na\"ive'' prediction method show that it may be
+possible that models trained on a drug--target split utilize certain
+biases in the dataset without necessarily producing novel biological
+insights.
 

### 2021-04-27 paper/main.tex
.

@@
 $MicroAUC_p$ score for all considered methods, leading to similar
-results as observed on the Yaminishi dataset. As \name{} considers
-less drugs than alternative approaches, we execute a one-tailed T-test
-with $\alpha=0.01$ over all considered methods. Considering Macro
-AUCs, \name{} improves significantly ($p<10^{-4}$, one-sided t-test)
-over MolTrans ($t=36.6$), DeepDTI ($t=40.5$), DeepDTA ($t=36.6$), and
-DeepConv-DTI ($t=40.5$) on the BioSnap dataset, and significantly
-($p<10^{-4}$, one-sided t-test) over DTINet ($t=20.8$) and DTIGEMS+
-($t=18.0$) on the Yamanishi dataset; DTI-CDF performs significantly
-better than \name{} ($p=0.02$, one-sided t-test). % As
-                                % we compare \name{} to 4 different
-                                % methods over BioSnap and 3 over
-                                % Yamanishi dataset, we add a
-                                % Bonferroni correction by multiplying
-                                % the $p$-values by 4 and 3,
-                                % respectively, leading to $p<0.0002$
-                                % ($\alpha=0.01$) for all previous
-                                % comparisons except DTI-CDF.
-We further perform another one-tailed t-test to compare $MicroAUC_p$
-performance. We find that \name{} improved significantly ($p <
-10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
+results as observed on the Yaminishi dataset. We test whether \name{}
+performs better than the methods we compare against or whether the
+observed values for $MicroAUC_p$ and $MacroAUC$ fall within expected
+variant; we use a one-tailed T-test with Bonferroni correction for
+this test.
+
+Considering Macro AUCs, \name{} improves significantly ($p<10^{-4}$,
+one-sided t-test) over MolTrans ($t=36.6$), DeepDTI ($t=40.5$),
+DeepDTA ($t=36.6$), and DeepConv-DTI ($t=40.5$) on the BioSnap
+dataset, and significantly ($p<10^{-4}$, one-sided t-test) over DTINet
+($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
+performs significantly better than \name{} ($p=0.02$, one-sided
+t-test).  We further perform another one-tailed t-test to compare
+$MicroAUC_p$ performance. We find that \name{} improved significantly
+($p < 10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
 ($t=33.4$), and DeepConv-DTI ($t=23.7$) on the BioSnap dataset, and

### 2021-04-27 paper/main.tex
.

@@
 ($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
-performs significantly better than \name{} ($p=0.02$, one-sided
-t-test).  We further perform another one-tailed t-test to compare
-$MicroAUC_p$ performance. We find that \name{} improved significantly
+performs better than \name{} ($p=0.02$, one-sided t-test).  We further
+perform another one-tailed t-test to compare $MicroAUC_p$
+performance. We find that \name{} improved significantly
 ($p < 10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA

### 2021-04-27 paper/main.tex
Update t-test for DTI-CDF

@@
 \author[Hinnerichs \& Hoehndorf]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert
-  Hoehndorf\,$^{\text{\sfb 2}}$}
-\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
-  $^{\text{\sf 2}}$Computational Bioscience Research Center, Computer,
+  Hoehndorf\,$^{\text{\sfb 1}}$}
+\address{$^{\text{\sf 1}}$Computational Bioscience Research Center, Computer,
   Electrical and Mathematical Sciences \& Engineering Division, King
@@
 ($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
-performs better than \name{} ($p=0.02$, one-sided t-test).  We further
+performs better than \name{} ($p=0.1, t=1.3$, one-sided t-test) \todo{significance}.  We further
 perform another one-tailed t-test to compare $MicroAUC_p$

### 2021-04-27 paper/cover_letter/cover_letter.tex
.

@@
 
-Please find our manuscript entitled "DTI-Voodoo: machine learning over interaction networks and
-ontology-based background knowledge predicts drug--target
-interactions" for your consideration.
-We therefore developed a novel method for drug--target interaction prediction (DTI) utilizing both ontology-encoded phenotypical and molecular features for both drugs and proteins.
-We further combine these methods with interaction networks, aiming for a more protein-centric prediction and evaluation.
+Please find our manuscript entitled "DTI-Voodoo: machine learning over
+interaction networks and ontology-based background knowledge predicts
+drug--target interactions" for your consideration.
 
-We designed DTI-Voodoo under usage of graph convolutional neural network, exploiting the ``guilt-by-association`` principle and incorporate both top-down and bottom-up features into protein-protein interaction networks.
-Furthermore, we show that common DTI datasets share immanent biases, majorly influencing predictive DTI model performance evaluations.
-Thus, we propose a modified evaluation scheme, showing that DTI-Voodoo advances significantly over other state of the art DTI prediction approaches over several datasets.
+DTI-Voodoo is a novel method for drug--target interaction (DTI)
+prediction utilizing both ontology-based information about phenotypic
+effects of drugs (side effects), and molecular features that encode
+for interactions between drugs and proteins. We compare DTI-Voodoo to
+seven state of the art DTI prediction methods on two datasets, and
+significantly improve over all of them in our evaluation.
 
-The importance of DTI prediction for other challenges such as drug repurposing and repositioning makes DTI-Voodoo highly influential for a variety of researchers.
-Eventually, DTI-Voodoo's versatile methodology may be used over alternative protein-based representations to improve performances within other predictive tasks in the realm of bioinformatics and computational biology.
+Moreover, we introduce several novel aspects to DTI prediction: First,
+we identify, analyze, and correct for a number of inherent biases in
+drug--target interaction prediction datasets; for this purpose, we
+propose a ``naive'' classifier which does not make predictions based
+on any biological information but achieves performance close to state
+of the art methods.  Second, DTI-Voodoo combines indirect information
+such as side effects and molecular properties of both drugs and
+proteins in a novel way using an interaction network. Using this
+network, we can demonstrate that side effect information ``localizes''
+on the interaction network, i.e., that side effects are enriched on
+modules consisting of interacting proteins, whereas molecular features
+do not benefit from information the graph.
 
-Therefore, we believe that
-DTI-Voodoo will be well suited for the broad readership of
-Bioinformatics.
+We believe that our results, as well as the novel methods we
+introduced in developing DTI-Voodoo, makes our manuscript highly
+suitable for the readers of Bioinformatics.
 

### 2021-04-27 paper/main.tex
.

@@
 	\centering
-	\includegraphics[width=0.9\columnwidth]{figures/model_workflow.png}
+	\includegraphics[width=0.8\columnwidth]{figures/model_workflow.png}
 	\caption{Full DTI prediction model based on the pretrained
@@
 ($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
-performs better than \name{} ($p=0.1, t=1.3$, one-sided t-test) \todo{significance}.  We further
+performs better than \name{} ($p=0.1, t=1.3$, one-sided t-test).  We further
 perform another one-tailed t-test to compare $MicroAUC_p$
-performance. We find that \name{} improved significantly
+performance. We find that \name{} improves significantly
 ($p < 10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
@@
 that target a specific protein more accurately than other
-methods. Several methods achieve a higher MacroAUC than \name{}, in
+methods. Some methods achieve a higher MacroAUC than \name{}, in
 particular when evaluated using a drug--target pair splitting scheme;
@@
 
-\todo[inline]{Why is having only having 20\% of drugs reasonable -> significance test in results}
+%\todo[inline]{Why is having only having 20\% of drugs reasonable -> significance test in results}
 One major component of our experiments was to determine how the

### 2021-04-27 paper/main.tex
.
Update faulty citations

--- 2021-04-27 paper/main.tex
@@
 connections and a normalization scheme \cite{DeepGCN2019,
-  DeeperGCN2020} as shown in Supplementary Figure 3.  The residual
+  GENConv2020} as shown in Supplementary Figure 3.  The residual
 blocks are reusable and can be stacked multiple times.
@@
 convolutional methods may be the result of numerical instability and
-fading signals \cite{DeeperGCN2020}.
+fading signals \cite{GENConv2020}.
 

### 2021-04-27 paper/main.tex

--- 2021-04-27 paper/main.tex
@@
   effects of drugs with protein--protein interaction networks, and
-  uses a graph convolutional neural network to predict DTIs.  We demonstrate that
-  drug effect features can exploit information in the interaction
-  network whereas molecular features do not.  \name{} is designed to
-  predict candidate drugs for a given protein; we use this formulation
-  to show that common DTI datasets contain intrinsic biases with major
-  affects on performance evaluation and comparison of DTI prediction
-  methods. Using a modified evaluation scheme, we demonstrate that
-  \name{} improves substantially over state of the art DTI prediction
-  methods. \\
+  uses a graph convolutional neural network to predict DTIs.  We
+  demonstrate that drug effect features can exploit information in the
+  interaction network whereas molecular features do not.  \name{} is
+  designed to predict candidate drugs for a given protein; we use this
+  formulation to show that common DTI datasets contain intrinsic
+  biases with major affects on performance evaluation and comparison
+  of DTI prediction methods. Using a modified evaluation scheme, we
+  demonstrate that \name{} improves significantly over state of the
+  art DTI prediction methods. \\
   \textbf{Availability:} \name{} source code and data necessary to

### 2021-04-29 paper/main.tex
.

@@
   formulation to show that common DTI datasets contain intrinsic
-  biases with major affects on performance evaluation and comparison
+  biases with major effects on performance evaluation and comparison
   of DTI prediction methods. Using a modified evaluation scheme, we

### 2021-06-04 paper/main.tex
.

@@
   potential drug targets, or they can be direct, bottom-up and use
-  molecular information to directly predict binding potentials.  Both
+  molecular information to directly predict binding affinities.  Both
   approaches can be combined with information about interaction

### 2021-06-29 paper/main.tex
Draft for cover letter response in main paper, added heatmap and interaction table

@@
 % made. 
-Identifying drug--target interactions (DTIs) is important in drug
-repurposing and repositioning, and in the exploration of novel drug
-candidates and their interaction partners. Computational methods are
+Identifying drug--target interactions (DTIs) is one of the initial steps in drug discovery and screening for disease treatment and thus helps to mitigate the the costs and risks of drug development.
+Further, it forms a generalization of both drug repurposing and repositioning, finding novel functions for already approved and extensively analyzed drugs with well explored properties.
+Inferring and deriving from those existing properties and interactions can help to analyze and identify novel drugs for known targets, leading to desirable therapeutic effects.
+Moreover, while \textit{in vitro} DTI prediction requires the availability of all substances, computational, \textit{in silico} DTI predictors interpret and anticipate millions of interactions within short time.
+
+Computational methods are
 widely applied to predict DTIs and many computational methods have
@@
 prediction bear some advantages and limitations. Generally, bottom-up
-methods face the challenge to predict whether a structure binds to a
+methods face the challenge to predict whether a chemical structure binds to a
 protein given their molecular properties; whether two entities
@@
 (where binding sites and molecular forces need to be determined for
-accurate prediction) but also properties namely where a protein is
-expressed or in which celltypes and anatomical structures. Top-down
+accurate prediction) but also properties namely in which celltypes and anatomical structures a protein is expressed. Top-down
 methods use information for DTI prediction such as side-effect
@@
 knowledge gained from molecular properties. While methods that rely on
-molecular information are directly predicting whether two molecular
+molecular information are directly predicting whether two molecules
 could interact, top-down methods base on more indirect means and 
@@
 
+We further investigate the influence of the more specific interaction types \textbf{inhibition} and \textit{activation} within STITCH database on the performance of \name{}.
+The results are summarized in Supplementary table 3, where we find no difference in comparison to the complete dataset.
+
 % despite predicting per protein is
@@
 consequently, we expect these features to benefit more from including
-information about protein--protein interactions. Our results (Table
-\ref{tab:Results}) confirm this hypothesis and demonstrate that
+information about protein--protein interactions. Moreover, as the protein--protein 
+interaction relation is not transitive (protein $P_1$ interacting with $P_2$, 
+and $P_2$ relating with $P_3$ does not imply $P_1$ interacting
+or influencing $P_3$), we can only infer from the direct neighborhood of each protein within the PPI graph. 
+Our results (Table
+\ref{tab:Results}) confirm the first hypothesis and demonstrate that
 molecular features do not benefit from including the interaction
@@
 propagation over the interaction network.
+While the exploitation of ontology based features over PPI graphs was already examined in \citet{Trebacz2020} for gene expression analysis, our work is the first to investigate and apply this technique to DTI prediction.
+Further, \name{} is the first work to combine both ``bottom-up'' and ``top-down'' features.
 
@@
 partners is not desirable when applying the model to novel
-entities. We have demonstrated that even a ``na\"ive'' classifier that
+entities. We have demonstrated that even the newly proposed ``na\"ive'' classifier that
 makes predictions only by exploiting the imbalanced number of
@@
 the second and third using a Micro AUC$_p$ and Micro AUC$_d$.
+
+\subsection{Pharmacological novelty}
+
+As the target-based predictive power of \name{} improves significantly over other methods, we utilized our model to predict novel drug classes for protein families.
+We therefore collected the 2nd level ATC (Anatomical Therapeutic Chemical) groups for each drug and all InterPro \cite{Interpro2020} top-level families for each protein.
+Utilizing the STITCH interactions, we followed a protein split within each InterPro family by predicting over all available drugs.
+We eventually normalized the number of novel interactions per class by the amount of proteins within the respective family.
+A heatmap showing the results of this analysis can be found in Supplementary figure 5.
+\dots \todo{Roberts class analysis}.
 

### 2021-06-29 paper/supplement/supplement.tex
.
Draft for cover letter response in main paper, added heatmap and interaction table

@@
 		\hline
-		\name{} & $0.89$& $0.91$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$&$\mathbf{0.94}$\\
+		\name{} & $0.89$& $0.91$ & $0.90$&$0.93$&$0.94$&$0.94$\\
 		\hline
@@
 	\caption{Extended results table for MolPred, OntoPred and
-          \name{} over STITCH with added $MicroAUC'_p$ for comparison
-          of both metrics.}
+          \name{} over STITCH with added $MicroAUC'_p$ (without
+          imputation) for comparison of both metrics.}
 \end{table}
@@
 		\hline
-		Approach&\multicolumn{6}{c|}{Graph convolution method}\\
+		Approach&\multicolumn{6}{c|}{Graph convolution method (MacroAUC)}\\
 		&without&without/ empty RGC&GCN-Conv&GEN-Conv&RGC + GCNConv&RGC + GENConv\\
@@
 
+\clearpage
+\begin{figure}[ht]%figure1
+	\centerline{\includegraphics[width=1.1\textwidth]{../figures/ATC_Interpro_heatmap.png}}
+	\caption{Heatmap showing the results of the ATC drug class and InterPro family prediction analysis}
+	\label{fig:ATCInterproHeatmap}
+\end{figure}
+
+\clearpage
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{|p{2.0cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
+		\hline
+		&\multicolumn{4}{c|}{Interaction type}\\
+		&\multicolumn{2}{c|}{Activation}&\multicolumn{2}{c|}{Inhibition}\\
+		\#Drugs & \multicolumn{2}{l|}{723} & \multicolumn{2}{l|}{954}\\
+		\#Proteins&\multicolumn{2}{l|}{4562}&\multicolumn{2}{l|}{6222}\\
+		\#Links&\multicolumn{2}{l|}{14174}&\multicolumn{2}{l|}{16364}\\
+		\hline
+		Metric&Macro AUC&Micro AUC$_p$&Macro AUC&Micro AUC$_p$\\
+		\name&$0.95$&$0.95$&$0.93$&$0.94$\\
+		\hline
+
+	\end{tabular}
+	\caption{Extended results table for MolPred, OntoPred and
+		\name{} over STITCH with added $MicroAUC'_p$ (without
+		imputation) for comparison of both metrics.}
+\end{table}
+
 \end{document}

### 2021-06-29 data/STITCH_data/download.sh

@@ -4,6 +4,11 @@ URL=http://stitch.embl.de/download/protein_chemical.links.transfer.v5.0/9606.pro
 curl -O "$URL" && gunzip -f "${URL##*/}"
 wait
 
+URL=http://stitch.embl.de/download/actions.v5.0/9606.actions.v5.0.tsv.gz
+curl -O "$URL" && gunzip -f "${URL##*/}"
+wait
+
+
 URL=http://stitch.embl.de/download/chemical.aliases.v5.0.tsv.gz
 curl -O "$URL" && gunzip -f "${URL##*/}"
 wait

### 2021-06-29 src/DDI_utils.py

@@ -6,9 +6,6 @@ from tqdm import tqdm
 
 import networkx as nx
 
-import similarity_measurement
-
-
 
 def get_db_PubChem_id_mapping_dict():
     STITCH_mapping = get_STITCH_db_Pubchem_mapping_dict()
@@ -385,59 +382,6 @@ def parse_SIDER_indications():
 
 
 
-
-
-
-
-def evaluate_dicts_and_graph():
-
-    '''
-    map_1 = get_db_PubChem_id_mapping_dict()
-    map_2 = get_db_PubChem_id_mapping_dict_mahmud()
-    map_3 = get_pddi_db_pubchem_mapping()
-
-    key_set1 = set(map_1.keys())
-    key_set2 = set(map_2.items())
-    key_set3 = set(map_3.items())
-
-    print(len(key_set1))
-    print(len(key_set2))
-    print(len(key_set3))
-
-    print(len(key_set2 & key_set3))
-
-    super_dict = dict(key_set2 | key_set2)
-
-    print(len(list(super_dict.keys())))
-    '''
-
-    drugbank_graph = get_DDI_drugbank_graph()
-    boyce_graph = get_DDI_Boyce_graph()
-
-    # print(len(drugbank_graph.nodes()))
-    print('boyce', len(boyce_graph.nodes()), len(boyce_graph.edges()))
-    print('db', len(drugbank_graph.nodes()), len(drugbank_graph.edges()))
-
-    intersect = set(drugbank_graph.nodes()) & set(boyce_graph.nodes())
-    print('intersect', len(intersect))
-
-    merged_graph = get_merged_DDI_graph()
-    print('merged', len(merged_graph.nodes()), len(merged_graph.edges()))
-
-    SIDER_drugs = set(similarity_measurement.get_SIDER_drug_list())
-
-    print(len(SIDER_drugs | intersect))
-
-
-    # SIDER_only_graph = get_SIDER_only_graph()
-    # drug_set = set(get_SIDER_drug_list())
-
-    # print(len(drug_set & key_set1))
-
-
-
-
-
 if __name__ == '__main__':
     # write_SITCH_db_Pubchem_mapping_dict()
     # get_STITCH_db_Pubchem_mapping_dict()
@@ -445,7 +389,6 @@ if __name__ == '__main__':
     # get_DDI_Boyce_graph()
 
     # write_DDI_drugbank_graph()
[…]

### 2021-06-29 src/DTI_data_preparation.py

@@ -20,11 +20,10 @@ import PhenomeNET_DL2vec_utils
 
 
 def write_human_DTI_graph(min_score=700,
-                          mode=''):
-    filename = "../data/STITCH_data/9606.protein_chemical.links.transfer.v5.0.tsv"
+                          mode='',
+                          interaction_type=''):
     dti_graph = nx.Graph()
 
-    # drug_set = similarity_measurement.get_SIDER_Boyce_Drubank_drug_intersection()
     # drug_set = get_drug_list()
 
     print("Loading chemical stereo to mono mapping...")
@@ -32,32 +31,53 @@ def write_human_DTI_graph(min_score=700,
     print("Done.\n")
 
     print("Parsing human drug-protein-links data ...")
-    with open(file=filename, mode='r') as f:
-        f.readline()
-        for line in tqdm(f, total=15473940):
-            split_line = line.split('\t')
-            drug = split_line[0].strip()
-            if 's' in drug:
-                drug = stereo_mono_mapping.get(drug, None)
-                if not drug:
-                    continue
-            # drug = split_line[0].replace('s','m')
-            target = split_line[1]
+    if not interaction_type:
+        filename = "../data/STITCH_data/9606.protein_chemical.links.transfer.v5.0.tsv"
+        with open(file=filename, mode='r') as f:
+            f.readline()
+            for line in tqdm(f, total=15473940):
+                split_line = line.split('\t')
+                drug = split_line[0].strip()
+                if 's' in drug:
+                    drug = stereo_mono_mapping.get(drug, None)
+                    if not drug:
+                        continue
+                # drug = split_line[0].replace('s','m')
+                target = split_line[1]
 
-            score = None
-            if mode=='experimental':
-                score = int((1- (1-int(split_line[2])/1000) * (1-int(split_line[3])/1000))*1000)
-            elif mode=='database':
-                score = int((1- (1-int(split_line[2])/1000) * (1-int(split_line[3])/1000) * (1-int(split_line[6])/1000) * (1-int(split_line[7])/1000))*1000)
-            else:
-                score = int(split_line[-1])
-            # if not drug in drug_set:
-                # continue
+                score = None
+                if mode=='experimental':
[…]

### 2021-06-29 src/HPO_predictor.py

@@ -1,3 +1,5 @@
+import subprocess
+
 import numpy as np
 import math
 from sklearn.model_selection import KFold
@@ -32,7 +34,7 @@ class HPODTIDataBuilder:
                 indication_drug_list = set(PhenomeNET_DL2vec_utils.get_UMLS_drug_list())
                 self.drug_list = np.array(list(set(self.drug_list) & indication_drug_list))
         else:
-            self.drug_list = np.array(DTI_data_preparation.get_drug_list(config.mode))
+            self.drug_list = np.array(DTI_data_preparation.get_drug_list(config.mode, config.interaction_type))
         print(len(self.drug_list), "drugs present")
 
         # get protein lists for each ontology
@@ -41,10 +43,17 @@ class HPODTIDataBuilder:
         MP_protein_list = PhenomeNET_DL2vec_utils.get_PhenomeNET_protein_list(mode='MP')
 
         dti_graph = DTI_data_preparation.get_human_DTI_graph(mode=config.mode)
-        PPI_graph = PPI_utils.get_PPI_graph(min_score=700)
+        self.PPI_graph = PPI_utils.get_PPI_graph(min_score=700)
+        self.protein_list = np.array(list(set(self.PPI_graph.nodes()) & set(dti_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
+
+        if config.ipro_class:
+            print('Fetching matching interpro proteins...')
+            print('InterPro class:', config.ipro_class)
+            ipro_prots = DTI_data_preparation.get_prots_per_IPRO_class(config.ipro_class)
+            self.protein_list = np.array(list(set(self.protein_list) & set(ipro_prots)))
 
         if config.yamanishi_test or config.biosnap_test:
-            self.protein_list = np.array(list(set(PPI_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
+            self.protein_list = np.array(list(set(self.PPI_graph.nodes()) & (set(uberon_protein_list) | set(GO_protein_list) | set(MP_protein_list))))
             if config.yamanishi_test:
                 print("Loading Yamanishi data ...")
                 self.drug_list, self.protein_list, self.y_dti_data = DTI_data_preparation.get_yamanishi_data(self.drug_list, self.protein_list)
@@ -58,11 +67,6 @@ class HPODTIDataBuilder:
                 raise ValueError
[…]

### 2021-06-29 src/results_evaluation.py

@@ -2,6 +2,8 @@ import pickle
 import numpy as np
 import random
 
+import matplotlib.pyplot as plt
+
 from tqdm import tqdm
 
 import DTI_data_preparation
@@ -10,7 +12,6 @@ import dti_utils
 
 import sklearn.metrics as metrics
 from sklearn.model_selection import KFold
-
 
 
 def write_predicted_DTIs(fold=3):
@@ -378,14 +379,77 @@ def naive_predictor_pair_split(dti_matrix):
         results = np.array(results)
         print(results.mean())
 
+def analyze_ATC_Ipro_rel():
+    # ipro_class_list = ['IPR013087', 'IPR000276', 'IPR000742', 'IPR003961', 'IPR007110', 'IPR000719', 'IPR002048', 'IPR001452', 'IPR000504', 'IPR001356']
+    ipro_class_list = ['IPR013087', 'IPR017452', 'IPR000276', 'IPR001909', 'IPR001841', 'IPR003961', 'IPR001806',
+                  'IPR001314', 'IPR001965', 'IPR001304', 'IPR002035', 'IPR000315', 'IPR001781', 'IPR000571']
+
+    ATC_mapping = DTI_data_preparation.get_drug_ATC_classification_mapping()
+    drug_list = None
+    with open(file='../results/ATC_Interpro_results/drug_list', mode='rb') as f:
+        drug_list = pickle.load(f)
+    drug_list = np.array([ATC_mapping.get(drug, None) for drug in drug_list])
+
+    present_ATC_classes = list(set(drug_list))
+    present_ATC_classes.remove(None)
+    print('Number of present level 2 ATC classes:', len(present_ATC_classes))
+
+    results_list = np.zeros((len(ipro_class_list), len(present_ATC_classes)))
+    for ipro_index, ipro_class in enumerate(ipro_class_list):
+        filename = f'../results/ATC_Interpro_results/{ipro_class}_'
+
+        with open(file=filename+'results', mode='r') as f:
+            # count lines of results file
+            num_prots = sum(1 for _ in f)
+
+        with open(file=filename + 'results', mode='r') as f:
+            for line in f:
+                prot, labels, predictions = line.strip().split('\t')
+                labels = np.array([int(float(i)) for i in labels[1:-1].split(',')])
+                predictions = np.array([round(float(i)) for i in predictions[1:-1].split(',')])
+
[…]

### 2021-06-29 src/torch_dti_predictor.py

@@ -20,12 +20,7 @@ import argparse
 from torch_dti_utils import *
 from torch_networks import *
 
-from protein_function_utils import ProteinFunctionPredNet, ProteinFunctionDTIDataBuilder
 import dti_utils
-
-import DTI_data_preparation
-import PPI_utils
-import DDI_utils
 
 
 def quickened_missing_target_predictor(config,
@@ -105,15 +100,6 @@ def quickened_missing_target_predictor(config,
         print("model total parameters", sum(p.numel() for p in model.parameters()))
 
         optimizer = torch.optim.Adam(model.parameters(), lr=config.lr)#, momentum=0.9)
-
-        # storing best results
-        best_loss = math.inf
-        best_test_loss = math.inf
-        best_epoch = -1
-        best_test_ci = 0
-
-        model_st = 'transductive_simple_node_feature'
-        # model_file_name = '../models/'+model_st+'_'+config.node_features + '_'+ str(config.num_proteins)+'_fold_'+str(fold)+'_model.model'
 
         sys.stdout.flush()
 
@@ -197,7 +183,6 @@ if __name__ == '__main__':
     parser.add_argument("--arch", type=str, default='GCNConv')
     parser.add_argument("--node_features", type=str, default='MolPred')
 
-
     parser.add_argument("--num_epochs", type=int, default=3)
     parser.add_argument("--batch_size", type=int, default=64)
     parser.add_argument("--num_folds", type=int, default=5)
@@ -215,6 +200,8 @@ if __name__ == '__main__':
     parser.add_argument("--biosnap_test", action='store_true')
     parser.add_argument("--include_indications", action='store_true')
 
+    parser.add_argument("--ipro_class", type=str, default='')
+    parser.add_argument("--interaction_type", type=str, default='')
 
     config = parser.parse_args()
 

### 2021-06-29 src/torch_dti_utils.py

@@ -27,7 +27,7 @@ class QuickProtFuncDTINetworkData:
         if config.yamanishi_test or config.biosnap_test:
             self.drug_list = np.array(PhenomeNET_DL2vec_utils.get_PhenomeNET_drug_list())
         else:
-            self.drug_list = np.array(DTI_data_preparation.get_drug_list(config.mode))
+            self.drug_list = np.array(DTI_data_preparation.get_drug_list(config.mode, config.interaction_type))
         print(len(self.drug_list), "drugs present")
 
         # self.protein_list = np.array(DTI_data_preparation.get_human_prot_func_proteins())[:config.num_proteins]
@@ -84,8 +84,6 @@ class QuickProtFuncDTINetworkData:
 
         print(len(self.protein_list), "proteins present\n")
 
-
-
         # PPI data
         print("Loading PPI graph ...")
         self.PPI_graph = self.PPI_graph.subgraph(self.protein_list)
@@ -120,7 +118,7 @@ class QuickProtFuncDTINetworkData:
         # DTI data
         if not config.yamanishi_test:
             print("Loading DTI links ...")
-            y_dti_data = DTI_data_preparation.get_DTIs(drug_list=self.drug_list, protein_list=self.protein_list, mode=config.mode)
+            y_dti_data = DTI_data_preparation.get_DTIs(drug_list=self.drug_list, protein_list=self.protein_list, mode=config.mode, interaction_type=config.interaction_type)
             self.y_dti_data = y_dti_data.reshape((len(self.drug_list), len(self.protein_list)))
         print(self.y_dti_data.shape)
 

### 2021-06-29 src/torch_networks.py

@@ -443,7 +443,7 @@ class QuickTemplateNodeFeatureNet(torch.nn.Module):
 
         # mask feature
 
-        # GCN laye4s
+        # GCN layers
         if 'GCNConv' in conv_method:
             self.conv1 = nn.GCNConv(200, 200, cached=True, improved=True)
             self.conv2 = nn.GCNConv(200, 200, cached=True, improved=True)

### 2021-07-01 paper/main.tex
.

@@
 % made. 
-Identifying drug--target interactions (DTIs) is one of the initial steps in drug discovery and screening for disease treatment and thus helps to mitigate the the costs and risks of drug development.
-Further, it forms a generalization of both drug repurposing and repositioning, finding novel functions for already approved and extensively analyzed drugs with well explored properties.
-Inferring and deriving from those existing properties and interactions can help to analyze and identify novel drugs for known targets, leading to desirable therapeutic effects.
-Moreover, while \textit{in vitro} DTI prediction requires the availability of all substances, computational, \textit{in silico} DTI predictors interpret and anticipate millions of interactions within short time.
-
-Computational methods are
-widely applied to predict DTIs and many computational methods have
-been developed.  These methods can be broadly classified into
-``top-down'' and ``bottom-up'' approaches. Top-down approaches start
-from observable characteristics resulting from a drug--target
-interaction, such as side-effects or the diseases treated by a drug,
-and infer likely molecular mechanisms (i.e., the interaction) using
-these observations.
+Identifying drug--target interactions (DTIs) is a crucial step in drug
+discovery; finding novel DTIs for approved drugs can be used for drug
+repurposing, either by finding new drugs for a known target or finding
+a drug for a novel target involved in a disease process.  Inferring
+the interactions between drugs and their targets can help to analyze
+and identify potential desired or adverse drug effects as well as
+desirable therapeutic effects. While \textit{in vitro} DTI prediction
+is time consuming, computational \textit{in silico} DTI predictors can
+screen for millions of interactions within a short time.  Determining
+DTIs computationally can therefore help to mitigate the costs and
+risks of drug development.
+
+Computational methods are widely applied to predict DTIs and many
+computational methods have been developed.  These methods can be
+broadly classified into ``top-down'' and ``bottom-up''
+approaches. Top-down approaches start from observable characteristics
+resulting from a drug--target interaction, such as side-effects or the
+diseases treated by a drug, and infer likely molecular mechanisms
+(i.e., the interaction) using these observations.
 % and indications represented by knowledge graphs or ontologies, induced by a drug and infer targets based on the likely molecular mechanisms that result in these phenotypes
@@
 prediction bear some advantages and limitations. Generally, bottom-up
-methods face the challenge to predict whether a chemical structure binds to a
-protein given their molecular properties; whether two entities
-interact depends not only on the molecular structure of the entities
-(where binding sites and molecular forces need to be determined for
-accurate prediction) but also properties namely in which celltypes and anatomical structures a protein is expressed. Top-down
-methods use information for DTI prediction such as side-effect
-similarity \cite{Campillos2008} that is largely complementary to
-knowledge gained from molecular properties. While methods that rely on
-molecular information are directly predicting whether two molecules
-could interact, top-down methods base on more indirect means and 
-infer a DTI from observable effects resulting from the interaction.
+methods face the challenge to predict whether a chemical structure
+binds to a protein given their molecular properties; whether two
+entities interact depends not only on the molecular structure of the
+entities (where binding sites and molecular forces need to be
+determined for accurate prediction) but also properties such as in
+which celltypes and anatomical structures a protein is
+expressed. Top-down methods use information about physiological
+effects of drugs for DTI prediction, such as side-effect similarity
+\cite{Campillos2008}, that is largely complementary to knowledge
+gained from molecular properties. While methods that rely on molecular
+information are directly predicting whether two molecules could
+interact, top-down methods base on more indirect means and infer a DTI
+from observable effects resulting from the interaction.
 
@@
 
-We further investigate the influence of the more specific interaction types \textbf{inhibition} and \textit{activation} within STITCH database on the performance of \name{}.
-The results are summarized in Supplementary table 3, where we find no difference in comparison to the complete dataset.
+We further investigate the performance on the specific interaction
+types {\em inhibition} and {\em activation} that are available within
+the STITCH database. The results are summarized in Supplementary Table
+3; we find no difference between the performance on specific
+interaction types and  the complete dataset where we do not separate
+types of interaction.
 

### 2021-07-01 paper/main.tex
.

@@
 Another view on these two approaches is as direct and indirect ways to
-predict drug--target interactions. On one hand, molecular information can be used
-to directly determine whether two molecules (such as a drug and
-protein) have the ability to interact, whereas information about
-phenotypic consequences of a drug (drug effects) or disruption of a
-protein function can be used to indirectly suggest candidate
+predict drug--target interactions. On one hand, molecular information
+can be used to directly determine whether two molecules (such as a
+drug and protein) have the ability to interact, whereas information
+about phenotypic consequences of a drug (drug effects) or disruption
+of a protein function can be used to indirectly suggest candidate
 drug--target interactions. Molecular features will be specific to a
@@
 consequently, we expect these features to benefit more from including
-information about protein--protein interactions. Moreover, as the protein--protein 
-interaction relation is not transitive (protein $P_1$ interacting with $P_2$, 
-and $P_2$ relating with $P_3$ does not imply $P_1$ interacting
-or influencing $P_3$), we can only infer from the direct neighborhood of each protein within the PPI graph. 
-Our results (Table
+information about protein--protein interactions. Moreover, as the
+protein--protein interaction relation is not transitive (protein $P_1$
+interacting with $P_2$, and $P_2$ relating with $P_3$ does not imply
+$P_1$ interacting with $P_3$), we can only infer from the direct
+neighborhood of each protein within the PPI graph. Our results (Table
 \ref{tab:Results}) confirm the first hypothesis and demonstrate that
@@
 propagation over the interaction network.
-While the exploitation of ontology based features over PPI graphs was already examined in \citet{Trebacz2020} for gene expression analysis, our work is the first to investigate and apply this technique to DTI prediction.
-Further, \name{} is the first work to combine both ``bottom-up'' and ``top-down'' features.
 
@@
 features in the future.
+
+\name{} introduces a number of novel
+
+While the exploitation of ontology based features over PPI graphs was already examined in \citet{Trebacz2020} for gene expression analysis, our work is the first to investigate and apply this technique to DTI prediction.
+Further, \name{} is the first work to combine both ``bottom-up'' and ``top-down'' features.
 

### 2021-07-01 paper/main.tex
.

@@
 
-\name{} introduces a number of novel
-
-While the exploitation of ontology based features over PPI graphs was already examined in \citet{Trebacz2020} for gene expression analysis, our work is the first to investigate and apply this technique to DTI prediction.
-Further, \name{} is the first work to combine both ``bottom-up'' and ``top-down'' features.
+To the best of our knowledge, \name{} is the first DTI prediction
+model that propagates ontology-based features of protein interaction
+networks; while a similar approach of combining ontology embeddings
+with interaction networks has previously been used for analyzing gene
+expression \citet{Trebacz2020}, \name{} extends this method to DTI
+prediction. Further, \name{} is novel in that it explicitly integrates
+``bottom-up'' and ``top-down'' features using a graph representation
+of interactions between proteins; this property of \name{} relies on
+the ability to integrate heterogeneous features using graphs, and the
+ability to exploit this graph through the use of graph neural
+networks.
 

### 2021-07-01 paper/main.tex
.

@@
 ``bottom-up'' and ``top-down'' features using a graph representation
-of interactions between proteins; this property of \name{} relies on
-the ability to integrate heterogeneous features using graphs, and the
-ability to exploit this graph through the use of graph neural
-networks.
+of interactions between proteins; while there are other DTI prediction
+methods that combine these features \cite{Gottlieb2011}, \name{}
+exploits the ability to integrate heterogeneous features using graphs
+and the ability to utilize this graph in machine learning through the
+use of graph neural networks.
 

### 2021-07-01 paper/main.tex
.

@@
 partners is not desirable when applying the model to novel
-entities. We have demonstrated that even the newly proposed ``na\"ive'' classifier that
-makes predictions only by exploiting the imbalanced number of
-interaction partners can achieve performance close to state of the art
-methods (when measuring Macro AUC). When training a machine learning
-model on such imbalanced data, it will eventually overfit to this
-imbalance.  Splitting by entity (protein or drug) can reduce the
-impact of this spurious correlation but not reduce it entirely,
-because similar entities will still exhibit similar interaction
-patterns. In our experiments, we observed the impact of splitting
-training and evaluation sets by protein as a decrease in overall
-performance (Macro AUC), providing some evidence that models trained
-using this splitting scheme are less sensitive to overfitting to this
-type of bias.
+entities. We have demonstrated that even the newly proposed
+``na\"ive'' classifier that makes predictions only by exploiting the
+imbalanced number of interaction partners can achieve performance
+close to state of the art methods (when measuring Macro AUC). When
+training a machine learning model on such imbalanced data, it will
+eventually overfit to this imbalance.  Splitting by entity (protein or
+drug) can reduce the impact of these spurious correlations but not
+reduce it entirely, because similar entities will still exhibit
+similar interaction patterns. In our experiments, we observed the
+impact of splitting training and evaluation sets by protein as a
+decrease in overall performance (Macro AUC), providing some evidence
+that models trained using this splitting scheme are less sensitive to
+overfitting to this type of bias.
 
@@
   specific} protein (e.g., a protein involved in a disease) needs to
-be identified. We use an evaluation measure based on micro-averages
-per protein (Micro AUC$_p$) to evaluate this scenario, and we often
-find substantial differences in predictive performance when evaluating
-with Macro AUC and Micro AUC$_p$; generally, models that are trained
-using a split over drug--target pairs perform worse in Micro AUC$_p$
-than models that use a protein-based split, further providing evidence
-that a drug--target split results in overfitting to dataset biases.
+be identified. We introduce an evaluation measure based on
+micro-averages per protein (Micro AUC$_p$) to evaluate this scenario,
+and we often find substantial differences in predictive performance
+when evaluating with Macro AUC and Micro AUC$_p$; generally, models
+that are trained using a split over drug--target pairs perform worse
+in Micro AUC$_p$ than models that use a protein-based split, further
+providing evidence that a drug--target split results in overfitting to
+dataset biases.
 

### 2021-07-01 paper/main.tex
.

@@
 
+Several of the biases we identify in evaluating DTI prediction methods
+have been observed previously. The performance difference based on how
+training and evaluation data is split (by interaction pair, by drug,
+or by protein) has been demonstrated before using a
+\todo{MacroAUC/MicroAUC???} measure \cite{}; we further extend on
+these results by introducing performance measures based on micro
+averages (Micro AUC$_p$ and Micro AUC_$d$) to further illustrate how
+prediction performance changes when evaluation data is imbalanced. We
+have further extended on prior results by introducing a ``na\"ive''
+classifier that explicitly exploits one data bias to make predictions,
+illustrating that this bias has a significant impact on DTI
+methods. The bias we identify is similar to a previous bias found in
+gene networks and methods that rely on the guilt-by-association
+approach
+\cite{}\todo{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002444}
+but which has, to our knowledge, not been demonstrated in the context
+of DTI prediction.
+
 In summary, drug--target interaction prediction is not a single

### 2021-07-01 paper/main.tex
.

@@
 gene networks and methods that rely on the guilt-by-association
-approach
-\cite{}\todo{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002444}
-but which has, to our knowledge, not been demonstrated in the context
-of DTI prediction.
+approach \cite{10.1371/journal.pcbi.1002444} but which has, to our
+knowledge, not been demonstrated in the context of DTI prediction.
 

### 2021-07-01 paper/main.tex
.

@@
 illustrating that this bias has a significant impact on DTI
-methods. The bias we identify is similar to a previous bias found in
-gene networks and methods that rely on the guilt-by-association
-approach \cite{10.1371/journal.pcbi.1002444} but which has, to our
-knowledge, not been demonstrated in the context of DTI prediction.
+methods. The bias we identify with the ``na\"ive'' classifier is
+similar to a previous bias found in gene networks when using methods
+that rely on the guilt-by-association principle
+\cite{10.1371/journal.pcbi.1002444} but which has, to our knowledge,
+not been demonstrated in the context of DTI prediction.
 

### 2021-07-01 paper/main.tex
.

@@
 these results by introducing performance measures based on micro
-averages (Micro AUC$_p$ and Micro AUC_$d$) to further illustrate how
+averages (Micro AUC$_p$ and Micro AUC$_d$) to further illustrate how
 prediction performance changes when evaluation data is imbalanced. We

### 2021-07-04 paper/main.tex
.

@@
 with interaction networks has previously been used for analyzing gene
-expression \citet{Trebacz2020}, \name{} extends this method to DTI
+expression \cite{Trebacz2020}, \name{} extends this method to DTI
 prediction. Further, \name{} is novel in that it explicitly integrates
@@
 that rely on the guilt-by-association principle
-\cite{10.1371/journal.pcbi.1002444} but which has, to our knowledge,
+\cite{Gillis2012} but which has, to our knowledge,
 not been demonstrated in the context of DTI prediction.

### 2021-07-04 paper/main.tex
.

@@
 training affects the performance of the model. Similarly to previous
-work \cite{MolTrans2020, DeepConvDTI2019}, we find significant
+work \cite{MolTrans2020, DeepConvDTI2019}\todo{These seem wrong, they
+  only use a single splitting scheme -- should this be cite:DTI-CDF2019?}, we find significant
 differences in predictive performance across different splitting

### 2021-07-12 paper/main.tex
.

@@
 interacting with $P_2$, and $P_2$ relating with $P_3$ does not imply
-$P_1$ interacting with $P_3$), we can only infer from the direct
-neighborhood of each protein within the PPI graph. Our results (Table
-\ref{tab:Results}) confirm the first hypothesis and demonstrate that
-molecular features do not benefit from including the interaction
+$P_1$ interacting with $P_3$), we mainly transfer information to the
+direct neighborhood of each protein within the PPI graph. Our results
+(Table \ref{tab:Results}) confirm the first hypothesis and demonstrate
+that molecular features do not benefit from including the interaction
 network whereas the indirect, top-down features benefit from the
-propagation over the interaction network.
+propagation over the interaction network. While the graph we chose is
+based on interactions between proteins, other types of entities and
+interactions can also be chosen, in particular similarity networks
+\cite{Gottlieb2011}; in such networks, information between nodes
+may be transmitted differently than in \name{}.
 
@@
 % \end{itemize}
+
+\todo{TODO}Most of novel predictions made by \name{} are for antineoplastic
+agents targeting proteins belonging to the Zinc finger, LIM-type
+family ({\tt IPR001781}).
 

### 2021-07-12 paper/main.tex
.

@@
 propagation over the interaction network. While the graph we chose is
-based on interactions between proteins, other types of entities and
-interactions can also be chosen, in particular similarity networks
-\cite{Gottlieb2011}; in such networks, information between nodes
-may be transmitted differently than in \name{}.
+based on interactions between proteins and our results hold true for
+such a graph, other types of entities and interactions can also be
+chosen, in particular similarity networks \cite{Gottlieb2011}; in such
+networks, information between nodes may be transmitted differently
+than in \name{}.
 

### 2021-07-12 paper/main.tex
.

@@
 
-\todo{TODO}Most of novel predictions made by \name{} are for antineoplastic
-agents targeting proteins belonging to the Zinc finger, LIM-type
-family ({\tt IPR001781}).
-
 \subsection{The challenge of evaluating drug--target interaction
@@
 
-As the target-based predictive power of \name{} improves significantly over other methods, we utilized our model to predict novel drug classes for protein families.
-We therefore collected the 2nd level ATC (Anatomical Therapeutic Chemical) groups for each drug and all InterPro \cite{Interpro2020} top-level families for each protein.
-Utilizing the STITCH interactions, we followed a protein split within each InterPro family by predicting over all available drugs.
-We eventually normalized the number of novel interactions per class by the amount of proteins within the respective family.
-A heatmap showing the results of this analysis can be found in Supplementary figure 5.
-\dots \todo{Roberts class analysis}.
+As the target-based predictive power of \name{} improves significantly
+over other methods, we utilized our model to predict novel drug
+classes for protein families. We therefore collected the 2nd level ATC
+(Anatomical Therapeutic Chemical) groups for each drug and all
+InterPro \cite{Interpro2020} top-level families for each
+protein. Utilizing the STITCH interactions, we followed a protein
+split within each InterPro family by predicting over all available
+drugs. We eventually normalized the number of novel interactions per
+class by the amount of proteins within the respective family. A
+heatmap showing the results of this analysis can be found in
+Supplementary Figure 5. \dots \todo{Roberts class analysis}.
+
+Most of novel predictions made by \name{} are for antineoplastic
+agents targeting proteins belonging to the Zinc finger, LIM-type
+family ({\tt IPR001781}).
 

### 2021-07-13 paper/supplement/supplement.tex
.
Update heatmap and distribution figures and their caption in supplement

@@
 \clearpage
-\begin{figure}[ht]%figure1
-	\centerline{\includegraphics[width=1.1\textwidth]{../figures/ATC_Interpro_heatmap.png}}
-	\caption{Heatmap showing the results of the ATC drug class and InterPro family prediction analysis}
-	\label{fig:ATCInterproHeatmap}
-\end{figure}
-
-\clearpage
 \begin{table}[ht]
@@
 
+\clearpage
+\begin{figure}[ht]%figure1
+	\centerline{\includegraphics[width=1.1\textwidth]{../figures/ATC_Interpro_heatmap.png}}
+	\caption{Heatmap showing the results of the ATC drug class and InterPro family prediction analysis.
+The figure shows the proportion of false negatives compared to the total amount of drugs in the corresponding level 2 ATC class.
+While few ATC classes show a high share of novel interactors, this is mainly due to the total amount of drugs in the ATC class.
+For more information note the correspondence with Supplementary Figure 6.}
+	\label{fig:ATCInterproHeatmap}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[ht]%figure1
+	\centerline{\includegraphics[width=1\textwidth]{../figures/ATC_bar_distr.png}}
+	\caption{Amount of drugs used in the ATC/InterPro analysis for each level 2 ATC class. Notably, some of the classes have only few or single drug members.}
+	\label{fig:ATCdistribution}
+\end{figure}
+
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2021-07-13 src/results_evaluation.py

@@ -15,17 +15,32 @@ from sklearn.model_selection import KFold
 
 
 def write_predicted_DTIs(fold=3):
+    '''
     filename = '../models/graph_models/PPI_network_model_with_mol_features_fold_'+str(fold)+'_predictions.pkl'
+
+    pred_list = None #[(drug, prot, label, pred)]
+    with open(file=filename, mode='rb') as f:
+        pred_list = pickle.load(f)
+    '''
+
+    pred_list = []
+    for i in range(1,6):
+        print(f'Parsing fold {str(i)}')
+        path = '../results/full_model_STITCH_preds/all_'
+        filename = path + 'results_fold_' + str(i)
+
+        with open(file=filename, mode='r') as f:
+            for line in f:
+                drug, prot, val = line.strip().split('\t')
+                pred_list.append((drug,prot,float(val)))
+
+    print('num triples:', len(pred_list))
 
     drug_mapping = DTI_data_preparation.get_drug_to_name_mapping()
     protein_mapping = DTI_data_preparation.get_protein_to_EnsemblProtein_id()
 
     print('drug_dict.keys()', len(list(drug_mapping.keys())))
     print('protein_dict.keys()', len(list(protein_mapping.keys())))
-
-    pred_list = None #[(drug, prot, label, pred)]
-    with open(file=filename, mode='rb') as f:
-        pred_list = pickle.load(f)
 
     drug_list = [tup[0] for tup in pred_list]
     protein_list = [tup[1] for tup in pred_list]
@@ -42,12 +57,13 @@ def write_predicted_DTIs(fold=3):
     # print('sanity_check', pos_sanity_list[:200])
 
 
-    pred_list = [tup for tup in pred_list if round(tup[2])==0 and tup[3] > 0.7]
+    pred_list = [tup for tup in pred_list if tup[2] > 0.7]
 
-    pred_list = sorted(pred_list, key=lambda tup: tup[3], reverse=True)
+    pred_list = sorted(pred_list, key=lambda tup: tup[2], reverse=True)
     print('len(pred_list):', len(pred_list))
 
     drug_list = list({drug for drug in drug_list if drug in drug_mapping.keys()})
+
     protein_list = list({protein for protein in protein_list if protein in protein_mapping.keys()})
 
     filename = '../data/PPI_data/protein_to_gene_dict.pkl'
@@ -78,18 +94,18 @@ def write_predicted_DTIs(fold=3):
 
     print('Num proteins that are driver genes:', len(protein_list))
 
[…]

### 2021-07-13 paper/main.tex
.

@@
 heatmap showing the results of this analysis can be found in
-Supplementary Figure 5. \dots \todo{Roberts class analysis}.
-
-Most of novel predictions made by \name{} are for antineoplastic
-agents targeting proteins belonging to the Zinc finger, LIM-type
-family ({\tt IPR001781}).
+Supplementary Figure 5. \name{} predicts novel candidate drug--target
+interactions for a broad range of ATC categories as well as protein
+families.
+
+For example, ATC group A07 (Antidiarrheals) has relatively few
+approved drugs in total (Supplementary Figure 6), but \name{} predicts
+several candidate targets from proteins with a PHD-type zinc finger
+domain ({\tt IPR001965}). For example, the drug mesalazine, used to
+treat inflammatory bowel disease but with an apoptosis-inducing and
+chemopreventative effect in colon cancer \cite{Bus1999, Ishikawa2021};
+\name{} predicts mesalazine to interact with five proteins with
+PHD-type zinc finger domain: BRPF1, TRIM33, BAZ1A, RSF1, and
+DPF2. Overexpression of RSF1 is associated with poor prognosis in
+colorectal cancer, and knock-down of RSF1 leads to decrease of cell
+proliferation \cite{Liu2012}, indicating that, in addition to its
+antiinflammatory effects, mesalazine may act through inhibition of
+RSF1 in its chemopreventative effects on colon cancer.
 

### 2021-07-13 paper/main.tex
.
Fixed TODOs with suitable citations

--- 2021-07-12 paper/main.tex
@@
 training affects the performance of the model. Similarly to previous
-work \cite{MolTrans2020, DeepConvDTI2019}\todo{These seem wrong, they
-  only use a single splitting scheme -- should this be cite:DTI-CDF2019?}, we find significant
+work \cite{MolTrans2020, DeepConvDTI2019, DTI-CDF2019}, we find significant
 differences in predictive performance across different splitting
@@
 training and evaluation data is split (by interaction pair, by drug,
-or by protein) has been demonstrated before using a
-\todo{MacroAUC/MicroAUC???} measure \cite{}; we further extend on
+or by protein) has been demonstrated before using the MacroAUC measure \cite{MolTrans2020, Pahikkala2014, vanLaarhoven2014}; we further extend on
 these results by introducing performance measures based on micro

### 2021-07-13 paper/main.tex

@@
 heatmap showing the results of this analysis can be found in
-Supplementary Figure 5. \dots \todo{Roberts class analysis}.
-
-Most of novel predictions made by \name{} are for antineoplastic
-agents targeting proteins belonging to the Zinc finger, LIM-type
-family ({\tt IPR001781}).
+Supplementary Figure 5. \name{} predicts novel candidate drug--target
+interactions for a broad range of ATC categories as well as protein
+families.
+
+For example, ATC group A07 (Antidiarrheals) has relatively few
+approved drugs in total (Supplementary Figure 6), but \name{} predicts
+several candidate targets from proteins with a PHD-type zinc finger
+domain ({\tt IPR001965}). For example, the drug mesalazine, used to
+treat inflammatory bowel disease but with an apoptosis-inducing and
+chemopreventative effect in colon cancer \cite{Bus1999, Ishikawa2021};
+\name{} predicts mesalazine to interact with five proteins with
+PHD-type zinc finger domain: BRPF1, TRIM33, BAZ1A, RSF1, and
+DPF2. Overexpression of RSF1 is associated with poor prognosis in
+colorectal cancer, and knock-down of RSF1 leads to decrease of cell
+proliferation \cite{Liu2012}, indicating that, in addition to its
+antiinflammatory effects, mesalazine may act through inhibition of
+RSF1 in its chemopreventative effects on colon cancer. We make all
+predictions including the ATC class of the drug and the Interpro
+family of the predicted target available on our website to allow
+further exploration of \name{}'s prediction results.
 
